### 2021-06-18 graphs_newt_snakes/Expected_results_fig.Rmd
updated help flie and added a place to make newt snake figures

---
title: "Fake Graphs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(viridis)
setwd("~/Desktop/Proposal Old Drafts")
```

## This R Markdown is for my expected results Figs

#### Aim 1

Genetic Architecture by Resistance - Toxicity
-concerns that it might be messy might be useful to do trait A and trait B

```{r, aim 1 Genetic Architecture}
GA <- c("A", "B", "C", "D")
# Group A Large effect – High Mutation 
# Group B Large effect – Low Mutation
# Group C Small effect – High Mutation
# Group D Small effect – Low Mutation
my_GA_data <- as.data.frame(GA)
my_GA_data$fake <-c(3,1,0,-2)

#pdf("Resistance_of_Populations_as_the_Genetic_Architecture_of_Snake_Changes.pdf")
ggplot(my_GA_data, aes(GA, fake))+
  geom_hline(yintercept=0, color ="red")+
  geom_point(size=3, color = "darkgreen")+
  ggtitle("Resistance of Populations as the Genetic Architecture of Snake Changes") + 
  xlab("Genetic Architecture of Snake") + 
  ylab("Resistance-Toxicity") +
  theme_bw()

#pdf("Genetic_Architecture_Map.pdf", width=4, height=2.9)
test_grid <- expand.grid(X = GA, Y = GA)
test_grid$Z <- c(0, -1, -2, -3, 1, 0, -1, -2, 2, 1, 0, -1, 3, 2, 1, 0)
ggplot(test_grid, aes(X, Y, fill= Z)) + 
  geom_tile()+
  scale_y_discrete(limits = rev(levels(test_grid$Y)))+
  ggtitle("Genetic Architecture Map") + 
  xlab("Genetic Architecture of Resistance") + 
  ylab("Genetic Architecture of Toxicity") +
  scale_fill_viridis(name = "Resistance-Toxicity")+
  theme_bw()

```

Density of the Newt increases selection pressure

```{r, aim 1 Density}

newt_density <-  as.data.frame(n_d<-c(1:10))

pdf("Interaction_Strength_as_Newt_Density_Increases.pdf")
ggplot(newt_density, aes(n_d, n_d))+
  geom_point(size=3, color = "darkred")+
  ggtitle("Interaction Strength as Newt Density Increases") + 
  xlab("Newt Density") + 
  ylab("Interaction Strength") +
  theme_bw()

```

Spatial component. the mismatch will decrease as spatial structure increases

```{r, aim 1 spatial aspect}
spatial_structure <- runif(1000, min = 0.001, max = 5) # this is sigma
mismatch <- (exp(-(spatial_structure/2)))
sigma_effects <-  as.data.frame(spatial_structure, mismatch)
pdf("SigmaCurve.pdf")
ggplot(sigma_effects, aes(spatial_structure, mismatch))+
  geom_point(color = "green")+
  ggtitle("Sigma Curve") + 
  xlab("Spatial_structure ") + 
  ylab("Mismatch") +
  theme_bw()

#well mixed when sigma is large

```

Simulation on how likely the snake or new will die depending on their phenotype

```{r, aim 1 Probablity of Death}
Resistance <- c(0:5)
Toxicity <- c(0:3)
E <- 0.1
c <- 0.0
w <- 10
map <- expand.grid(Res = seq(min(Resistance), max(Resistance), length.out = 100), Tox = seq(min(Toxicity), max(Toxicity), length.out = 100))
map$r_t <- map$Res-map$Tox
map$outcome <- 1/(1+exp(-(map$r_t-c)/w))

test <- c(0, 0.0001, 0.001, 0.01, 0.1, 1, 10, 100, -0.0001, -0.001, -0.01, -0.1, -1, -10, -100)

test2 <- runif(1000, min = -100, max = 100)
survival_df <- as.data.frame(test2)
survival_df$outcome2 <- 1/(1+exp(-((survival_df$test2-c)/w)))
pdf("SurvivalCurve.pdf", width=2.5, height=2.5)
ggplot(survival_df, aes(test2, outcome2))+
  geom_point(color = "darkred")+
  ggtitle("Survival Curve") + 
  xlab("Snake-Newt") + 
  ylab("Survival") +
  theme_bw()

```

### Aim 2

Diversity! Trait and whole genome 
large/small dispersal 
interaction/ no interaction

```{r geographical distance vs gentic distance}
#dispersal <- runif(1000, min = 0, max = 3) # this is a range for large and small dispersal 
distance <- seq(0,35,3)  # this is distances that might be probable 
diversity_in_general <- as.data.frame(sort(distance))
names(diversity_in_general)<- "distance"
diversity_in_general$Genetic_i <- diversity_in_general$distance * 1/35 # finding Genetic Distance
diversity_in_general$Genetic_in <- diversity_in_general$distance * 1/70 + 0.05# finding Genetic Distance
diversity_in_general$Genetic_inn <- diversity_in_general$distance * 1/185 + 0.15 # finding Genetic Distance
# y = mx + b y ranges from 0 to 1
diversity_in_general$Genetic_no_interact <- diversity_in_general$distance * 1/60 
diversity_in_general$Genetic_no_interactn <- diversity_in_general$distance * 1/360 + 0.20
diversity_in_general$Genetic_no_interactnn <- 0.5

pdf("General_Diversity_Trend_For_Interactaction.pdf", width=4, height=2.9)
ggplot(diversity_in_general, aes(distance, Genetic_i, color = "darkred" ))+
  geom_point(size=5)+
  geom_point(data = diversity_in_general, aes(distance, Genetic_in, color = "darkgreen"), size=5)+
  geom_point(data = diversity_in_general, aes(distance, Genetic_inn, color = "purple"), size=5)+
  ggtitle("General Diversity Trend For Interactaction") + 
  xlab("Geographical Distance") + 
  ylab("Genetic Distance") +
  scale_color_identity(name = "Dispersal",
                          breaks = c("darkred", "darkgreen", "purple"),
                          labels = c("Small", "Medium", "Large"),
                          guide = "legend")+
  theme_bw()

pdf("General_Diversity_Trend_For_No_Interactaction.pdf")
ggplot(diversity_in_general, aes(distance, Genetic_no_interact, color = "darkred"))+
  geom_point( size=5)+
  geom_point(data = diversity_in_general, aes(distance, Genetic_no_interactn, color = "darkgreen"), size=5)+
  geom_point(data = diversity_in_general, aes(distance, Genetic_no_interactnn, color = "purple"), size=5)+
  ylim(0,1)+
  ggtitle("General Diversity Trend For No Interactaction") + 
  xlab("Geographical Distance") + 
  ylab("Genetic Distance") +
  theme_bw()+
  scale_color_identity(name = "Dispersal",
                        breaks = c("darkred", "darkgreen", "purple"),
                        labels = c("Small", "Medium", "Large"),
                        guide = "legend")
```

```{r geographical distance vs  pi for trait and for whole genome (focus on the interactions)}
library(reshape)
dispersal_distance <- rep(seq(0.35, 1.75, 0.35) ,5)
dispersal_diversity <- as.data.frame(sort(dispersal_distance))
names(dispersal_diversity)<- "dispersal_distance"
# interaction cause diversity of the trait to be lower and the genome to be higher than the trait but lower then no interaction
# no_interact both trait and genome will have similar levels of genetic diversity
# pi values (nucleotide diversity) range from 0 to 1, 0 low diversity 1 high diversity
# as dispersal distance increases pi goes down (sweeps)
 # dispersal distance 0.35 to 5.25
# high .6 low .2
# y = mx + b

exp(unique(-dispersal_distance))

trait_pi_i <- c()
whole_genome_pi_i <- c()
trait_pi_n <- c()
whole_genome_pi_n <- c()
for(i in unique(dispersal_distance)){
  trait_pi_i <- c(trait_pi_i, exp(-i) + c(0.05, 0.02, 0, -0.02, -0.05))
 whole_genome_pi_i <- c(whole_genome_pi_i, exp(-i) + c(0.1, 0.07, 0.05, 0.03, 0)- exp(-i)*.2)
 trait_pi_n <- c(trait_pi_n, exp(-i) + c(0.01, -0.01, -0.03, -0.05, -0.06))
whole_genome_pi_n <- c(whole_genome_pi_n, exp(-i) + c(0.01, -0.01, -0.03, -0.05, -0.06))
}

dispersal_diversity$trait_pi_i <-trait_pi_i
dispersal_diversity$whole_genome_pi_i <-whole_genome_pi_i
dispersal_diversity$trait_pi_n <- trait_pi_n 
dispersal_diversity$whole_genome_pi_n <- whole_genome_pi_n

dispersal_diversity_t<-t(dispersal_diversity)
dispersal_diversity_m <- melt(dispersal_diversity, id=c("dispersal_distance"))

pdf("dispersal_diversit.pdf", width=4.7, height=2.7)

ggplot(dispersal_diversity_m, aes(factor(dispersal_distance), value, fill=variable))+
  geom_boxplot(data=dispersal_diversity_m, aes(factor(dispersal_distance), value, fill=variable))+
  ggtitle("Levels of Genetic Diversity and Dispersal Distance") + 
  xlab("Dispersal Distance") + 
  ylab("Diversity (pi)") +
  theme_bw()+
  scale_fill_manual(name = "Legend", labels = c("Interaction/Trait", "Interaction/Genome", "No Interaction/Trait", "No Interaction/Genome"), values = c("lightblue", "steelblue", "greenyellow", "forestgreen") )
  #scale_fill_viridis_d(option="plasma", name = "Group", labels = c("Trait/Interaction", "Genome/Interaction", "Trait/No Interaction", "Genome/No Interaction"))
# position = position_dodge(-0.3)

```

```{r local vs global diversity levels}
large_dis
small_dis
interact
no_interact
trait 
whole_genome

```

### Aim 3

How convolution effects the genome 
machine learning ~ neural network

```{r manhatten plot or like gwas?}

```