### 2021-08-03 graphs_newt_snakes/find_params.Rmd
I added a legend to the find param r markdown (today I will be adding more explination). I am working ou the the population size problem in nospace slim. How many of the newts interact with the snakes?

---
title: "Finding Sigma and Mu"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(viridis)
library(colorspace)
library(gridExtra)

#setwd("~/Desktop/Proposal Old Drafts")
```

## This is a r file to explore paramater values for my coevolution simulation

I ran a few experiments in the 1000 generation simulation and discovered that there were some combinations of sigma and mu that created "extreme" phenotypes (but not outside of the mathematical expectations). This markdown will hopefully help me visualize different values of sigma and mu and their resulting phenotypes.

### Math

$$
\text{Population size, } N = 10^4 
$$
$$
\text{Time to most recent comman ancestor, } 2N 
$$
$$
\text{Number of Mutations differ between two, } \sim 4N\mu L
$$
$$
\mu = \text{mutation rate / bp, My simulation } 10^-12 \text{ to } 10^-8
$$
$$
L = \text{sequence length }
$$
$$
\sigma = \text{effect size SD, My simulation } 0.01 \text{ to } 0.5
$$
$$
Var[phenotype] \sim 4N\mu L\sigma^2
$$
$$
SD[phenotype] \sim \sigma \sqrt{4N\mu L} -> \sigma \sqrt{\mu}*2*10^6
$$
$$
\mu = 10^-12 \sigma = 0.01, = 0.02
$$
$$
\mu = 10^-8 \sigma = 0.5, = 100
$$

# Different Combinations

Here I am using different values with the above formula to try to get the phenotypes of 5 and 2

```{r find sigma and mu, echo=FALSE}

sigma = seq(0.01, 0.5, length.out=100)
mu = seq(10^-12, 10^-8, length.out=100)

mat = mapply(rep, 1:4, 4:1)
my_grid <- as.data.frame(expand.grid(sigma, mu))
my_grid$results = mapply(`*`,sigma*2*10^6,rep(sqrt(mu),each=length(sigma)))

ggplot(my_grid, aes(x = Var1, y=Var2, colour = results)) +
  geom_point(data=my_grid[my_grid$results <=5, ], aes(x = Var1, y=Var2, colour="Less than 5"))+
  geom_point(data=my_grid[my_grid$results <=2, ], aes(x = Var1, y=Var2, colour="Less than 2"))+
  ggtitle("Value Combinations That Might Work") +
  xlab("Sigma") + ylab("Mu")+  
  theme_bw()+
  scale_color_manual(name = "Fake Phenotypes", values = c("Less than 5" = "blue", "Less than 2" = "green"))

```

I think it would be cool to draw them as lines (like take the max Mu that would produce the correct phenotype for every sigma value)

```{r max unique values, echo=FALSE}

#2=sigma*sqrt(mu)*2*10^6
y_mu_2 = (2/(sigma*2*10^6))^2
y_mu_5 = (5/(sigma*2*10^6))^2
x_sigma_2 = 2/(sqrt(mu)*2*10^6)
x_sigma_5 = 5/(sqrt(mu)*2*10^6)
ggplot() +
  geom_line(aes(x = sigma, y=y_mu_5, colour="Less than 5"))+
  geom_line(aes(x = sigma, y=y_mu_2, colour="Less than 2"))+
  #lines are out of order, but the points line up. Looks Good!
  #geom_point(aes(x = x_sigma_5, y=mu), col="blue")+
  #geom_point(aes(x = x_sigma_2, y=mu), col="green")+
  scale_y_log10()+
  #geom_point(data=my_grid[my_grid$results <=2, ], aes(x = Var1, y=Var2), col="green")+
  ggtitle("Value Combinations That Might Work") +
  xlab("Sigma") + ylab("Mu (log)")+ #xlim(0, 0.6)+
  theme_bw()+
  scale_color_manual(name = "Fake Phenotypes", values = c("Less than 5" = "blue", "Less than 2" = "green"))

```