### 2021-08-11 graphs_newt_snakes/narmal_sim.Rmd
working on making figures to see what a normal simulation looks like. Added a bunch of new plots that compare the simulations with the same params to each other

---
title: "What is a Normal Simulation Like"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(viridis)
library(colorspace)
library(gridExtra)
library(tidyverse)
require(reshape2)

#setwd("~/Desktop/Proposal Old Drafts")
```

# Goals

I will try to discern what a typical simulation looks like. I attempt to do this by looking at many simulations under the same parameters. There are 110 simulations for a combination. I don't think 100 repeated simulations will cover all the possible outcomes, but it will be a good beginning.

The first set of simulations I will look at parameters of sigma = 0.2, mu = 1.5625e-10. I will look at the means and standard deviation for each of the simulations over the 1000 generations. I am going to specifically look at the similarities and differences between each of theses simulations. I predict that they  will look similar at the beginning and vary as the amount of generations increases.

Newts and  snakes should evenly have the larger phenotype (be about 50-50).

- collect the means, highs, lows, and sds for the phenotypes of newts and snakes

thinking of making a data frame for each thing to look at columns -> generation, rows -> trials

```{r, making data frames}

making_data_frames <- function(files, spe_col) {
i=1
for (current_file in files){
  #print(current_file)
  if(i ==1){
    file <- read.table(current_file, header = TRUE)

    file[[spe_col]][file[[spe_col]] == "NULL"] <- NA
      # might want to change to all_of(spe_col)
    file %>% 
      select(spe_col, gen) %>%
      #rownames_to_column() %>% 
      gather(variable, value, -gen) %>% 
      spread(gen, value) -> new_file

    new_file$variable[i] <- paste0(spe_col,"_trial_", i)
    i = i + 1
    next
    }
  file <- read.table(current_file, header = TRUE)

  file[[spe_col]][file[[spe_col]] == "NULL"] <- NA

file %>% 
  select(spe_col, gen) %>%
  #rownames_to_column() %>% 
  gather(variable, value, -gen) %>% 
  spread(gen, value) -> temp_file

temp_file$variable[1] <- paste0(spe_col,"_trial_", i)

new_file<-rbind(new_file, temp_file)
i = i + 1
}
return(new_file)
}

```

```{r, making figures}
phenotype_boxplot <- function(Newt_file, Snake_file, title, subt, boxy=FALSE){
  data_frame_newt_temp <- Newt_file
  data_frame_newt_temp <- data_frame_newt_temp[ , !(names(data_frame_newt_temp) %in% "variable")]

  data_frame_snake_temp <- Snake_file
  data_frame_snake_temp <- data_frame_snake_temp[ , !(names(data_frame_snake_temp) %in% "variable")]

  Number <- colnames(data_frame_newt_temp)
  data_frame_newt <- gather(data_frame_newt_temp, Number, value)
  data_frame_newt$type <- "Newt"
  data_frame_snake <- gather(data_frame_snake_temp, Number, value)
  data_frame_snake$type <- "Snake"

 combo <- rbind(data_frame_newt, data_frame_snake)

 small_combo <- aggregate(as.numeric(combo$value),by=list(Number=as.numeric(combo$Number),type=combo$type),data=combo,FUN=mean, na.action = TRUE)

if(boxy==TRUE){
  p <- ggplot(combo) +
  #xlim(0, 1000)+
    ggtitle(title, subtitle = subt) +
    xlab("Generation") + ylab("Average Phenotypes")+
    scale_color_manual(name = "Species", values = c("Newt" = "darkred", "Snake" = "steelblue"))+
    theme(plot.title = element_text(size=5)) +
    theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))+
    geom_boxplot(aes(x=factor(as.numeric(Number)), y=as.numeric(value), color=type) )
  }
if(boxy==FALSE){
  p <- ggplot(small_combo) +
  #xlim(0, 1000)+
    ggtitle(title, subtitle = subt) +
    xlab("Generation") + ylab("Average Phenotypes")+
    scale_color_manual(name = "Species", values = c("Newt" = "darkred", "Snake" = "steelblue"))+
    theme(plot.title = element_text(size=5)) +
    theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))+
    geom_point(aes(x=factor(as.numeric(Number)), y=as.numeric(x), color=type) )
}

return(p)
}

starting_end_comparison <- function(Newt_file, Snake_file, title, subt, f_point, l_point){

  test_newt_mean<-dplyr::select(Newt_file, c(f_point,l_point))
  test_newt_mean %>% 
  rename(
    "newt_start" = f_point,
    "newt_end" = l_point
    ) -> test_newt_mean

  test_snake_mean<-dplyr::select(Snake_file, c(f_point, l_point))
  test_snake_mean %>% 
  rename(
    "snake_start" = f_point,
    "snake_end" = l_point
    ) -> test_snake_mean
  all_points <- cbind(test_newt_mean, test_snake_mean)
  newt_adv_beg<- as.numeric(all_points$newt_start)- as.numeric(all_points$snake_start) >0
  newt_adv_end<- as.numeric(all_points$newt_end)- as.numeric(all_points$snake_end) >0

  data_frame_all_points<- data.frame(newt_adv_beg, newt_adv_end)
  data_frame_all_points <- gather(data_frame_all_points, )

  all_points.count<-as.data.frame(table(data_frame_all_points))

  ggplot(all_points.count, aes(factor(key), (Freq/ nrow(test_newt_mean)), fill = value)) + geom_bar(stat="identity", position = "dodge") +
    ggtitle(title, subtitle = subt) +
    scale_fill_manual(name = "Species", values = c('TRUE' = "darkred", 'FALSE' = "steelblue"), labels=c("Snakes", "Newts"))+
    xlab("Counts") + ylab("Frequency")+ scale_x_discrete(labels=c("Begining", "Ending"))+
    theme_bw()

}

phenotype_trends <- function(file, title, subt, my_color, is_log=FALSE){
  data_frame_temp <- as.data.frame(t(file))

  colnames(data_frame_temp) <- data_frame_temp[1,]
  names(data_frame_temp) <- colnames(data_frame_temp)
  steps <- colnames(file[ , !(names(file) %in% "variable")])
data_frame_temp <- data_frame_temp[-1, ] 
data_frame_temp$steps <- as.numeric(unlist(steps))

  graph_file <- melt(data_frame_temp,  id.vars = 'steps', variable.name = 'series')

    p <- ggplot(graph_file, aes(steps, as.numeric(value))) + geom_line(aes(colour = series))+ 
    ggtitle(title, subtitle = subt) +
    scale_color_manual(values = rep(my_color, nrow(file)))+
    xlab("Generations") + ylab("Phenotype")+
    theme_bw() + theme(legend.position="none")

    if(is_log==TRUE){
  p = p + scale_y_log10() + ylab("Phenotype (Log)")
    }
    return(p)
}

phenotype_dif_boxplot <- function(Newt_file, Snake_file, title, subt, boxy=FALSE){
  data_frame_newt_temp <- Newt_file
  data_frame_newt_temp <- data_frame_newt_temp[ , !(names(data_frame_newt_temp) %in% "variable")]

  data_frame_snake_temp <- Snake_file
  data_frame_snake_temp <- data_frame_snake_temp[ , !(names(data_frame_snake_temp) %in% "variable")]

  Number <- colnames(data_frame_newt_temp)
  data_frame_newt <- gather(data_frame_newt_temp, Number, value)
  data_frame_snake <- gather(data_frame_snake_temp, Number, value)
  dif <- as.numeric(data_frame_snake$value)-as.numeric(data_frame_newt$value)

  df_dif <- as.data.frame(dif)
  df_dif$Number <- as.numeric(data_frame_newt$Number)
  df_dif$trial <- rep(Newt_mean$variable, ncol(Newt_mean)-1)
  if(boxy==TRUE){
 p <- ggplot(df_dif) +
    ggtitle(title, subtitle = subt) +
    xlab("Generation") + ylab("Average Phenotype Differences")+
    theme(plot.title = element_text(size=5)) +
    theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))+
    geom_boxplot(aes(x=factor(as.numeric(Number)), y=as.numeric(dif)), color="darkolivegreen" )
 }
 if(boxy==FALSE){
   p <- ggplot(df_dif) +
    ggtitle(title, subtitle = subt) +
    xlab("Generation") + ylab("Average Phenotype Differences")+
    theme(plot.title = element_text(size=5)) +
    theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))+
    geom_line(aes(x=as.numeric(Number), y=as.numeric(dif), color=trial ))+ 
     theme(legend.position="none")+
     scale_color_manual(values = rep("darkolivegreen", nrow(Newt_file)))

 }
  return(p)
}

```

```{r, data collection}

files <- list.files(path="~/Desktop/findparams_u5", pattern="nu_*1.56", full.names=TRUE, recursive=FALSE)

file <- read.table(files[1], header = TRUE)

Newt_mean <- making_data_frames(files, all_of("Newt_mean_Pheno"))
Newt_min <- making_data_frames(files, all_of("Newt_min_Pheno"))
Newt_max <- making_data_frames(files, all_of("Newt_max_Pheno"))
Newt_sd <- making_data_frames(files, all_of("Newt_sd_Pheno"))
Snake_mean <- making_data_frames(files, all_of("Snake_mean_Pheno"))
Snake_min <- making_data_frames(files, all_of("Snake_min_Pheno"))
Snake_max <- making_data_frames(files, all_of("Snake_max_Pheno"))
Snake_sd <- making_data_frames(files, all_of("Snake_sd_Pheno"))
```

```{r, comparative figures for sigma = 0.2, mu = 1.5625e-10}
sp<-phenotype_boxplot(Newt_mean, Snake_mean, title="Variation in the Mean Phenotype For Newts and Snakes by Generation", subt="sigma = 0.2, mu = 1.5625e-10", boxy=TRUE)
sp

sdp<-phenotype_boxplot(Newt_sd, Snake_sd, title="Variation in the Sd of Phenotype For Newts and Snakes by Generation", subt="sigma = 0.2, mu = 1.5625e-10", boxy=TRUE)
sdp

sp2<-phenotype_boxplot(Newt_mean, Snake_mean, title="Averaging the Mean Phenotype For Newts and Snakes by Generation", subt="sigma = 0.2, mu = 1.5625e-10", boxy=FALSE)
sp2

sdp2<-phenotype_boxplot(Newt_sd, Snake_sd, title="Averaging the Sd of Phenotype For Newts and Snakes by Generation", subt="sigma = 0.2, mu = 1.5625e-10", boxy=FALSE)
sdp2

sec<-starting_end_comparison(Newt_mean, Snake_mean, title="Comparison of Starting and Ending Newt and Snake Phenotypes (Larger, Mean)", subt="sigma = 0.2, mu = 1.5625e-10", f_point="1", l_point="1000")
sec

sdsec<-starting_end_comparison(Newt_sd, Snake_sd, title="Comparison of Starting and Ending Newt and Snake Phenotypes (Larger, SD)", subt="sigma = 0.2, mu = 1.5625e-10", f_point="1", l_point="1000")
sdsec

n<-phenotype_trends(Newt_mean, title="Newt Average Phenotype Trends By Generations", subt="sigma = 0.2, mu = 1.5625e-10", my_color="darkred", is_log=FALSE)
n

s<-phenotype_trends(Snake_mean, title="Snake Average Phenotype Trends By Generations", subt="sigma = 0.2, mu = 1.5625e-10", my_color="Steelblue", is_log=FALSE)
s

bp <- phenotype_dif_boxplot(Newt_mean, Snake_mean, title="Differences", subt="sigma = 0.2, mu = 1.5625e-10",boxy=TRUE)
bp

bp2 <- phenotype_dif_boxplot(Newt_mean, Snake_mean, title="Differences", subt="sigma = 0.2, mu = 1.5625e-10",boxy=FALSE)
bp2
```