### 2021-08-11 graphs_newt_snakes/narmal_sim.Rmd
worked on the what is normal r markdown and talked about the results with Peter. Very tired today and have lots of new things to do. I really wish it would cool down.

---
title: "What is a Normal Simulation Like"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(viridis)
library(colorspace)
library(gridExtra)
library(tidyverse)
require(reshape2)

#setwd("~/Desktop/Proposal Old Drafts")
```

# Goals

I will try to discern what a typical simulation looks like. I attempt to do this by looking at many simulations under the same parameters. There are 110 simulations for a combination. I don't think 100 repeated simulations will cover all the possible outcomes, but it will be a good beginning.

The first set of simulations I will look at parameters of sigma = 0.2, mu = 1.5625e-10. I will look at the means and standard deviation for each of the simulations over the 1000 generations. I am going to specifically look at the similarities and differences between each of theses simulations. I predict that they  will look similar at the beginning and vary as the amount of generations increases.

Newts and  snakes should evenly have the larger phenotype (be about 50-50).

- collect the means, highs, lows, and sds for the phenotypes of newts and snakes

thinking of making a data frame for each thing to look at columns -> generation, rows -> trials

```{r, making data frames}

making_data_frames <- function(files, spe_col) {
i=1
for (current_file in files){
  #print(current_file)
  if(i ==1){
    file <- read.table(current_file, header = TRUE)

    file[[spe_col]][file[[spe_col]] == "NULL"] <- NA
      # might want to change to all_of(spe_col)
    file %>% 
      select(all_of(spe_col), gen) %>%
      #rownames_to_column() %>% 
      gather(variable, value, -gen) %>% 
      spread(gen, value) -> new_file

    new_file$variable[i] <- paste0(spe_col,"_trial_", i)
    i = i + 1
    next
    }
  file <- read.table(current_file, header = TRUE)

  file[[spe_col]][file[[spe_col]] == "NULL"] <- NA

file %>% 
  select(all_of(spe_col), gen) %>%
  #rownames_to_column() %>% 
  gather(variable, value, -gen) %>% 
  spread(gen, value) -> temp_file

temp_file$variable[1] <- paste0(spe_col,"_trial_", i)

new_file<-rbind(new_file, temp_file)
i = i + 1
}
return(new_file)
}

```

```{r, making figures}
phenotype_boxplot <- function(Newt_file, Snake_file, title, subt, boxy=FALSE, section=FALSE, section_num=0){
  data_frame_newt_temp <- Newt_file
  data_frame_newt_temp <- data_frame_newt_temp[ , !(names(data_frame_newt_temp) %in% "variable")]

  data_frame_snake_temp <- Snake_file
  data_frame_snake_temp <- data_frame_snake_temp[ , !(names(data_frame_snake_temp) %in% "variable")]

   if(section==TRUE){

   selcol = seq(1,ncol(data_frame_newt_temp), section_num)
   data_frame_newt_temp <- data_frame_newt_temp %>% select(all_of(selcol))
   data_frame_snake_temp <- data_frame_snake_temp %>% select(all_of(selcol))

 }

  Number <- colnames(data_frame_newt_temp)
  data_frame_newt <- gather(data_frame_newt_temp, Number, value)
  data_frame_newt$type <- "Newt"
  data_frame_snake <- gather(data_frame_snake_temp, Number, value)
  data_frame_snake$type <- "Snake"

 combo <- rbind(data_frame_newt, data_frame_snake)

 small_combo <- aggregate(as.numeric(combo$value),by=list(Number=as.numeric(combo$Number),type=combo$type),data=combo,FUN=mean, na.action = TRUE)

if(boxy==TRUE){
  p <- ggplot(combo) +
  #xlim(0, 1000)+
    ggtitle(title, subtitle = subt) +
    xlab("Generation") + ylab("Average Phenotypes")+
    scale_color_manual(name = "Species", values = c("Newt" = "darkred", "Snake" = "steelblue"))+
    theme(plot.title = element_text(size=5)) +
    theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))+
    geom_boxplot(aes(x=factor(as.numeric(Number)), y=as.numeric(value), color=type) )
  }
if(boxy==FALSE){
  p <- ggplot(small_combo) +
  #xlim(0, 1000)+
    ggtitle(title, subtitle = subt) +
    xlab("Generation") + ylab("Average Phenotypes")+
    scale_color_manual(name = "Species", values = c("Newt" = "darkred", "Snake" = "steelblue"))+
    theme(plot.title = element_text(size=5)) +
    theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))+
    geom_point(aes(x=factor(as.numeric(Number)), y=as.numeric(x), color=type) )
}

return(p)
}

starting_end_comparison <- function(Newt_file, Snake_file, title, subt, f_point, l_point){

  test_newt_mean<-dplyr::select(Newt_file, c(f_point,l_point))
  test_newt_mean %>% 
  rename(
    "newt_start" = all_of(f_point),
    "newt_end" = all_of(l_point)
    ) -> test_newt_mean

  test_snake_mean<-dplyr::select(Snake_file, c(f_point, l_point))
  test_snake_mean %>% 
  rename(
    "snake_start" = all_of(f_point),
    "snake_end" = all_of(l_point)
    ) -> test_snake_mean
  all_points <- cbind(test_newt_mean, test_snake_mean)
  newt_adv_beg<- as.numeric(all_points$newt_start)- as.numeric(all_points$snake_start) >0
  newt_adv_end<- as.numeric(all_points$newt_end)- as.numeric(all_points$snake_end) >0

  data_frame_all_points<- data.frame(newt_adv_beg, newt_adv_end)
  data_frame_all_points <- gather(data_frame_all_points, )

  all_points.count<-as.data.frame(table(data_frame_all_points))

  ggplot(all_points.count, aes(factor(key), (Freq/ nrow(test_newt_mean)), fill = value)) + geom_bar(stat="identity", position = "dodge") +
    ggtitle(title, subtitle = subt) +
    scale_fill_manual(name = "Species", values = c('TRUE' = "darkred", 'FALSE' = "steelblue"), labels=c("Snakes", "Newts"))+
    xlab("Counts") + ylab("Frequency")+ scale_x_discrete(labels=c("Beginning ", "Ending"))+
    theme_bw()

}

phenotype_trends <- function(file, title, subt, my_color, is_log=FALSE){
  data_frame_temp <- as.data.frame(t(file))

  colnames(data_frame_temp) <- data_frame_temp[1,]
  names(data_frame_temp) <- colnames(data_frame_temp)
  steps <- colnames(file[ , !(names(file) %in% "variable")])
data_frame_temp <- data_frame_temp[-1, ] 
data_frame_temp$steps <- as.numeric(unlist(steps))

  graph_file <- melt(data_frame_temp,  id.vars = 'steps', variable.name = 'series')

    p <- ggplot(graph_file, aes(steps, as.numeric(value))) + geom_line(aes(colour = series))+ 
    ggtitle(title, subtitle = subt) +
    scale_color_manual(values = rep(my_color, nrow(file)))+
    xlab("Generations") + ylab("Phenotype")+
    theme_bw() + theme(legend.position="none")

    if(is_log==TRUE){
  p = p + scale_y_log10() + ylab("Phenotype (Log)")
    }
    return(p)
}

phenotype_dif_boxplot <- function(Newt_file, Snake_file, title, subt, boxy=FALSE){
  data_frame_newt_temp <- Newt_file
  data_frame_newt_temp <- data_frame_newt_temp[ , !(names(data_frame_newt_temp) %in% "variable")]

  data_frame_snake_temp <- Snake_file
  data_frame_snake_temp <- data_frame_snake_temp[ , !(names(data_frame_snake_temp) %in% "variable")]

  Number <- colnames(data_frame_newt_temp)
  data_frame_newt <- gather(data_frame_newt_temp, Number, value)
  data_frame_snake <- gather(data_frame_snake_temp, Number, value)
  dif <- as.numeric(data_frame_snake$value)-as.numeric(data_frame_newt$value)

  df_dif <- as.data.frame(dif)
  df_dif$Number <- as.numeric(data_frame_newt$Number)
  df_dif$trial <- rep(Newt_mean$variable, ncol(Newt_mean)-1)
  if(boxy==TRUE){
 p <- ggplot(df_dif) +
    ggtitle(title, subtitle = subt) +
    xlab("Generation") + ylab("Average Phenotype Differences")+
    theme(plot.title = element_text(size=5)) +
    theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))+
    geom_boxplot(aes(x=factor(as.numeric(Number)), y=as.numeric(dif)), color="darkolivegreen" )
 }
 if(boxy==FALSE){
   p <- ggplot(df_dif) +
    ggtitle(title, subtitle = subt) +
    xlab("Generation") + ylab("Average Phenotype Differences")+
    theme(plot.title = element_text(size=5)) +
    theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))+
    geom_line(aes(x=as.numeric(Number), y=as.numeric(dif), color=trial ))+ 
     theme(legend.position="none")+
     scale_color_manual(values = rep("darkolivegreen", nrow(Newt_file)))

 }
  return(p)
}

```

```{r, data collection}

files <- list.files(path="~/Desktop/findparams_u5", pattern="nu_*1.56", full.names=TRUE, recursive=FALSE)

file <- read.table(files[1], header = TRUE)

Newt_mean <- making_data_frames(files, all_of("Newt_mean_Pheno"))
Newt_min <- making_data_frames(files, all_of("Newt_min_Pheno"))
Newt_max <- making_data_frames(files, all_of("Newt_max_Pheno"))
Newt_sd <- making_data_frames(files, all_of("Newt_sd_Pheno"))
Snake_mean <- making_data_frames(files, all_of("Snake_mean_Pheno"))
Snake_min <- making_data_frames(files, all_of("Snake_min_Pheno"))
Snake_max <- making_data_frames(files, all_of("Snake_max_Pheno"))
Snake_sd <- making_data_frames(files, all_of("Snake_sd_Pheno"))
```

# Comparisons

To start I am going to compare newts and snakes starting and ending phenotypes. Put more specifically, I will look at each simulation and record which species had a higher phenotype at the beginning and end of each of the simulations. The hope is that both newts and snakes have an equal (random) opportunity to have a higher phenotype. The ending could be skewed if coevoultion took place (but it is unknown, and I predict that it will be close 1000gen isn't a long time). I will look at the difference in the mean phenotype and the difference in standard deviations.

```{r, begining and end comparison }

sec<-starting_end_comparison(Newt_mean, Snake_mean, title="Comparison of Starting and Ending Newt and Snake Phenotypes (Larger, Mean)", subt="sigma = 0.2, mu = 1.5625e-10", f_point="1", l_point="1000")
sec

sdsec<-starting_end_comparison(Newt_sd, Snake_sd, title="Comparison of Starting and Ending Newt and Snake Phenotypes (Larger, SD)", subt="sigma = 0.2, mu = 1.5625e-10", f_point="1", l_point="1000")
sdsec

```

It looks like newts had a higher mean and sd in the beginning of the simulation more than half of the trials that I ran. Probably not statistically significant (thought is to run more simulations to see or look at the other combos that I ran and see if it still happens there). The ending phenotypes are strange. I do not really know if there is something happening or if the advantage from the beginning of the run is leading to a newt advantage. I want to look at the other parameters to see if this is a reoccurring trend.

Next, I want to look at the distribution of mean phenotype values for every 100th generation. This will look into how the mean phenotype changes over time. I can look at the distributions of the mean and the standard devation of the phenotype.

```{r, comparative figures for sigma = 0.2, mu = 1.5625e-10 change in phenotype}
sp<-phenotype_boxplot(Newt_mean, Snake_mean, title="Variation in the Mean Phenotype For Newts and Snakes by Generation", subt="sigma = 0.2, mu = 1.5625e-10", boxy=TRUE, section=TRUE, section_num=10)
sp

sdp<-phenotype_boxplot(Newt_sd, Snake_sd, title="Variation in the Sd of Phenotype For Newts and Snakes by Generation", subt="sigma = 0.2, mu = 1.5625e-10", boxy=TRUE, section=TRUE, section_num=10)
sdp

```

It seems like the mean phenotype of the newts and snakes are not exploding when using this set of parameters. It does look like there is more variation in the average phenotypes and sd as the number of generations increased. Nothing out of the ordinary.

Here, I will look at the barest of trends. Specifically, I will look at the trend of the mean of means and the mean of the standard deviation. This will show the average trend that my simulations are taking.

```{r, general trends}
sp2<-phenotype_boxplot(Newt_mean, Snake_mean, title="Averaging the Mean Phenotype For Newts and Snakes by Generation", subt="sigma = 0.2, mu = 1.5625e-10", boxy=FALSE)
sp2

sdp2<-phenotype_boxplot(Newt_sd, Snake_sd, title="Averaging the Sd of Phenotype For Newts and Snakes by Generation", subt="sigma = 0.2, mu = 1.5625e-10", boxy=FALSE)
sdp2

```

The phenotypes of newts and snakes are multiplicative, so the sd might go up faster? Is this why the standard devation has the weird curve?

```{r, other figs}
n<-phenotype_trends(Newt_mean, title="Newt Average Phenotype Trends By Generations", subt="sigma = 0.2, mu = 1.5625e-10", my_color="darkred", is_log=FALSE)
n

s<-phenotype_trends(Snake_mean, title="Snake Average Phenotype Trends By Generations", subt="sigma = 0.2, mu = 1.5625e-10", my_color="Steelblue", is_log=FALSE)
s

bp <- phenotype_dif_boxplot(Newt_mean, Snake_mean, title="Differences", subt="sigma = 0.2, mu = 1.5625e-10",boxy=TRUE)
bp

bp2 <- phenotype_dif_boxplot(Newt_mean, Snake_mean, title="Differences", subt="sigma = 0.2, mu = 1.5625e-10",boxy=FALSE)
bp2
```