### 2021-08-12 graphs_newt_snakes/Space_display.Rmd
Made space r markdown that talkes generations and phenotypes and make spatial plots and distributions

---
title: "Space Display"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(viridis)
library(colorspace)
library(gridExtra)
library(tidyverse)
require(reshape2)
library(plyr)

#setwd("~/Desktop/Proposal Old Drafts")
```

# Goals

I am trying to see if there are any spatial correlations between newts and snakes. Like any areas with elevated phenotypes with the newts and snakes

MSprime: both_1000_su_1.11e-11_nu_1.11e-11_sue_0.3_nue_0.3_17018532.init.trees
text: file_redo_su_1.11e-11_nu_1.11e-11_sue_0.3_nue_0.3_17018532.txt
SLiM: mu_rate_1.11e-11_newt_mu_rate_1.11e-11_snake_mu_effect_sd_0.3_newt_mu_effect_sd_0.3_sigma_0.5_ID_3913371281942_late_1000_.trees
* working on renaming the slim files to have the FN number, but right now I can connect them from the slim id in the .txt file

```{r space plot, echo=FALSE}
space_data <- read.table("~/Desktop/pulled_info_test.txt", header = TRUE)

ggplot(space_data[which(space_data$gen == 1),], aes(x=loc_x, y=loc_y, color=type)) +
    ggtitle("Start Space: Generation 1", subtitle = "file_redo_su_1.11e-11_nu_1.11e-11_sue_0.3_nue_0.3_17018532.txt") +
    xlab("Longitude") + ylab("Latitude")+
    #theme(plot.title = element_text(size=5)) +
    theme_bw() +  #theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))+
    scale_color_manual(name = "Species", values = c('Newt' = "darkred", 'Snake' = "steelblue"), labels=c("Snakes", "Newts"))+
    geom_point(aes(alpha=ind_phy))

ggplot(space_data[which(space_data$gen == 1000),], aes(x=loc_x, y=loc_y, color=type)) +
    ggtitle("End Space: Generation 1000", subtitle = "file_redo_su_1.11e-11_nu_1.11e-11_sue_0.3_nue_0.3_17018532.txt") +
    xlab("Longitude") + ylab("Latitude")+
    #theme(plot.title = element_text(size=5)) +
    scale_fill_manual(name = "Species", values = c('TRUE' = "darkred", 'FALSE' = "steelblue"), labels=c("Snakes", "Newts"))+
    scale_color_manual(name = "Species", values = c('Newt' = "darkred", 'Snake' = "steelblue"), labels=c("Snakes", "Newts"))+
    theme_bw() +  #theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))+
    geom_point(aes(alpha=ind_phy))

```
# Distributions

Look at the phenotype distributions of newts and snakes in the 1st and 1000th generation. The dashed lines are the mean phenotype values

```{r distribution plot, echo=FALSE}

mu_1gen <- ddply(space_data[which(space_data$gen == 1),], "type", summarise, grp.mean=mean(ind_phy))
mu_1000gen <- ddply(space_data[which(space_data$gen == 1),], "type", summarise, grp.mean=mean(ind_phy))

ggplot(space_data[which(space_data$gen == 1),], aes(x=ind_phy, color=type, fill=type), alpha=0.3) +
  ggtitle("Start Space: Generation 1", subtitle = "file_redo_su_1.11e-11_nu_1.11e-11_sue_0.3_nue_0.3_17018532.txt") +
  xlab("Phenotype") + ylab("Count")+
  geom_histogram(position="identity", alpha=0.3)+
  geom_vline(data=mu_1gen, aes(xintercept=grp.mean, color=type),
             linetype="dashed")+
  scale_color_manual(name = "Species", values = c('Newt' = "darkred", 'Snake' = "steelblue"), labels=c("Snakes", "Newts"))+ 
  scale_fill_manual(name = "Species", values = c('Newt' = "darkred", 'Snake' = "steelblue"), labels=c("Snakes", "Newts"))+ 
  theme_bw()

ggplot(space_data[which(space_data$gen == 1000),], aes(x=ind_phy, color=type, fill=type), alpha=0.3) +
  ggtitle("Start Space: Generation 1000", subtitle = "file_redo_su_1.11e-11_nu_1.11e-11_sue_0.3_nue_0.3_17018532.txt") +
  xlab("Phenotype") + ylab("Count")+
  geom_histogram(position="identity", alpha=0.3)+
  geom_vline(data=mu_1000gen, aes(xintercept=grp.mean, color=type),
             linetype="dashed")+
  scale_color_manual(name = "Species", values = c('Newt' = "darkred", 'Snake' = "steelblue"), labels=c("Snakes", "Newts"))+ 
  scale_fill_manual(name = "Species", values = c('Newt' = "darkred", 'Snake' = "steelblue"), labels=c("Snakes", "Newts"))+ 
  theme_bw()

```