### 2021-10-04 graphs_newt_snakes/nospce_10000.Rmd
I looked the the results from letting nospace runt tell 10000 generations.

---
title: "nospace_10000"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(viridis)
library(colorspace)
library(gridExtra)
library(tidyverse)
require(reshape2)
library(plyr)
library(gridGraphics)
library(grid)
```

## What is No Space gen_10000?

This is my no space experiment out to 10,000!

## My Goals

## The Specifics of the Simulations

To make these simulations similar to the previous simulation experiments I use the same msprime files to start each of the SLiM simulations. I also keep the parameters the same.

```{r, making data frames}

making_data_frames <- function(files, spe_col) {
i=1
for (current_file in files){
 #print(current_file)

  if(i ==1){
    file <- read.table(current_file, header = TRUE)

    file[[spe_col]][file[[spe_col]] == "NULL"] <- NA
      # might want to change to all_of(spe_col)
    file %>% 
      select(all_of(spe_col), gen) %>%
      #rownames_to_column() %>% 
      gather(variable, value, -gen) %>% 
      spread(gen, value) -> new_file

    new_file$variable[i] <- paste0(spe_col,"_trial_", i)
    i = i + 1
    next
    }
  file <- read.table(current_file, header = TRUE)

  file[[spe_col]][file[[spe_col]] == "NULL"] <- NA

file %>% 
  select(all_of(spe_col), gen) %>%
  #rownames_to_column() %>% 
  gather(variable, value, -gen) %>% 
  spread(gen, value) -> temp_file

temp_file$variable[1] <- paste0(spe_col,"_trial_", i)

new_file<-rbind(new_file, temp_file)
i = i + 1
}
return(new_file)
}

```

```{r, making resueable figures, results='hide', message=FALSE, warning=FALSE}

phenotype_trends <- function(file, title, subt, my_color, is_log=FALSE, leg_off=FALSE){
  data_frame_temp <- as.data.frame(t(file))

  colnames(data_frame_temp) <- data_frame_temp[1,]
  names(data_frame_temp) <- colnames(data_frame_temp)
  steps <- colnames(file[ , !(names(file) %in% "variable")])
data_frame_temp <- data_frame_temp[-1, ] 
data_frame_temp$steps <- as.numeric(unlist(steps))

  graph_file <- melt(data_frame_temp,  id.vars = 'steps', variable.name = 'series')

    p <- ggplot(graph_file, aes(steps, as.numeric(value))) + geom_line(aes(colour = series))+ 
    ggtitle(title, subtitle = subt) +
    scale_color_manual(values = rep(my_color, nrow(file)))+
    xlab("Generations") + ylab("Interaction")+
    theme_bw() + theme(legend.position="none")

    if(is_log==TRUE){
  p = p + scale_y_log10() + ylab("Interaction (Log)")
    }
     if(leg_off==TRUE){
 p = p+ theme(legend.position = "none")}
    return(p)
}

var_by_var <- function(Newt_file_x, Snake_file_x, Newt_file_y, Snake_file_y, title, subt, x_label, y_label, logy=FALSE, logx=FALSE, subtract_x=FALSE, by_gen=FALSE, just_two=FALSE, colored=FALSE, color_label=NULL, highcol="darkseagreen"){

  data_frame_newt_temp_x <- Newt_file_x
  data_frame_newt_temp_x <- data_frame_newt_temp_x[ , !(names(data_frame_newt_temp_x) %in% "variable")]

  data_frame_snake_temp_x <- Snake_file_x
  data_frame_snake_temp_x <- data_frame_snake_temp_x[ , !(names(data_frame_snake_temp_x) %in% "variable")]

  data_frame_newt_temp_y <- Newt_file_y
  data_frame_newt_temp_y <- data_frame_newt_temp_y[ , !(names(data_frame_newt_temp_y) %in% "variable")]

  data_frame_snake_temp_y <- Snake_file_y
  data_frame_snake_temp_y <- data_frame_snake_temp_y[ , !(names(data_frame_snake_temp_y) %in% "variable")]

  Number <- colnames(data_frame_newt_temp_x)
  data_frame_newt_x <- gather(data_frame_newt_temp_x, Number, value_x)
  data_frame_newt_x$type <- "Newt"
  data_frame_newt_y <- gather(data_frame_newt_temp_y, Number, value_y)
  data_frame_newt <- cbind(data_frame_newt_x, data_frame_newt_y)
  data_frame_newt <- data_frame_newt[-4]
  data_frame_snake_x <- gather(data_frame_snake_temp_x, Number, value_x)
  data_frame_snake_x$type <- "Snake"
  data_frame_snake_y <- gather(data_frame_snake_temp_y, Number, value_y)
  data_frame_snake <- cbind(data_frame_snake_x, data_frame_snake_y)
  data_frame_snake <- data_frame_snake[-4]

 if(subtract_x==FALSE){
   combo <- rbind(data_frame_newt, data_frame_snake)
   combo_s <- combo[sample(nrow(combo), nrow(combo)), ]
   p <- ggplot(combo_s) +
    #xlim(0, 1000)+
      ggtitle(title, subtitle = subt) +
      xlab(x_label) + ylab(y_label)+
      geom_point(aes(x=as.numeric(value_x), y=as.numeric(value_y), color=type), alpha=0.1)+
      scale_color_manual(name = "Species", values = c("Newt" = "darkred", "Snake" = "steelblue"))+
      theme(plot.title = element_text(size=5)) +
      theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
 }
 if(subtract_x==TRUE){
    subtract_val <- data_frame_snake_x$value_x - data_frame_newt_x$value_x
    total_sub_dataframe <- data.frame(subtract_val, data_frame_newt_y$value_y, data_frame_snake_y$value_y)
    total_sub_dataframe <- total_sub_dataframe[sample(nrow(total_sub_dataframe), nrow(total_sub_dataframe)), ]
    colnames(total_sub_dataframe) <- c("SnakeMisNewt_Pheno","Newts","Snakes")
     p <- ggplot(total_sub_dataframe) +
      #xlim(0, 1000)+
       ggtitle(title, subtitle = subt) +
       xlab(x_label) + ylab(y_label)+
       geom_point(aes(x=as.numeric(SnakeMisNewt_Pheno), y=as.numeric(Newts), color="Newts"), alpha=0.1)+
       geom_point(aes(x=as.numeric(SnakeMisNewt_Pheno), y=as.numeric(Snakes), color="Snakes"), alpha=0.1)+
        scale_colour_manual(name = "Species", values = c("Newts" = "darkred", "Snakes" = "steelblue"))+
        theme(plot.title = element_text(size=5)) +
        theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
 }
  if(by_gen==TRUE){
    combo <- rbind(data_frame_newt, data_frame_snake)
    combo_s <- combo[sample(nrow(combo), nrow(combo)), ]
       p <- ggplot(combo_s) +
    #xlim(0, 1000)+
      ggtitle(title, subtitle = subt) +
      xlab(x_label) + ylab(y_label)+
      geom_point(aes(x=as.numeric(Number), y=as.numeric(value_y), color=type), alpha=0.1)+
      scale_color_manual(name = "Species", values = c("Newt" = "darkred", "Snake" = "steelblue"))+
      theme(plot.title = element_text(size=5)) +
      theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
  }

   if(just_two==TRUE){
     just2 <- cbind(data_frame_newt_x, data_frame_snake_y)
     just2 <- just2[-4]
       p <- ggplot(just2) +
    #xlim(0, 1000)+
      ggtitle(title, subtitle = subt) +
      xlab(x_label) + ylab(y_label)+
      geom_point(aes(x=as.numeric(value_x), y=as.numeric(value_y)), color = "darkgreen", alpha=0.1)+
      theme(plot.title = element_text(size=5)) +
      theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
   }

   if(colored==TRUE){
     color3 <- cbind(data_frame_newt_x, data_frame_snake_y, data_frame_newt_y)
     color3 <- color3[-c(1,3,4,6)]
     suppressWarnings(suppressMessages(suppressPackageStartupMessages({require(scales)})))
     amin <- 0.1
     lowcol.hex <- as.hexmode(round(col2rgb(highcol) * amin + 255 * (1 - amin)))
     lowcol <- paste0("#",   sep = "",
                 paste(format(lowcol.hex, width = 2), collapse = ""))
       p <- ggplot(color3) +
    #xlim(0, 1000)+
      ggtitle(title, subtitle = subt) +
      xlab(x_label) + ylab(y_label)+
      geom_point(aes(x=as.numeric(value_x), y=as.numeric(value_y), color = value_y.1))+
      theme(plot.title = element_text(size=5)) +
      scale_colour_gradient(high = highcol, low = lowcol) + 
      labs(colour = color_label) +
      theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
  }
  if(logy==TRUE){
    p = p + scale_y_log10()
  }
   if(logx==TRUE){
    p = p + scale_x_log10()
  }

  return(p)

}

phenotype_boxplot <- function(Newt_file, Snake_file, title, subt, boxy=FALSE, section=FALSE, section_num=0, leg_off=FALSE, ylim_max=100){
  data_frame_newt_temp <- Newt_file
  data_frame_newt_temp <- data_frame_newt_temp[ , !(names(data_frame_newt_temp) %in% "variable")]

  data_frame_snake_temp <- Snake_file
  data_frame_snake_temp <- data_frame_snake_temp[ , !(names(data_frame_snake_temp) %in% "variable")]

   if(section==TRUE){

   selcol = seq(1,ncol(data_frame_newt_temp), section_num)
   data_frame_newt_temp <- data_frame_newt_temp %>% select(all_of(selcol))
   data_frame_snake_temp <- data_frame_snake_temp %>% select(all_of(selcol))

   }
   Number <- colnames(data_frame_newt_temp)
  data_frame_newt <- gather(data_frame_newt_temp, Number, value)
  data_frame_newt$type <- "Newt"
  data_frame_snake <- gather(data_frame_snake_temp, Number, value)
  data_frame_snake$type <- "Snake"

 combo <- rbind(data_frame_newt, data_frame_snake)

 small_combo <- aggregate(as.numeric(combo$value),by=list(Number=as.numeric(combo$Number),type=combo$type),data=combo,FUN=mean, na.action = TRUE)

if(boxy==TRUE){
  p <- ggplot(combo) +
    ylim(0, ylim_max)+
    ggtitle(title, subtitle = subt) +
    xlab("Generation") + ylab("Average Phenotypes")+
    scale_color_manual(name = "Species", values = c("Newt" = "darkred", "Snake" = "steelblue"))+
    theme(plot.title = element_text(size=5)) +
    theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))+
    geom_boxplot(aes(x=factor(as.numeric(Number)), y=as.numeric(value), color=type) )
  }
if(boxy==FALSE){
  p <- ggplot(small_combo) +
  #xlim(0, 1000)+
    ggtitle(title, subtitle = subt) +
    xlab("Generation") + ylab("Average Phenotypes")+
    scale_color_manual(name = "Species", values = c("Newt" = "darkred", "Snake" = "steelblue"))+
    theme(plot.title = element_text(size=5)) +
    theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))+
    geom_point(aes(x=factor(as.numeric(Number)), y=as.numeric(x), color=type) )
}

 if(leg_off==TRUE){
 p = p+ theme(legend.position = "none")}

return(p)
}

phenotype_dif_boxplot <- function(Newt_file, Snake_file, title, subt, boxy=FALSE, section=FALSE, section_num=0, leg_off=FALSE){
  data_frame_newt_temp <- Newt_file
  data_frame_newt_temp <- data_frame_newt_temp[ , !(names(data_frame_newt_temp) %in% "variable")]

  data_frame_snake_temp <- Snake_file 
  data_frame_snake_temp <- data_frame_snake_temp[ , !(names(data_frame_snake_temp) %in% "variable")]

  if(section==TRUE){

   selcol = seq(1,ncol(data_frame_newt_temp), section_num)
   data_frame_newt_temp <- data_frame_newt_temp %>% select(all_of(selcol))
   data_frame_snake_temp <- data_frame_snake_temp %>% select(all_of(selcol))

 }

  Number <- colnames(data_frame_newt_temp)
  data_frame_newt <- gather(data_frame_newt_temp, Number, value)
  data_frame_snake <- gather(data_frame_snake_temp, Number, value)
  dif <- as.numeric(data_frame_snake$value)-as.numeric(data_frame_newt$value)

  df_dif <- as.data.frame(dif)
  df_dif$Number <- as.numeric(data_frame_newt$Number)

  if(boxy==TRUE){
 p <- ggplot(df_dif) +
    ggtitle(title, subtitle = subt) +
    xlab("Generation") + ylab("Average Phenotype Differences")+
    theme(plot.title = element_text(size=5)) +
    theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))+
    geom_boxplot(aes(x=factor(as.numeric(Number)), y=as.numeric(dif)), color="darkolivegreen" )
 }
 if(boxy==FALSE){
   df_dif$trial <- rep(Newt_file$variable, ncol(Newt_file)-1)
   p <- ggplot(df_dif) +
    ggtitle(title, subtitle = subt) +
    xlab("Generation") + ylab("Average Phenotype Differences")+
    theme(plot.title = element_text(size=5)) +
    theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))+
    geom_line(aes(x=as.numeric(Number), y=as.numeric(dif), color=trial ))+ 
     theme(legend.position="none")+
     scale_color_manual(values = rep("darkolivegreen", nrow(data_frame_newt_temp)))
 }
   if(leg_off==TRUE){
 p = p+ theme(legend.position = "none")}
  return(p)
}

phenotype_change_graph_s <- function(files, title, is_log=FALSE, xlim_value){
file1 <- read.table(files[1], header = TRUE)
p <- ggplot(file1, aes(x=1:144)) +
  xlim(0, xlim_value)+
  ggtitle(title) +
  xlab("Simulation number") + ylab("Phenotype")+
  theme_bw()

if(is_log==TRUE){
  p = p + scale_y_log10() + ylab("Phenotype  (Log)")

}

i=1
for (current_file in files){

  file <- read.table(current_file, header = TRUE)
  file$Newt_mean_Pheno[file$Newt_mean_Pheno == "NULL"] <- 0
  file$Snake_mean_Pheno[file$Snake_mean_Pheno == "NULL"] <- 0
  file <- file[complete.cases(file), ]
 p <- p + geom_point(data= file, x = i, aes(y = as.numeric(Newt_mean_Pheno[1])+0.1), color= darken("darkred")) +
   geom_point(data= file,x = i, aes(y = as.numeric(Newt_mean_Pheno[50])+0.1), color=lighten("darkred"))  + 
  geom_point(data= file, x = i, aes(y = as.numeric(Snake_mean_Pheno[1])+0.1), color=darken("steelblue")) +
  geom_point(data= file, x = i, aes(y = as.numeric(Snake_mean_Pheno[50])+0.1), color=lighten("steelblue"))+
  geom_segment(data = file, x=i, xend=i, aes( y=as.numeric(Newt_mean_Pheno[1])+0.1, yend=as.numeric(Newt_mean_Pheno[50])+0.1, colour="Newt"), size = 0.3, arrow = arrow(length = unit(0.5, "cm")))+
   geom_segment(data = file, x=i, xend=i, aes( y=as.numeric(Snake_mean_Pheno[1])+0.1, yend=as.numeric(Snake_mean_Pheno[50])+0.1, colour="Snake"), size = 0.3, arrow = arrow(length = unit(0.5, "cm")))

 i = i+1
}
p <- p +
 scale_colour_manual(name = "Species", values = c("Newt" = "darkred", "Snake"="steelblue"))
return(p)
}

```

```{r, data info}
# Files for No space gen_10000 results
files <- list.files(path="~/Desktop/nospace/gen_10000", pattern="nu_*1.56", full.names=TRUE, recursive=FALSE)
file <- read.table(files[51], header = TRUE)
beta_n <- making_data_frames(files, all_of("beta_n"))
beta_s <- making_data_frames(files, all_of("beta_s"))
Newt_mean_Pheno <- making_data_frames(files, all_of("Newt_mean_Pheno"))
Snake_mean_Pheno <- making_data_frames(files, all_of("Snake_mean_Pheno"))
Newt_sd_Pheno <- making_data_frames(files, all_of("Newt_sd_Pheno"))
Snake_sd_Pheno <- making_data_frames(files, all_of("Snake_sd_Pheno"))
mean_newts_eaten <- making_data_frames(files, all_of("mean_newts_eaten"))
mean_snakes_eaten <- making_data_frames(files, all_of("mean_snakes_eaten"))
newt_deaths <- making_data_frames(files, all_of("newt_deaths"))
snake_deaths <- making_data_frames(files, all_of("snake_deaths"))
Newt_age <- making_data_frames(files, all_of("Newt_age"))
Snake_age <- making_data_frames(files, all_of("Snake_age"))
Newt_pop_size <- making_data_frames(files, all_of("Newt_pop_size"))
Snake_pop_size <- making_data_frames(files, all_of("Snake_pop_size"))

# Files for space results
files_u5 <- list.files(path="~/Desktop/beta_u5", pattern="nu_*1.56", full.names=TRUE, recursive=FALSE)

beta_n_s <- making_data_frames(files_u5, all_of("beta_n"))
beta_s_s <- making_data_frames(files_u5, all_of("beta_s"))
Newt_mean_Pheno_s <- making_data_frames(files_u5, all_of("Newt_mean_Pheno"))
Snake_mean_Pheno_s <- making_data_frames(files_u5, all_of("Snake_mean_Pheno"))
Newt_sd_Pheno_s <- making_data_frames(files_u5, all_of("Newt_sd_Pheno"))
Snake_sd_Pheno_s <- making_data_frames(files_u5, all_of("Snake_sd_Pheno"))
mean_newts_eaten_s <- making_data_frames(files_u5, all_of("mean_newts_eaten"))
mean_snakes_eaten_s <- making_data_frames(files_u5, all_of("mean_snakes_eaten"))
newt_deaths_s <- making_data_frames(files_u5, all_of("newt_deaths"))
snake_deaths_s <- making_data_frames(files_u5, all_of("snake_deaths"))
Newt_age_s <- making_data_frames(files_u5, all_of("Newt_age"))
Snake_age_s <- making_data_frames(files_u5, all_of("Snake_age"))
Newt_pop_size_s <- making_data_frames(files_u5, all_of("Newt_pop_size"))
Snake_pop_size_s <- making_data_frames(files_u5, all_of("Snake_pop_size"))

```

## The distribution of Phenotypes

## How Restance and Toxin Levels Change Overtime

### Phenotype Change

```{r, phenotype change}

w1 <- phenotype_change_graph_s(files_u5, title="Change in Phenotype Space", is_log=TRUE, xlim_value=104)

w2 <- phenotype_change_graph_s(files, title="Change in Phenotype NoSpace", is_log=TRUE, xlim_value=104)

w1
w2
```

### Phenotype Variation

```{r, comparative figures this params}

sp_s<-phenotype_boxplot(Newt_mean_Pheno_s, Snake_mean_Pheno_s, title="Space: Variation in the Mean Phenotype For Newts and Snakes by Generation", subt="sigma = 0.2, mu = 1.5625e-10", boxy=TRUE, section=TRUE, section_num=10, ylim_max=30)

sdp_s<-phenotype_boxplot(Newt_sd_Pheno_s, Snake_sd_Pheno_s, title="Space: Variation in the Sd of Phenotype For Newts and Snakes by Generation", subt="sigma = 0.2, mu = 1.5625e-10", boxy=TRUE, section=TRUE, section_num=10, ylim_max=30)

grid.arrange(sp_s, sdp_s, nrow=1)

sp<-phenotype_boxplot(Newt_mean_Pheno, Snake_mean_Pheno, title="NoSpace: Variation in the Mean Phenotype For Newts and Snakes by Generation", subt="sigma = 0.2, mu = 1.5625e-10", boxy=TRUE, section=TRUE, section_num=50, ylim_max=30)

sdp<-phenotype_boxplot(Newt_sd_Pheno, Snake_sd_Pheno, title="NoSpace: Variation in the Sd of Phenotype For Newts and Snakes by Generation", subt="sigma = 0.2, mu = 1.5625e-10", boxy=TRUE, section=TRUE, section_num=50, ylim_max=30)

grid.arrange(sp, sdp, nrow=1)
```

### Phenotype Differences

```{r, the differences}
bp_s<- phenotype_dif_boxplot(Newt_mean_Pheno_s, Snake_mean_Pheno_s, title="Space: Differences", subt="sigma = 0.2, mu = 1.5625e-10",boxy=TRUE, section=TRUE, section_num=10)

bp2_s <- phenotype_dif_boxplot(Newt_mean_Pheno_s, Snake_mean_Pheno_s, title="Space: Differences", subt="sigma = 0.2, mu = 1.5625e-10",boxy=FALSE, section=FALSE, section_num=0)

grid.arrange(bp_s, bp2_s, nrow=1)

bp <- phenotype_dif_boxplot(Newt_mean_Pheno, Snake_mean_Pheno, title="NoSpace: Differences", subt="sigma = 0.2, mu = 1.5625e-10",boxy=TRUE, section=TRUE, section_num=50)

bp2 <- phenotype_dif_boxplot(Newt_mean_Pheno, Snake_mean_Pheno, title="NoSpace: Differences", subt="sigma = 0.2, mu = 1.5625e-10",boxy=FALSE, section=FALSE, section_num=0)

grid.arrange(bp, bp2, nrow=1)
```

## The Interaction

### Differnece in mean phenotype by Deaths/Age or Pouplation Size  (from Selection coef)

```{r, sel coef graphs}

ps1_s <- var_by_var( Newt_mean_Pheno_s, Snake_mean_Pheno_s, mean_newts_eaten_s, mean_snakes_eaten_s, title="Space: Differnece in mean phenotype by Deaths", subt="sigma = 0.2, mu = 1.5625e-10", x_label="Dif in Mean Pheno", y_label="Deaths/age", subtract_x=TRUE)

ps2_s <- var_by_var( Newt_mean_Pheno_s, Snake_mean_Pheno_s, Newt_pop_size_s, Snake_pop_size_s, title="Space: Differnece in mean phenotype by PopSize", subt="sigma = 0.2, mu = 1.5625e-10", x_label="Dif in Mean Pheno", y_label="PopSize", subtract_x=TRUE)

grid.arrange(ps1_s, ps2_s, nrow=1)

ps1 <- var_by_var( Newt_mean_Pheno, Snake_mean_Pheno, mean_newts_eaten, mean_snakes_eaten, title="NoSpace: Differnece in mean phenotype by Deaths", subt="sigma = 0.2, mu = 1.5625e-10", x_label="Dif in Mean Pheno", y_label="Deaths/age", subtract_x=TRUE)

ps2 <- var_by_var( Newt_mean_Pheno, Snake_mean_Pheno, Newt_pop_size, Snake_pop_size, title="NoSpace: Differnece in mean phenotype by PopSize", subt="sigma = 0.2, mu = 1.5625e-10", x_label="Dif in Mean Pheno", y_label="PopSize", subtract_x=TRUE)

grid.arrange(ps1, ps2, nrow=1)

```

### Population size of Newts by Snakes with phenotype as color

```{r, popsize of newts and snakes with color of phenotype}

vc1_s <- var_by_var(Newt_pop_size_s, Newt_mean_Pheno_s, Newt_mean_Pheno_s, Snake_pop_size_s, title="Space: Newt VS Snake Population Size Color=Phenotype", subt="sigma = 0.1, mu = 1.0e-10", y_label="Snake Popsize", x_label="Newt Popsize", colored=TRUE, color_label="Newt Phenotype" , highcol="darkred")

vc2_s <- var_by_var(Newt_pop_size_s, Snake_mean_Pheno_s, Snake_mean_Pheno_s, Snake_pop_size_s, title="Space: Newt VS Snake Population Size Color=Phenotype", subt="sigma = 0.1, mu = 1.0e-10", y_label="Snake Popsize", x_label="Newt Popsize", colored=TRUE, color_label="Snake Phenotype", highcol="steelblue")

vc1_s 
vc2_s

vc1 <- var_by_var(Newt_pop_size, Newt_mean_Pheno, Newt_mean_Pheno, Snake_pop_size, title="NoSpace: Newt VS Snake Population Size Color=Phenotype", subt="sigma = 0.1, mu = 1.0e-10", y_label="Snake Popsize", x_label="Newt Popsize", colored=TRUE, color_label="Newt Phenotype" , highcol="darkred")

vc2 <- var_by_var(Newt_pop_size, Snake_mean_Pheno, Snake_mean_Pheno, Snake_pop_size, title="NoSpace: Newt VS Snake Population Size Color=Phenotype", subt="sigma = 0.1, mu = 1.0e-10", y_label="Snake Popsize", x_label="Newt Popsize", colored=TRUE, color_label="Snake Phenotype", highcol="steelblue")

vc1 
vc2
```

### Beta Distributions

```{r, beta distribution}

hist(unlist(beta_s_s[ , 2:52]), col=rgb(173,216,230,max = 255, alpha = 80, names = "lt.blue"), main="Space: Beta Values", xlab="Beta", breaks=15) 
hist(unlist(beta_n_s[ , 2:52]), col=rgb(255,192,203, max = 255, alpha = 80, names = "lt.pink"), add = TRUE, breaks=15) #ylim=c(0,2100)

abline(v = mean(unlist(beta_n_s[ , 2:52])), col = rgb(255,192,203, max = 255), lwd = 2)
abline(v = mean(unlist(beta_s_s[ , 2:52])), col = rgb(173,216,230, max = 255), lwd = 2)
legend("topright", c("Newt", "Snake"), fill=c(rgb(255,192,203, max = 255), rgb(173,216,230, max = 255)))

hist(unlist(beta_s[ , 2:52]), col=rgb(173,216,230,max = 255, alpha = 80, names = "lt.blue"), main="NoSpace: Beta Values", xlab="Beta", breaks=15) 
hist(unlist(beta_n[ , 2:52]), col=rgb(255,192,203, max = 255, alpha = 80, names = "lt.pink"), add = TRUE, breaks=15) #ylim=c(0,2100)

abline(v = mean(unlist(beta_n[ , 2:52])), col = rgb(255,192,203, max = 255), lwd = 2)
abline(v = mean(unlist(beta_s[ , 2:52])), col = rgb(173,216,230, max = 255), lwd = 2)
legend("topright", c("Newt", "Snake"), fill=c(rgb(255,192,203, max = 255), rgb(173,216,230, max = 255)))

```

```{r, comp betas}
var_by_var(beta_n_s, beta_s_s, beta_n_s, beta_s_s, title="Space: Newt Betas by Snake Betas", subt="sigma = 0.2, mu = 1.5625e-10", x_label="Newt betas", y_label="Snake Betas", just_two = TRUE)

var_by_var(beta_n, beta_s, beta_n, beta_s, title="NoSpace: Newt Betas by Snake Betas", subt="sigma = 0.2, mu = 1.5625e-10", x_label="Newt betas", y_label="Snake Betas", just_two = TRUE)

```

### Extra Figures

```{r, beta figures}

g1_s <- var_by_var( beta_n_s, beta_s_s, beta_n_s, beta_s_s, title="Space: Betas By Generation", subt="sigma = 0.2, mu = 1.5625e-10", x_label="Generation", y_label="Betas", by_gen=TRUE)

g2_s <- var_by_var( Newt_mean_Pheno_s, Snake_mean_Pheno_s, beta_n_s, beta_s_s, title="Space: Differnece in mean phenotype by Betas", subt="sigma = 0.2, mu = 1.5625e-10", x_label="Dif in Pheno", y_label="Betas", subtract_x=TRUE)

g3_s <- var_by_var( Newt_mean_Pheno_s, Snake_mean_Pheno_s, Newt_age_s, Snake_age_s, title="Space: Differnece in mean phenotype by Age", subt="sigma = 0.2, mu = 1.5625e-10", x_label="Dif in Pheno", y_label="Age", subtract_x=TRUE)

g4_s <- var_by_var(Newt_mean_Pheno_s, beta_s_s, beta_n_s, Snake_mean_Pheno_s, title="Space: Newt Pheno by Snake Pheno", subt="sigma = 0.2, mu = 1.5625e-10", x_label="Newt Mean Phenotype", y_label="Snake Mean Phenotype", just_two = TRUE)

grid.arrange(g1_s, g2_s, g3_s, g4_s, nrow=2)

g1 <- var_by_var( beta_n, beta_s, beta_n, beta_s, title="NoSpace: Betas By Generation", subt="sigma = 0.2, mu = 1.5625e-10", x_label="Generation", y_label="Betas", by_gen=TRUE)

g2 <- var_by_var( Newt_mean_Pheno, Snake_mean_Pheno, beta_n, beta_s, title="NoSpace: Differnece in mean phenotype by Betas", subt="sigma = 0.2, mu = 1.5625e-10", x_label="Dif in Pheno", y_label="Betas", subtract_x=TRUE)

g3 <- var_by_var( Newt_mean_Pheno, Snake_mean_Pheno, Newt_age, Snake_age, title="NoSpace: Differnece in mean phenotype by Age", subt="sigma = 0.2, mu = 1.5625e-10", x_label="Dif in Pheno", y_label="Age", subtract_x=TRUE)

g4 <- var_by_var(Newt_mean_Pheno, beta_s, beta_n, Snake_mean_Pheno, title="NoSpace: Newt Pheno by Snake Pheno", subt="sigma = 0.2, mu = 1.5625e-10", x_label="Newt Mean Phenotype", y_label="Snake Mean Phenotype", just_two = TRUE)

grid.arrange(g1, g2, g3, g4, nrow=2)

```