### 2022-01-24 graphs_newt_snakes/Are_we_at_steadystate.Rmd
I am currently running a very long simulation to to see what happens overtime. I have also added some new graphs to fitness cost by looking at how different fitness cost effects the phenotype of the population. Even when there are different cost values all of the simulation reach a point where the it equiliborates on the time scale of drift. I have also started two new rmarkdowns/ experiments. One is to compare different genetic architectures (which I just finished running on the cluster) and the other is to look at steady-state of my co-evo cost simulation. I look forward to the results

---
title: "Are We At SteadyState Yet"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(viridis)
library(colorspace)
library(gridExtra)
library(tidyverse)
require(reshape2)
library(plyr)
library(gridGraphics)
library(grid)
```

## Goal

My aim is to run one cost_1on1 simulation for a really long time to determine if I am able to tell if the simulation has gone into steady state. The simulation the cost value will be 100 and I will run it for 10 million generations. The goal is to collect whole population data every 20th generation and individual data every 1000th generation

Approximation of data points:
gen/when data is collected 10000000/20 = 
gen/when data is collected * number of ind 10000000/1000*1500 = 1.5e+07

These files might be too big (we will see).

Things should stop changing by 2Ne generations. So maybe around 1,400 generations if my population size is around 700.

I have noticed that there seems to be no more variation in the populations. On SLiM all of the individuals in their respective populations have the same color indicating same phenotype