### 2022-01-25 graphs_newt_snakes/Are_we_at_steadystate.Rmd
I am running cost1on1 for very long periods of time and am prining out the values. I want to see what happens when i change interaction rate and population size. Does a population go extinct? is ther coexistance or coevo? I am also writing and making more graphs in the rmarkdowns

---
title: "Are We At SteadyState Yet"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(viridis)
library(colorspace)
library(gridExtra)
library(tidyverse)
require(reshape2)
library(plyr)
library(gridGraphics)
library(grid)
```

## Goal

My aim is to run one cost_1on1 simulation for a really long time to determine if I am able to tell if the simulation has gone into steady state. The simulation the cost value will be 100 and I will run it for 10 million generations. The goal is to collect whole population data every 20th generation and individual data every 1000th generation

Approximation of data points:
gen/when data is collected 10000000/20 = 
gen/when data is collected * number of ind 10000000/1000*1500 = 1.5e+07

These files might be too big (we will see).

Things should stop changing by 2Ne generations. So maybe around 1,400 generations if my population size is around 700.

I have noticed that there seems to be no more variation in the populations. On SLiM all of the individuals in their respective populations have the same color indicating same phenotype. The beta values are all 0's. All of the phenotypes have gotten really small (very close to 0). There is no longer co-evolution, but maybe coexistence.

More things that I have noticed: If interaction rate is higher like 0.15 or 0.10 then one of the populations will go extinct. This might change if the population size is larger