### 2022-02-01 graphs_newt_snakes/Comparing_GA.Rmd
made more graphs, wrote down new values to try for mu and sigma

---
title: "Comparing Genetic Architectures"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(viridis)
library(colorspace)
library(gridExtra)
library(tidyverse)
require(reshape2)
library(plyr)
library(gridGraphics)
library(grid)
library("reshape")
```

## Purpose

Here I am doing a small experiment to compare different genetic architectures between snakes and newts. I will use the cost_1on1 simulation and my snake_newt_gv msprime simulation.

Outline

-Pick 4 different mu sigma combinations where mu*sigma^2 is constant (mu=1e-11, sigma=5) to (mu=1e-8, sigma=0.005) i.e ((mu=1e-11, sigma=5(A)), (mu=1e-10, sigma=0.5(B)), (mu=1e-9, sigma=0.05(C)), (mu=1e-8, sigma=0.005(D))) The ABCD represents the group of mu and sigma. When in the newt and snake simulation the groups will be paired to make 16 pairs. 
-In total that will be 16 msprime simulations and 16 slim simulations
-See if there are any obvious winners
-make a co-evolutionary heatmap

I ran msprime simulations and the slim simulations on cluster in GA_lt folder.
Mt goal is to first see if there is an obvious "winner" in the co-evolutionary arms race between these species.

First, I want to look at the mean pheotypes of newts and snakes as the generations increase. Then I want to look at the difference between the mean phenotypes for each of the 16 simulations. I need to make the data frame to compare all 16 simulations. I noticed in the mu=1e-11, sigma=5 simulations that the staring pheotypes of some of the individulas were really high and sometimes the population crashed and died out.

```{r, making data frames}

making_data_frames <- function(files, spe_col) {
i=1
for (current_file in files){
 #print(current_file)

  if(i ==1){
    file <- read.table(current_file, header = TRUE)

    file[[spe_col]][file[[spe_col]] == "NULL"] <- NA
      # might want to change to all_of(spe_col)
    file %>% 
      select(all_of(spe_col), gen) %>%
      #rownames_to_column() %>% 
      gather(variable, value, -gen) %>% 
      spread(gen, value) -> new_file

    new_file$variable[i] <- paste0(spe_col,"_output_", i)
    i = i + 1
    next
    }
  file <- read.table(current_file, header = TRUE)

  file[[spe_col]][file[[spe_col]] == "NULL"] <- NA

file %>% 
  select(all_of(spe_col), gen) %>%
  #rownames_to_column() %>% 
  gather(variable, value, -gen) %>% 
  spread(gen, value) -> temp_file

temp_file$variable[1] <- paste0(spe_col,"_output_", i)

new_file<-rbind(new_file, temp_file)
i = i + 1
}
return(new_file)
}

```

```{r, making resueable figures, results='hide', message=FALSE, warning=FALSE}

phenotype_trends <- function(file, title, subt, my_color, is_log=FALSE, leg_off=FALSE){
  # Looks at the mean phenotype of all of the simulations and graphs them as lines together
  # good for seeing what all of the simulations are doing
  # it can log the y axis and turn off the legend
  data_frame_temp <- as.data.frame(t(file))

  colnames(data_frame_temp) <- data_frame_temp[1,]
  names(data_frame_temp) <- colnames(data_frame_temp)
  steps <- colnames(file[ , !(names(file) %in% "variable")])
data_frame_temp <- data_frame_temp[-1, ] 
data_frame_temp$steps <- as.numeric(unlist(steps))

  graph_file <- melt(data_frame_temp,  id.vars = 'steps', variable.name = 'series')

  graph_file$series<-unlist(str_split_fixed(graph_file$series, "Pheno_", 2))[,2]

    p <- ggplot(graph_file, aes(steps, as.numeric(value))) + geom_line(aes(colour = series))+ 
    #ggtitle(title, subtitle = subt) +
    #scale_color_manual(values = rep(my_color, nrow(file)))+
    xlab("Generations") + ylab("Phenotype")+
    theme_bw() + theme(legend.position="none")

    if(is_log==TRUE){
  p = p + scale_y_log10() + ylab("Phenotype (Log)")
    }
     if(leg_off==TRUE){
 p = p+ theme(legend.position = "none")}
    return(p)
}

var_by_var <- function(Newt_file_x, Snake_file_x, Newt_file_y, Snake_file_y, title, subt, x_label, y_label, logy=FALSE, subtract_x=FALSE, by_gen=FALSE, just_two=FALSE, colored=FALSE, color_label=NULL, highcol="darkseagreen", facet=FALSE){

  # very powerful and confusing graph. It can plot a max of 4 variables at a time!
  # I use it to compare newts and snakes to each other
  # I can log the y axis
  # I can look at things by generation (uses just the newt x and snake x)
  # I can compare newt to snake (uses the newt x and snake y)
  # I can subtract (snake - newt) (subtracts the x value and plots the y value)
  # Color can plot var by var and pick a 3rd var to color the points (use thenewt x, snake y and newt file y for the color)

  data_frame_newt_temp_x <- Newt_file_x
  data_frame_newt_temp_x <- data_frame_newt_temp_x[ , !(names(data_frame_newt_temp_x) %in% "variable")]

  data_frame_snake_temp_x <- Snake_file_x
  data_frame_snake_temp_x <- data_frame_snake_temp_x[ , !(names(data_frame_snake_temp_x) %in% "variable")]

  data_frame_newt_temp_y <- Newt_file_y
  data_frame_newt_temp_y <- data_frame_newt_temp_y[ , !(names(data_frame_newt_temp_y) %in% "variable")]

  data_frame_snake_temp_y <- Snake_file_y
  data_frame_snake_temp_y <- data_frame_snake_temp_y[ , !(names(data_frame_snake_temp_y) %in% "variable")]

  Number <- colnames(data_frame_newt_temp_x)
  data_frame_newt_x <- gather(data_frame_newt_temp_x, Number, value_x)
  data_frame_newt_x$type <- "Newt"
  data_frame_newt_y <- gather(data_frame_newt_temp_y, Number, value_y)
  data_frame_newt <- cbind(data_frame_newt_x, data_frame_newt_y)
  data_frame_newt <- data_frame_newt[-4]
  data_frame_snake_x <- gather(data_frame_snake_temp_x, Number, value_x)
  data_frame_snake_x$type <- "Snake"
  data_frame_snake_y <- gather(data_frame_snake_temp_y, Number, value_y)
  data_frame_snake <- cbind(data_frame_snake_x, data_frame_snake_y)
  data_frame_snake <- data_frame_snake[-4]

 if(subtract_x==FALSE){
   combo <- rbind(data_frame_newt, data_frame_snake)
   combo_s <- combo[sample(nrow(combo), nrow(combo)), ]
   p <- ggplot(combo_s) +
    #xlim(0, 1000)+
      ggtitle(title, subtitle = subt) +
      xlab(x_label) + ylab(y_label)+
      geom_point(aes(x=as.numeric(value_x), y=as.numeric(value_y), color=type), alpha=0.1)+
      scale_color_manual(name = "Species", values = c("Newt" = "darkred", "Snake" = "steelblue"))+
      theme(plot.title = element_text(size=5)) +
      theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
 }
 if(subtract_x==TRUE){
    subtract_val <- data_frame_snake_x$value_x - data_frame_newt_x$value_x
    total_sub_dataframe <- data.frame(subtract_val, data_frame_newt_y$value_y, data_frame_snake_y$value_y)
    total_sub_dataframe$wrap <- rep(1:16, length(Number))
    total_sub_dataframe <- total_sub_dataframe[sample(nrow(total_sub_dataframe), nrow(total_sub_dataframe)), ]
    colnames(total_sub_dataframe) <- c("SnakeMisNewt_Pheno","Newts","Snakes")
     p <- ggplot(total_sub_dataframe) +
      #xlim(0, 1000)+
       ggtitle(title, subtitle = subt) +
       xlab(x_label) + ylab(y_label)+
       geom_point(aes(x=as.numeric(SnakeMisNewt_Pheno), y=as.numeric(Newts), color="Newts"), alpha=0.1)+
       geom_point(aes(x=as.numeric(SnakeMisNewt_Pheno), y=as.numeric(Snakes), color="Snakes"), alpha=0.1)+
        scale_colour_manual(name = "Species", values = c("Newts" = "darkred", "Snakes" = "steelblue"))+
        theme(plot.title = element_text(size=5)) +
        theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))

 }
  if(by_gen==TRUE){
    combo <- rbind(data_frame_newt, data_frame_snake)
    combo_s <- combo[sample(nrow(combo), nrow(combo)), ]
       p <- ggplot(combo_s) +
    #xlim(0, 1000)+
      ggtitle(title, subtitle = subt) +
      xlab(x_label) + ylab(y_label)+
      geom_point(aes(x=as.numeric(Number), y=as.numeric(value_y), color=type), alpha=0.1)+
      scale_color_manual(name = "Species", values = c("Newt" = "darkred", "Snake" = "steelblue"))+
      theme(plot.title = element_text(size=5)) +
      theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
  }

   if(just_two==TRUE){
     just2 <- cbind(data_frame_newt_x, data_frame_snake_y)
     just2 <- just2[-4]
       p <- ggplot(just2) +
    #xlim(0, 1000)+
      ggtitle(title, subtitle = subt) +
      xlab(x_label) + ylab(y_label)+
      geom_point(aes(x=as.numeric(value_x), y=as.numeric(value_y)), color = "darkgreen", alpha=0.1)+
      theme(plot.title = element_text(size=5)) +
      theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
   }

   if(colored==TRUE){

     color3 <- cbind(data_frame_newt_x, data_frame_snake_y, data_frame_newt_y)
     color3 <- color3[-c(1,3,4,6)]
     color3$wrap <- rep(1:16, length(Number))
     suppressWarnings(suppressMessages(suppressPackageStartupMessages({require(scales)})))
     amin <- 0.1
     lowcol.hex <- as.hexmode(round(col2rgb(highcol) * amin + 255 * (1 - amin)))
     lowcol <- paste0("#",   sep = "",
                 paste(format(lowcol.hex, width = 2), collapse = ""))
       p <- ggplot(color3) +
    #xlim(0, 1000)+
      ggtitle(title, subtitle = subt) +
      xlab(x_label) + ylab(y_label)+
      geom_point(aes(x=as.numeric(value_x), y=as.numeric(value_y), color = value_y.1))+
      theme(plot.title = element_text(size=5)) +
      scale_colour_gradient(high = highcol, low = lowcol) + 
      labs(colour = color_label) +
      theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
       if(facet==TRUE){

     p = p + facet_wrap(~wrap)
  }
  }
  if(logy==TRUE){
    p = p + scale_y_log10()
  }

  return(p)

}

phenotype_boxplot <- function(Newt_file, Snake_file, title, subt, boxy=FALSE, section=FALSE, section_num=0, leg_off=FALSE, ylim_max=100){
  #this function can graph phenotype of all of the simulations (as a mean or boxplot) by generations (I can specify the amount of generations to show)
  data_frame_newt_temp <- Newt_file
  data_frame_newt_temp <- data_frame_newt_temp[ , !(names(data_frame_newt_temp) %in% "variable")]

  data_frame_snake_temp <- Snake_file
  data_frame_snake_temp <- data_frame_snake_temp[ , !(names(data_frame_snake_temp) %in% "variable")]

   if(section==TRUE){

   selcol = seq(1,ncol(data_frame_newt_temp), section_num)
   data_frame_newt_temp <- data_frame_newt_temp %>% select(all_of(selcol))
   data_frame_snake_temp <- data_frame_snake_temp %>% select(all_of(selcol))

   }
   Number <- colnames(data_frame_newt_temp)
  data_frame_newt <- gather(data_frame_newt_temp, Number, value)
  data_frame_newt$type <- "Newt"
  data_frame_snake <- gather(data_frame_snake_temp, Number, value)
  data_frame_snake$type <- "Snake"

 combo <- rbind(data_frame_newt, data_frame_snake)

 small_combo <- aggregate(as.numeric(combo$value),by=list(Number=as.numeric(combo$Number),type=combo$type),data=combo,FUN=mean, na.action = TRUE)

if(boxy==TRUE){
  p <- ggplot(combo) +
    ylim(0, ylim_max)+
    ggtitle(title, subtitle = subt) +
    xlab("Generation") + ylab("Average Phenotypes")+
    scale_color_manual(name = "Species", values = c("Newt" = "darkred", "Snake" = "steelblue"))+
    theme(plot.title = element_text(size=5)) +
    theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))+
    geom_boxplot(aes(x=factor(as.numeric(Number)), y=as.numeric(value), color=type) )
  }
if(boxy==FALSE){
  p <- ggplot(small_combo) +
  #xlim(0, 1000)+
    ggtitle(title, subtitle = subt) +
    xlab("Generation") + ylab("Average Phenotypes")+
    scale_color_manual(name = "Species", values = c("Newt" = "darkred", "Snake" = "steelblue"))+
    theme(plot.title = element_text(size=5)) +
    theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))+
    geom_point(aes(x=factor(as.numeric(Number)), y=as.numeric(x), color=type) )
}

 if(leg_off==TRUE){
 p = p+ theme(legend.position = "none")}

return(p)
}

phenotype_dif_boxplot <- function(Newt_file, Snake_file, title, subt, boxy=FALSE, section=FALSE, section_num=0, leg_off=FALSE){
  #this function can graph to difference between phenotypes of all of the simulations (as lines or boxplot) by generations
  #(I can specify the amount of generations to show)

  data_frame_newt_temp <- Newt_file
  data_frame_newt_temp <- data_frame_newt_temp[ , !(names(data_frame_newt_temp) %in% "variable")]

  data_frame_snake_temp <- Snake_file 
  data_frame_snake_temp <- data_frame_snake_temp[ , !(names(data_frame_snake_temp) %in% "variable")]

  if(section==TRUE){

   selcol = seq(1,ncol(data_frame_newt_temp), section_num)
   data_frame_newt_temp <- data_frame_newt_temp %>% select(all_of(selcol))
   data_frame_snake_temp <- data_frame_snake_temp %>% select(all_of(selcol))

 }

  Number <- colnames(data_frame_newt_temp)
  data_frame_newt <- gather(data_frame_newt_temp, Number, value)
  data_frame_snake <- gather(data_frame_snake_temp, Number, value)
  dif <- as.numeric(data_frame_snake$value)-as.numeric(data_frame_newt$value)

  df_dif <- as.data.frame(dif)
  df_dif$Number <- as.numeric(data_frame_newt$Number)

  if(boxy==TRUE){
 p <- ggplot(df_dif) +
    ggtitle(title, subtitle = subt) +
    xlab("Generation") + ylab("Average Phenotype Differences")+
    theme(plot.title = element_text(size=5)) +
    theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))+
    geom_boxplot(aes(x=factor(as.numeric(Number)), y=as.numeric(dif)), color="darkolivegreen" )
 }
 if(boxy==FALSE){
   usims<-length(unique(Newt_file$variable))
   df_dif$trial <- rep(paste0("sim_",LETTERS[1:usims]), ncol(Newt_file)-1)
   p <- ggplot(df_dif) +
    ggtitle(title, subtitle = subt) +
    #xlim(100, 5000)+
    ylim(-5, 5)+
    xlab("Generation") + ylab("Average Phenotype Differences")+
    theme(plot.title = element_text(size=5)) +
    theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))+
    geom_line(aes(x=as.numeric(Number), y=as.numeric(dif), color=factor(trial)))
    #scale_color_brewer(palette="viridis")
    #theme(legend.position="none")
    #scale_color_manual(values = rep("darkolivegreen", nrow(data_frame_newt_temp)))
   p
 }
   if(leg_off==TRUE){
 p = p+ theme(legend.position = "none")}
  return(p)

}

phenotype_change_graph_s <- function(files, title, is_log=FALSE, xlim_value){
  # this shows me how the phenotype change in one single for each simulation
  # it takes all of the files and reads all of them in to take the first and last point
  # highly inefficient
file1 <- read.table(files[1], header = TRUE)
p <- ggplot(file1, aes(x=1:144)) +
  xlim(0, xlim_value)+
  ggtitle(title) +
  xlab("Simulation number") + ylab("Phenotype")+
  theme_bw()

if(is_log==TRUE){
  p = p + scale_y_log10() + ylab("Phenotype  (Log)")

}

i=1
for (current_file in files){

  file <- read.table(current_file, header = TRUE)
  file$Newt_mean_Pheno[file$Newt_mean_Pheno == "NULL"] <- 0
  file$Snake_mean_Pheno[file$Snake_mean_Pheno == "NULL"] <- 0
  file <- file[complete.cases(file), ]
 p <- p + geom_point(data= file, x = i, aes(y = as.numeric(Newt_mean_Pheno[1])+0.1), color= darken("darkred")) +
   geom_point(data= file,x = i, aes(y = as.numeric(Newt_mean_Pheno[50])+0.1), color=lighten("darkred"))  + 
  geom_point(data= file, x = i, aes(y = as.numeric(Snake_mean_Pheno[1])+0.1), color=darken("steelblue")) +
  geom_point(data= file, x = i, aes(y = as.numeric(Snake_mean_Pheno[50])+0.1), color=lighten("steelblue"))+
  geom_segment(data = file, x=i, xend=i, aes( y=as.numeric(Newt_mean_Pheno[1])+0.1, yend=as.numeric(Newt_mean_Pheno[50])+0.1, colour="Newt"), size = 0.3, arrow = arrow(length = unit(0.5, "cm")))+
   geom_segment(data = file, x=i, xend=i, aes( y=as.numeric(Snake_mean_Pheno[1])+0.1, yend=as.numeric(Snake_mean_Pheno[50])+0.1, colour="Snake"), size = 0.3, arrow = arrow(length = unit(0.5, "cm")))

 i = i+1
}
p <- p +
 scale_colour_manual(name = "Species", values = c("Newt" = "darkred", "Snake"="steelblue"))
return(p)
}

```

```{r, create dataframe}
 files <- list.files(path="~/Desktop/GA_lt/gen_5000/lit/text_files", pattern="GA_lit_", full.names=TRUE, recursive=FALSE)
 file <- read.table(files[1], header = TRUE)
 Newt_mean_Pheno <- making_data_frames(files, all_of("Newt_mean_Pheno"))
 Snake_mean_Pheno <- making_data_frames(files, all_of("Snake_mean_Pheno"))
 Newt_pop_size <- making_data_frames(files, all_of("Newt_pop_size"))
 Snake_pop_size <- making_data_frames(files, all_of("Snake_pop_size"))
```

```{r, plots compairing sims}

# a1<-phenotype_trends(Newt_mean_Pheno, "Newt Phenotype Simulation Trend", "16 trials", "darkred", is_log=T, leg_off=FALSE)
# a2<-phenotype_trends(Snake_mean_Pheno, "Snake Phenotype Simulation Trend", "16 trial", "steelblue", is_log=T, leg_off=FALSE)
# 
# grid.arrange(a1, a2, nrow=1)
# 
# 
# bp2 <- phenotype_dif_boxplot(Newt_mean_Pheno, Snake_mean_Pheno, title="Space: Differences", subt="16 trial",boxy=FALSE, section=FALSE, section_num=0)
# 
# grid.arrange(bp2, nrow=1)
```

```{r, phenotypes at differnt costs data gather}

files <- list.files(path="~/Desktop/GA_lt/gen_5000/lit/text_files", pattern="GA_lit_", full.names=TRUE, recursive=FALSE)
number=1
GA_file <- read.table(files[number], header = TRUE)
#file_name <- substr(files[number], 36, 186)
#print(paste0("File Name: ", file_name))

GA_file$sim <-"A"

for(i in seq(2, length(files), 1)){
  temp_file <- read.table(files[i], header = TRUE)
  temp_file$sim <-LETTERS[i]
  GA_file <- rbind(GA_file, temp_file)

}
GA_file$dif <- GA_file$Snake_mean_Pheno-GA_file$Newt_mean_Pheno
```

```{r easier dif plot}

pl1 <- ggplot(GA_file, aes(gen, Newt_mean_Pheno, col=factor(sim))) + geom_line()+ 
    ggtitle("Newt Mean Pheno by Gen", subtitle = "Different sim") +
    ylim(0,10)+
    xlab("Generations") + ylab("Phenotype")+
    theme_bw()

pl2 <- ggplot(GA_file, aes(gen, Snake_mean_Pheno, col=factor(sim))) + geom_line()+ 
    ggtitle("Snake Mean Pheno by Gen", subtitle = "Different Cost") +
    ylim(0,10)+
    xlab("Generations") + ylab("Phenotype")+
    theme_bw() 
grid.arrange(pl1, pl2, nrow=1)

ggplot(GA_file, aes(gen, dif, col=factor(sim))) + geom_line()+ 
    ggtitle("Snake Mean Pheno by Gen", subtitle = "Different Sim") +
    ylim(-5,5)+
    xlab("Generations") + ylab("Phenotype")+
    theme_bw()

```

First things that I noticed it that there are some newt and snake phenotypes that start of really high

lets look at other things like population size and newt deaths

```{r population size}

sl1 <- ggplot(GA_file, aes(gen, Newt_pop_size, col=factor(sim))) + geom_line()+ 
    ggtitle("Newt pop-size by Gen", subtitle = "Different sim") +
    #ylim(0,10)+
    xlab("Generations") + ylab("Phenotype")+
    theme_bw()

sl2 <- ggplot(GA_file, aes(gen, Snake_pop_size, col=factor(sim))) + geom_line()+ 
    ggtitle("Snake pop-size by Gen", subtitle = "Different Cost") +
    #ylim(0,10)+
    xlab("Generations") + ylab("Phenotype")+
    theme_bw() 
grid.arrange(sl1, sl2, nrow=1)

vcf1 <- var_by_var(subset(Newt_pop_size, select = -c(2)), subset(Newt_mean_Pheno, select = -c(2)), subset(Newt_mean_Pheno, select = -c(2)), subset(Snake_pop_size, select = -c(2)), title="Space: Newt VS Snake Population Size Color=Phenotype", subt="sigma = 0.1, mu = 1.0e-10", y_label="Snake Popsize", x_label="Newt Popsize", colored=TRUE, color_label="Newt Phenotype" , highcol="darkred", facet=FALSE)

vcf2 <- var_by_var(subset(Newt_pop_size, select = -c(2)), subset(Snake_mean_Pheno, select = -c(2)), subset(Snake_mean_Pheno, select = -c(2)), subset(Snake_pop_size, select = -c(2)), title="Space: Newt VS Snake Population Size Color=Phenotype", subt="sigma = 0.1, mu = 1.0e-10", y_label="Snake Popsize", x_label="Newt Popsize", colored=TRUE, color_label="Snake Phenotype", highcol="steelblue", facet=FALSE)

vc1 <- var_by_var(subset(Newt_pop_size, select = -c(2)), subset(Newt_mean_Pheno, select = -c(2)), subset(Newt_mean_Pheno, select = -c(2)), subset(Snake_pop_size, select = -c(2)), title="Space: Newt VS Snake Population Size Color=Phenotype", subt="sigma = 0.1, mu = 1.0e-10", y_label="Snake Popsize", x_label="Newt Popsize", colored=TRUE, color_label="Newt Phenotype" , highcol="darkred", facet=TRUE)

vc2 <- var_by_var(subset(Newt_pop_size, select = -c(2)), subset(Snake_mean_Pheno, select = -c(2)), subset(Snake_mean_Pheno, select = -c(2)), subset(Snake_pop_size, select = -c(2)), title="Space: Newt VS Snake Population Size Color=Phenotype", subt="sigma = 0.1, mu = 1.0e-10", y_label="Snake Popsize", x_label="Newt Popsize", colored=TRUE, color_label="Snake Phenotype", highcol="steelblue", facet=TRUE)

vcf1 
vcf2
vc1 
vc2
```

```{r, whale}
title="Space: Newt VS Snake Population Size Color=S-N Phenotype"
subt="16sims"
y_label="Snake Popsize"
x_label="Newt Popsize"
dif <- ggplot(GA_file[ which(GA_file$gen!=1),]) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = subt) +
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=as.numeric(Newt_pop_size), y=as.numeric(Snake_pop_size), color = dif))+
  theme(plot.title = element_text(size=5)) +
  scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
dif
```

```{r,pheno by popsize}

svc1 <- var_by_var(subset(Newt_mean_Pheno, select = -c(2)), subset(Newt_pop_size, select = -c(2)), subset(Newt_pop_size, select = -c(2)), subset(Snake_mean_Pheno, select = -c(2)), title="Space: Newt VS Snake Phenotype Color=Population Size", subt="Newt", y_label="Snake Phenotype", x_label="Newt Phenotype", colored=TRUE, color_label="Newt Popsize" , highcol="darkred", facet=TRUE)

svc2 <- var_by_var(subset(Newt_mean_Pheno, select = -c(2)), subset(Snake_pop_size, select = -c(2)), subset(Snake_pop_size, select = -c(2)), subset(Snake_mean_Pheno, select = -c(2)), title="Space: Newt VS Snake Phenotype Color=Population Size", subt="Snake", y_label="Snake Phenotype", x_label="Newt Phenotype", colored=TRUE, color_label="Snake Popsize", highcol="steelblue", facet=TRUE)

svc1 
svc2

```
```{r, mean predation}
GA_file$mean_predation <- GA_file$newt_deaths/GA_file$Newt_pop_size

title="Space: Newt VS Snake Population Size Color=Mean_predation"
subt="16sims newt_deaths/Newt_pop_size"
y_label="Snake Popsize"
x_label="Newt Popsize"
pre <- ggplot(GA_file[ which(GA_file$gen!=1),]) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = subt) +
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=as.numeric(Newt_pop_size), y=as.numeric(Snake_pop_size), color = mean_predation))+
  theme(plot.title = element_text(size=5)) +
  scale_colour_gradient2(high = "darkolivegreen4", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
pre

```

```{r, 20000 gen}
files <- list.files(path="~/Desktop/GA_lt/gen_20000/text_files", pattern="GA_lit_", full.names=TRUE, recursive=FALSE)
number=1
GA_file_20 <- read.table(files[number], header = TRUE)
#file_name <- substr(files[number], 36, 186)
#print(paste0("File Name: ", file_name))

GA_file_20$sim <-"A"

for(i in seq(2, length(files), 1)){
  temp_file <- read.table(files[i], header = TRUE)
  temp_file$sim <-LETTERS[i]
  GA_file_20 <- rbind(GA_file_20, temp_file)

}
GA_file_20$dif <- GA_file_20$Snake_mean_Pheno-GA_file_20$Newt_mean_Pheno

title="Space: Newt VS Snake Population Size Color=S-N Phenotype"
subt="20000 Gen 16sims"
y_label="Snake Popsize"
x_label="Newt Popsize"
dif <- ggplot(GA_file_20[ which(GA_file_20$gen!=1),]) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = subt) +
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=as.numeric(Newt_pop_size), y=as.numeric(Snake_pop_size), color = dif))+
  theme(plot.title = element_text(size=5)) +
  scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
dif
```

```{r, 100000 gen}
files <- list.files(path="~/Desktop/GA_lt/gen_100000/lit/text_files", pattern="GA_lit_", full.names=TRUE, recursive=FALSE)
number=1
GA_file_100 <- read.table(files[number], header = TRUE)
#file_name <- substr(files[number], 36, 186)
#print(paste0("File Name: ", file_name))

GA_file_100$sim <-"A"

for(i in seq(2, length(files), 1)){

  temp_file <- read.table(files[i], header = TRUE)
  temp_file$sim <-LETTERS[i]
  GA_file_100 <- rbind(GA_file_100, temp_file)

}
GA_file_100$dif <- GA_file_100$Snake_mean_Pheno-GA_file_100$Newt_mean_Pheno

title="Space: Newt VS Snake Population Size Color=S-N Phenotype"
subt="100000 Gen 16sims"
y_label="Snake Popsize"
x_label="Newt Popsize"
dif <- ggplot(GA_file_100[ which(GA_file_100$gen!=1),]) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = subt) +
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=as.numeric(Newt_pop_size), y=as.numeric(Snake_pop_size), color = dif))+
  theme(plot.title = element_text(size=5)) +
  scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
dif
```

```{r, what does the params have to do with it}
params <- read.csv("/Users/victoria/Desktop/Coevo/data/GAtestvar_lit.csv", header = TRUE)
params$Newt_mean_Pheno<-0
params$Snake_mean_Pheno<-0
params$Newt_pop_size<-0
params$Snake_pop_size<-0
params$dif<-0
params$Newt_mean_Pheno_20<-0
params$Snake_mean_Pheno_20<-0
params$Newt_pop_size_20<-0
params$Snake_pop_size_20<-0
params$dif_20<-0

params$Newt_mean_Pheno_100<-0
params$Snake_mean_Pheno_100<-0
params$Newt_pop_size_100<-0
params$Snake_pop_size_100<-0
params$dif_100<-0

for(i in seq(1, nrow(params), 1)){

  sumitup <- GA_file[ which(GA_file$gen>=1500 & GA_file$sim ==LETTERS[i]),]
  sumitup_20 <- GA_file_20[ which(GA_file_20$gen>=3500 & GA_file_20$sim ==LETTERS[i]),]
  params$Newt_mean_Pheno[i]<-mean(sumitup$Newt_mean_Pheno)
  params$Snake_mean_Pheno[i]<-mean(sumitup$Snake_mean_Pheno)
  params$Newt_pop_size[i]<-mean(sumitup$Newt_pop_size)
  params$Snake_pop_size[i]<-mean(sumitup$Snake_pop_size)
  params$dif[i]<-mean(sumitup$dif)
  params$Newt_mean_Pheno_20[i]<-mean(sumitup_20$Newt_mean_Pheno)
  params$Snake_mean_Pheno_20[i]<-mean(sumitup_20$Snake_mean_Pheno)
  params$Newt_pop_size_20[i]<-mean(sumitup_20$Newt_pop_size)
  params$Snake_pop_size_20[i]<-mean(sumitup_20$Snake_pop_size)
  params$dif_20[i]<-mean(sumitup_20$dif)

}

title="Space: Newt VS Snake Population Size"
subt="Gen 16sims"
y_label="Snake Popsize"
x_label="Newt Popsize"

params$snake <- paste0(params$snake_mu_rate, "_",params$snake_mu_effect_sd)
params$newt <- paste0(params$newt_mu_rate, "_",params$newt_mu_effect_sd)

t16 <- ggplot(params) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = subt) +
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=as.numeric(Newt_pop_size), y=as.numeric(Snake_pop_size), color = factor(snake), shape=factor(newt)),size=3)+
#  geom_point(aes(x=as.numeric(Newt_pop_size_20), y=as.numeric(Snake_pop_size_20), color = factor(snake), shape=factor(newt)))+
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  #facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
t16

t16d <- ggplot(params) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = subt) +
  xlab(x_label) + ylab("DIF: s-n")+
  #geom_point(aes(x=as.numeric(Newt_pop_size), y=as.numeric(Snake_pop_size), color = factor(snake), shape=factor(newt)))+
  geom_point(aes(x=as.numeric(Newt_pop_size), y=as.numeric(dif), color = factor(snake), shape=factor(newt)), size=3)+
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  #facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
t16d

t16d2 <- ggplot(params) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = subt) +
  xlab("Snake pop size") + ylab("DIF: s-n")+
  #geom_point(aes(x=as.numeric(Newt_pop_size), y=as.numeric(Snake_pop_size), color = factor(snake), shape=factor(newt)))+
  geom_point(aes(x=as.numeric(Snake_pop_size), y=as.numeric(dif), color = factor(snake), shape=factor(newt)), size=3)+
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  #facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
t16d2

```
```{r, 20000gen simple params}

t16g <- ggplot(params) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = "Gen 16sims 20000Gen") +
  xlab(x_label) + ylab(y_label)+
  #geom_point(aes(x=as.numeric(Newt_pop_size), y=as.numeric(Snake_pop_size), color = factor(snake), shape=factor(newt)))+
  geom_point(aes(x=as.numeric(Newt_pop_size_20), y=as.numeric(Snake_pop_size_20), color = factor(snake), shape=factor(newt)), size=3)+
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  #facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
t16g

```

```{r, 1000000gen}

```

```{r, building a heatmap}
                                      # Install reshape package

params_small <- params[-c(1:7, 9:20)]

data_melt <- melt(params_small)

ggp <- ggplot(data_melt, aes(newt, snake)) +                           # Create heatmap with ggplot2
  ggtitle("Newt GA by Snake GA", subtitle = "16 Sim Heatmap") +
  xlab("Newt GA") + ylab("Snake GA")+
  geom_tile(aes(fill = value))+ 
  scale_fill_gradient(low = "green", high = "black", name = "Newt Population Size") 
ggp

params_small <- params[-c(1:8, 10:20)]

data_melt <- melt(params_small)

ggp21 <- ggplot(data_melt, aes(newt, snake)) +    # Create heatmap with ggplot2
  ggtitle("Newt GA by Snake GA", subtitle = "16 Sim Heatmap") +
  xlab("Newt GA") + ylab("Snake GA")+
  geom_tile(aes(fill = value))+ 
  scale_fill_viridis_c(name = "Snake Population Size")
  #scale_fill_gradient(low = "green", high = "black") 
ggp21

params_small <- params[-c(1:9, 11:20)]
data_melt <- melt(params_small)

gdif <- ggplot(data_melt, aes(newt, snake)) +                           # Create heatmap with ggplot2
  ggtitle("Newt GA by Snake GA", subtitle = "16 Sim Heatmap") +
  xlab("Newt GA") + ylab("Snake GA")+
  geom_tile(aes(fill = value))+ 
  scale_fill_gradient2(low = "red",mid = "beige", high = "blue", name = "Snake-Newt") 
gdif

params_small <- params[-c(1:14, 16:20)]
data_melt <- melt(params_small)

gdif <- ggplot(data_melt, aes(newt, snake)) +                           # Create heatmap with ggplot2
  ggtitle("Newt GA by Snake GA", subtitle = "16 Sim Heatmap 20000") +
  xlab("Newt GA") + ylab("Snake GA")+
  geom_tile(aes(fill = value))+ 
  scale_fill_gradient2(low = "red",mid = "beige", high = "blue", name = "Snake-Newt") 
gdif

#TESTING
GA_file_20$Newt_mean_Pheno
ggplot(GA_file_20, aes(gen, dif, col=factor(sim))) + geom_line()+ 
  facet_wrap(~sim)+
    ggtitle("Snake Mean Pheno by Gen", subtitle = "Different Sim") +
    ylim(-5,5)+
    xlab("Generations") + ylab("Phenotype")+
    theme_bw()

ggplot(GA_file_20, aes(gen, Newt_mean_Pheno, col=factor(sim))) + geom_line()+ 
  facet_wrap(~sim)+
    ggtitle("Newt_mean_Pheno by Gen", subtitle = "Different Sim") +
    ylim(-5,5)+
    xlab("Generations") + ylab("Phenotype")+
    theme_bw()

ggplot(GA_file_20, aes(gen, Snake_mean_Pheno, col=factor(sim))) + geom_line()+ 
  facet_wrap(~sim)+
    ggtitle("Snake_mean_Pheno by Gen", subtitle = "Different Sim") +
    ylim(-5,5)+
    xlab("Generations") + ylab("Phenotype")+
    theme_bw()

ggplot(GA_file_20, aes(gen, Newt_pop_size, col=factor(sim))) + geom_line()+ 
    ggtitle("Snake Mean Pheno by Gen", subtitle = "Different Sim") +

    xlab("Generations") + ylab("Newt_pop_size")+
    theme_bw()

ggplot(GA_file_20, aes(gen, Snake_pop_size, col=factor(sim))) + geom_line()+ 
    ggtitle("Snake Mean Pheno by Gen", subtitle = "Different Sim") +

    xlab("Generations") + ylab("Snake_pop_size")+
    theme_bw()

```

thoughts

a simulation with no interaction

check that that column is the variance 
mu*sig^2 = 2.5e-13
2.5e-12 vs 2.5e-13

try to check out the sigma values 0.015
1e-8*(0.005)^2=2.5e-13
1e-9*(0.05)^2=2.5e-12
1e-10*(0.5)^2=2.5e-11
1e-11*(5)^2=2.5e-10

lets set everything to 2.5e-11

Now making the smprime with these
1e-12*(5)^2 = 2.5e-11
1e-10*(0.5)^2 = 2.5e-11
1e-8*(0.05)^2 = 2.5e-11
1e-6*(0.005)^2 = 2.5e-11