### 2022-03-14 graphs_newt_snakes/Correlation.Rmd
i am making more figures for the comparing GA specifically boxplots. I also plotted all the data from the cor datasets and they look like fuzzy catterplilars

---
title: "Correlation"
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE,echo = FALSE)

library(ggplot2)
library(viridis)
library(colorspace)
library(gridExtra)
library(tidyverse)
require(reshape2)
library(plyr)
library(gridGraphics)
library(grid)
library("reshape")
```

## Goals

I am starting to investigate what is going on in small scales in my co-evolution simulation. I decided to get newt and snake information and see in they were correlated in 5-5 grids. I collected correlated data from the GA experiments from the 5,000, 20,000, and 100,000 simulations. I also will examine how the correlation values change when the grid has more boxes (smaller area to compare).

GA1 experiment values

- 1e-8*(0.005)^2 = 2.5e-13
- 1e-9*(0.05)^2 = 2.5e-12
- 1e-10*(0.5)^2 = 2.5e-11
- 1e-11*(5.0)^2 = 2.5e-10

GA2 experiment values

- 1e-8*(0.05)^2 = 2.5e-11
- 1e-9*(0.1581139)^2 = 2.5e-11
- 1e-10*(0.5)^2 = 2.5e-11
- 1e-11*(1.581139)^2 = 2.5e-11
- 1e-12*(5)^2 = 2.5e-11

collected info:
	mean_newt_pheno_By_mean_snake_pheno = cor(c(newt_mean_pheno), c(snake_mean_pheno));
	num_newts_By_num_snakes = cor(c(newt_sum_ind), c(snake_sum_ind));
	sum_newt_pheno_By_num_snake = cor(c(newt_sum_pheno), c(snake_sum_ind));
	sum_snake_pheno_By_num_newt = cor(c(snake_sum_pheno), c(newt_sum_ind));
	sum_newt_pheno_By_num_newt = cor(c(newt_sum_pheno), c(newt_sum_ind));
	sum_snake_pheno_By_snake_newt = cor(c(snake_sum_pheno), c(snake_sum_ind));

Tested Grids

- (5, 5)
- (8, 8)
- (10, 10)
- (20, 20)

```{r, Cor data gather}
#Reading in the names of files to cycle through them
lit_5000_files <- list.files(path="~/Desktop/cor_5000/text_files", pattern="GA6_lit_", full.names=TRUE, recursive=FALSE)
cor_5000_files <- list.files(path="~/Desktop/cor_5000/text_files", pattern="GA6_cor_", full.names=TRUE, recursive=FALSE)

lit_20000_files <- list.files(path="~/Desktop/cor_20000/text_files", pattern="GA6_lit_", full.names=TRUE, recursive=FALSE)
cor_20000_files <- list.files(path="~/Desktop/cor_20000/text_files", pattern="GA6_cor_", full.names=TRUE, recursive=FALSE)

lit_100000_files <- list.files(path="~/Desktop/cor_20000/text_files", pattern="GA6_lit_", full.names=TRUE, recursive=FALSE)
cor_100000_files <- list.files(path="~/Desktop/cor_20000/text_files", pattern="GA6_cor_", full.names=TRUE, recursive=FALSE)

lit_5000_files_88 <- list.files(path="~/Desktop/cor_5000/text_files", pattern="GA88_lit_", full.names=TRUE, recursive=FALSE)
cor_5000_files_88 <- list.files(path="~/Desktop/cor_5000/text_files", pattern="GA88_cor_", full.names=TRUE, recursive=FALSE)

lit_5000_files_1010 <- list.files(path="~/Desktop/cor_5000/text_files", pattern="GA1010_lit_", full.names=TRUE, recursive=FALSE)
cor_5000_files_1010 <- list.files(path="~/Desktop/cor_5000/text_files", pattern="GA1010_cor_", full.names=TRUE, recursive=FALSE)

GA2_lit_5000_files <- list.files(path="~/Desktop/cor_GA2_5000/text_files", pattern="GA55_lit_", full.names=TRUE, recursive=FALSE)
GA2_cor_5000_files <- list.files(path="~/Desktop/cor_GA2_5000/text_files", pattern="GA55_cor_", full.names=TRUE, recursive=FALSE)

#files <- list.files(path="~/Desktop/GA_lt/doover/text_files", pattern="GA_lit_", full.names=TRUE, recursive=FALSE)
# Reading in the first file from the list to add a new column. To compare simulations with different params
number=1
cor_5000_file <- read.table(cor_5000_files[number], header = TRUE)
GA_5000_file <- read.table(lit_5000_files[number], header = TRUE)

cor_20000_file <- read.table(cor_20000_files[number], header = TRUE)
GA_20000_file <- read.table(lit_20000_files[number], header = TRUE)

cor_100000_file <- read.table(cor_100000_files[number], header = TRUE)
GA_100000_file <- read.table(lit_100000_files[number], header = TRUE)

cor_5000_file_88 <- read.table(cor_5000_files_88[number], header = TRUE)
GA_5000_file_88 <- read.table(lit_5000_files_88[number], header = TRUE)

cor_5000_file_1010 <- read.table(cor_5000_files_1010[number], header = TRUE)
GA_5000_file_1010 <- read.table(lit_5000_files_1010[number], header = TRUE)

GA2_cor_5000_file <- read.table(GA2_cor_5000_files[number], header = TRUE)
GA2_GA_5000_file <- read.table(GA2_lit_5000_files[number], header = TRUE)

GA_5000_file$sim <-"A"
cor_5000_file$sim <-"A"

GA_20000_file$sim <-"A"
cor_20000_file$sim <-"A"

GA_100000_file$sim <-"A"
cor_100000_file$sim <-"A"

GA_5000_file_88$sim <-"A"
cor_5000_file_88$sim <-"A"

GA_5000_file_1010$sim <-"A"
cor_5000_file_1010$sim <-"A"

GA2_GA_5000_file$sim <-"A"
GA2_cor_5000_file$sim <-"A"

#Cycle through the different simulations to make one file. Not recommended for large files, but right now mine are small
for(i in seq(1, length(GA2_lit_5000_files), 1)){
  myletter <- LETTERS[i]

  GA2_lit_temp_file <- read.table(GA2_lit_5000_files[i], header = TRUE)
  GA2_lit_temp_file$sim <-LETTERS[i]
  GA2_cor_temp_file <- read.table(GA2_cor_5000_files[i], header = TRUE)
  GA2_cor_temp_file$sim <-LETTERS[i]

  GA2_GA_5000_file <- rbind(GA2_GA_5000_file, GA2_lit_temp_file)
  GA2_cor_5000_file <- rbind(GA2_cor_5000_file, GA2_cor_temp_file)

  if (i<=length(lit_5000_files)){
  lit_temp_file <- read.table(lit_5000_files[i], header = TRUE)
  lit_temp_file$sim <-LETTERS[i]

  lit_temp_file_2 <- read.table(lit_20000_files[i], header = TRUE)
  lit_temp_file_2$sim <-LETTERS[i]

  lit_temp_file_3 <- read.table(lit_100000_files[i], header = TRUE)
  lit_temp_file_3$sim <-LETTERS[i]

  lit_temp_file_4 <- read.table(lit_5000_files_88[i], header = TRUE)
  lit_temp_file_4$sim <-LETTERS[i]

  lit_temp_file_5 <- read.table(lit_5000_files_1010[i], header = TRUE)
  lit_temp_file_5$sim <-LETTERS[i]

  cor_temp_file <- read.table(cor_5000_files[i], header = TRUE)
  cor_temp_file$sim <-LETTERS[i]

  cor_temp_file_2 <- read.table(cor_20000_files[i], header = TRUE)
  cor_temp_file_2$sim <-LETTERS[i]

  cor_temp_file_3 <- read.table(cor_100000_files[i], header = TRUE)
  cor_temp_file_3$sim <-LETTERS[i]

  cor_temp_file_4 <- read.table(cor_5000_files_88[i], header = TRUE)
  cor_temp_file_4$sim <-LETTERS[i]

  cor_temp_file_5 <- read.table(cor_5000_files_1010[i], header = TRUE)
  cor_temp_file_5$sim <-LETTERS[i]

  GA_5000_file <- rbind(GA_5000_file, lit_temp_file)
  GA_20000_file <- rbind(GA_20000_file, lit_temp_file_2)
  GA_100000_file <- rbind(GA_100000_file, lit_temp_file_3)
  GA_5000_file_88 <- rbind(GA_5000_file_88, lit_temp_file_4)
  GA_5000_file_1010 <- rbind(GA_5000_file_1010, lit_temp_file_5)

  cor_5000_file <- rbind(cor_5000_file, cor_temp_file)
  cor_20000_file <- rbind(cor_20000_file, cor_temp_file_2)
  cor_100000_file <- rbind(cor_100000_file, cor_temp_file_3)
  cor_5000_file_88 <- rbind(cor_5000_file_88, cor_temp_file_4)
  cor_5000_file_1010 <- rbind(cor_5000_file_1010, cor_temp_file_5)

  snake_mu_rate <- unlist(str_split(unlist(str_split(lit_5000_files[i],"snake_mu_rate_"))[2], "_"))[1]
  newt_mu_rate <- unlist(str_split(unlist(str_split(lit_5000_files[i],"newt_mu_rate_"))[2], "_"))[1]
  snake_mu_effect_sd <- unlist(str_split(unlist(str_split(lit_5000_files[i],"snake_mu_effect_sd_"))[2], "_"))[1]
  newt_mu_effect_sd <- unlist(str_split(unlist(str_split(lit_5000_files[i],"newt_mu_effect_sd_"))[2], "_"))[1]
  print(paste0("Simulation ", LETTERS[i],": Snake mu-rate & effect sd (", snake_mu_rate,", ", snake_mu_effect_sd, ") Newt mu-rate & effect sd (", newt_mu_rate, ", ", newt_mu_effect_sd, ")" ))
  }

}

# Calculating the difference between mean snake pheno and mean newt pheno

GA_5000_file$dif <- GA_5000_file$Snake_mean_Pheno-GA_5000_file$Newt_mean_Pheno
GA_20000_file$dif <- GA_20000_file$Snake_mean_Pheno-GA_20000_file$Newt_mean_Pheno
GA_100000_file$dif <- GA_100000_file$Snake_mean_Pheno-GA_100000_file$Newt_mean_Pheno
GA_5000_file_88$dif <- GA_5000_file_88$Snake_mean_Pheno-GA_5000_file_88$Newt_mean_Pheno
GA_5000_file_1010$dif <- GA_5000_file_1010$Snake_mean_Pheno-GA_5000_file_1010$Newt_mean_Pheno
GA2_GA_5000_file$dif <- GA2_GA_5000_file$Snake_mean_Pheno-GA2_GA_5000_file$Newt_mean_Pheno
GA2_GA_5000_file$dif <- GA2_GA_5000_file$Snake_mean_Pheno-GA2_GA_5000_file$Newt_mean_Pheno
```

```{r, reusable figures, results='hide', message=FALSE, warning=FALSE}

cors_by_gen <- function(file, title, subt, x_label, y_label){

plot <- ggplot(file) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = subt) +
  xlab(x_label) + ylab(y_label)+
  geom_line(aes(x=as.numeric(gen), y=as.numeric(mean_newt_pheno_By_mean_snake_pheno), color = "mean_newt_pheno_By_mean_snake_pheno"))+
  geom_line(aes(x=as.numeric(gen), y=as.numeric(num_newts_By_num_snakes), color = "num_newts_By_num_snakes"))+
  geom_line(aes(x=as.numeric(gen), y=as.numeric(sum_newt_pheno_By_num_snake), color = "sum_newt_pheno_By_num_snake"))+
  geom_line(aes(x=as.numeric(gen), y=as.numeric(sum_snake_pheno_By_num_newt), color = "sum_snake_pheno_By_num_newt"))+
  theme(plot.title = element_text(size=5)) +

  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))

  return(plot)
}
trends_by_gen <- function(file, title, subt, x_label, y_label){
  plot <- ggplot(file) +
    facet_wrap(~sim)+
    geom_line(aes(gen, dif, col="dif"))+
    ggtitle(title, subtitle = subt) +
    geom_line(aes(gen, Snake_mean_Pheno, color="Snake_mean_Pheno"),)+
    geom_line(aes(gen, Newt_mean_Pheno, color="Newt_mean_Pheno"))+
    ylim(-5,10)+
    xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake_mean_Pheno" = "steelblue", "Newt_mean_Pheno" = "darkred", "dif" = "black"),
labels = c("Snake", "Newt", "Dif"))+
    theme_bw() +theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
  return(plot)
}

```

```{r, lest see that correlation}

title="Space: Newt VS Snake"
subt="5,000 Gen Grid 5 by 5"
y_label="Generation"
x_label="Correlation"
cors_by_gen(cor_5000_file, title, subt, x_label, y_label)

y_label="Phenotype"
title="Phenotpe by Gen For all Sims"
subt="GA1 5000 Gen"
trends_by_gen(GA_5000_file, title, subt, x_label, y_label)

```
```{r, cor 5000 zooom}

title="ZOOM Space: Newt VS Snake"
subt="5,000 Gen Grid 5 by 5"
y_label="Generation"
x_label="Correlation"
cors_by_gen(cor_5000_file[ which(cor_5000_file$gen<200),], title, subt, x_label, y_label)

title="ZOOM Phenotpe by Gen For all Sims"
subt="GA1 5000 Gen"
y_label="Phenotype"
trends_by_gen(GA_5000_file[ which(GA_5000_file$gen<200),], title, subt, x_label, y_label)

```

```{r, correlation 20000}

title="Space: Newt VS Snake"
subt="20,000 Gen Grid 5 by 5"
y_label="Generation"
x_label="Correlation"
cors_by_gen(cor_20000_file, title, subt, x_label, y_label)

title="Phenotpe by Gen For all Sims"
subt="GA1 20000 Gen"
y_label="Phenotype"
trends_by_gen(GA_20000_file, title, subt, x_label, y_label)

```

```{r, correlation 100000}

title="Space: Newt VS Snake"
subt="100,000 Gen Grid 5 by 5"
y_label="Generation"
x_label="Correlation"
cors_by_gen(cor_100000_file, title, subt, x_label, y_label)

title="Phenotpe by Gen For all Sims"
subt="GA1 100000 Gen"
y_label="Phenotype"
trends_by_gen(GA_100000_file, title, subt, x_label, y_label)

```

```{r, correlation 5000 88}

title="Space: Newt VS Snake"
subt="5,000 Gen Grid 8 by 8"
y_label="Generation"
x_label="Correlation"
cors_by_gen(cor_5000_file_88, title, subt, x_label, y_label)

title="Phenotpe by Gen For all Sims"
subt="GA1 5,000 Gen"
y_label="Phenotype"
trends_by_gen(GA_5000_file_88, title, subt, x_label, y_label)

```

```{r, correlation 5000 1010}

title="Space: Newt VS Snake"
subt="5,000 Gen Grid 10 by 10"
y_label="Generation"
x_label="Correlation"
cors_by_gen(cor_5000_file_1010, title, subt, x_label, y_label)

title="Phenotpe by Gen For all Sims"
subt="GA1 5,000 Gen"
y_label="Phenotype"
trends_by_gen(GA_5000_file_1010, title, subt, x_label, y_label)

```

```{r, GA2 correlation}

title="GA2 Space: Newt VS Snake"
subt="5,000 Gen Grid 5 by 5"
y_label="Generation"
x_label="Correlation"
cors_by_gen(GA2_cor_5000_file, title, subt, x_label, y_label)

title="Phenotpe by Gen For all Sims"
subt="GA2 5,000 Gen"
y_label="Phenotype"
trends_by_gen(GA2_GA_5000_file, title, subt, x_label, y_label)

```

```{r, cor 5000 GA2 zooom}

title="GA2 ZOOM Space: Newt VS Snake"
subt="5,000 Gen Grid 5 by 5"
y_label="Generation"
x_label="Correlation"
cors_by_gen(GA2_cor_5000_file[ which(GA2_cor_5000_file$gen<200),], title, subt, x_label, y_label)

title="ZOOM Phenotpe by Gen For all Sims"
subt="GA2 5,000 Gen"
y_label="Phenotype"
trends_by_gen(GA2_GA_5000_file[ which(GA2_GA_5000_file$gen<200),], title, subt, x_label, y_label)

```