### 2022-03-23 graphs_newt_snakes/Correlation.Rmd
I put in the wild cards for the snakemake file. I also looked more at the correlations

---
title: "Correlation"
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE,echo = FALSE)

library(ggplot2)
library(viridis)
library(colorspace)
library(gridExtra)
library(tidyverse)
require(reshape2)
library(plyr)
library(gridGraphics)
library(grid)
library("reshape")
```

## Goals

I am starting to investigate what is going on in small scales in my co-evolution simulation. I decided to get newt and snake information and see in they were correlated in n-n grids. I collected correlated data from the GA experiments from the 5,000, 20,000, and 100,000 simulations. (I also tested the correlations with and without perUnitArea=T. It turns out when you don't have perUnitArea=T the edge of the simulation has a greater effect on the calculations turning the correlation positive.) I also will examine how the correlation values change when the grid has more boxes (smaller area to compare). I will specifically, test 5 by 5, 8 by 8, 10 by 10 and 20 by 20 grids

#Simulation Background

My simulation has newts and snakes coevolving in an arms race! The newts are toxic and interact with resistant snakes. Snakes are attacking and eating the newts, and in order to eat more newts they are becoming more resistant. However, if a snake eats a newt that is more toxic the snake dies. As snakes become more resistant, less toxic newts are being devoured, leaving only the most toxic newts. These toxic newts reproduce and create even more toxic newts. The cycle contentiously repeats. My simulation is run in a two-dimensional space with a x and y axis.

GA1 experiment values

- 1e-8*(0.005)^2 = 2.5e-13
- 1e-9*(0.05)^2 = 2.5e-12
- 1e-10*(0.5)^2 = 2.5e-11
- 1e-11*(5.0)^2 = 2.5e-10

GA2 experiment values

- 1e-8*(0.05)^2 = 2.5e-11
- 1e-9*(0.1581139)^2 = 2.5e-11
- 1e-10*(0.5)^2 = 2.5e-11
- 1e-11*(1.581139)^2 = 2.5e-11
- 1e-12*(5)^2 = 2.5e-11

collected info:
	mean_newt_pheno_By_mean_snake_pheno = cor(c(newt_mean_pheno), c(snake_mean_pheno));
	num_newts_By_num_snakes = cor(c(newt_sum_ind), c(snake_sum_ind));
	sum_newt_pheno_By_num_snake = cor(c(newt_sum_pheno), c(snake_sum_ind));
	sum_snake_pheno_By_num_newt = cor(c(snake_sum_pheno), c(newt_sum_ind));
	sum_newt_pheno_By_num_newt = cor(c(newt_sum_pheno), c(newt_sum_ind));
	sum_snake_pheno_By_snake_newt = cor(c(snake_sum_pheno), c(snake_sum_ind));

Tested Grids

- (5, 5)
- (8, 8)
- (10, 10)
- (20, 20)

Something I noticed when playing around with different variables in my simulation was that there was a large effect when I was changing interaction rate. So I will also look at that here. Special note when interaction rate is increased newts and snake may avoid each other.

```{r, Cor data gather}
#Reading in the names of files to cycle through them
lit_5000_files <- list.files(path="~/Desktop/cor_5000/cor_byUnit/text_files", pattern="GA5_lit_", full.names=TRUE, recursive=FALSE)
cor_5000_files <- list.files(path="~/Desktop/cor_5000/cor_byUnit/text_files", pattern="GA5_cor_", full.names=TRUE, recursive=FALSE)

lit_5000_files_HI <- list.files(path="~/Desktop/cor_5000/cor_byUnit/text_files", pattern="GA55_lit_", full.names=TRUE, recursive=FALSE)
cor_5000_files_HI <- list.files(path="~/Desktop/cor_5000/cor_byUnit/text_files", pattern="GA55_cor_", full.names=TRUE, recursive=FALSE)

lit_20000_files <- list.files(path="~/Desktop/cor_20000/cor_byUnit/text_files", pattern="GA55_lit_", full.names=TRUE, recursive=FALSE)
cor_20000_files <- list.files(path="~/Desktop/cor_20000/cor_byUnit/text_files", pattern="GA55_cor_", full.names=TRUE, recursive=FALSE)

lit_100000_files <- list.files(path="~/Desktop/cor_20000/cor_byUnit/text_files", pattern="GA55_lit_", full.names=TRUE, recursive=FALSE)
cor_100000_files <- list.files(path="~/Desktop/cor_20000/cor_byUnit/text_files", pattern="GA55_cor_", full.names=TRUE, recursive=FALSE)

lit_5000_files_88 <- list.files(path="~/Desktop/cor_5000/cor_byUnit/text_files", pattern="GA88_lit_", full.names=TRUE, recursive=FALSE)
cor_5000_files_88 <- list.files(path="~/Desktop/cor_5000/cor_byUnit/text_files", pattern="GA88_cor_", full.names=TRUE, recursive=FALSE)

lit_5000_files_1010 <- list.files(path="~/Desktop/cor_5000/cor_byUnit/text_files", pattern="GA1010_lit_", full.names=TRUE, recursive=FALSE)
cor_5000_files_1010 <- list.files(path="~/Desktop/cor_5000/cor_byUnit/text_files", pattern="GA1010_cor_", full.names=TRUE, recursive=FALSE)

lit_5000_files_2020 <- list.files(path="~/Desktop/cor_5000/cor_byUnit/text_files", pattern="GA2020_lit_", full.names=TRUE, recursive=FALSE)
cor_5000_files_2020 <- list.files(path="~/Desktop/cor_5000/cor_byUnit/text_files", pattern="GA2020_cor_", full.names=TRUE, recursive=FALSE)

GA2_lit_5000_files <- list.files(path="~/Desktop/cor_GA2_5000/cor_byUnit/text_files", pattern="GA55_lit_", full.names=TRUE, recursive=FALSE)
GA2_cor_5000_files <- list.files(path="~/Desktop/cor_GA2_5000/cor_byUnit/text_files", pattern="GA55_cor_", full.names=TRUE, recursive=FALSE)

#files <- list.files(path="~/Desktop/GA_lt/doover/text_files", pattern="GA_lit_", full.names=TRUE, recursive=FALSE)
# Reading in the first file from the list to add a new column. To compare simulations with different params
number=1
cor_5000_file <- read.table(cor_5000_files[number], header = TRUE)
GA_5000_file <- read.table(lit_5000_files[number], header = TRUE)

cor_20000_file <- read.table(cor_20000_files[number], header = TRUE)
GA_20000_file <- read.table(lit_20000_files[number], header = TRUE)

cor_100000_file <- read.table(cor_100000_files[number], header = TRUE)
GA_100000_file <- read.table(lit_100000_files[number], header = TRUE)

cor_5000_file_88 <- read.table(cor_5000_files_88[number], header = TRUE)
GA_5000_file_88 <- read.table(lit_5000_files_88[number], header = TRUE)

cor_5000_file_1010 <- read.table(cor_5000_files_1010[number], header = TRUE)
GA_5000_file_1010 <- read.table(lit_5000_files_1010[number], header = TRUE)

cor_5000_file_2020 <- read.table(cor_5000_files_2020[number], header = TRUE)
GA_5000_file_2020 <- read.table(lit_5000_files_2020[number], header = TRUE)

GA2_cor_5000_file <- read.table(GA2_cor_5000_files[number], header = TRUE)
GA2_GA_5000_file <- read.table(GA2_lit_5000_files[number], header = TRUE)

GA_5000_file$sim <-"A"
cor_5000_file$sim <-"A"

GA_20000_file$sim <-"A"
cor_20000_file$sim <-"A"

GA_100000_file$sim <-"A"
cor_100000_file$sim <-"A"

GA_5000_file_88$sim <-"A"
cor_5000_file_88$sim <-"A"

GA_5000_file_1010$sim <-"A"
cor_5000_file_1010$sim <-"A"

GA_5000_file_2020$sim <-"A"
cor_5000_file_2020$sim <-"A"

GA2_GA_5000_file$sim <-"A"
GA2_cor_5000_file$sim <-"A"

#Cycle through the different simulations to make one file. Not recommended for large files, but right now mine are small
for(i in seq(1, length(GA2_lit_5000_files), 1)){
  myletter <- LETTERS[i]

  GA2_lit_temp_file <- read.table(GA2_lit_5000_files[i], header = TRUE)
  GA2_lit_temp_file$sim <-LETTERS[i]
  GA2_cor_temp_file <- read.table(GA2_cor_5000_files[i], header = TRUE)
  GA2_cor_temp_file$sim <-LETTERS[i]

  GA2_GA_5000_file <- rbind(GA2_GA_5000_file, GA2_lit_temp_file)
  GA2_cor_5000_file <- rbind(GA2_cor_5000_file, GA2_cor_temp_file)

  if (i<=length(lit_5000_files)){
  lit_temp_file <- read.table(lit_5000_files[i], header = TRUE)
  lit_temp_file$sim <-LETTERS[i]

  lit_temp_file_2 <- read.table(lit_20000_files[i], header = TRUE)
  lit_temp_file_2$sim <-LETTERS[i]

  lit_temp_file_3 <- read.table(lit_100000_files[i], header = TRUE)
  lit_temp_file_3$sim <-LETTERS[i]

  lit_temp_file_4 <- read.table(lit_5000_files_88[i], header = TRUE)
  lit_temp_file_4$sim <-LETTERS[i]

  lit_temp_file_5 <- read.table(lit_5000_files_1010[i], header = TRUE)
  lit_temp_file_5$sim <-LETTERS[i]

  lit_temp_file_6 <- read.table(lit_5000_files_2020[i], header = TRUE)
  lit_temp_file_6$sim <-LETTERS[i]

  cor_temp_file <- read.table(cor_5000_files[i], header = TRUE)
  cor_temp_file$sim <-LETTERS[i]

  cor_temp_file_2 <- read.table(cor_20000_files[i], header = TRUE)
  cor_temp_file_2$sim <-LETTERS[i]

  cor_temp_file_3 <- read.table(cor_100000_files[i], header = TRUE)
  cor_temp_file_3$sim <-LETTERS[i]

  cor_temp_file_4 <- read.table(cor_5000_files_88[i], header = TRUE)
  cor_temp_file_4$sim <-LETTERS[i]

  cor_temp_file_5 <- read.table(cor_5000_files_1010[i], header = TRUE)
  cor_temp_file_5$sim <-LETTERS[i]

  cor_temp_file_6 <- read.table(cor_5000_files_2020[i], header = TRUE)
  cor_temp_file_6$sim <-LETTERS[i]

  GA_5000_file <- rbind(GA_5000_file, lit_temp_file)
  GA_20000_file <- rbind(GA_20000_file, lit_temp_file_2)
  GA_100000_file <- rbind(GA_100000_file, lit_temp_file_3)
  GA_5000_file_88 <- rbind(GA_5000_file_88, lit_temp_file_4)
  GA_5000_file_1010 <- rbind(GA_5000_file_1010, lit_temp_file_5)
  GA_5000_file_2020 <- rbind(GA_5000_file_2020, lit_temp_file_6)

  cor_5000_file <- rbind(cor_5000_file, cor_temp_file)
  cor_20000_file <- rbind(cor_20000_file, cor_temp_file_2)
  cor_100000_file <- rbind(cor_100000_file, cor_temp_file_3)
  cor_5000_file_88 <- rbind(cor_5000_file_88, cor_temp_file_4)
  cor_5000_file_1010 <- rbind(cor_5000_file_1010, cor_temp_file_5)
  cor_5000_file_2020 <- rbind(cor_5000_file_2020, cor_temp_file_6)

  snake_mu_rate <- unlist(str_split(unlist(str_split(lit_5000_files[i],"snake_mu_rate_"))[2], "_"))[1]
  newt_mu_rate <- unlist(str_split(unlist(str_split(lit_5000_files[i],"newt_mu_rate_"))[2], "_"))[1]
  snake_mu_effect_sd <- unlist(str_split(unlist(str_split(lit_5000_files[i],"snake_mu_effect_sd_"))[2], "_"))[1]
  newt_mu_effect_sd <- unlist(str_split(unlist(str_split(lit_5000_files[i],"newt_mu_effect_sd_"))[2], "_"))[1]
  print(paste0("Simulation ", LETTERS[i],": Snake mu-rate & effect sd (", snake_mu_rate,", ", snake_mu_effect_sd, ") Newt mu-rate & effect sd (", newt_mu_rate, ", ", newt_mu_effect_sd, ")" ))
  }

}

# Calculating the difference between mean snake pheno and mean newt pheno

GA_5000_file$dif <- GA_5000_file$Snake_mean_Pheno-GA_5000_file$Newt_mean_Pheno
GA_20000_file$dif <- GA_20000_file$Snake_mean_Pheno-GA_20000_file$Newt_mean_Pheno
GA_100000_file$dif <- GA_100000_file$Snake_mean_Pheno-GA_100000_file$Newt_mean_Pheno
GA_5000_file_88$dif <- GA_5000_file_88$Snake_mean_Pheno-GA_5000_file_88$Newt_mean_Pheno
GA_5000_file_1010$dif <- GA_5000_file_1010$Snake_mean_Pheno-GA_5000_file_1010$Newt_mean_Pheno
GA_5000_file_2020$dif <- GA_5000_file_2020$Snake_mean_Pheno-GA_5000_file_2020$Newt_mean_Pheno
GA2_GA_5000_file$dif <- GA2_GA_5000_file$Snake_mean_Pheno-GA2_GA_5000_file$Newt_mean_Pheno
GA2_GA_5000_file$dif <- GA2_GA_5000_file$Snake_mean_Pheno-GA2_GA_5000_file$Newt_mean_Pheno
```

```{r, reusable figures, results='hide', message=FALSE, warning=FALSE}

cors_by_gen <- function(file, title, subt, x_label, y_label){
  print("TA = mean_newt_pheno_By_mean_snake_pheno")
  print("TB = num_newts_By_num_snakes")
  print("TC = sum_newt_pheno_By_num_snake")
  print("TD = sum_snake_pheno_By_num_newt")

  plot <- ggplot(file) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = subt) +
  xlab(x_label) + ylab(y_label)+
  geom_line(aes(x=as.numeric(gen), y=as.numeric(mean_newt_pheno_By_mean_snake_pheno), color = "TA"))+
  geom_line(aes(x=as.numeric(gen), y=as.numeric(num_newts_By_num_snakes), color = "TB"))+
  geom_line(aes(x=as.numeric(gen), y=as.numeric(sum_newt_pheno_By_num_snake), color = "TC"))+
  geom_line(aes(x=as.numeric(gen), y=as.numeric(sum_snake_pheno_By_num_newt), color = "TD"))+
  theme(plot.title = element_text(size=5)) +

  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))

  return(plot)
}
trends_by_gen <- function(file, title, subt, x_label, y_label){
  plot <- ggplot(file) +
    facet_wrap(~sim)+
    geom_line(aes(gen, dif, col="dif"))+
    ggtitle(title, subtitle = subt) +
    geom_line(aes(gen, Snake_mean_Pheno, color="Snake_mean_Pheno"),)+
    geom_line(aes(gen, Newt_mean_Pheno, color="Newt_mean_Pheno"))+
    ylim(-5,10)+
    xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake_mean_Pheno" = "steelblue", "Newt_mean_Pheno" = "darkred", "dif" = "black"),
labels = c("Snake", "Newt", "Dif"))+
    theme_bw() +theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
  return(plot)
}

```

```{r, lest see that correlation}

title="Space: Newt VS Snake"
subt="5,000 Gen Grid 5 by 5"
y_label="Generation"
x_label="Correlation"
cors_by_gen(cor_5000_file, title, subt, x_label, y_label)

y_label="Phenotype"
title="Phenotpe by Gen For all Sims"
subt="GA1 5000 Gen"
trends_by_gen(GA_5000_file, title, subt, x_label, y_label)

title="Space: Newt VS Snake"
subt="5,000 Gen Grid 5 by 5"
y_label="Generation"
x_label="Correlation"
my_letter="A"
subt=paste0("Sim ", subt, " ", my_letter)
cors_by_gen(cor_5000_file[which(cor_5000_file$sim==my_letter), ], title, subt, x_label, y_label)

```
```{r, cor 5000 zooom}

title="ZOOM Space: Newt VS Snake"
subt="5,000 Gen Grid 5 by 5"
y_label="Generation"
x_label="Correlation"
cors_by_gen(cor_5000_file[ which(cor_5000_file$gen<200),], title, subt, x_label, y_label)

title="ZOOM Phenotpe by Gen For all Sims"
subt="GA1 5000 Gen"
y_label="Phenotype"
trends_by_gen(GA_5000_file[ which(GA_5000_file$gen<200),], title, subt, x_label, y_label)

```

```{r, correlation 20000}

title="Space: Newt VS Snake"
subt="20,000 Gen Grid 5 by 5"
y_label="Generation"
x_label="Correlation"
cors_by_gen(cor_20000_file, title, subt, x_label, y_label)

title="Phenotpe by Gen For all Sims"
subt="GA1 20000 Gen"
y_label="Phenotype"
trends_by_gen(GA_20000_file, title, subt, x_label, y_label)

```

```{r, correlation 100000}

title="Space: Newt VS Snake"
subt="100,000 Gen Grid 5 by 5"
y_label="Generation"
x_label="Correlation"
cors_by_gen(cor_100000_file, title, subt, x_label, y_label)

title="Phenotpe by Gen For all Sims"
subt="GA1 100000 Gen"
y_label="Phenotype"
trends_by_gen(GA_100000_file, title, subt, x_label, y_label)

```

```{r, correlation 5000 88}

title="Space: Newt VS Snake"
subt="5,000 Gen Grid 8 by 8"
y_label="Generation"
x_label="Correlation"
cors_by_gen(cor_5000_file_88, title, subt, x_label, y_label)

title="Phenotpe by Gen For all Sims"
subt="GA1 5,000 Gen"
y_label="Phenotype"
trends_by_gen(GA_5000_file_88, title, subt, x_label, y_label)

```

```{r, correlation 5000 1010}

title="Space: Newt VS Snake"
subt="5,000 Gen Grid 10 by 10"
y_label="Generation"
x_label="Correlation"
cors_by_gen(cor_5000_file_1010, title, subt, x_label, y_label)

title="Phenotpe by Gen For all Sims"
subt="GA1 5,000 Gen"
y_label="Phenotype"
trends_by_gen(GA_5000_file_1010, title, subt, x_label, y_label)

```

```{r, correlation 5000 2020}

title="Space: Newt VS Snake"
subt="5,000 Gen Grid 20 by 20"
y_label="Generation"
x_label="Correlation"
cors_by_gen(cor_5000_file_2020, title, subt, x_label, y_label)

title="Phenotpe by Gen For all Sims"
subt="GA1 5,000 Gen"
y_label="Phenotype"
trends_by_gen(GA_5000_file_2020, title, subt, x_label, y_label)

```

```{r, GA2 correlation}

title="GA2 Space: Newt VS Snake"
subt="5,000 Gen Grid 5 by 5"
y_label="Generation"
x_label="Correlation"
cors_by_gen(GA2_cor_5000_file, title, subt, x_label, y_label)

title="Phenotpe by Gen For all Sims"
subt="GA2 5,000 Gen"
y_label="Phenotype"
trends_by_gen(GA2_GA_5000_file, title, subt, x_label, y_label)

```

```{r, cor 5000 GA2 zooom}

title="GA2 ZOOM Space: Newt VS Snake"
subt="5,000 Gen Grid 5 by 5"
y_label="Generation"
x_label="Correlation"
cors_by_gen(GA2_cor_5000_file[ which(GA2_cor_5000_file$gen<200),], title, subt, x_label, y_label)

title="ZOOM Phenotpe by Gen For all Sims"
subt="GA2 5,000 Gen"
y_label="Phenotype"
trends_by_gen(GA2_GA_5000_file[ which(GA2_GA_5000_file$gen<200),], title, subt, x_label, y_label)

```