### 2022-04-12 graphs_newt_snakes/correlation_part2.Rmd
I started a cor part 2 file to look more at the individual locations (both at previous data and slim data. I also am trying to figure out logs and functions to help me make better files

---
title: "correlation_part2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggmap)
library(ggplot2)
library(sf)
library(mapview)
```

## Comparing newt and snake phenotype correlation

newt and snake data from Hanifin et al. 2008

```{r, reak data cor}
snewt <- read.csv("~/Desktop/snake_newt_real_data.csv", header = TRUE)
snewt$mean_res = snewt$mean_toxin+snewt$dif
cor(snewt$mean_toxin, snewt$mean_res)
(sites <- data.frame(longitude = snewt$lon, latitude = snewt$lat))
just_loc <- data.frame(snewt$lon, snewt$lat)
snewt$lon<- snewt$lon *-1
locations_sf <- st_as_sf(snewt, coords = c("lon", "lat"), crs = 4326)
mapview(locations_sf)
```

```{r, more maps}
library("rnaturalearth")
library("rnaturalearthdata")
library("tools")
library("maps")

world <- ne_countries(scale = "medium", returnclass = "sf")
states <- st_as_sf(maps::map("state", plot = FALSE, fill = TRUE))
states$ID <- toTitleCase(states$ID)

ggplot(data = world) +
  geom_sf() +
  geom_sf(data = states, fill = NA)+
  geom_point(data = snewt, aes(x = lon, y = lat, fill = mean_res), size = 5, shape = 22, ) +
  geom_point(data = snewt, aes(x = lon, y = lat, fill = mean_toxin), size = 3, shape = 21, ) +
  scale_fill_viridis_c()+
  coord_sf(xlim = c(-130.4494, -111.6764), ylim = c(29.742, 54.9415), expand = FALSE)
```

## More data from the slim simulation

might be a specific generation or a small set of generations, If I can get the data to be saved and then interpreted by R
make a list in a column 
how to proportion and recreate what I see in the slim map (the different box and their dims)