### 2022-04-19 graphs_newt_snakes/correlation_part2.Rmd

---
title: "correlation_part2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggmap)
library(ggplot2)
library(sf)
library(mapview)
```

## Comparing newt and snake phenotype correlation

newt and snake data from Hanifin et al. 2008

```{r, reak data cor}
snewt <- read.csv("~/Desktop/snake_newt_real_data.csv", header = TRUE)
snewt$mean_res = snewt$mean_toxin+snewt$dif
cor(snewt$mean_toxin, snewt$mean_res)
(sites <- data.frame(longitude = snewt$lon, latitude = snewt$lat))
just_loc <- data.frame(snewt$lon, snewt$lat)
snewt$lon<- snewt$lon *-1
locations_sf <- st_as_sf(snewt, coords = c("lon", "lat"), crs = 4326)
mapview(locations_sf)
```

```{r, more maps}
library("rnaturalearth")
library("rnaturalearthdata")
library("tools")
library("maps")

world <- ne_countries(scale = "medium", returnclass = "sf")
states <- st_as_sf(maps::map("state", plot = FALSE, fill = TRUE))
states$ID <- toTitleCase(states$ID)

ggplot(data = world) +
  geom_sf() +
  geom_sf(data = states, fill = NA)+
  geom_point(data = snewt, aes(x = lon, y = lat, fill = mean_res), size = 5, shape = 22, ) +
  geom_point(data = snewt, aes(x = lon, y = lat, fill = mean_toxin), size = 3, shape = 21, ) +
  scale_fill_viridis_c()+
  coord_sf(xlim = c(-130.4494, -111.6764), ylim = c(29.742, 54.9415), expand = FALSE)
```

## More data from the slim simulation

might be a specific generation or a small set of generations, If I can get the data to be saved and then interpreted by R
make a list in a column 
how to proportion and recreate what I see in the slim map (the different box and their dims)

GA1 experiment values:

- 1e-8*(0.005)^2 = 2.5e-13
- 1e-9*(0.05)^2 = 2.5e-12
- 1e-10*(0.5)^2 = 2.5e-11
- 1e-11*(5.0)^2 = 2.5e-10

GA2 experiment values:

- 1e-8*(0.05)^2 = 2.5e-11
- 1e-9*(0.1581139)^2 = 2.5e-11
- 1e-10*(0.5)^2 = 2.5e-11
- 1e-11*(1.581139)^2 = 2.5e-11
- 1e-12*(5)^2 = 2.5e-11

```{r, cor by location}

cor_grid_30 <- list.files(path="~/Desktop/GA1_rep", pattern="_grid_", full.names=TRUE, recursive=FALSE)
cor_grid_30_G2 <- list.files(path="~/Desktop/GA2_rep", pattern="_grid_", full.names=TRUE, recursive=FALSE)

cor_grid_30_file<-read.csv(cor_grid_30[1], header = TRUE)
replica <- unlist(str_split(unlist(str_split(cor_grid_30[1],"rep_"))[3], "_"))[1]
cor_grid_30_file$rep<-replica
cor_grid_30_file$sim <-"A"
cor_grid_30_file$rep <-unlist(str_split(unlist(str_split(cor_grid_30[i],"rep_"))[3], "_"))[1]
cor_grid_30_file$snake_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30[1],"snake_mu_rate_"))[2], "_"))[1]
cor_grid_30_file$newt_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30[1],"newt_mu_rate_"))[2], "_"))[1]
cor_grid_30_file$snake_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30[1],"snake_mu_effect_sd_"))[2], "_"))[1]
cor_grid_30_file$newt_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30[1],"newt_mu_effect_sd_"))[2], "_"))[1]

cor_grid_30_G2_file<-read.csv(cor_grid_30_G2[1], header = TRUE)
replica <- unlist(str_split(unlist(str_split(cor_grid_30_G2[1],"rep_"))[3], "_"))[1]
cor_grid_30_G2_file$rep<-replica
cor_grid_30_G2_file$sim <-"A"
cor_grid_30_G2_file$rep <-unlist(str_split(unlist(str_split(cor_grid_30_G2[i],"rep_"))[3], "_"))[1]
cor_grid_30_G2_file$snake_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30_G2[1],"snake_mu_rate_"))[2], "_"))[1]
cor_grid_30_G2_file$newt_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30_G2[1],"newt_mu_rate_"))[2], "_"))[1]
cor_grid_30_G2_file$snake_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30_G2[1],"snake_mu_effect_sd_"))[2], "_"))[1]
cor_grid_30_G2_file$newt_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30_G2[1],"newt_mu_effect_sd_"))[2], "_"))[1]

i=2
oldstand=0
letter=1
for(i in seq(2, length(cor_grid_30_G2), 1)){
  stand <- unlist(str_split(unlist(str_split(cor_grid_30_G2[i],"newt_mu_rate_"))[2], "_"))[1]
  if (stand!=oldstand){
    oldstand=stand
    letter = letter + 1
  }

  GA2_lit_temp_file <- read.csv(cor_grid_30_G2[i], header = TRUE)
  GA2_lit_temp_file$sim <-LETTERS[letter]
  GA2_lit_temp_file$rep <-unlist(str_split(unlist(str_split(cor_grid_30_G2[i],"rep_"))[3], "_"))[1]
  GA2_lit_temp_file$snake_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30_G2[i],"snake_mu_rate_"))[2], "_"))[1]
GA2_lit_temp_file$newt_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30[i],"newt_mu_rate_"))[2], "_"))[1]
GA2_lit_temp_file$snake_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30[i],"snake_mu_effect_sd_"))[2], "_"))[1]
GA2_lit_temp_file$newt_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30[i],"newt_mu_effect_sd_"))[2], "_"))[1]

  cor_grid_30_G2_file <- rbind(cor_grid_30_G2_file, GA2_lit_temp_file)

  if (i<=length(cor_grid_30)){
  cor_temp_file <- read.csv(cor_grid_30[i], header = TRUE)
  cor_temp_file$sim <-LETTERS[letter]
  cor_temp_file$rep <-unlist(str_split(unlist(str_split(cor_grid_30[i],"rep_"))[3], "_"))[1]
  cor_temp_file$snake_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30[i],"snake_mu_rate_"))[2], "_"))[1]
cor_temp_file$newt_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30[i],"newt_mu_rate_"))[2], "_"))[1]
cor_temp_file$snake_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30[i],"snake_mu_effect_sd_"))[2], "_"))[1]
cor_temp_file$newt_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30[i],"newt_mu_effect_sd_"))[2], "_"))[1]
  cor_grid_30_file <- rbind(cor_grid_30_file, cor_temp_file)
  rep <-unlist(str_split(unlist(str_split(cor_grid_30[i],"rep_"))[3], "_"))[1]

  snake_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30[i],"snake_mu_rate_"))[2], "_"))[1]
  newt_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30[i],"newt_mu_rate_"))[2], "_"))[1]
  snake_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30[i],"snake_mu_effect_sd_"))[2], "_"))[1]
  newt_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30[i],"newt_mu_effect_sd_"))[2], "_"))[1]
  print(paste0("Simulation ", LETTERS[letter],": Snake mu-rate & effect sd (", snake_mu_rate,", ", snake_mu_effect_sd, ") Newt mu-rate & effect sd (", newt_mu_rate, ", ", newt_mu_effect_sd, ")",  " rep: ", rep))
  }

}

```

```{r, small sample}

GA1<-cor_grid_30_file[which(cor_grid_30_file$generation>=10000 & cor_grid_30_file$gen<=110000 ),]
GA2<-cor_grid_30_G2_file[which(cor_grid_30_G2_file$generation>=10000 & cor_grid_30_G2_file$gen<=110000 ),]
length(GA1$generation)

GA1<-cor_grid_30_file[which(cor_grid_30_file$generation==10001 & rep==0),]
GA1_mean = GA1[ , grepl( "mean" , names( GA1 ) ) ]
GA1T<-data.frame(t(GA1_mean))
GA1T
```