### 2022-05-19 graphs_newt_snakes/Dac_2022.Rmd
I am improving my work and figures. No huge changes in slim. I am fixing some of the graphs in the r markdown and I fixed one huge mistage in the GAs conclusion. I am current working on the cor part 2 and will have a cool graph by the end of tonight hopefully

---
title: "graph dor DAC"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, fitness graph}
setCOST=100
Phenotype= seq(1, 20, 0.5)
the_y = exp(-(Phenotype/setCOST)^2)

plot(Phenotype, the_y, type="l", lwd=3)

Phenotype= seq(1, 20, 0.5)
Cost = 1-exp(-(Phenotype/setCOST)^2)

plot(Phenotype, Cost, type="l", lwd=3)
#how to change this to the probably of survival

```