### 2022-05-19 graphs_newt_snakes/correlation_part2.Rmd

---
title: "correlation_part2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggmap)
library(ggplot2)
library(sf)
library(mapview)
```

## Comparing newt and snake phenotype correlation

newt and snake data from Hanifin et al. 2008

```{r, reak data cor}
snewt <- read.csv("~/Desktop/snake_newt_real_data.csv", header = TRUE)

snewt$mean_res = snewt$mean_toxin+snewt$dif
cor(snewt$mean_toxin, snewt$mean_res)
(sites <- data.frame(longitude = snewt$lon, latitude = snewt$lat))
just_loc <- data.frame(snewt$lon, snewt$lat)
snewt$lon<- snewt$lon *-1
locations_sf <- st_as_sf(snewt, coords = c("lon", "lat"), crs = 4326)
mapview(locations_sf)
```

```{r, more maps}
library("rnaturalearth")
library("rnaturalearthdata")
library("tools")
library("maps")

world <- ne_countries(scale = "medium", returnclass = "sf")
states <- st_as_sf(maps::map("state", plot = FALSE, fill = TRUE))
states$ID <- toTitleCase(states$ID)

ggplot(data = world) +
  geom_sf() +
  geom_sf(data = states, fill = NA)+
  geom_point(data = snewt, aes(x = lon, y = lat, fill = mean_res), size = 5, shape = 22, ) +
  geom_point(data = snewt, aes(x = lon, y = lat, fill = mean_toxin), size = 3, shape = 21, ) +
  scale_fill_viridis_c()+
  coord_sf(xlim = c(-130.4494, -111.6764), ylim = c(29.742, 54.9415), expand = FALSE)
```

## More data from the slim simulation

might be a specific generation or a small set of generations, If I can get the data to be saved and then interpreted by R
make a list in a column 
how to proportion and recreate what I see in the slim map (the different box and their dims)

GA1 experiment values:

- 1e-8*(0.005)^2 = 2.5e-13
- 1e-9*(0.05)^2 = 2.5e-12
- 1e-10*(0.5)^2 = 2.5e-11
- 1e-11*(5.0)^2 = 2.5e-10

GA2 experiment values:

- 1e-8*(0.05)^2 = 2.5e-11
- 1e-9*(0.1581139)^2 = 2.5e-11
- 1e-10*(0.5)^2 = 2.5e-11
- 1e-11*(1.581139)^2 = 2.5e-11
- 1e-12*(5)^2 = 2.5e-11

```{r, cor by location}

#cor_grid_30 <- list.files(path="~/Desktop/GA1_rep", pattern="_grid_", full.names=TRUE, recursive=FALSE)
#cor_grid_30_G2 <- list.files(path="~/Desktop/GA2_rep", pattern="_grid_", full.names=TRUE, recursive=FALSE)
cor_grid_30 <- list.files(path="~/Desktop/GA1_csv_notperarea", pattern="_grid_", full.names=TRUE, recursive=FALSE)
cor_grid_30_G2 <- list.files(path="~/Desktop/GA2_csv_notperarea", pattern="_grid_", full.names=TRUE, recursive=FALSE)

cor_grid_30_file<-read.csv(cor_grid_30[1], header = TRUE)
replica <- unlist(str_split(unlist(str_split(cor_grid_30[1],"rep_"))[3], "_"))[1]
cor_grid_30_file$rep<-replica
cor_grid_30_file$sim <-"A"
cor_grid_30_file$rep <-unlist(str_split(unlist(str_split(cor_grid_30[1],"rep_"))[3], "_"))[1]
cor_grid_30_file$snake_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30[1],"snake_mu_rate_"))[2], "_"))[1]
cor_grid_30_file$newt_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30[1],"newt_mu_rate_"))[2], "_"))[1]
cor_grid_30_file$snake_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30[1],"snake_mu_effect_sd_"))[2], "_"))[1]
cor_grid_30_file$newt_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30[1],"newt_mu_effect_sd_"))[2], "_"))[1]

cor_grid_30_G2_file<-read.csv(cor_grid_30_G2[1], header = TRUE)
replica <- unlist(str_split(unlist(str_split(cor_grid_30_G2[1],"rep_"))[3], "_"))[1]
cor_grid_30_G2_file$rep<-replica
cor_grid_30_G2_file$sim <-"A"
cor_grid_30_G2_file$rep <-unlist(str_split(unlist(str_split(cor_grid_30_G2[1],"rep_"))[3], "_"))[1]
cor_grid_30_G2_file$snake_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30_G2[1],"snake_mu_rate_"))[2], "_"))[1]
cor_grid_30_G2_file$newt_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30_G2[1],"newt_mu_rate_"))[2], "_"))[1]
cor_grid_30_G2_file$snake_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30_G2[1],"snake_mu_effect_sd_"))[2], "_"))[1]
cor_grid_30_G2_file$newt_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30_G2[1],"newt_mu_effect_sd_"))[2], "_"))[1]

oldstand=0
letter=1
for(i in seq(1, length(cor_grid_30_G2), 1)){
  stand <- unlist(str_split(unlist(str_split(cor_grid_30_G2[i],"newt_mu_rate_"))[2], "_"))[1]
  if (stand!=oldstand){
    oldstand=stand
    #letter = letter + 1
  }
  letter = i
  GA2_lit_temp_file <- read.csv(cor_grid_30_G2[i], header = TRUE)
  GA2_lit_temp_file$sim <-LETTERS[letter]
  GA2_lit_temp_file$rep <-unlist(str_split(unlist(str_split(cor_grid_30_G2[i],"rep_"))[3], "_"))[1]
  GA2_lit_temp_file$snake_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30_G2[i],"snake_mu_rate_"))[2], "_"))[1]
GA2_lit_temp_file$newt_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30[i],"newt_mu_rate_"))[2], "_"))[1]
GA2_lit_temp_file$snake_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30[i],"snake_mu_effect_sd_"))[2], "_"))[1]
GA2_lit_temp_file$newt_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30[i],"newt_mu_effect_sd_"))[2], "_"))[1]

  cor_grid_30_G2_file <- rbind(cor_grid_30_G2_file, GA2_lit_temp_file)

  if (i<=length(cor_grid_30)){
  cor_temp_file <- read.csv(cor_grid_30[i], header = TRUE)
  cor_temp_file$sim <-LETTERS[letter]
  cor_temp_file$rep <-unlist(str_split(unlist(str_split(cor_grid_30[i],"rep_"))[3], "_"))[1]
  cor_temp_file$snake_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30[i],"snake_mu_rate_"))[2], "_"))[1]
cor_temp_file$newt_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30[i],"newt_mu_rate_"))[2], "_"))[1]
cor_temp_file$snake_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30[i],"snake_mu_effect_sd_"))[2], "_"))[1]
cor_temp_file$newt_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30[i],"newt_mu_effect_sd_"))[2], "_"))[1]
  cor_grid_30_file <- rbind(cor_grid_30_file, cor_temp_file)
  rep <-unlist(str_split(unlist(str_split(cor_grid_30[i],"rep_"))[3], "_"))[1]

  snake_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30[i],"snake_mu_rate_"))[2], "_"))[1]
  newt_mu_rate <- unlist(str_split(unlist(str_split(cor_grid_30[i],"newt_mu_rate_"))[2], "_"))[1]
  snake_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30[i],"snake_mu_effect_sd_"))[2], "_"))[1]
  newt_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_grid_30[i],"newt_mu_effect_sd_"))[2], "_"))[1]
  print(paste0("Simulation ", LETTERS[letter],": Snake mu-rate & effect sd (", snake_mu_rate,", ", snake_mu_effect_sd, ") Newt mu-rate & effect sd (", newt_mu_rate, ", ", newt_mu_effect_sd, ")",  " rep: ", rep))
  }

}

```

```{r, small sample}

cor_grid_30_file$snake <- paste0(cor_grid_30_file$snake_mu_rate, "_",cor_grid_30_file$snake_mu_effect_sd)
cor_grid_30_file$newt <- paste0(cor_grid_30_file$newt_mu_rate, "_",cor_grid_30_file$newt_mu_effect_sd)

cor_grid_30_file$newt_snake_rep <- paste0("Newt_",cor_grid_30_file$newt_mu_rate, "_", cor_grid_30_file$newt_mu_effect_sd, "_Snake_",cor_grid_30_file$snake_mu_rate, "_", cor_grid_30_file$snake_mu_effect_sd, "_", cor_grid_30_file$rep)

#GA1_mean = cor_grid_30_file[ , grepl( "mean" , names( cor_grid_30_file ) ) ]

#unique(cor_grid_30_file$newt_snake_rep)

GA1<-cor_grid_30_file[which(cor_grid_30_file$generation>=10000 & cor_grid_30_file$gen<=110000 ),]
#GA2<-cor_grid_30_G2_file[which(cor_grid_30_G2_file$generation>=10000 & cor_grid_30_G2_file$gen<=110000 ),]
#length(GA1$generation)

#GA1<-cor_grid_30_file[which(cor_grid_30_file$generation<=50 & cor_grid_30_file$newt_snake_rep=="Newt_1.0e-08_0.005_Snake_1.0e-09_0.05_2"),]

GA1_mean = GA1[,grepl("generation|mean", names(GA1)) ]

#How to remake the dataframe into gen, newt_value, snake_value, location (x, y)?

GA1_t <- data.frame(t(GA1_mean))

loc_x = 1
loc_y = 5
skip = 1
GA1_t$loc_x = 0
GA1_t$loc_y = 0
GA1_t$species = 0
grid_info <- as.character(rownames(GA1_t))
for (i in 2:nrow(GA1_t)){
   #print(as.numeric(unlist(str_split(i,"_pheno_"))[2]))

  GA1_t$species[i] = unlist(str_split(grid_info[i],"_mean"))[1]

  if(loc_y==0){
    loc_x = loc_x + 1
    loc_y = 5
  }
  #print(paste0(loc_x, " ", loc_y))
  GA1_t$loc_x[i]= loc_x
GA1_t$loc_y[i] = loc_y

  if (skip %% 2 == 0 ){
    loc_y = loc_y - 1
  } 
  skip = skip +1

}

GA1_t$names<- rownames(GA1_t)

GA1_mean = cor_grid_30_file[,grepl( "mean" , names(cor_grid_30_file)) ]
GA1_mean$generation <- cor_grid_30_file$generation
GA1_mean$rep <- cor_grid_30_file$rep
GA1_mean$snake_mu_rate <- cor_grid_30_file$snake_mu_rate
GA1_mean$newt_mu_rate <- cor_grid_30_file$newt_mu_rate
GA1_mean$snake_mu_effect_sd <- cor_grid_30_file$snake_mu_effect_sd
GA1_mean$newt_mu_effect_sd <- cor_grid_30_file$newt_mu_effect_sd
GA1_mean$snake <- cor_grid_30_file$snake
GA1_mean$newt <- cor_grid_30_file$newt
GA1_mean$newt_snake_rep <- cor_grid_30_file$newt_snake_rep

data_long_mean <- gather(GA1_mean, condition, measurement, newt_mean_pheno_0:snake_mean_pheno_24, factor_key=TRUE)

data_long_mean$lox <- GA1_t[match(data_long_mean$condition, GA1_t$names),"loc_x"]
data_long_mean$loy <- GA1_t[match(data_long_mean$condition, GA1_t$names),"loc_y"]
data_long_mean$species <- GA1_t[match(data_long_mean$condition, GA1_t$names),"species"]

GA1_max = cor_grid_30_file[,grepl( "max" , names(cor_grid_30_file)) ]
GA1_max$generation <- cor_grid_30_file$generation
GA1_max$rep <- cor_grid_30_file$rep
GA1_max$snake_mu_rate <- cor_grid_30_file$snake_mu_rate
GA1_max$newt_mu_rate <- cor_grid_30_file$newt_mu_rate
GA1_max$snake_mu_effect_sd <- cor_grid_30_file$snake_mu_effect_sd
GA1_max$newt_mu_effect_sd <- cor_grid_30_file$newt_mu_effect_sd
GA1_max$snake <- cor_grid_30_file$snake
GA1_max$newt <- cor_grid_30_file$newt
GA1_max$newt_snake_rep <- cor_grid_30_file$newt_snake_rep

data_long_max <- gather(GA1_max, condition, measurement, newt_max_pheno_0:snake_max_pheno_24, factor_key=TRUE)

data_long_max$lox <- data_long_mean$lox 
data_long_max$loy <- data_long_mean$loy
data_long_max$species <- data_long_mean$species

```

```{r, grid graph values}

GA1_onefile<-data_long_mean[which(data_long_mean$newt_snake_rep=="Newt_1.0e-10_0.5_Snake_1.0e-10_0.5_0" & data_long_max$gen==29981 ),]

#GA1_onefile<-data_long_mean[which(data_long_mean$newt_snake_rep==unique(data_long_mean$newt_snake_rep)[16] & data_long_max$gen==29981 ),]

#GA1_onefile<-data_long_max[which(data_long_max$newt_snake_rep=="Newt_1.0e-10_0.5_Snake_1.0e-10_0.5_0" & data_long_max$gen==29051 ),]

ggplot(GA1_onefile)+

  geom_point(data = GA1_onefile[which(GA1_onefile$species=="snake"),], aes(x = lox, y = loy, fill = measurement), size = 10, shape = 22, ) +
  geom_point(data = GA1_onefile[which(GA1_onefile$species=="newt"),], aes(x = lox, y = loy, fill = measurement), size = 7, shape = 21, ) +
  scale_fill_viridis_c()+
  theme_bw()

```

```{r, corelation plots}

cor_files <- list.files(path="~/Desktop/GA1_rep", pattern="_cor_", full.names=TRUE, recursive=FALSE)
cor_GA2_files <- list.files(path="~/Desktop/GA2_rep", pattern="_cor_", full.names=TRUE, recursive=FALSE)
lit_30 <- list.files(path="~/Desktop/GA1_rep", pattern="_lit_", full.names=TRUE, recursive=FALSE)
lit_30_G2 <- list.files(path="~/Desktop/GA2_rep", pattern="_lit_", full.names=TRUE, recursive=FALSE)

lit_30_file<- data.frame()
lit_30_G2_file <- data.frame()

number=1
#cor_file <- read.csv(cor_files[number], header = TRUE)
#cor_file$sim <-"A"
#cor_GA2_file <- read.csv(cor_GA2_files[number], header = TRUE)
#cor_GA2_file$sim <-"A"

cor_file<- data.frame()
cor_GA2_file<- data.frame()
oldstand=0
letter=0
newt_unique=0
newt_numb = 0
for(i in seq(1, length(cor_GA2_files), 1)){
  stand <- unlist(str_split(unlist(str_split(cor_GA2_files[i],"newt_mu_rate_"))[2], "_"))[1]
  up <- unlist(str_split(unlist(str_split(cor_GA2_files[i],"newt_mu_effect_sd_"))[2], "_"))[1]
  phrase = paste0(stand, "_", up)
  if (stand!=oldstand){
    oldstand=stand
    letter = letter + 1
  }

    if (phrase=="1.0e-08_0.05" ){
    newt_numb =  1
   }
    if (phrase=="1.0e-09_0.158114"){
    newt_numb =  2
   }
      if (phrase=="1.0e-10_0.5"){
    newt_numb =  3
   }
    if (phrase=="1.0e-11_1.58114"){
    newt_numb =  4
   }

      if (phrase=="1.0e-12_5.0"){
    newt_numb =  5
   }

  GA2_cor_temp_file <- read.csv(cor_GA2_files[i], header = TRUE)
  GA2_cor_temp_file$sim <-LETTERS[letter]
  GA2_cor_temp_file$size <-newt_numb
  GA2_cor_temp_file$rep <-unlist(str_split(unlist(str_split(cor_GA2_files[i],"rep_"))[3], "_"))[1]
  GA2_cor_temp_file$snake_mu_rate <- unlist(str_split(unlist(str_split(cor_GA2_files[i],"snake_mu_rate_"))[2], "_"))[1]
  GA2_cor_temp_file$newt_mu_rate <- unlist(str_split(unlist(str_split(cor_GA2_files[i],"newt_mu_rate_"))[2], "_"))[1]
  GA2_cor_temp_file$snake_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_GA2_files[i],"snake_mu_effect_sd_"))[2], "_"))[1]
  GA2_cor_temp_file$newt_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_GA2_files[i],"newt_mu_effect_sd_"))[2], "_"))[1]

  cor_GA2_file <- rbind(cor_GA2_file, GA2_cor_temp_file)

  GA2_lit_temp_file <- read.csv(lit_30_G2[i], header = TRUE)
  GA2_lit_temp_file$sim <-LETTERS[letter]
  GA2_lit_temp_file$size <-newt_numb
  GA2_lit_temp_file$rep <-unlist(str_split(unlist(str_split(lit_30_G2[i],"rep_"))[3], "_"))[1]
  GA2_lit_temp_file$snake_mu_rate <- unlist(str_split(unlist(str_split(lit_30_G2[i],"snake_mu_rate_"))[2], "_"))[1]
  GA2_lit_temp_file$newt_mu_rate <- unlist(str_split(unlist(str_split(lit_30_G2[i],"newt_mu_rate_"))[2], "_"))[1]
  GA2_lit_temp_file$snake_mu_effect_sd <- unlist(str_split(unlist(str_split(lit_30_G2[i],"snake_mu_effect_sd_"))[2], "_"))[1]
  GA2_lit_temp_file$newt_mu_effect_sd <- unlist(str_split(unlist(str_split(lit_30_G2[i],"newt_mu_effect_sd_"))[2], "_"))[1]

  lit_30_G2_file <- rbind(lit_30_G2_file, GA2_lit_temp_file)

  if (i<=length(cor_files)){
  stand <- unlist(str_split(unlist(str_split(cor_files[i],"newt_mu_rate_"))[2], "_"))[1]
  up <- unlist(str_split(unlist(str_split(cor_files[i],"newt_mu_effect_sd_"))[2], "_"))[1]
  phrase = paste0(stand, "_", up)

   if (phrase=="1.0e-08_0.005"){
    newt_numb =  1
   }
    if (phrase=="1.0e-09_0.05"){
    newt_numb =  2
   }
      if (phrase=="1.0e-10_0.5"){
    newt_numb =  3
   }
    if (phrase=="1.0e-11_5.0"){
    newt_numb =  4
   }
  cor_temp_file <- read.csv(cor_files[i], header = TRUE)
  cor_temp_file$sim <-LETTERS[letter]
  cor_temp_file$size <-newt_numb
  cor_temp_file$rep <-unlist(str_split(unlist(str_split(cor_files[i],"rep_"))[3], "_"))[1]
  cor_temp_file$snake_mu_rate <- unlist(str_split(unlist(str_split(cor_files[i],"snake_mu_rate_"))[2], "_"))[1]
  cor_temp_file$newt_mu_rate <- unlist(str_split(unlist(str_split(cor_files[i],"newt_mu_rate_"))[2], "_"))[1]
  cor_temp_file$snake_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_files[i],"snake_mu_effect_sd_"))[2], "_"))[1]
  cor_temp_file$newt_mu_effect_sd <- unlist(str_split(unlist(str_split(cor_files[i],"newt_mu_effect_sd_"))[2], "_"))[1]
  cor_file <- rbind(cor_file, cor_temp_file)

  GA1_temp_file <- read.csv(lit_30[i], header = TRUE)
  GA1_temp_file$sim <-LETTERS[letter]
  GA1_temp_file$size <-newt_numb
  GA1_temp_file$rep <-unlist(str_split(unlist(str_split(lit_30[i],"rep_"))[3], "_"))[1]
  GA1_temp_file$snake_mu_rate <- unlist(str_split(unlist(str_split(lit_30[i],"snake_mu_rate_"))[2], "_"))[1]
  GA1_temp_file$newt_mu_rate <- unlist(str_split(unlist(str_split(lit_30[i],"newt_mu_rate_"))[2], "_"))[1]
  GA1_temp_file$snake_mu_effect_sd <- unlist(str_split(unlist(str_split(lit_30[i],"snake_mu_effect_sd_"))[2], "_"))[1]
  GA1_temp_file$newt_mu_effect_sd <- unlist(str_split(unlist(str_split(lit_30[i],"newt_mu_effect_sd_"))[2], "_"))[1]
  lit_30_file <- rbind(lit_30_file, GA1_temp_file)

  #print(paste0("Simulation ", LETTERS[letter],": Snake mu-rate & effect sd (", cor_temp_file$snake_mu_rate[i],", ", cor_temp_file$snake_mu_effect_sd[i], ") Newt mu-rate & effect sd (", cor_temp_file$newt_mu_rate[i], ", ",cor_temp_file$ newt_mu_effect_sd[i], ")" ))
  }

}
lit_30_file$dif <- lit_30_file$Snake_mean_Pheno-lit_30_file$Newt_mean_Pheno
lit_30_G2_file$dif <- lit_30_G2_file$Snake_mean_Pheno-lit_30_G2_file$Newt_mean_Pheno

```

```{r, cor box set up}

boxplot_cors_by_gen <- function(file, title, subt, x_label, y_label, min_gen, max_gen){
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
  file$newt_snake_rep <- paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$rep)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ylim(-1.1, 1.1)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor(rep), y=as.numeric(mean_newt_pheno_By_mean_snake_pheno), color = "CA"))+
  geom_hline(yintercept=real_snewt_cor)+
  #geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  #geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  #geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  theme(plot.title = element_text(size=5)) +
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

  return(plot)
}

trends_by_gen <- function(file, title, subt, x_label, y_label, min_gen, max_gen){
 file$newt_snake_rep <- paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$rep)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
    facet_wrap(~sim)+
    geom_line(aes(generation, dif, group=rep, col="dif"))+
    ggtitle(title, subtitle = subt) +
    geom_line(aes(generation, Snake_mean_Pheno, group=rep, color="Snake_mean_Pheno"),)+
    geom_line(aes(generation, Newt_mean_Pheno, group=rep, color="Newt_mean_Pheno"))+
    ylim(-5,10)+
    xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake_mean_Pheno" = "steelblue", "Newt_mean_Pheno" = "darkred", "dif" = "black"),
labels = c("Snake", "Newt", "Dif"))+
    theme_bw() +theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
  return(plot)
}

boxplot_cors_by_gen(cor_file, "Simulation Correlation Comparied to Real Correlation", "GA1", "newt/snake GAs&rep", "Cor", 10000, 15000)

boxplot_cors_by_gen(cor_GA2_file, "Simulation Correlation Comparied to Real Correlation", "GA2", "newt/snake GAs&rep", "Cor", 10000, 15000)

trends_by_gen(lit_30_file, "test", "GA1", "newt/snake GAs&rep", "Cor", 10000, 15000)

```