### 2022-05-27 graphs_newt_snakes/running_100000gen.Rmd
looking at correlations, currtly things are not spatially correlated. this might be due to the small map and population size. I added H with increases the length in the simulation. I am excited to see the results

---
title: "Running 100,000 Gen"
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE,echo = FALSE)

library(ggplot2)
library(viridis)
library(colorspace)
library(gridExtra)
library(tidyverse)
require(reshape2)
library(plyr)
library(gridGraphics)
library(grid)
library(knitr)
library(kableExtra)
library("reshape")
library(ggplot2)
library(ggpubr)
theme_set(theme_pubr())
```

### Goal

figure out a main conclusion for the poster!

GA1 experiment values:

- 1e-8*(0.005)^2 = 2.5e-13
- 1e-9*(0.05)^2 = 2.5e-12
- 1e-10*(0.5)^2 = 2.5e-11
- 1e-11*(5.0)^2 = 2.5e-10

GA2 experiment values:

- 1e-8*(0.05)^2 = 2.5e-11
- 1e-9*(0.1581139)^2 = 2.5e-11
- 1e-10*(0.5)^2 = 2.5e-11
- 1e-11*(1.581139)^2 = 2.5e-11
- 1e-12*(5)^2 = 2.5e-11

```{r, data tables}

para_table <- read.csv("/Users/victoria/Desktop/Coevo/data/GA1_values.csv", header = TRUE)
para_table$trial<-LETTERS[para_table$trial]
para_table <- data.frame("Trial"=para_table$trial, "Snake Mu Rate"=para_table$snake_mu_rate, "Snake effect size"= para_table$snake_mu_effect_sd, "Newt Mu Rate"=para_table$newt_mu_rate, "Newt effect size"= para_table$newt_mu_effect_sd)
#kable(para_table, caption="GA2 Params")

para_table %>% dplyr::mutate_if(is.numeric, funs(as.character(signif(., 3)))) %>%
    kable(., caption="GA1 Params") %>%
 kable_styling("striped")

para_table <- read.csv("/Users/victoria/Desktop/Coevo/data/GAtestvar_lit_25.csv", header = TRUE)
para_table$trial<-LETTERS[para_table$trial]
para_table <- data.frame("Trial"=para_table$trial, "Snake Mu Rate"=para_table$snake_mu_rate, "Snake effect size"= para_table$snake_mu_effect_sd, "Newt Mu Rate"=para_table$newt_mu_rate, "Newt effect size"= para_table$newt_mu_effect_sd)
#kable(para_table, caption="GA2 Params")

para_table %>% dplyr::mutate_if(is.numeric, funs(as.character(signif(., 3)))) %>%
    kable(., caption="GA2 Params") %>%
 kable_styling("striped")
```

```{r, data gathering}

snewt <- read.csv("~/Desktop/snake_newt_real_data.csv", header = TRUE)

GA1_long <- list.files(path="~/Desktop/GA1_long_csv", pattern="_lit_", full.names=TRUE, recursive=FALSE)
GA2_long <- list.files(path="~/Desktop/GA2_long_csv", pattern="_lit_", full.names=TRUE, recursive=FALSE)
GA1_cor_files <- list.files(path="~/Desktop/GA1_long_csv", pattern="_cor_", full.names=TRUE, recursive=FALSE)
GA2_cor_files <- list.files(path="~/Desktop/GA2_long_csv", pattern="_cor_", full.names=TRUE, recursive=FALSE)

GA1_long_file<- data.frame()
GA2_long_file <- data.frame()
GA1_cor_file<- data.frame()
GA2_cor_file<- data.frame()
oldstand=0
letter=0
newt_unique=0
newt_numb = 0
for(i in seq(1, length(GA2_long), 1)){
  stand <- unlist(str_split(unlist(str_split(GA2_long[i],"newt_mu_rate_"))[2], "_"))[1]
  up <- unlist(str_split(unlist(str_split(GA2_long[i],"newt_mu_effect_sd_"))[2], "_"))[1]
  phrase = paste0(stand, "_", up)
  if (stand!=oldstand){
    oldstand=stand
    letter = letter + 1
  }

    if (phrase=="1.0e-08_0.05" ){
    newt_numb =  1
   }
    if (phrase=="1.0e-09_0.158114"){
    newt_numb =  2
   }
      if (phrase=="1.0e-10_0.5"){
    newt_numb =  3
   }
    if (phrase=="1.0e-11_1.58114"){
    newt_numb =  4
   }

      if (phrase=="1.0e-12_5.0"){
    newt_numb =  5
   }

  GA2_lit_temp_file <- read.csv(GA2_long[i], header = TRUE)
  GA2_lit_temp_file$sim <-LETTERS[letter]
   GA2_lit_temp_file$size <-newt_numb
  GA2_lit_temp_file$rep <-unlist(str_split(unlist(str_split(GA2_long[i],"rep_"))[3], "_"))[1]
  GA2_lit_temp_file$snake_mu_rate <- unlist(str_split(unlist(str_split(GA2_long[i],"snake_mu_rate_"))[2], "_"))[1]
  GA2_lit_temp_file$newt_mu_rate <- unlist(str_split(unlist(str_split(GA2_long[i],"newt_mu_rate_"))[2], "_"))[1]
  GA2_lit_temp_file$snake_mu_effect_sd <- unlist(str_split(unlist(str_split(GA2_long[i],"snake_mu_effect_sd_"))[2], "_"))[1]
  GA2_lit_temp_file$newt_mu_effect_sd <- unlist(str_split(unlist(str_split(GA2_long[i],"newt_mu_effect_sd_"))[2], "_"))[1]

  GA2_cor_temp_file <- read.csv(GA2_cor_files[i], header = TRUE)
  GA2_cor_temp_file$sim <-LETTERS[letter]
  GA2_cor_temp_file$size <-newt_numb
  GA2_cor_temp_file$rep <-unlist(str_split(unlist(str_split(GA2_cor_files[i],"rep_"))[3], "_"))[1]
  GA2_cor_temp_file$snake_mu_rate <- unlist(str_split(unlist(str_split(GA2_cor_files[i],"snake_mu_rate_"))[2], "_"))[1]
  GA2_cor_temp_file$newt_mu_rate <- unlist(str_split(unlist(str_split(GA2_cor_files[i],"newt_mu_rate_"))[2], "_"))[1]
  GA2_cor_temp_file$snake_mu_effect_sd <- unlist(str_split(unlist(str_split(GA2_cor_files[i],"snake_mu_effect_sd_"))[2], "_"))[1]
  GA2_cor_temp_file$newt_mu_effect_sd <- unlist(str_split(unlist(str_split(GA2_cor_files[i],"newt_mu_effect_sd_"))[2], "_"))[1]

  GA2_cor_file <- rbind(GA2_cor_file, GA2_cor_temp_file)

  GA2_long_file <- rbind(GA2_long_file, GA2_lit_temp_file)

  if (i<=length(GA1_long)){

  stand <- unlist(str_split(unlist(str_split(GA1_long[i],"newt_mu_rate_"))[2], "_"))[1]
  up <- unlist(str_split(unlist(str_split(GA1_long[i],"newt_mu_effect_sd_"))[2], "_"))[1]
  phrase = paste0(stand, "_", up)

   if (phrase=="1.0e-08_0.005"){
    newt_numb =  1
   }
    if (phrase=="1.0e-09_0.05"){
    newt_numb =  2
   }
      if (phrase=="1.0e-10_0.5"){
    newt_numb =  3
   }
    if (phrase=="1.0e-11_5.0"){
    newt_numb =  4
   }

  GA1_temp_file <- read.csv(GA1_long[i], header = TRUE)
  GA1_temp_file$sim <-LETTERS[letter]
  GA1_temp_file$size <-newt_numb
  GA1_temp_file$rep <-unlist(str_split(unlist(str_split(GA1_long[i],"rep_"))[3], "_"))[1]
  GA1_temp_file$snake_mu_rate <- unlist(str_split(unlist(str_split(GA1_long[i],"snake_mu_rate_"))[2], "_"))[1]
  GA1_temp_file$newt_mu_rate <- unlist(str_split(unlist(str_split(GA1_long[i],"newt_mu_rate_"))[2], "_"))[1]
  GA1_temp_file$snake_mu_effect_sd <- unlist(str_split(unlist(str_split(GA1_long[i],"snake_mu_effect_sd_"))[2], "_"))[1]
  GA1_temp_file$newt_mu_effect_sd <- unlist(str_split(unlist(str_split(GA1_long[i],"newt_mu_effect_sd_"))[2], "_"))[1]
  GA1_long_file <- rbind(GA1_long_file, GA1_temp_file)

  cor_temp_file <- read.csv(GA1_cor_files[i], header = TRUE)
  cor_temp_file$sim <-LETTERS[letter]
  cor_temp_file$size <-newt_numb
  cor_temp_file$rep <-unlist(str_split(unlist(str_split(GA1_cor_files[i],"rep_"))[3], "_"))[1]
  cor_temp_file$snake_mu_rate <- unlist(str_split(unlist(str_split(GA1_cor_files[i],"snake_mu_rate_"))[2], "_"))[1]
  cor_temp_file$newt_mu_rate <- unlist(str_split(unlist(str_split(GA1_cor_files[i],"newt_mu_rate_"))[2], "_"))[1]
  cor_temp_file$snake_mu_effect_sd <- unlist(str_split(unlist(str_split(GA1_cor_files[i],"snake_mu_effect_sd_"))[2], "_"))[1]
  cor_temp_file$newt_mu_effect_sd <- unlist(str_split(unlist(str_split(GA1_cor_files[i],"newt_mu_effect_sd_"))[2], "_"))[1]
  GA1_cor_file <- rbind(GA1_cor_file, cor_temp_file)

  }

}

GA1_long_file$dif <- GA1_long_file$Snake_mean_Pheno-GA1_long_file$Newt_mean_Pheno
GA2_long_file$dif <- GA2_long_file$Snake_mean_Pheno-GA2_long_file$Newt_mean_Pheno
```

```{r, cor box set up}

boxplot_cors_by_gen <- function(file, title, subt, x_label, y_label, min_gen, max_gen){
 snewt$mean_res = snewt$mean_toxin+snewt$dif
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
  file$newt_snake_rep <- paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$rep)
  file$newt_snake <- paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  unique(file$sim)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ylim(-1.1, 1.1)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor(rep), y=as.numeric(mean_newt_pheno_By_mean_snake_pheno), color = "CA"))+
  geom_hline(yintercept=real_snewt_cor, linetype="dashed")+
    geom_hline(yintercept=0, linetype="solid")+
  #geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  #geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  #geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  theme(plot.title = element_text(size=5)) +
  facet_wrap(~newt_snake)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

  return(plot)
}

hist_cors_by_gen_col_rep <- function(file, title, subt, x_label, y_label, min_gen, max_gen){
 snewt$mean_res = snewt$mean_toxin+snewt$dif
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
  file$newt_snake_rep <- paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$rep)
  file$newt_snake <- paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  unique(file$sim)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=as.numeric(mean_newt_pheno_By_mean_snake_pheno),fill=rep))+
  geom_vline(xintercept=real_snewt_cor, linetype="dashed")+
  geom_vline(xintercept=0, linetype="solid")+
  #geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  #geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  #geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  theme(plot.title = element_text(size=5)) +
  facet_wrap(~newt_snake)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

  return(plot)
}

hist_cors_by_gen_col_rep_nofill <- function(file, title, subt, x_label, y_label, min_gen, max_gen){
 snewt$mean_res = snewt$mean_toxin+snewt$dif
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
  file$newt_snake_rep <- paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$rep)
  file$newt_snake <- paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  unique(file$sim)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=as.numeric(mean_newt_pheno_By_mean_snake_pheno)))+
  geom_vline(xintercept=real_snewt_cor, linetype="dashed")+
  geom_vline(xintercept=0, linetype="solid")+
  #geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  #geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  #geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  theme(plot.title = element_text(size=5)) +
  facet_wrap(~newt_snake)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

  return(plot)
}

var_cor <- function(file, cfile, title, subt, x_label, y_label, min_gen, max_gen){
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
  real_snewt_var = var(snewt$dif)
  cfile$newt_snake_rep <- paste0("Newt_", cfile$newt_mu_rate, "_", cfile$newt_mu_effect_sd, "_Snake_", cfile$snake_mu_rate, "_", cfile$snake_mu_effect_sd, "_", cfile$rep)
  var_cor_table=data.frame()

  for(i in 1:length(unique(file$newt_snake_rep))){
    pattern<-unique(file$newt_snake_rep)[i]
    temp_file<-file[which(file$gen>=min_gen & file$gen<=max_gen & file$newt_snake_rep==pattern),]
    temp_cor_file<-cfile[which(cfile$gen>=min_gen & cfile$gen<=max_gen & cfile$newt_snake_rep==pattern),]

    var_cor_table_temp<- data.frame("pattern"=pattern, "var"= var(temp_file$dif),"cor" = cor(temp_file$newt_toxin, temp_file$snake_res), "mean_corr" = mean(temp_cor_file$mean_newt_pheno_By_mean_snake_pheno))
    var_cor_table <- rbind(var_cor_table, var_cor_table_temp)
  }

  var_cor_table_temp<- data.frame("pattern"="real", "var"= real_snewt_var, "cor" = real_snewt_cor, "mean_corr" = real_snewt_cor)
  var_cor_table <- rbind(var_cor_table, var_cor_table_temp)

  plot <- ggplot(var_cor_table) +
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=as.numeric(var), y=as.numeric(cor), color = pattern), shape=17, size=3)+
  geom_point(aes(x=as.numeric(var), y=as.numeric(mean_corr), color = pattern), size=3)+
  theme(plot.title = element_text(size=5)) +
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))

  return(plot)
}

cor_trends_by_gen <- function(file, cfile, title, subt, x_label, x_label_cor, y_label, min_gen, max_gen, pattern){
 file$newt_snake_rep <- paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$rep)
  cfile$newt_snake_rep <- paste0(cfile$newt_mu_rate, "_", cfile$newt_mu_effect_sd, "_", cfile$snake_mu_rate, "_", cfile$snake_mu_effect_sd, "_", cfile$rep)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen & file$newt_snake_rep==pattern),]) +
    #geom_line(aes(generation, dif, group=rep, col="dif"))+
      ggtitle(title, subtitle = paste0(subt," pattern ", pattern, " From ", min_gen, " gen to ", max_gen, " gen"))+
    geom_line(aes(generation, Snake_mean_Pheno, group=rep, color="Snake_mean_Pheno"),)+
    geom_line(aes(generation, Newt_mean_Pheno, group=rep, color="Newt_mean_Pheno"))+
    #ylim(0,10)+
    #xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake_mean_Pheno" = "steelblue", "Newt_mean_Pheno" = "darkred"),
labels = c("Snake", "Newt"))+
    theme_bw() +theme(legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank(), axis.ticks.x = element_blank(), axis.text.x = element_blank(),
                      plot.margin=unit(c(1,1,+0.3,1), "cm"))

  cor_plot <- ggplot(cfile[which(cfile$gen>=min_gen & cfile$gen<=max_gen & cfile$newt_snake_rep==pattern),]) +
    geom_line(aes(generation, mean_newt_pheno_By_mean_snake_pheno, col="CA"))+
    #ggtitle(title, subtitle = subt) +
    ylim(-1.1, 1.1)+
    #xlab(x_label_cor) + ylab(y_label)+
    theme_bw() +theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank(),
                      plot.margin=unit(c(-0.5,1,1,1), "cm"))

  file<-file[which(file$gen>=min_gen & file$gen<=max_gen & file$newt_snake_rep==pattern),]
  cfile<-cfile[which(cfile$gen>=min_gen & cfile$gen<=max_gen & cfile$newt_snake_rep==pattern),]
  toDelete <- seq(0, nrow(cfile), 2)
  cfile_short <-  cfile[-toDelete, ]
  figure <- ggarrange(plot, cor_plot + font("x.text", size = 10),heights = c(2, 1.5),ncol = 1, nrow = 2, align = "v")
    print(paste0("pattern ", pattern))
  print(paste0("Cor between average snake pheno and local cor ", cor(file$Snake_mean_Pheno, cfile_short$mean_newt_pheno_By_mean_snake_pheno)))
  print(paste0("Cor between average newt pheno and local cor ", cor(file$Newt_mean_Pheno, cfile_short$mean_newt_pheno_By_mean_snake_pheno)))
  print(paste0("Cor between average dif pheno and local cor ", cor(file$dif, cfile_short$mean_newt_pheno_By_mean_snake_pheno)))
  print(paste0("Cor between newt pheno and snake ", cor(file$Newt_mean_Pheno, file$Snake_mean_Pheno)))
  return(figure)
}

four_types_boxplot_cors_by_gen <- function(file,title, subt, x_label, y_label, min_gen, max_gen){
  file$newt_snake <- paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ylim(-1.1, 1.1)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor("CA"), y=as.numeric(mean_newt_pheno_By_mean_snake_pheno), color = "CA"))+
  geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  geom_hline(yintercept=0, linetype="solid")+
  theme(plot.title = element_text(size=5)) +
  facet_wrap(~newt_snake)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

  return(plot)
}

```

```{r, general file and graph making}

#the functions to make the params file and the summary and heatmap figures

create_parm_file <- function(file, min_gen, max_gen){
params <- data_frame()

for(i in 1:length(unique(file$newt_snake_rep))){

  pattern=unique(file$newt_snake_rep)[i]
  sumitup <- file[which(file$gen>=min_gen & file$gen<=max_gen & file$newt_snake_rep ==pattern),]

  params_temp<- data_frame(Newt_mean_Pheno=mean(sumitup$Newt_mean_Pheno),
                           Snake_mean_Pheno=mean(sumitup$Snake_mean_Pheno), 
                           Newt_mean_Pheno_sd=sd(sumitup$Newt_mean_Pheno),
                           Snake_mean_Pheno_sd=sd(sumitup$Snake_mean_Pheno),
                           Newt_pop_size=mean(sumitup$Newt_pop_size),
                           Snake_pop_size=mean(sumitup$Snake_pop_size),
                           dif=mean(sumitup$dif),
                           snake_mu_rate=sumitup$snake_mu_rate[1],
                           newt_mu_rate=sumitup$newt_mu_rate[1],
                           snake_mu_effect_sd=sumitup$snake_mu_effect_sd[1],
                           newt_mu_effect_sd=sumitup$newt_mu_effect_sd[1],
                           rep=sumitup$rep[1])

  params <- rbind(params, params_temp)
}

params$snake <- paste0(params$snake_mu_rate, "_",params$snake_mu_effect_sd)
params$newt <- paste0(params$newt_mu_rate, "_",params$newt_mu_effect_sd)
params$newt_snake <- paste0(params$newt, "_", params$snake)
return(params)
}

GA_summary <- function(file, min_gen, max_gen, x_line, y_line, title, subt, y_label, x_label){

file$newt_snake_rep <- paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$rep)

params<-create_parm_file(file, min_gen, max_gen)

my_plot <- ggplot(params) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=as.numeric(get(x_line)), y=as.numeric(get(y_line)), color = factor(snake), shape=factor(newt)),size=3)+
#  geom_point(aes(x=as.numeric(Newt_pop_size_20), y=as.numeric(Snake_pop_size_20), color = factor(snake), shape=factor(newt)))+
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  #facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text( hjust=1, size = 8))

return(my_plot)
}

build_heatmap <- function(file, min_gen, max_gen, fill_line, fill_line_text, title, subt, y_label, x_label, red_blue=F){
  file$newt_snake_rep <- paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$rep)

  params<-create_parm_file(file, min_gen, max_gen)

 my_plot <- ggplot(params, aes(newt, paste0(snake, "_", rep))) +    # Create heatmap with ggplot2
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_tile(aes(fill = get(fill_line)))+theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
 if(red_blue==FALSE){
  my_plot<- my_plot + scale_fill_viridis_c(name = fill_line_text)
  }
 if(red_blue==TRUE){
  my_plot<- my_plot + scale_fill_gradient2(low = "red",mid = "beige", high = "blue", name = fill_line_text)
 }
  return(my_plot)
}

trends_by_gen <- function(file, title, subt, x_label, y_label){

  plot <- ggplot(file) +
    facet_wrap(~sim)+
    geom_line(aes(generation, dif, group=rep, col="dif"))+
    ggtitle(title, subtitle = subt) +
    geom_line(aes(generation, Snake_mean_Pheno, group=rep, color="Snake_mean_Pheno"),)+
    geom_line(aes(generation, Newt_mean_Pheno, group=rep, color="Newt_mean_Pheno"))+
    ylim(-5,10)+
    xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake_mean_Pheno" = "steelblue", "Newt_mean_Pheno" = "darkred", "dif" = "black"),
labels = c("Snake", "Newt", "Dif"))+
    theme_bw() +theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))
  return(plot)
}

whale_gen <- function(file, title, subt, x_label, y_label,min_gen, max_gen, rep){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)
dif <- ggplot(file[ which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep),]) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=as.numeric(Newt_pop_size), y=as.numeric(Snake_pop_size), color = dif))+
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      # strip.background = element_blank(), 
                      # strip.text.x = element_blank(), 
                      # panel.grid.major = element_blank(), 
                      # panel.grid.minor = element_blank(),
                      # panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

pop_size_hist <- function(file, title, subt, x_label, y_label,min_gen, max_gen, rep){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  file<-file[ which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep),]
  themin <- min(file$Snake_pop_size-file$Newt_pop_size) 
  themax <- max(file$Snake_pop_size-file$Newt_pop_size)
  true_max <- max(abs(themin), abs(themax))
dif <- ggplot(file) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(Snake_pop_size-Newt_pop_size), fill="darkolivegreen4")+
  xlim((-1*themax)-5, themax+5)+
  theme(plot.title = element_text(size=5)) +
  facet_wrap(~sim)+
  geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      # strip.background = element_blank(), 
                      # strip.text.x = element_blank(), 
                      # panel.grid.major = element_blank(), 
                      # panel.grid.minor = element_blank(),
                      # panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)
}

pheno_dif_hist <- function(file, title, subt, x_label, y_label,min_gen, max_gen, rep){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  file<-file[ which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep),]

  themin <- min(file$dif) 
  themax <- max(file$dif)
  true_max <- max(abs(themin), abs(themax))
dif <- ggplot(file) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen, rep ",rep)) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(dif), fill="aquamarine4")+
  xlim((-1*true_max)-1, true_max+1)+

  # y=as.numeric(Snake_pop_size-Newt_pop_size), color = snake_GA, shape = factor(newt_GA)
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      # strip.background = element_blank(), 
                      # strip.text.x = element_blank(), 
                      # panel.grid.major = element_blank(), 
                      # panel.grid.minor = element_blank(),
                      # panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

two_hist <- function(file, title, subt, x_label, y_label,min_gen, max_gen, rep){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)

  file<-file[ which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep),]

  thepopmin <- min(file$Snake_pop_size-file$Newt_pop_size)/100 
  thepopmax <- max(file$Snake_pop_size-file$Newt_pop_size)/100
  thedifmin <- min(file$dif) 
  thedifmax <- max(file$dif)
  true_max <- max(abs(thepopmin), abs(thepopmax),abs(thedifmin), abs(thedifmax))

  data_points <- nrow(file[which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep &  file$snake_GA==file$snake_GA[1] &  file$newt_GA==file$newt_GA[1]),])
dif <- ggplot(file) +

  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen, rep ",rep)) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=dif, y=..count../data_points), fill="#8B008B",  alpha = 0.65)+
  geom_histogram(aes(x=(Snake_pop_size-Newt_pop_size)/100, y=..count../data_points), fill="#008B00",  alpha = 0.65)+
  scale_y_continuous(labels = scales::percent)+
  xlim((-1*true_max)-1, true_max+1)+
  # y=as.numeric(Snake_pop_size-Newt_pop_size), color = snake_GA, shape = factor(newt_GA)
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      # strip.background = element_blank(), 
                      # strip.text.x = element_blank(), 
                      # panel.grid.major = element_blank(), 
                      # panel.grid.minor = element_blank(),
                      # panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

```

First, I will look at a plot of how the mean phenotype of newts and snakes changes overtime. Each of these plots contains four replicas for all GA combinations (letter label is the experiment that was run, 16 Sim is from GA1 and 25 Sim is from GA2). Each of these plots has three lines, red for the mean newt phenotype, blue for the mean snake phenotype, and black for the difference between mean snake and mean newt phenotype. Note this is a mean for the entire population.

```{r, graphing by gen}
trends_by_gen(GA1_long_file, title="Space: Newt VS Snake Population Size GA1", subt="Gen 16sims", y_label="Mean Phenotype", x_label="Generations")

trends_by_gen(GA2_long_file, title="Space: Newt VS Snake Population Size GA2", subt="Gen 25sims", y_label="Mean Phenotype", x_label="Generations")

```

```{r, whale}
min_gen=5000
max_gen=10000
rep=0

whale_gen(GA1_long_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake-Newt Popsize", x_label="Snake-Newt Phenotype", min_gen=min_gen, max_gen=max_gen, rep=rep )

#pop_size_hist(GA1_long_file, title="Snake-Newt Popsize Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Popsize",  min_gen=min_gen, max_gen=max_gen, rep=rep )

#pheno_dif_hist(GA1_long_file, title="Snake-Newt Phenotype Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype",  min_gen=min_gen, max_gen=max_gen, rep=rep )

two_hist(GA1_long_file, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype/100 (green) & Snake-Newt Popsize (purple)",  min_gen=min_gen, max_gen=max_gen, rep=rep )

```

```{r, combo myplots}
#files
#GA1_long_file
#GA2_long_file 
#GA1_cor_file
#GA2_cor_file

ginmin = 9000
ginmax = 10000

boxplot_cors_by_gen(GA1_cor_file, "Simulation Correlation Comparied to Real Correlation", "GA1", "newt/snake GAs&rep", "Cor", ginmin, ginmax)

boxplot_cors_by_gen(GA2_cor_file, "Simulation Correlation Comparied to Real Correlation", "GA2", "newt/snake GAs&rep", "Cor", ginmin, ginmax)

four_types_boxplot_cors_by_gen(GA1_cor_file, "Simulation four Correlation", "GA1", "newt/snake GAs&rep", "Cor", ginmin, ginmax)

four_types_boxplot_cors_by_gen(GA2_cor_file, "Simulation four Correlation", "GA2", "newt/snake GAs&rep", "Cor", ginmin, ginmax)

hist_cors_by_gen_col_rep(GA1_cor_file, "Simulation Correlation Comparied to Real Correlation", "GA1", "newt/snake GAs&rep", "Cor", ginmin, ginmax)

hist_cors_by_gen_col_rep(GA2_cor_file, "Simulation Correlation Comparied to Real Correlation", "GA1", "newt/snake GAs&rep", "Cor", ginmin, ginmax)

hist_cors_by_gen_col_rep_nofill(GA1_cor_file, "Simulation Correlation Comparied to Real Correlation", "GA1", "newt/snake GAs&rep", "Cor", ginmin, ginmax)

```

```{r, ga1 cor and pheno}
GA1_pattern <- unique(paste0(GA1_long_file$snake_mu_rate, "_", GA1_long_file$snake_mu_effect_sd,"_",GA1_long_file$newt_mu_rate, "_", GA1_long_file$newt_mu_effect_sd, "_", GA1_long_file$rep))
for(i in 1:length(GA1_pattern)){
  myine<-cor_trends_by_gen(GA1_long_file, GA1_cor_file, "test", "GA1", "newt/snake GAs&rep","other", "Cor", min_gen=0, max_gen=15000, pattern=GA1_pattern[i])
  plot(myine)
}

```

```{r, ga2 cor}
GA2_pattern <- unique(paste0(GA2_long_file$newt_mu_rate, "_", GA2_long_file$newt_mu_effect_sd, "_", GA2_long_file$snake_mu_rate, "_", GA2_long_file$snake_mu_effect_sd, "_", GA2_long_file$rep))

for(i in 1:length(GA2_pattern)){
  myine<-cor_trends_by_gen(GA2_long_file, GA2_cor_file, "test", "GA1", "newt/snake GAs&rep","other", "Cor", min_gen=0, max_gen=15000, pattern=GA2_pattern[i])
  plot(myine)
}

```

```{r, early plots}

min_gen=5000
max_gen=10000

for(i in 1:length(unique(GA1_long_file$rep))){
  myine<-whale_gen(GA1_long_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake-Newt Popsize", x_label="Snake-Newt Phenotype", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_long_file$rep)[i])
  plot(myine)
  myine<-two_hist(GA1_long_file, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype/100 (green) & Snake-Newt Popsize (purple)",  min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_long_file$rep)[i] )
  plot(myine)
}

cor_trends_by_gen(GA1_long_file, GA1_cor_file, "test", "GA1", "newt/snake GAs&rep","other", "Cor", min_gen=min_gen, max_gen=max_gen, pattern="1.0e-08_0.005_1.0e-09_0.05_0")

```

```{r, mis plots}

min_gen=20000
max_gen=25000
rep=0

for(i in 1:length(unique(GA1_long_file$rep))){
  myine<-whale_gen(GA1_long_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake-Newt Popsize", x_label="Snake-Newt Phenotype", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_long_file$rep)[i])
  plot(myine)
  myine<-two_hist(GA1_long_file, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype/100 (green) & Snake-Newt Popsize (purple)",  min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_long_file$rep)[i] )
  plot(myine)
}

cor_trends_by_gen(GA1_long_file, GA1_cor_file, "test", "GA1", "newt/snake GAs&rep","other", "Cor", min_gen=min_gen, max_gen=max_gen, pattern="1.0e-08_0.005_1.0e-09_0.05_0")

```

```{r, late plots}

min_gen=95000
max_gen=100000
rep=0

for(i in 1:length(unique(GA1_long_file$rep))){
  myine<-whale_gen(GA1_long_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake-Newt Popsize", x_label="Snake-Newt Phenotype", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_long_file$rep)[i])
  plot(myine)
  myine<-two_hist(GA1_long_file, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype/100 (green) & Snake-Newt Popsize (purple)",  min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_long_file$rep)[i] )
  plot(myine)
}

cor_trends_by_gen(GA1_long_file, GA1_cor_file, "test", "GA1", "newt/snake GAs&rep","other", "Cor", min_gen=min_gen, max_gen=max_gen, pattern="1.0e-08_0.005_1.0e-09_0.05_0")

```

```{r, really cool square plots}
file<-GA1_long_file[ which(GA1_long_file$generation>=min_gen & GA1_long_file$generation<=max_gen & GA1_long_file$rep==rep),]
file$newt_snake_rep <- paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$rep)
plot.x <- ggplot(file) + geom_boxplot(aes(newt_snake_rep, dif))
plot.y <- ggplot(file) + geom_boxplot(aes(newt_snake_rep, Snake_pop_size-Newt_pop_size))

grid.arrange(plot.x, plot.y, ncol=2)

plot.x <- layer_data(plot.x)[,1:6]
plot.y <- layer_data(plot.y)[,1:6]
colnames(plot.x) <- paste0("x.", gsub("y", "", colnames(plot.x)))
colnames(plot.y) <- paste0("y.", gsub("y", "", colnames(plot.y)))
df <- cbind(plot.x, plot.y); rm(plot.x, plot.y)
df$category <- sort(unique(file$newt_snake_rep))
df.outliers <- df %>%
  select(category, x.middle, x.outliers, y.middle, y.outliers) %>%
  data.table::data.table()
#df.outliers <- df.outliers[, list(x.outliers = unlist(x.outliers), y.outliers = unlist(y.outliers)), by = list(category, x.middle, y.middle)]
ggplot(df, aes(fill = category, color = category)) +

  # 2D box defined by the Q1 & Q3 values in each dimension, with outline
  geom_rect(aes(xmin = x.lower, xmax = x.upper, ymin = y.lower, ymax = y.upper), alpha = 0.3) +
  geom_rect(aes(xmin = x.lower, xmax = x.upper, ymin = y.lower, ymax = y.upper), 
            color = "black", fill = NA) +

  # whiskers for x-axis dimension with ends
  geom_segment(aes(x = x.min, y = y.middle, xend = x.max, yend = y.middle)) + #whiskers
  geom_segment(aes(x = x.min, y = y.lower, xend = x.min, yend = y.upper)) + #lower end
  geom_segment(aes(x = x.max, y = y.lower, xend = x.max, yend = y.upper)) + #upper end

  # whiskers for y-axis dimension with ends
  geom_segment(aes(x = x.middle, y = y.min, xend = x.middle, yend = y.max)) + #whiskers
  geom_segment(aes(x = x.lower, y = y.min, xend = x.upper, yend = y.min)) + #lower end
  geom_segment(aes(x = x.lower, y = y.max, xend = x.upper, yend = y.max)) + #upper end

  # outliers
  #geom_point(data = df.outliers, aes(x = x.outliers, y = y.middle), size = 3, shape = 1) + # x-direction
  #geom_point(data = df.outliers, aes(x = x.middle, y = y.outliers), size = 3, shape = 1) + # y-direction

  xlab("s-n mean pheno") + ylab("s-n pop size") +  
  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      strip.background = element_blank(), 
                      strip.text.x = element_blank(), 
                      legend.position="right",
                      legend.text=element_text(size=5),
                      panel.grid.major = element_blank(), 
                      panel.grid.minor = element_blank(),
                      panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

```

Next, I will discuss a summary plot for the different GA sets. The first plot has the newt population size by snake population size (a larger population size might indicate a co-evolution advantage) and the second plot is the difference between snake and newt mean phenotype and snake population size. There are four points for each GA set (one for each replica). Color will represent the four GAs of the snake and the shapes will represent the four GAs of newts. Each plot will have a starting generation and an ending generation (that can easily be changed).

```{r, graphing the summary}

GA_summary(GA1_long_file, min_gen=5000, max_gen=10000, x_line="Newt_pop_size", y_line="Snake_pop_size", title="Space: Newt VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize")

GA_summary(GA1_long_file, min_gen=5000, max_gen=10000, x_line="dif", y_line="Snake_pop_size", title="Space: Dif VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Dif")

GA_summary(GA2_long_file, min_gen=5000, max_gen=10000, x_line="Newt_pop_size", y_line="Snake_pop_size", title="Space: Newt VS Snake Population Size", subt="Gen 25sims", y_label="Snake Popsize", x_label="Newt Popsize")

GA_summary(GA2_long_file, min_gen=5000, max_gen=10000, x_line="dif", y_line="Snake_pop_size", title="Space: Dif VS Snake Population Size", subt="Gen 25sims", y_label="Snake Popsize", x_label="Dif")

```

This next section is a more visual representation of the summary plots. Here, there will be four heatmaps. The first two show the results from the GA1 experiment with the last two showing the results from the GA2 experiment. The heatmaps have the newt GAs on the x-axis and the snake GAs and the simulation replication number on the y-axis. The colors in the heat might will either be yellow to purple for the snake population size or blue to red for the difference in mean phenotypes.

```{r, heatmap}

#params_GA1

build_heatmap(GA1_long_file, min_gen=5000, max_gen=10000, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

build_heatmap(GA1_long_file, min_gen=5000, max_gen=10000, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)

#build_heatmap(GA1_long_file, min_gen=5000, max_gen=10000, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=F)

build_heatmap(GA2_long_file, min_gen=10000, max_gen=11000, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="25 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

build_heatmap(GA2_long_file, min_gen=000, max_gen=5000, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="25 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

build_heatmap(GA2_long_file, min_gen=15000, max_gen=20000, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="25 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

build_heatmap(GA2_long_file, min_gen=20000, max_gen=25000, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="25 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

build_heatmap(GA2_long_file, min_gen=25000, max_gen=30000, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="25 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

build_heatmap(GA2_long_file, min_gen=5000, max_gen=10000, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="25 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)

```