### 2022-05-29 graphs_newt_snakes/Dac_2022.Rmd
made some more figures for the poster. started to look at the lond file named tall data. Made a back up snakemake file

---
title: "graph dor DAC"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, fitness graph}
setCOST=100
Phenotype= seq(1, 20, 0.5)
the_y = exp(-(Phenotype/setCOST)^2)

plot(Phenotype, the_y, type="l", lwd=3)

Phenotype= seq(1, 20, 0.5)
Cost = 1-exp(-(Phenotype/setCOST)^2)

plot(Phenotype, Cost, type="l", lwd=3)
#how to change this to the probably of survival

```

```{r, my plots for poster}

min_gen=100
max_gen=20000
rep=0
unique(GA1_long_file$snake_mu_rate)
ggplot(GA1_long_file[ which(GA1_long_file$generation>=min_gen & GA1_long_file$generation<=max_gen & GA1_long_file$rep==rep & GA1_long_file$snake_mu_rate=="1.0e-10" & GA1_long_file$newt_mu_rate=="1.0e-10"),]) +
    #facet_wrap(~sim)+
    geom_line(aes(generation, dif, col="dif"), size=1 )+
    ggtitle("test", subtitle = paste0(" From ", min_gen, " gen to ", max_gen, " gen")) +
    geom_line(aes(generation, Snake_mean_Pheno, color="Snake_mean_Pheno"), size=1 )+
    geom_line(aes(generation, Newt_mean_Pheno, color="Newt_mean_Pheno"), size=1 )+
    geom_hline(yintercept=0, linetype="dashed", color="gray54", size=1 )+
    ylim(-5,8)+
    #xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake_mean_Pheno" = "steelblue", "Newt_mean_Pheno" = "darkred", "dif" = "black"),
labels = c("Snake", "Newt", "Dif"))+
    theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())
ggsave(
  "~/Desktop/newt_snake_coevo_example.pdf",
  plot = last_plot(),
  scale = 1,
  width = 7,
  height = 5,
  #units = c("in"),
  dpi = 300,
  limitsize = TRUE,
  bg = NULL,
)
```