### 2022-06-24 graphs_newt_snakes/Dac_2022.Rmd
looking at some new data

---
title: "graph dor DAC"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, fitness graph}
setCOST=100
Phenotype= seq(1, 20, 0.5)
the_y = exp(-(Phenotype/setCOST)^2)

plot(Phenotype, the_y, type="l", lwd=3)

Phenotype= seq(1, 20, 0.5)
Cost = 1-exp(-(Phenotype/setCOST)^2)

plot(Phenotype, Cost, type="l", lwd=3)
#how to change this to the probably of survival

```

```{r, my plots for poster}

min_gen=100
max_gen=20000
rep=0
unique(GA1_long_file$snake_mu_rate)
ggplot(GA1_long_file[ which(GA1_long_file$generation>=min_gen & GA1_long_file$generation<=max_gen & GA1_long_file$rep==rep & GA1_long_file$snake_mu_rate=="1.0e-10" & GA1_long_file$newt_mu_rate=="1.0e-10"),]) +
    #facet_wrap(~sim)+
    geom_line(aes(generation, dif, col="dif"), size=1 )+
    ggtitle("test", subtitle = paste0(" From ", min_gen, " gen to ", max_gen, " gen")) +
    geom_line(aes(generation, Snake_mean_Pheno, color="Snake_mean_Pheno"), size=1 )+
    geom_line(aes(generation, Newt_mean_Pheno, color="Newt_mean_Pheno"), size=1 )+
    geom_hline(yintercept=0, linetype="dashed", color="gray54", size=1 )+
    ylim(-5,8)+
    #xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake_mean_Pheno" = "steelblue", "Newt_mean_Pheno" = "darkred", "dif" = "black"),
labels = c("Snake", "Newt", "Dif"))+
    theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())
# ggsave(
#   "~/Desktop/newt_snake_coevo_example.pdf",
#   plot = last_plot(),
#   scale = 1,
#   width = 7,
#   height = 5,
#   #units = c("in"),
#   dpi = 300,
#   limitsize = TRUE,
#   bg = NULL,
# )

title="Snake-Newt Winning Hist"
subt="Gen 16sims"
y_label="Count"
x_label="Snake-Newt Phenotype/100 (green) & Snake-Newt Popsize (purple)"
min_gen=5000
max_gen=15000

GA1_long_file$snake_GA<-paste0(GA1_long_file$snake_mu_rate, "_", GA1_long_file$snake_mu_effect_sd)
  GA1_long_file$newt_GA<-paste0(GA1_long_file$newt_mu_rate, "_", GA1_long_file$newt_mu_effect_sd)

  GA1_fig_file<-GA1_long_file[ which(GA1_long_file$generation>=min_gen & GA1_long_file$generation<=max_gen),]

  thepopmin <- min(GA1_fig_file$Snake_pop_size-GA1_fig_file$Newt_pop_size)/100 
  thepopmax <- max(GA1_fig_file$Snake_pop_size-GA1_fig_file$Newt_pop_size)/100
  thedifmin <- min(GA1_fig_file$dif) 
  thedifmax <- max(GA1_fig_file$dif)
  true_max <- max(abs(thepopmin), abs(thepopmax),abs(thedifmin), abs(thedifmax))

  data_points <- nrow(GA1_fig_file[which(GA1_fig_file$generation>=min_gen & GA1_fig_file$generation<=max_gen  &  GA1_fig_file$snake_GA==GA1_fig_file$snake_GA[1] &  GA1_fig_file$newt_GA==GA1_fig_file$newt_GA[1]),])
  print(data_points)
  GA1_fig_file$dif_pop<- GA1_fig_file$Snake_pop_size - GA1_fig_file$Newt_pop_size

ggplot(GA1_fig_file) +

  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=dif, y=..count../data_points), fill="#8B008B",  alpha = 0.65, bins = 20)+
  geom_histogram(aes(x=(dif_pop)/100, y=..count../data_points), fill="#008B00",  alpha = 0.65, bins = 20)+
  scale_y_continuous(labels = scales::percent)+
#limits=c(0, .6)
  # y=as.numeric(Snake_pop_size-Newt_pop_size), color = snake_GA, shape = factor(newt_GA)
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      #strip.background = element_blank(), 
                      strip.text.x = element_blank(), 
                      #panel.grid.major = element_blank(), 
                      #panel.grid.minor = element_blank(),
                      #panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

ggsave(
  "~/Desktop/newt_snake_hist.pdf",
  plot = last_plot(),
  scale = 1,
  width = 14,
  height = 6,
  #units = c("in"),
  dpi = 300,
  limitsize = TRUE,
  bg = NULL,
)

title="Snake-Newt Winning Hist"
subt="Gen 16sims"
y_label="Count"
x_label="Snake-Newt Phenotype/100 (green) & Snake-Newt Popsize (purple)"
min_gen=85000
max_gen=100000

GA1_long_file$snake_GA<-paste0(GA1_long_file$snake_mu_rate, "_", GA1_long_file$snake_mu_effect_sd)
  GA1_long_file$newt_GA<-paste0(GA1_long_file$newt_mu_rate, "_", GA1_long_file$newt_mu_effect_sd)

  GA1_fig_file<-GA1_long_file[ which(GA1_long_file$generation>=min_gen & GA1_long_file$generation<=max_gen),]

  thepopmin <- min(GA1_fig_file$Snake_pop_size-GA1_fig_file$Newt_pop_size)/100 
  thepopmax <- max(GA1_fig_file$Snake_pop_size-GA1_fig_file$Newt_pop_size)/100
  thedifmin <- min(GA1_fig_file$dif) 
  thedifmax <- max(GA1_fig_file$dif)
  true_max <- max(abs(thepopmin), abs(thepopmax),abs(thedifmin), abs(thedifmax))

  data_points <- nrow(GA1_fig_file[which(GA1_fig_file$generation>=min_gen & GA1_fig_file$generation<=max_gen  &  GA1_fig_file$snake_GA==GA1_fig_file$snake_GA[1] &  GA1_fig_file$newt_GA==GA1_fig_file$newt_GA[1]),])
  print(data_points)
ggplot(GA1_fig_file) +

  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=dif, y=..count../data_points), fill="#8B008B",  alpha = 0.65, bins = 20)+
  geom_histogram(aes(x=(Snake_pop_size-Newt_pop_size)/100, y=..count../data_points), fill="#008B00",  alpha = 0.65, bins = 20)+
  scale_y_continuous(labels = scales::percent)+
  #xlim((-1*true_max)-1, true_max+1)+
  # y=as.numeric(Snake_pop_size-Newt_pop_size), color = snake_GA, shape = factor(newt_GA)
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      #strip.background = element_blank(), 
                      strip.text.x = element_blank(), 
                      #panel.grid.major = element_blank(), 
                      #panel.grid.minor = element_blank(),
                      #panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

ggsave(
  "~/Desktop/newt_snake_hist_late.pdf",
  plot = last_plot(),
  scale = 1,
  width = 7,
  height = 5,
  #units = c("in"),
  dpi = 300,
  limitsize = TRUE,
  bg = NULL,
)

fakedata<-as.data.frame(rnorm(1000,0, 2))

ggplot(fakedata) +
  ggtitle("Phenotype Dif")+
geom_histogram(aes(x=`rnorm(1000, 0, 2)`), fill="#8B008B",  alpha = 0.65, bins = 20)+
theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      #strip.background = element_blank(), 
                      strip.text.x = element_blank(), 
                      #panel.grid.major = element_blank(), 
                      #panel.grid.minor = element_blank(),
                      #panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

# ggsave(
#   "~/Desktop/hist_pheno_dif.pdf",
#   plot = last_plot(),
#   scale = 1,
#   width = 7,
#   height = 5,
#   #units = c("in"),
#   dpi = 300,
#   limitsize = TRUE,
#   bg = NULL,
# )

fakedata$popsize<-rnorm(1000,0, 4)

ggplot(fakedata) +
  ggtitle("Phenotype Dif")+
geom_histogram(aes(x=popsize), fill="#008B00",  alpha = 0.65, bins = 20)+
theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      #strip.background = element_blank(), 
                      strip.text.x = element_blank(), 
                      #panel.grid.major = element_blank(), 
                      #panel.grid.minor = element_blank(),
                      #panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

# ggsave(
#   "~/Desktop/hist_popsize_dif.pdf",
#   plot = last_plot(),
#   scale = 1,
#   width = 7,
#   height = 5,
#   #units = c("in"),
#   dpi = 300,
#   limitsize = TRUE,
#   bg = NULL,
# )
```

```{r, more plots for poster}
min_gen=100
max_gen=6000
rep=0
pattern="1.0e-09_0.5_1.0e-10_0.05_0"
GA1_long_file$snake_GA
unique(GA1_long_file$snake_mu_rate)
ggplot(GA1_long_file[ which(GA1_long_file$generation>=min_gen & GA1_long_file$generation<=max_gen & GA1_long_file$rep==rep & GA1_long_file$snake_mu_rate=="1.0e-09" & GA1_long_file$newt_mu_rate=="1.0e-10"),]) +
    #facet_wrap(~sim)+
    ggtitle("test", subtitle = paste0(" From ", min_gen, " gen to ", max_gen, " gen")) +
    geom_line(aes(generation, Snake_mean_Pheno, color="Snake_mean_Pheno"), size=1 )+
    geom_line(aes(generation, Newt_mean_Pheno, color="Newt_mean_Pheno"), size=1 )+
    ylim(-5,8)+
    #xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake_mean_Pheno" = "steelblue", "Newt_mean_Pheno" = "darkred", "dif" = "black"),
labels = c("Snake", "Newt", "Dif"))+
    theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

# ggsave(
#   "~/Desktop/catch_up_part1.pdf",
#   plot = last_plot(),
#   scale = 1,
#   width = 7,
#   height = 5,
#   #units = c("in"),
#   dpi = 300,
#   limitsize = TRUE,
#   bg = NULL,
# )

min_gen=4500
max_gen=85000
rep=0
pattern="1.0e-09_0.5_1.0e-10_0.05_0"
GA1_long_file$snake_GA
unique(GA1_long_file$snake_mu_rate)
ggplot(GA1_long_file[ which(GA1_long_file$generation>=min_gen & GA1_long_file$generation<=max_gen & GA1_long_file$rep==rep & GA1_long_file$snake_mu_rate=="1.0e-09" & GA1_long_file$newt_mu_rate=="1.0e-10"),]) +
    #facet_wrap(~sim)+
    ggtitle("test", subtitle = paste0(" From ", min_gen, " gen to ", max_gen, " gen")) +
    geom_line(aes(generation, Snake_mean_Pheno, color="Snake_mean_Pheno"), size=1 )+
    geom_line(aes(generation, Newt_mean_Pheno, color="Newt_mean_Pheno"), size=1 )+
    ylim(-5,8)+
    #xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake_mean_Pheno" = "steelblue", "Newt_mean_Pheno" = "darkred", "dif" = "black"),
labels = c("Snake", "Newt", "Dif"))+
    theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

# ggsave(
#   "~/Desktop/catch_up_part2.pdf",
#   plot = last_plot(),
#   scale = 1,
#   width = 7,
#   height = 5,
#   #units = c("in"),
#   dpi = 300,
#   limitsize = TRUE,
#   bg = NULL,
# )

min_gen=80000
max_gen=99999
rep=0
pattern="1.0e-09_0.5_1.0e-10_0.05_0"
GA1_long_file$snake_GA
unique(GA1_long_file$snake_mu_rate)
ggplot(GA1_long_file[ which(GA1_long_file$generation>=min_gen & GA1_long_file$generation<=max_gen & GA1_long_file$rep==rep & GA1_long_file$snake_mu_rate=="1.0e-09" & GA1_long_file$newt_mu_rate=="1.0e-10"),]) +
    #facet_wrap(~sim)+
    ggtitle("test", subtitle = paste0(" From ", min_gen, " gen to ", max_gen, " gen")) +
    geom_line(aes(generation, Snake_mean_Pheno, color="Snake_mean_Pheno"), size=1 )+
    geom_line(aes(generation, Newt_mean_Pheno, color="Newt_mean_Pheno"), size=1 )+
    ylim(-5,8)+
    #xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake_mean_Pheno" = "steelblue", "Newt_mean_Pheno" = "darkred", "dif" = "black"),
labels = c("Snake", "Newt", "Dif"))+
    theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

# ggsave(
#   "~/Desktop/catch_up_part3.pdf",
#   plot = last_plot(),
#   scale = 1,
#   width = 7,
#   height = 5,
#   #units = c("in"),
#   dpi = 300,
#   limitsize = TRUE,
#   bg = NULL,
# )
min_gen=100
max_gen=99999
rep=0
pattern="1.0e-09_0.5_1.0e-10_0.05_0"
GA1_long_file$snake_GA
unique(GA1_long_file$snake_mu_rate)
ggplot(GA1_long_file[ which(GA1_long_file$generation>=min_gen & GA1_long_file$generation<=max_gen & GA1_long_file$rep==rep & GA1_long_file$snake_mu_rate=="1.0e-09" & GA1_long_file$newt_mu_rate=="1.0e-10"),]) +
    #facet_wrap(~sim)+
    ggtitle("test", subtitle = paste0(" From ", min_gen, " gen to ", max_gen, " gen")) +
    geom_line(aes(generation, Snake_mean_Pheno, color="Snake_mean_Pheno"), size=1 )+
    geom_line(aes(generation, Newt_mean_Pheno, color="Newt_mean_Pheno"), size=1 )+
    #xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake_mean_Pheno" = "steelblue", "Newt_mean_Pheno" = "darkred", "dif" = "black"),
labels = c("Snake", "Newt", "Dif"))+
    theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

# ggsave(
#   "~/Desktop/catch_up_all.pdf",
#   plot = last_plot(),
#   scale = 1,
#   width = 7,
#   height = 5,
#   #units = c("in"),
#   dpi = 300,
#   limitsize = TRUE,
#   bg = NULL,
# )
```