### 2022-08-29 graphs_newt_snakes/GA3.Rmd
I am writing and making my markdown more readable

---
title: "GA3"
output: html_document
date: '2022-07-21'
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE,echo = FALSE)

library(ggplot2)
library(viridis)
library(colorspace)
library(gridExtra)
library(tidyverse)
require(reshape2)
library(plyr)
library(gridGraphics)
library(grid)
library(knitr)
library(kableExtra)
library("reshape")
library(ggplot2)
library(ggpubr)
theme_set(theme_pubr())
library(stringr)
library(patchwork)
```

GA3 file will be the slow evolution file. The main goal will be testing to see if there is an effect caused by mutation rate or effect size when the mutational variance between two species is the same.

Prediction: Higher mutation rate will be advantages in the beginning of a simulation. A higher effect size might lead to more spatial co-evolution. Space might separate the effects.

For the initial part of this experiment I will only be running one trial per GA combination

GA combinations:

- 1e-8*(0.005)^2 = 2.5e-13
- 1e-9*(0.0158)^2 = 2.5e-13
- 1e-10*(0.05)^2 = 2.5e-13
- 1e-11*(0.158)^2 = 2.5e-13

## Data Colllection

I have created some bash script to run through all the files in my data folders and add some columns and coagulate all of the files into one file per data set. In this case, there are three data sets; one for whole population data (lit), one for correlation data (cor), and one for individual grids (grid). After these all file is created, I load it into R and finish adding in some comparative columns. I also load in my figure creation functions. There originally was a problem with naming where I was putting the newt information in first (fixed, but is something to be aware of moving forward).

```{bash, data file gathering}
#!/usr/bin/env bash

# Run these comands while in The the folder that contains the file you would like to condence
# Goal is to condence all the text files into one file 
# Requirements: need to have the same name and need to add additional columns
thepath=/Users/victoria/Desktop/
foldername=GA3
thefolder=${thepath}${foldername}/*

add_in_column=",rep,snake_mu_rate,snake_mu_effect_sd,newt_mu_rate,newt_mu_effect_sd"

cor_head_file=$(find ${thepath}${foldername} -name "*_cor_*" | head -n 1)
cor_head=$(head -n 1 "${cor_head_file}")

grid_head_file=$(find ${thepath}${foldername} -name "*_grid_*" | head -n 1)
grid_head=$(head -n 1 "${grid_head_file}")

lit_head_file=$(find ${thepath}${foldername} -name "*_lit_*" | head -n 1)
lit_head=$(head -n 1 "${lit_head_file}")

#test -f /Users/victoria/Desktop/bash_test/all_lit.txt && echo "all lit file exists."
#test -f /Users/victoria/Desktop/bash_test/all_grid.txt && echo "all grid file exists."
#test -f /Users/victoria/Desktop/bash_test/all_cor.txt && echo "all cor file exists."

if [[ -f ${thepath}${foldername}/all_lit.txt ]] && [[ -f ${thepath}${foldername}/all_grid.txt ]] && [[ -f ${thepath}${foldername}/all_cor.txt ]]
then
  echo "All cor, lit, and grid files exist!"
  echo "This program will now end!"
  exit 0
fi

#creating the files
echo ${lit_head}${add_in_column} > ${thepath}${foldername}/all_lit.txt
echo ${grid_head}${add_in_column} > ${thepath}${foldername}/all_grid.txt
echo ${cor_head}${add_in_column} > ${thepath}${foldername}/all_cor.txt

for filename in $thefolder*.text;
  do 
#getting the variables 
#my_type=$(echo ${filename} |sed 's/.*GA205_//; s/_.*//')
my_other_type=$(echo ${filename} |sed 's/.*GA520_//; s/_.*//')
snake_mu_rate=$(echo ${filename} |sed 's/.*snake_mu_rate_//; s/_.*//')
snake_mu_effect_sd=$(echo ${filename} |sed 's/.*snake_mu_effect_sd_//; s/_.*//')
newt_mu_rate=$(echo ${filename} |sed 's/.*newt_mu_rate_//; s/_.*//')
newt_mu_effect_sd=$(echo ${filename} |sed 's/.*newt_mu_effect_sd_//; s/_.*//')
rep=$(echo ${filename} | sed "s/.*rep_//; s/_.*//")
#echo ${newt_mu_rate}	
# echo $my_type
# echo $my_other_type
# these new columns adding them to the data
sed -i '' -e 's/$/,'${rep}'/' $filename
sed -i '' -e 's/$/,'${snake_mu_rate}'/' $filename
sed -i '' -e 's/$/,'${snake_mu_effect_sd}'/' $filename
sed -i '' -e 's/$/,'${newt_mu_rate}'/' $filename
sed -i '' -e 's/$/,'${newt_mu_effect_sd}'/' $filename
#infile >outfile
if [[ $my_type == "lit" ]]
then
    tail -n +2 -q $filename >> ${thepath}${foldername}/all_lit.txt
fi

if [[ $my_type == "grid" ]]
then
    tail -n +2 -q $filename >> ${thepath}${foldername}/all_grid.txt
fi

if [[ $my_other_type == "cor" ]]
then
    tail -n +2 -q $filename >> ${thepath}${foldername}/all_cor.txt
fi

done

```

```{r, data gathering}

folder="GA3"
filelocation=paste0("~/Desktop/",folder)

snewt <- read.csv("~/Desktop/snake_newt_real_data.csv", header = TRUE)

GA3_tall_tall_file<- read.csv(paste0(filelocation, "/all_lit.txt") , header = TRUE)

GA3_tall_tall_file$group_rep <- paste0(GA3_tall_tall_file$snake_mu_rate, "_", GA3_tall_tall_file$snake_mu_effect_sd,"_",GA3_tall_tall_file$newt_mu_rate, "_", GA3_tall_tall_file$newt_mu_effect_sd, "_", GA3_tall_tall_file$rep)

GA3_tall_tall_file$group <- paste0(GA3_tall_tall_file$snake_mu_rate, "_", GA3_tall_tall_file$snake_mu_effect_sd,"_",GA3_tall_tall_file$newt_mu_rate, "_", GA3_tall_tall_file$newt_mu_effect_sd)

GA3_tall_tall_file$sim=LETTERS[factor(GA3_tall_tall_file$group)]
GA3_tall_tall_file$dif <- GA3_tall_tall_file$Snake_mean_Pheno-GA3_tall_tall_file$Newt_mean_Pheno

GA3_cor_tall_file<- read.csv(paste0(filelocation,"/all_cor.txt") , header = TRUE)
GA3_cor_tall_file$group_rep <- paste0(GA3_cor_tall_file$snake_mu_rate, "_", GA3_cor_tall_file$snake_mu_effect_sd,"_",GA3_cor_tall_file$newt_mu_rate, "_", GA3_cor_tall_file$newt_mu_effect_sd, "_", GA3_cor_tall_file$rep)
GA3_cor_tall_file$group <- paste0(GA3_cor_tall_file$snake_mu_rate, "_", GA3_cor_tall_file$snake_mu_effect_sd,"_",GA3_cor_tall_file$newt_mu_rate, "_", GA3_cor_tall_file$newt_mu_effect_sd)
GA3_cor_tall_file$sim=LETTERS[factor(GA3_cor_tall_file$group)]

GA3_GA3_grid_tall_file<-read.csv(paste0(filelocation,"/all_grid.txt") , header = TRUE)
GA3_GA3_grid_tall_file$group <- paste0(GA3_GA3_grid_tall_file$snake_mu_rate, "_", GA3_GA3_grid_tall_file$snake_mu_effect_sd,"_",GA3_GA3_grid_tall_file$newt_mu_rate, "_", GA3_GA3_grid_tall_file$newt_mu_effect_sd)
GA3_GA3_grid_tall_file$group_rep <- paste0(GA3_GA3_grid_tall_file$snake_mu_rate, "_", GA3_GA3_grid_tall_file$snake_mu_effect_sd,"_",GA3_GA3_grid_tall_file$newt_mu_rate, "_", GA3_GA3_grid_tall_file$newt_mu_effect_sd, "_", GA3_GA3_grid_tall_file$rep)

GA3_GA3_grid_tall_file$sim=LETTERS[factor(GA3_GA3_grid_tall_file$group)]
GA3_GA3_grid_tall_file$snake_newt_rep <- paste0("Snake_",GA3_GA3_grid_tall_file$snake_mu_rate, "_", GA3_GA3_grid_tall_file$snake_mu_effect_sd, "_Newt_",GA3_GA3_grid_tall_file$newt_mu_rate, "_", GA3_GA3_grid_tall_file$newt_mu_effect_sd, "_", GA3_GA3_grid_tall_file$rep)

```

```{r, cor box set up}

boxplot_cors_by_gen <- function(file, title, subt, x_label, y_label, min_gen, max_gen){
 snewt$mean_res = snewt$mean_toxin+snewt$dif
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
  file$snake_newt <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd)

  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ylim(-1.1, 1.1)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor(rep), y=as.numeric(mean_newt_pheno_By_mean_snake_pheno), color = "CA"))+
  geom_hline(yintercept=real_snewt_cor, linetype="dashed")+
    geom_hline(yintercept=0, linetype="solid")+
  #geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  #geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  #geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  theme(plot.title = element_text(size=5)) +
  facet_wrap(~snake_newt)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

  return(plot)
}

hist_cors_by_gen_col_rep <- function(file, title, subt, x_label, y_label, min_gen, max_gen){
 snewt$mean_res = snewt$mean_toxin+snewt$dif
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
   file$snake_newt <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  #unique(file$sim)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=as.numeric(mean_newt_pheno_By_mean_snake_pheno),fill=factor(rep)))+
  geom_vline(xintercept=real_snewt_cor, linetype="dashed")+
  geom_vline(xintercept=0, linetype="solid")+
  #geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  #geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  #geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  theme(plot.title = element_text(size=5)) +
  xlim(-1.1, 1.1)+
  facet_wrap(~snake_newt)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

  return(plot)
}

hist_cors_by_gen_col_rep_nofill <- function(file, title, subt, x_label, y_label, min_gen, max_gen){
 snewt$mean_res = snewt$mean_toxin+snewt$dif
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
  file$snake_newt <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  unique(file$sim)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=as.numeric(mean_newt_pheno_By_mean_snake_pheno)))+
  geom_vline(xintercept=real_snewt_cor, linetype="dashed")+
  geom_vline(xintercept=0, linetype="solid")+
  #geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  #geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  #geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  theme(plot.title = element_text(size=5)) +
  #stat_bin(bins=30)+
  facet_wrap(~snake_newt)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

suppressMessages(print(plot))

  return(plot)
}

var_cor <- function(file, cfile, title, subt, x_label, y_label, min_gen, max_gen){
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
  real_snewt_var = var(snewt$dif)
  var_cor_table=data.frame()

  for(i in 1:length(unique(file$snake_newt_rep))){
    pattern<-unique(file$snake_newt_rep)[i]
    temp_file<-file[which(file$gen>=min_gen & file$gen<=max_gen & file$snake_newt_rep==pattern),]
    temp_cor_file<-cfile[which(cfile$gen>=min_gen & cfile$gen<=max_gen & cfile$snake_newt_rep==pattern),]

    var_cor_table_temp<- data.frame("pattern"=pattern, "var"= var(temp_file$dif),"cor" = cor(temp_file$newt_toxin, temp_file$snake_res), "mean_corr" = mean(temp_cor_file$mean_newt_pheno_By_mean_snake_pheno))
    var_cor_table <- rbind(var_cor_table, var_cor_table_temp)
  }

  var_cor_table_temp<- data.frame("pattern"="real", "var"= real_snewt_var, "cor" = real_snewt_cor, "mean_corr" = real_snewt_cor)
  var_cor_table <- rbind(var_cor_table, var_cor_table_temp)

  plot <- ggplot(var_cor_table) +
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=as.numeric(var), y=as.numeric(cor), color = pattern), shape=17, size=3)+
  geom_point(aes(x=as.numeric(var), y=as.numeric(mean_corr), color = pattern), size=3)+
  theme(plot.title = element_text(size=5)) +
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))

  return(plot)
}

cor_trends_by_gen <- function(file, cfile, title, subt, x_label, x_label_cor, y_label, min_gen, max_gen, pattern){
#file=GA3_tall_tall_file
#cfile=GA3_cor_tall_file 
 # pattern="1.0e-10_0.05_1.0e-08_0.005_3"
 file$snake_newt_rep <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd,"_", file$rep)
  cfile$snake_newt_rep <- paste0(cfile$snake_mu_rate, "_", cfile$snake_mu_effect_sd, "_", cfile$newt_mu_rate, "_", cfile$newt_mu_effect_sd, "_", cfile$rep)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen & file$snake_newt_rep==pattern),]) +
    #geom_line(aes(generation, dif, group=rep, col="dif"))+
      ggtitle(title, subtitle = paste0(subt," pattern ", pattern, " From ", min_gen, " gen to ", max_gen, " gen"))+
    geom_line(aes(generation, Snake_mean_Pheno, group=rep, color="Snake_mean_Pheno"),)+
    geom_line(aes(generation, Newt_mean_Pheno, group=rep, color="Newt_mean_Pheno"))+
    #ylim(0,10)+
    #xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake_mean_Pheno" = "steelblue", "Newt_mean_Pheno" = "darkred"),
labels = c("Snake", "Newt"))+
    theme_bw() +theme(legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank(), axis.ticks.x = element_blank(), axis.text.x = element_blank(),
                      plot.margin=unit(c(1,1,+0.3,1), "cm"))

  cor_plot <- ggplot(cfile[which(cfile$gen>=min_gen & cfile$gen<=max_gen & cfile$snake_newt_rep==pattern),]) +
    geom_line(aes(generation, mean_newt_pheno_By_mean_snake_pheno, col="CA"))+
    #ggtitle(title, subtitle = subt) +
    ylim(-1.1, 1.1)+
    #xlab(x_label_cor) + ylab(y_label)+
    theme_bw() +theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank(),
                      plot.margin=unit(c(-0.5,1,1,1), "cm"))

  file<-file[which(file$gen>=min_gen & file$gen<=max_gen & file$snake_newt_rep==pattern),]
  cfile<-cfile[which(cfile$gen>=min_gen & cfile$gen<=max_gen & cfile$snake_newt_rep==pattern),]

  if(nrow(cfile)>nrow(file)){
    toDelete <- seq(0, nrow(cfile), 2)
    cfile_short <-  cfile[-toDelete, ]
  }else{
    cfile_short <-  cfile
  }

  if(nrow(cfile_short)>nrow(file)){
      toDelete <- seq(0, nrow(cfile_short), 2)
      cfile_short <-  cfile_short[-toDelete, ]
    }
  figure <- ggarrange(plot, cor_plot + font("x.text", size = 10),heights = c(2, 1.5),ncol = 1, nrow = 2, align = "v")
    print(paste0("pattern ", pattern))
  print(paste0("Cor between average snake pheno and local cor ", cor(file$Snake_mean_Pheno, cfile_short$mean_newt_pheno_By_mean_snake_pheno)))
  print(paste0("Cor between average newt pheno and local cor ", cor(file$Newt_mean_Pheno, cfile_short$mean_newt_pheno_By_mean_snake_pheno)))
  print(paste0("Cor between average dif pheno and local cor ", cor(file$dif, cfile_short$mean_newt_pheno_By_mean_snake_pheno)))
  print(paste0("Cor between newt pheno and snake ", cor(file$Newt_mean_Pheno, file$Snake_mean_Pheno)))
  return(figure)
}

four_types_boxplot_cors_by_gen <- function(file,title, subt, x_label, y_label, min_gen, max_gen){
  file$snake_newt <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ylim(-1.1, 1.1)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor("CA"), y=as.numeric(mean_newt_pheno_By_mean_snake_pheno), color = "CA"))+
  geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  geom_hline(yintercept=0, linetype="solid")+
  theme(plot.title = element_text(size=5)) +
  facet_wrap(~snake_newt)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

  return(plot)
}

```

```{r, general file and graph making}

#the functions to make the params file and the summary and heatmap figures

create_parm_file <- function(file, min_gen, max_gen){
params <- data_frame()

for(i in 1:length(unique(file$snake_newt_rep))){

  pattern=unique(file$snake_newt_rep)[i]
  sumitup <- file[which(file$gen>=min_gen & file$gen<=max_gen & file$snake_newt_rep ==pattern),]

  params_temp<- data_frame(Newt_mean_Pheno=mean(sumitup$Newt_mean_Pheno),
                           Snake_mean_Pheno=mean(sumitup$Snake_mean_Pheno), 
                           Newt_mean_Pheno_sd=sd(sumitup$Newt_mean_Pheno),
                           Snake_mean_Pheno_sd=sd(sumitup$Snake_mean_Pheno),
                           Newt_pop_size=mean(sumitup$Newt_pop_size),
                           Snake_pop_size=mean(sumitup$Snake_pop_size),
                           dif=mean(sumitup$dif),
                           snake_mu_rate=sumitup$snake_mu_rate[1],
                           newt_mu_rate=sumitup$newt_mu_rate[1],
                           snake_mu_effect_sd=sumitup$snake_mu_effect_sd[1],
                           newt_mu_effect_sd=sumitup$newt_mu_effect_sd[1],
                           rep=sumitup$rep[1])

  params <- rbind(params, params_temp)
}

params$snake <- paste0(params$snake_mu_rate, "_",params$snake_mu_effect_sd)
params$newt <- paste0(params$newt_mu_rate, "_",params$newt_mu_effect_sd)
params$snake_newt <- paste0(params$snake, "_", params$newt)
return(params)
}

GA_summary <- function(file, min_gen, max_gen, x_line, y_line, title, subt, y_label, x_label){

file$snake_newt_rep <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$rep)

params<-create_parm_file(file, min_gen, max_gen)

my_plot <- ggplot(params) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=as.numeric(get(x_line)), y=as.numeric(get(y_line)), color = factor(snake), shape=factor(newt)),size=3)+
#  geom_point(aes(x=as.numeric(Newt_pop_size_20), y=as.numeric(Snake_pop_size_20), color = factor(snake), shape=factor(newt)))+
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  #facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text( hjust=1, size = 8))

return(my_plot)
}

build_heatmap <- function(file, min_gen, max_gen, fill_line, fill_line_text, title, subt, y_label, x_label, red_blue=F){
  file$snake_newt_rep <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$rep)
  params<-create_parm_file(file, min_gen, max_gen)

 my_plot <- ggplot(params, aes(newt, paste0(snake, "_", rep))) +    # Create heatmap with ggplot2
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_tile(aes(fill = get(fill_line)))+theme(axis.text.x = element_text(angle = 45, hjust=1, size = 10), axis.text.y = element_text(size = 10))

 if(red_blue==FALSE){
  my_plot<- my_plot + scale_fill_viridis_c(name = fill_line_text) + theme(legend.text=element_text(angle = 45, hjust=1, size=10))
  }
 if(red_blue==TRUE){
  my_plot<- my_plot + scale_fill_gradient2(low = "red",mid = "beige", high = "blue", name = fill_line_text) + theme(legend.text=element_text(size=5))
 }

  return(my_plot)
}

trends_by_gen <- function(file, title, subt, x_label, y_label, min_gen, max_gen){

  plot <- ggplot(file[ which(file$generation>=min_gen & file$generation<=max_gen),]) +
    facet_wrap(~sim)+
    geom_line(aes(generation, dif, group=rep, col="dif"))+
    ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
    geom_line(aes(generation, Snake_mean_Pheno, group=rep, color="Snake_mean_Pheno"),)+
    geom_line(aes(generation, Newt_mean_Pheno, group=rep, color="Newt_mean_Pheno"))+
    geom_hline(yintercept=0, linetype="dashed", color="gray54")+
    ylim(-1.,2.5)+
    xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake_mean_Pheno" = "steelblue", "Newt_mean_Pheno" = "darkred", "dif" = "black"),
labels = c("Snake", "Newt", "Dif"))+
    theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())
  return(plot)
}

whale_gen <- function(file, title, subt, x_label, y_label,min_gen, max_gen, rep){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)
dif <- ggplot(file[ which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep),]) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen, rep ",rep)) +
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=as.numeric(Newt_pop_size), y=as.numeric(Snake_pop_size), color = dif))+
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      # strip.background = element_blank(), 
                      # strip.text.x = element_blank(), 
                      # panel.grid.major = element_blank(), 
                      # panel.grid.minor = element_blank(),
                      # panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

pop_size_hist <- function(file, title, subt, x_label, y_label,min_gen, max_gen, rep){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  file<-file[ which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep),]
  themin <- min(file$Snake_pop_size-file$Newt_pop_size) 
  themax <- max(file$Snake_pop_size-file$Newt_pop_size)
  true_max <- max(abs(themin), abs(themax))
dif <- ggplot(file) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(Snake_pop_size-Newt_pop_size), fill="darkolivegreen4")+
  xlim((-1*themax)-5, themax+5)+
  theme(plot.title = element_text(size=5)) +
  facet_wrap(~sim)+
  geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      # strip.background = element_blank(), 
                      # strip.text.x = element_blank(), 
                      # panel.grid.major = element_blank(), 
                      # panel.grid.minor = element_blank(),
                      # panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)
}

pheno_dif_hist <- function(file, title, subt, x_label, y_label,min_gen, max_gen, rep){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  file<-file[ which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep),]

  themin <- min(file$dif) 
  themax <- max(file$dif)
  true_max <- max(abs(themin), abs(themax))
dif <- ggplot(file) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen, rep ",rep)) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(dif), fill="aquamarine4")+
  xlim((-1*true_max)-1, true_max+1)+

  # y=as.numeric(Snake_pop_size-Newt_pop_size), color = snake_GA, shape = factor(newt_GA)
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      # strip.background = element_blank(), 
                      # strip.text.x = element_blank(), 
                      # panel.grid.major = element_blank(), 
                      # panel.grid.minor = element_blank(),
                      # panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

two_hist <- function(file, title, subt, x_label, y_label,min_gen, max_gen, rep){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)

  file<-file[ which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep),]

  thepopmin <- min(file$Snake_pop_size-file$Newt_pop_size)/100 
  thepopmax <- max(file$Snake_pop_size-file$Newt_pop_size)/100
  thedifmin <- min(file$dif) 
  thedifmax <- max(file$dif)
  true_max <- max(abs(thepopmin), abs(thepopmax),abs(thedifmin), abs(thedifmax))

  data_points <- nrow(file[which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep &  file$snake_GA==file$snake_GA[1] &  file$newt_GA==file$newt_GA[1]),])
dif <- ggplot(file) +

  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen, rep ",rep)) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=dif, y=..count../data_points), fill="#8B008B",  alpha = 0.65)+
  geom_histogram(aes(x=(Snake_pop_size-Newt_pop_size)/100, y=..count../data_points), fill="#008B00",  alpha = 0.65)+
  scale_y_continuous(labels = scales::percent)+
  xlim((-1*true_max)-1, true_max+1)+
  # y=as.numeric(Snake_pop_size-Newt_pop_size), color = snake_GA, shape = factor(newt_GA)
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      # strip.background = element_blank(), 
                      # strip.text.x = element_blank(), 
                      # panel.grid.major = element_blank(), 
                      # panel.grid.minor = element_blank(),
                      # panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

two_hist_all_rep <- function(file, title, subt, x_label, y_label,min_gen, max_gen){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)

  file<-file[ which(file$generation>=min_gen & file$generation<=max_gen),]

  thepopmin <- min(file$Snake_pop_size-file$Newt_pop_size)/100 
  thepopmax <- max(file$Snake_pop_size-file$Newt_pop_size)/100
  thedifmin <- min(file$dif) 
  thedifmax <- max(file$dif)
  true_max <- max(abs(thepopmin), abs(thepopmax),abs(thedifmin), abs(thedifmax))

  data_points <- nrow(file[which(file$generation>=min_gen & file$generation<=max_gen  &  file$snake_GA==file$snake_GA[1] &  file$newt_GA==file$newt_GA[1]),])

dif <- ggplot(file) +

  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=dif, y=..count../data_points), fill="#8B008B",  alpha = 0.65, bins = 20)+ #purple
  geom_histogram(aes(x=(Snake_pop_size-Newt_pop_size)/100, y=..count../data_points), fill="#008B00",  alpha = 0.65, bins = 20)+ #green
  scale_y_continuous(labels = scales::percent)+
  xlim((-1*true_max)-1, true_max+1)+
  # y=as.numeric(Snake_pop_size-Newt_pop_size), color = snake_GA, shape = factor(newt_GA)
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      #strip.background = element_blank(), 
                      strip.text.x = element_blank(), 
                      #panel.grid.major = element_blank(), 
                      #panel.grid.minor = element_blank(),
                      #panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

two_hist_death_all_rep <- function(file, title, subt, x_label, y_label,min_gen, max_gen){
  #file=GA3_tall_tall_file
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)

  file<-file[ which(file$generation>=min_gen & file$generation<=max_gen),]

  thepopmin <- min(file$newt_deaths)
  thepopmax <- max(file$newt_deaths)
  thedifmin <- min(file$snake_deaths) 
  thedifmax <- max(file$snake_deaths)
  true_max <- max(abs(thepopmin), abs(thepopmax),abs(thedifmin), abs(thedifmax))

  data_points <- nrow(file[which(file$generation>=min_gen & file$generation<=max_gen  &  file$snake_GA==file$snake_GA[1] &  file$newt_GA==file$newt_GA[1]),])

dif <- ggplot(file) +

  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=newt_deaths), fill="red",  alpha = 0.65, bins = 20)+
  geom_histogram(aes(x=snake_deaths), fill="blue",  alpha = 0.65, bins = 20)+
  #scale_y_continuous(labels = scales::percent)+
  #xlim((-1*true_max)-1, true_max+1)+
  # y=as.numeric(Snake_pop_size-Newt_pop_size), color = snake_GA, shape = factor(newt_GA)
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  #geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      #strip.background = element_blank(), 
                      strip.text.x = element_blank(), 
                      #panel.grid.major = element_blank(), 
                      #panel.grid.minor = element_blank(),
                      #panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

#two_hist_death_all_rep(GA3_tall_tall_file, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=20000, max_gen=25000)

```

## Mean Phenotype Whole Simulation {.tabset}

Information to be typed

### Phenotype differences
```{r, graphing by gen}
trends_by_gen(GA3_tall_tall_file, title="Space Tall: Newt VS Snake Population Size GA3", subt="Gen 16sims", y_label="Mean Phenotype", x_label="Generations", 100, 50000)

#aggregate(GA3_tall_tall_file[which(GA3_tall_tall_file$generation>=100 & GA3_tall_tall_file$generation<=50000),]$dif, list(GA3_tall_tall_file[which(GA3_tall_tall_file$generation>=100 & GA3_tall_tall_file$generation<=50000),]$group_rep), FUN=mean)

```

### Table of average Differences

```{r, average difference}
aggregate(GA3_tall_tall_file[which(GA3_tall_tall_file$generation>=100 & GA3_tall_tall_file$generation<=50000),]$dif, list(GA3_tall_tall_file[which(GA3_tall_tall_file$generation>=100 & GA3_tall_tall_file$generation<=50000),]$group), FUN=mean) 
```

## {-}

explanation

## Connection between higher phenotype and population {.tabset}

more info

### Phenotype differences
```{r, whale}
min_gen=5000
max_gen=10000
rep=0

whale_gen(GA3_tall_tall_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake-Newt Popsize", x_label="Snake-Newt Phenotype", min_gen=min_gen, max_gen=max_gen, rep=rep )
```

### Phenotype & Populationsize differences
```{r, popsize and pheno}

suppressMessages(print(two_hist(GA3_tall_tall_file, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Popsize/100 (green) & Snake-Newt Phenotype (purple)",  min_gen=min_gen, max_gen=max_gen, rep=rep )))

#suppressMessages(print(two_hist_all_rep(GA3_tall_tall_file, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Popsize/100 (green) & Snake-Newt Phenotype (purple)",  min_gen=min_gen, max_gen=max_gen)))

```

## {-}

explanation

## Correlation

Info

```{r, combo myplots}
#files
#GA3_tall_tall_file
#GA2_long_file 
#GA3_cor_tall_file
#GA2_cor_file

# Legend: Not needed currently
# 
# - "CA" = mean_newt_pheno_By_mean_snake_pheno (Pink)
# - "CB" = num_newts_By_num_snakes (Green)
# - "CC" = sum_newt_pheno_By_num_snake (Blue)
# - "CD" = sum_snake_pheno_By_num_newt (Purple)

ginmin = 10000
ginmax = 15000

boxplot_cors_by_gen(GA3_cor_tall_file, "Simulation Correlation Comparied to Real Correlation Trials", "GA3", "newt/snake GAs&rep", "Cor", ginmin, ginmax)

#four_types_boxplot_cors_by_gen(GA3_cor_tall_file, "Simulation four Correlation", "GA3", "newt/snake GAs&rep", "Cor", ginmin, ginmax)

```

Explanation

## Correlation Histograms {.tabset}

what did I do and why Time slices to see general trends

### Plot 1
```{r, GA3 cor by dif windows}
ginmin = 0
ginmax = 5000
suppressMessages(print(hist_cors_by_gen_col_rep(GA3_cor_tall_file, "Simulation Correlation Comparied to Real Correlation", "GA3", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 2

```{r, GA3 cor by dif windows 2}
ginmin = 5000
ginmax = 10000
suppressMessages(print(hist_cors_by_gen_col_rep(GA3_cor_tall_file, "Simulation Correlation Comparied to Real Correlation", "GA3", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 3
```{r, GA3 cor by dif windows 3}
ginmin = 10000
ginmax = 15000
suppressMessages(print(hist_cors_by_gen_col_rep(GA3_cor_tall_file, "Simulation Correlation Comparied to Real Correlation", "GA3", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 4
```{r, GA3 cor by dif windows 4}
ginmin = 15000
ginmax = 20000
suppressMessages(print(hist_cors_by_gen_col_rep(GA3_cor_tall_file, "Simulation Correlation Comparied to Real Correlation", "GA3", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 5
```{r, GA3 cor by dif windows 5}
ginmin = 20000
ginmax = 25000
suppressMessages(print(hist_cors_by_gen_col_rep(GA3_cor_tall_file, "Simulation Correlation Comparied to Real Correlation", "GA3", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 6
```{r, GA3 cor by dif windows 6}
ginmin = 25000
ginmax = 30000
suppressMessages(print(hist_cors_by_gen_col_rep(GA3_cor_tall_file, "Simulation Correlation Comparied to Real Correlation", "GA3", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 7
```{r, GA3 cor by dif windows 7}
ginmin = 30000
ginmax = 35000
suppressMessages(print(hist_cors_by_gen_col_rep(GA3_cor_tall_file, "Simulation Correlation Comparied to Real Correlation", "GA3", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 8
```{r, GA3 cor by dif windows 8}
ginmin = 35000
ginmax = 40000
suppressMessages(print(hist_cors_by_gen_col_rep(GA3_cor_tall_file, "Simulation Correlation Comparied to Real Correlation", "GA3", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 9
```{r, GA3 cor by dif windows 9}
ginmin = 40000
ginmax = 45000
suppressMessages(print(hist_cors_by_gen_col_rep(GA3_cor_tall_file, "Simulation Correlation Comparied to Real Correlation", "GA3", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 10
```{r, GA3 cor by dif windows 10}
ginmin = 45000
ginmax = 50000
suppressMessages(print(hist_cors_by_gen_col_rep(GA3_cor_tall_file, "Simulation Correlation Comparied to Real Correlation", "GA3", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

## {-}

What does it mean?

## Correlation across time {.tabset}

All of the correlations across time look very similar. Generally, mean newt and snake phenotypes (red and blue) go up overtime while the local correlation between the two fluctuates. Sometime one mean phenotype stays flat. The local correlations range from being very positive to being very negative. How can the correlation start to suddenly go up and then fall? It is hard to notice any differences between the GA combinations. For space I have chosen 3 random figures.

### Random 1
```{r, GA3 cor and pheno, results="hide"}
GA3_pattern <- unique(paste0(GA3_tall_tall_file$snake_mu_rate, "_", GA3_tall_tall_file$snake_mu_effect_sd,"_",GA3_tall_tall_file$newt_mu_rate, "_", GA3_tall_tall_file$newt_mu_effect_sd, "_", GA3_tall_tall_file$rep))

myrandomlist=sample.int(length(GA3_pattern), 3)

# for(i in myrandomlist){
#   myine<-cor_trends_by_gen(GA3_tall_tall_file, GA3_cor_tall_file, "test", "GA3", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA3_pattern[i])
#   plot(myine)
# }

cor_trends_by_gen(GA3_tall_tall_file, GA3_cor_tall_file, "test", "GA3", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA3_pattern[myrandomlist[1]])

```

### Random 2
```{r, random cor-pheno plot 2}

cor_trends_by_gen(GA3_tall_tall_file, GA3_cor_tall_file, "test", "GA3", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA3_pattern[myrandomlist[2]])
```

### Random 3
```{r, random cor-pheno plot 3}

cor_trends_by_gen(GA3_tall_tall_file, GA3_cor_tall_file, "test", "GA3", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA3_pattern[myrandomlist[3]])
```

## {-}

Explain general figures

## What happens over time (looking at the beginning. middle, and late part of my simulations) {.tabset}

### Pheno Beginning

```{r, early plots}

min_gen=100
max_gen=5000

# for(i in 1:length(unique(GA3_tall_tall_file$rep))){
#   myine<-whale_gen(GA3_tall_tall_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA3_tall_tall_file$rep)[i])
#   plot(myine)
# 
# }
whale_gen(GA3_tall_tall_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA3_tall_tall_file$rep)[1])
```

### Pheno Middle

```{r, middle plots}

min_gen=25000
max_gen=30000

# for(i in 1:length(unique(GA3_tall_tall_file$rep))){
#   myine<-whale_gen(GA3_tall_tall_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA3_tall_tall_file$rep)[i])
#   plot(myine)
# 
# }
whale_gen(GA3_tall_tall_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA3_tall_tall_file$rep)[1])
```

### Pheno End

```{r, end plots}

min_gen=45000
max_gen=50000

# for(i in 1:length(unique(GA3_tall_tall_file$rep))){
#   myine<-whale_gen(GA3_tall_tall_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA3_tall_tall_file$rep)[i])
#   plot(myine)
# 
# }
whale_gen(GA3_tall_tall_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA3_tall_tall_file$rep)[1])
```

### Dif Beginning 
```{r, early plots 1a}

min_gen=100
max_gen=5000

# for(i in 1:length(unique(GA3_tall_tall_file$rep))){
#   myine<-two_hist(GA3_tall_tall_file, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(GA3_tall_tall_file$rep)[i] )
#   suppressMessages(plot(myine))
# }

two_hist(GA3_tall_tall_file, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(GA3_tall_tall_file$rep)[1] )

```

### Dif Middle 
```{r, early plots 2a}

min_gen=25000
max_gen=30000

# for(i in 1:length(unique(GA3_tall_tall_file$rep))){
#   myine<-two_hist(GA3_tall_tall_file, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(GA3_tall_tall_file$rep)[i] )
#   suppressMessages(plot(myine))
# }

two_hist(GA3_tall_tall_file, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(GA3_tall_tall_file$rep)[1] )

```

### Dif End 
```{r, early plots 3a}

min_gen=45000
max_gen=50000

# for(i in 1:length(unique(GA3_tall_tall_file$rep))){
#   myine<-two_hist(GA3_tall_tall_file, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(GA3_tall_tall_file$rep)[i] )
#   suppressMessages(plot(myine))
# }

two_hist(GA3_tall_tall_file, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(GA3_tall_tall_file$rep)[1] )

```

## {-}

## Summary {.tabset}

Describe

### Early-Sim Population Size Summary

```{r, early plots summary}

min_gen=100
max_gen=5000
GA_summary(GA3_tall_tall_file, min_gen=min_gen, max_gen=max_gen, x_line="Newt_pop_size", y_line="Snake_pop_size", title="Space: Newt VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize")
```

### Mid-Sim Population Size Summary
```{r, plots summary 2}
# summary

min_gen=25000
max_gen=30000

# summary
GA_summary(GA3_tall_tall_file, min_gen=min_gen, max_gen=max_gen, x_line="Newt_pop_size", y_line="Snake_pop_size", title="Space: Newt VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize")
```

### Late-Sim Population Size Summary
```{r, plots summary 3}
min_gen=45000
max_gen=50000

# summary
GA_summary(GA3_tall_tall_file, min_gen=min_gen, max_gen=max_gen, x_line="Newt_pop_size", y_line="Snake_pop_size", title="Space: Newt VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize")

```

### Early Difference Summary
```{r, dif plots summary 1}

min_gen=100
max_gen=5000

GA_summary(GA3_tall_tall_file, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="Snake_pop_size", title="Space: Dif VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Dif")
```

### Mid Difference Summary
```{r, dif plots summary 2}
min_gen=25000
max_gen=30000

GA_summary(GA3_tall_tall_file, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="Snake_pop_size", title="Space: Dif VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Dif")

min_gen=45000
max_gen=50000
```

### Late Difference Summary
```{r, dif plots summary 3}
GA_summary(GA3_tall_tall_file, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="Snake_pop_size", title="Space: Dif VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Dif")

```

## {-}

Explain

## Heatmap {.tabset}

What should I be focusing on here?

### Population Size (Early)

```{r, early plots Heatmap}

min_gen=100
max_gen=5000
#heatmap
build_heatmap(GA3_tall_tall_file, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)
```

### Population Size (Mid)  
```{r, early plots Heatmap 2}
min_gen=25000
max_gen=30000
#heatmap
build_heatmap(GA3_tall_tall_file, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

```

### Population Size (Late) 
```{r, early plots Heatmap 3}
min_gen=45000
max_gen=50000
#heatmap
build_heatmap(GA3_tall_tall_file, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)
```

### Phenotype (Early) 
```{r, early plots Heatmap pheno 1}
min_gen=100
max_gen=5000
build_heatmap(GA3_tall_tall_file, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### Phenotype (Mid) 
```{r, early plots Heatmap pheno 2}

min_gen=25000
max_gen=30000
#heatmap

build_heatmap(GA3_tall_tall_file, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### Phenotype (Late) 
```{r, early plots Heatmap pheno 3}
min_gen=45000
max_gen=50000
#heatmap

build_heatmap(GA3_tall_tall_file, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)

```

## {-}

Explanations

## What is up with the correlations

Need to add tabs to show the mean, max, min, and population size. #3 high + cor, 0 cor, -cor

```{r, cor grid by location 1}

#GA3_grid_tall_file

file_name= "Snake_1e-11_0.158_Newt_1e-10_0.05_0"

#Snake_1.0e-08_0.005_Newt_1.0e-11_0.158_0

spegen=20001
#unique(GA3_grid_tall_file$snake_newt_rep)
GA3<-GA3_GA3_grid_tall_file[which(GA3_GA3_grid_tall_file$snake_newt_rep==file_name & GA3_GA3_grid_tall_file$generation==spegen),]

GA3_mean = GA3[,grepl("generation|mean", names(GA3)) ]

GA3_data_long_mean <- gather(GA3_mean, condition, measurement, newt_mean_pheno_0:snake_mean_pheno_99, factor_key=TRUE)

GA3_data_long_mean[,c("species","type", "type2","location")]<- str_split_fixed(GA3_data_long_mean$condition, "_", 4)

GA3_data_long_mean_short<-GA3_data_long_mean

GA3_data_long_mean_short$y_loc<- rep(c(rep.int(20, 2),rep.int(19, 2),rep.int(18, 2),rep.int(17, 2),rep.int(16, 2),rep.int(15, 2),rep.int(14, 2),rep.int(13, 2),rep.int(12, 2),rep.int(11, 2),rep.int(10, 2),rep.int(9, 2),rep.int(8, 2),rep.int(7, 2),rep.int(6, 2),rep.int(5, 2),rep.int(4, 2),rep.int(3, 2),rep.int(2, 2),rep.int(1, 2)),5)
GA3_data_long_mean_short$x_loc<-c(rep.int(1, 40),rep.int(2, 40),rep.int(3, 40),rep.int(4, 40),rep.int(5, 40))

main_plot<-ggplot(GA3_data_long_mean_short)+
  labs(title = "Mean Phenotype of Locations",
              subtitle = paste0(file_name, ", Generation: ", spegen),
              fill = "Toxicity/Resistance")+
  geom_point(data = GA3_data_long_mean_short[which(GA3_data_long_mean_short$species=="snake"),], aes(x = x_loc, y = y_loc, fill = as.numeric(measurement)), size = 6, shape = 22, ) +
  geom_point(data = GA3_data_long_mean_short[which(GA3_data_long_mean_short$species=="newt"),], aes(x = x_loc, y = y_loc, fill = as.numeric(measurement)), size = 4, shape = 21, ) +
  scale_fill_viridis_c()+
  theme_bw()+
  theme(legend.justification = "top")

snake<-GA3_data_long_mean_short[which(GA3_data_long_mean_short$species=="snake"),]
newt_GA3<-GA3_data_long_mean_short[which(GA3_data_long_mean_short$species=="newt"),]
newt_GA3$snake_value <- snake$measurement
cor(as.numeric(snake$measurement), as.numeric(newt_GA3$measurement), use="complete.obs")

newt_GA3$info[newt_GA3$location %in% list(1, 2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,20, 39, 40, 59, 60, 79, 81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98)]   <- "edge"
newt_GA3$info[newt_GA3$location %in% list(0, 19, 80, 99)] <- "corner"
newt_GA3$info[is.na(newt_GA3$info)]<- "middle"

extra_plot <- ggplot(newt_GA3)+
  ggtitle("Mean Phenotype") +
  xlab("Newt") + ylab("Snake")+
  geom_point(aes(x = as.numeric(measurement), y = as.numeric(snake_value), color=info), size = 1 ) +
  theme_bw()+
   theme(text = element_text(size=10),title =element_text(size=8),
         legend.text = element_text(size = 8),
         legend.position ="bottom",
         legend.key.size = unit(0.1, "lines"),
         legend.title=element_blank())+
  guides(shape = guide_legend(override.aes = list(size = 1)),
         color = guide_legend(override.aes = list(size = 1)))

#try to arrage the grids

main_plot +  inset_element(extra_plot, left = 1., bottom = -0.12, right = 1.35, top = 0.5)

```

### file

```{r, cor grid by location 2}

#GA3_grid_tall_file

file_name= "Snake_1e-08_0.005_Newt_1e-11_0.158_0"
#Snake_1.0e-08_0.005_Newt_1.0e-11_0.158_0

spegen=30001
#unique(GA3_grid_tall_file$snake_newt_rep)
GA3<-GA3_GA3_grid_tall_file[which(GA3_GA3_grid_tall_file$snake_newt_rep==file_name & GA3_GA3_grid_tall_file$generation==spegen),]
GA3<-GA3[1,]
GA3_mean = GA3[,grepl("generation|mean", names(GA3)) ]

GA3_data_long_mean <- gather(GA3_mean, condition, measurement, newt_mean_pheno_0:snake_mean_pheno_99, factor_key=TRUE)

GA3_data_long_mean[,c("species","type", "type2","location")]<- str_split_fixed(GA3_data_long_mean$condition, "_", 4)

GA3_data_long_mean_short<-GA3_data_long_mean

GA3_data_long_mean_short$y_loc<- rep(c(rep.int(20, 2),rep.int(19, 2),rep.int(18, 2),rep.int(17, 2),rep.int(16, 2),rep.int(15, 2),rep.int(14, 2),rep.int(13, 2),rep.int(12, 2),rep.int(11, 2),rep.int(10, 2),rep.int(9, 2),rep.int(8, 2),rep.int(7, 2),rep.int(6, 2),rep.int(5, 2),rep.int(4, 2),rep.int(3, 2),rep.int(2, 2),rep.int(1, 2)),5)
GA3_data_long_mean_short$x_loc<-c(rep.int(1, 40),rep.int(2, 40),rep.int(3, 40),rep.int(4, 40),rep.int(5, 40))

main_plot<-ggplot(GA3_data_long_mean_short)+
  labs(title = "Mean Phenotype of Locations",
              subtitle = paste0(file_name, ", Generation: ", spegen),
              fill = "Toxicity/Resistance")+
  geom_point(data = GA3_data_long_mean_short[which(GA3_data_long_mean_short$species=="snake"),], aes(x = x_loc, y = y_loc, fill = as.numeric(measurement)), size = 6, shape = 22, ) +
  geom_point(data = GA3_data_long_mean_short[which(GA3_data_long_mean_short$species=="newt"),], aes(x = x_loc, y = y_loc, fill = as.numeric(measurement)), size = 4, shape = 21, ) +
  scale_fill_viridis_c()+
  theme_bw()+
  theme(legend.justification = "top")

snake<-GA3_data_long_mean_short[which(GA3_data_long_mean_short$species=="snake"),]
newt_GA3<-GA3_data_long_mean_short[which(GA3_data_long_mean_short$species=="newt"),]
newt_GA3$snake_value <- snake$measurement
cor(as.numeric(snake$measurement), as.numeric(newt_GA3$measurement), use="complete.obs")

newt_GA3$info[newt_GA3$location %in% list(1, 2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,20, 39, 40, 59, 60, 79, 81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98)]   <- "edge"
newt_GA3$info[newt_GA3$location %in% list(0, 19, 80, 99)] <- "corner"
newt_GA3$info[is.na(newt_GA3$info)]<- "middle"

extra_plot <- ggplot(newt_GA3)+
  ggtitle("Mean Phenotype") +
  xlab("Newt") + ylab("Snake")+
  geom_point(aes(x = as.numeric(measurement), y = as.numeric(snake_value), color=info), size = 1 ) +
  theme_bw()+
   theme(text = element_text(size=10),title =element_text(size=8),
         legend.text = element_text(size = 8),
         legend.position ="bottom",
         legend.key.size = unit(0.1, "lines"),
         legend.title=element_blank())+
  guides(shape = guide_legend(override.aes = list(size = 1)),
         color = guide_legend(override.aes = list(size = 1)))

#try to arrage the grids

main_plot +  inset_element(extra_plot, left = 1., bottom = -0.12, right = 1.35, top = 0.5)

```

```{r, cor grid by location 3}

#GA3_grid_tall_file

file_name= "Snake_1e-10_0.05_Newt_1e-11_0.158_3"
#Snake_1.0e-08_0.005_Newt_1.0e-11_0.158_0

spegen=14001
#unique(GA3_grid_tall_file$snake_newt_rep)
GA3<-GA3_GA3_grid_tall_file[which(GA3_GA3_grid_tall_file$snake_newt_rep==file_name & GA3_GA3_grid_tall_file$generation==spegen),]

GA3_mean = GA3[,grepl("generation|mean", names(GA3)) ]

GA3_data_long_mean <- gather(GA3_mean, condition, measurement, newt_mean_pheno_0:snake_mean_pheno_99, factor_key=TRUE)

GA3_data_long_mean[,c("species","type", "type2","location")]<- str_split_fixed(GA3_data_long_mean$condition, "_", 4)

GA3_data_long_mean_short<-GA3_data_long_mean

GA3_data_long_mean_short$y_loc<- rep(c(rep.int(20, 2),rep.int(19, 2),rep.int(18, 2),rep.int(17, 2),rep.int(16, 2),rep.int(15, 2),rep.int(14, 2),rep.int(13, 2),rep.int(12, 2),rep.int(11, 2),rep.int(10, 2),rep.int(9, 2),rep.int(8, 2),rep.int(7, 2),rep.int(6, 2),rep.int(5, 2),rep.int(4, 2),rep.int(3, 2),rep.int(2, 2),rep.int(1, 2)),5)
GA3_data_long_mean_short$x_loc<-c(rep.int(1, 40),rep.int(2, 40),rep.int(3, 40),rep.int(4, 40),rep.int(5, 40))

main_plot<-ggplot(GA3_data_long_mean_short)+
  labs(title = "Mean Phenotype of Locations",
              subtitle = paste0(file_name, ", Generation: ", spegen),
              fill = "Toxicity/Resistance")+
  geom_point(data = GA3_data_long_mean_short[which(GA3_data_long_mean_short$species=="snake"),], aes(x = x_loc, y = y_loc, fill = as.numeric(measurement)), size = 6, shape = 22, ) +
  geom_point(data = GA3_data_long_mean_short[which(GA3_data_long_mean_short$species=="newt"),], aes(x = x_loc, y = y_loc, fill = as.numeric(measurement)), size = 4, shape = 21, ) +
  scale_fill_viridis_c()+
  theme_bw()+
  theme(legend.justification = "top")

snake<-GA3_data_long_mean_short[which(GA3_data_long_mean_short$species=="snake"),]
newt_GA3<-GA3_data_long_mean_short[which(GA3_data_long_mean_short$species=="newt"),]
newt_GA3$snake_value <- snake$measurement
cor(as.numeric(snake$measurement), as.numeric(newt_GA3$measurement), use="complete.obs")

newt_GA3$info[newt_GA3$location %in% list(1, 2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,20, 39, 40, 59, 60, 79, 81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98)]   <- "edge"
newt_GA3$info[newt_GA3$location %in% list(0, 19, 80, 99)] <- "corner"
newt_GA3$info[is.na(newt_GA3$info)]<- "middle"

extra_plot <- ggplot(newt_GA3)+
  ggtitle("Mean Phenotype") +
  xlab("Newt") + ylab("Snake")+
  geom_point(aes(x = as.numeric(measurement), y = as.numeric(snake_value), color=info), size = 1 ) +
  theme_bw()+
   theme(text = element_text(size=10),title =element_text(size=8),
         legend.text = element_text(size = 8),
         legend.position ="bottom",
         legend.key.size = unit(0.1, "lines"),
         legend.title=element_blank())+
  guides(shape = guide_legend(override.aes = list(size = 1)),
         color = guide_legend(override.aes = list(size = 1)))

#try to arrage the grids

main_plot +  inset_element(extra_plot, left = 1., bottom = -0.12, right = 1.35, top = 0.5)

```

```{r, cor grid by location 4}

#GA3_grid_tall_file
file_name= "Snake_1e-08_0.005_Newt_1e-08_0.005_0"
#Snake_1.0e-08_0.005_Newt_1.0e-11_0.158_0

spegen=15001
#unique(GA3_grid_tall_file$snake_newt_rep)
GA3<-GA3_GA3_grid_tall_file[which(GA3_GA3_grid_tall_file$snake_newt_rep==file_name & GA3_GA3_grid_tall_file$generation==spegen),]
GA3<-GA3[1,]
GA3_mean = GA3[,grepl("generation|mean", names(GA3)) ]

GA3_data_long_mean <- gather(GA3_mean, condition, measurement, newt_mean_pheno_0:snake_mean_pheno_99, factor_key=TRUE)

GA3_data_long_mean[,c("species","type", "type2","location")]<- str_split_fixed(GA3_data_long_mean$condition, "_", 4)

GA3_data_long_mean_short<-GA3_data_long_mean

GA3_data_long_mean_short$y_loc<- rep(c(rep.int(20, 2),rep.int(19, 2),rep.int(18, 2),rep.int(17, 2),rep.int(16, 2),rep.int(15, 2),rep.int(14, 2),rep.int(13, 2),rep.int(12, 2),rep.int(11, 2),rep.int(10, 2),rep.int(9, 2),rep.int(8, 2),rep.int(7, 2),rep.int(6, 2),rep.int(5, 2),rep.int(4, 2),rep.int(3, 2),rep.int(2, 2),rep.int(1, 2)),5)
GA3_data_long_mean_short$x_loc<-c(rep.int(1, 40),rep.int(2, 40),rep.int(3, 40),rep.int(4, 40),rep.int(5, 40))

main_plot<-ggplot(GA3_data_long_mean_short)+
  labs(title = "Mean Phenotype of Locations",
              subtitle = paste0(file_name, ", Generation: ", spegen),
              fill = "Toxicity/Resistance")+
  geom_point(data = GA3_data_long_mean_short[which(GA3_data_long_mean_short$species=="snake"),], aes(x = x_loc, y = y_loc, fill = as.numeric(measurement)), size = 6, shape = 22, ) +
  geom_point(data = GA3_data_long_mean_short[which(GA3_data_long_mean_short$species=="newt"),], aes(x = x_loc, y = y_loc, fill = as.numeric(measurement)), size = 4, shape = 21, ) +
  scale_fill_viridis_c()+
  theme_bw()+
  theme(legend.justification = "top")

snake<-GA3_data_long_mean_short[which(GA3_data_long_mean_short$species=="snake"),]
newt_GA3<-GA3_data_long_mean_short[which(GA3_data_long_mean_short$species=="newt"),]
newt_GA3$snake_value <- snake$measurement
cor(as.numeric(snake$measurement), as.numeric(newt_GA3$measurement), use="complete.obs")

newt_GA3$info[newt_GA3$location %in% list(1, 2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,20, 39, 40, 59, 60, 79, 81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98)]   <- "edge"
newt_GA3$info[newt_GA3$location %in% list(0, 19, 80, 99)] <- "corner"
newt_GA3$info[is.na(newt_GA3$info)]<- "middle"

extra_plot <- ggplot(newt_GA3)+
  ggtitle("Mean Phenotype") +
  xlab("Newt") + ylab("Snake")+
  geom_point(aes(x = as.numeric(measurement), y = as.numeric(snake_value), color=info), size = 1 ) +
  theme_bw()+
   theme(text = element_text(size=10),title =element_text(size=8),
         legend.text = element_text(size = 8),
         legend.position ="bottom",
         legend.key.size = unit(0.1, "lines"),
         legend.title=element_blank())+
  guides(shape = guide_legend(override.aes = list(size = 1)),
         color = guide_legend(override.aes = list(size = 1)))

#try to arrage the grids

main_plot +  inset_element(extra_plot, left = 1., bottom = -0.12, right = 1.35, top = 0.5)

#spegen=30001
#unique(GA3_grid_tall_file$snake_newt_rep)
GA3<-GA3_GA3_grid_tall_file[which(GA3_GA3_grid_tall_file$snake_newt_rep==file_name & GA3_GA3_grid_tall_file$generation==spegen),]
GA3<-GA3[1,]
GA3_ind = GA3[,grepl("generation|num_ind", names(GA3)) ]

GA3_data_long_ind <- gather(GA3_ind, condition, measurement, newt_num_ind_0:snake_num_ind_99, factor_key=TRUE)

GA3_data_long_ind[,c("species","type", "type2","location")]<- str_split_fixed(GA3_data_long_ind$condition, "_", 4)

GA3_data_long_ind_short<-GA3_data_long_ind

GA3_data_long_ind_short$y_loc<- rep(c(rep.int(20, 2),rep.int(19, 2),rep.int(18, 2),rep.int(17, 2),rep.int(16, 2),rep.int(15, 2),rep.int(14, 2),rep.int(13, 2),rep.int(12, 2),rep.int(11, 2),rep.int(10, 2),rep.int(9, 2),rep.int(8, 2),rep.int(7, 2),rep.int(6, 2),rep.int(5, 2),rep.int(4, 2),rep.int(3, 2),rep.int(2, 2),rep.int(1, 2)),5)
GA3_data_long_ind_short$x_loc<-c(rep.int(1, 40),rep.int(2, 40),rep.int(3, 40),rep.int(4, 40),rep.int(5, 40))

GA3_data_long_ind_short$measurement[GA3_data_long_ind_short$measurement == 0]  <- NA

main_plot<-ggplot(GA3_data_long_ind_short)+
  labs(title = "Locations of Ind",
              subtitle = paste0(file_name, ", Generation: ", spegen),
              fill = "Number of Individuals")+
  geom_point(data = GA3_data_long_ind_short[which(GA3_data_long_ind_short$species=="snake"),], aes(x = x_loc, y = y_loc, fill = as.numeric(measurement)), size = 6, shape = 22, ) +
  geom_point(data = GA3_data_long_ind_short[which(GA3_data_long_ind_short$species=="newt"),], aes(x = x_loc, y = y_loc, fill = as.numeric(measurement)), size = 4, shape = 21, ) +
  scale_fill_viridis_c(option = "plasma")+
  theme_bw()+
  theme(legend.justification = "top")

snake<-GA3_data_long_ind_short[which(GA3_data_long_ind_short$species=="snake"),]
newt_GA3<-GA3_data_long_ind_short[which(GA3_data_long_ind_short$species=="newt"),]
newt_GA3$snake_value <- snake$measurement
cor(as.numeric(snake$measurement), as.numeric(newt_GA3$measurement), use="complete.obs")

newt_GA3$info[newt_GA3$location %in% list(1, 2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,20, 39, 40, 59, 60, 79, 81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98)]   <- "edge"
newt_GA3$info[newt_GA3$location %in% list(0, 19, 80, 99)] <- "corner"
newt_GA3$info[is.na(newt_GA3$info)]<- "middle"

extra_plot <- ggplot(newt_GA3)+
  ggtitle("Num of Ind") +
  xlab("Newt") + ylab("Snake")+
  geom_point(aes(x = as.numeric(measurement), y = as.numeric(snake_value), color=info), size = 1 ) +
  theme_bw()+
   theme(text = element_text(size=10),title =element_text(size=8),
         legend.text = element_text(size = 8),
         legend.position ="bottom",
         legend.key.size = unit(0.1, "lines"),
         legend.title=element_blank())+
  guides(shape = guide_legend(override.aes = list(size = 1)),
         color = guide_legend(override.aes = list(size = 1)))

#try to arrage the grids

main_plot +  inset_element(extra_plot, left = 1., bottom = -0.12, right = 1.35, top = 0.5)

```

```{r, cor grid by location 5}

#GA3_grid_tall_file

file_name= "Snake_1e-10_0.05_Newt_1e-09_0.0158_0"

#Snake_1.0e-08_0.005_Newt_1.0e-11_0.158_0

#Snake_1.0e-08_0.005_Newt_1.0e-11_0.158_0

spegen=49991
#spegen=30001
#unique(GA3_grid_tall_file$snake_newt_rep)
GA3<-GA3_GA3_grid_tall_file[which(GA3_GA3_grid_tall_file$snake_newt_rep==file_name & GA3_GA3_grid_tall_file$generation==spegen),]

GA3_ind = GA3[,grepl("generation|num_ind", names(GA3)) ]

GA3_data_long_ind <- gather(GA3_ind, condition, measurement, newt_num_ind_0:snake_num_ind_99, factor_key=TRUE)

GA3_data_long_ind[,c("species","type", "type2","location")]<- str_split_fixed(GA3_data_long_ind$condition, "_", 4)

GA3_data_long_ind_short<-GA3_data_long_ind

GA3_data_long_ind_short$y_loc<- rep(c(rep.int(20, 2),rep.int(19, 2),rep.int(18, 2),rep.int(17, 2),rep.int(16, 2),rep.int(15, 2),rep.int(14, 2),rep.int(13, 2),rep.int(12, 2),rep.int(11, 2),rep.int(10, 2),rep.int(9, 2),rep.int(8, 2),rep.int(7, 2),rep.int(6, 2),rep.int(5, 2),rep.int(4, 2),rep.int(3, 2),rep.int(2, 2),rep.int(1, 2)),5)
GA3_data_long_ind_short$x_loc<-c(rep.int(1, 40),rep.int(2, 40),rep.int(3, 40),rep.int(4, 40),rep.int(5, 40))

GA3_data_long_ind_short$measurement[GA3_data_long_ind_short$measurement == 0]  <- NA

main_plot<-ggplot(GA3_data_long_ind_short)+
  labs(title = "Locations of Ind",
              subtitle = paste0(file_name, ", Generation: ", spegen),
              fill = "Number of Individuals")+
  geom_point(data = GA3_data_long_ind_short[which(GA3_data_long_ind_short$species=="snake"),], aes(x = x_loc, y = y_loc, fill = as.numeric(measurement)), size = 6, shape = 22, ) +
  geom_point(data = GA3_data_long_ind_short[which(GA3_data_long_ind_short$species=="newt"),], aes(x = x_loc, y = y_loc, fill = as.numeric(measurement)), size = 4, shape = 21, ) +
  scale_fill_viridis_c(option = "plasma")+
  theme_bw()+
  theme(legend.justification = "top")

snake<-GA3_data_long_ind_short[which(GA3_data_long_ind_short$species=="snake"),]
newt_GA3<-GA3_data_long_ind_short[which(GA3_data_long_ind_short$species=="newt"),]
newt_GA3$snake_value <- snake$measurement
cor(as.numeric(snake$measurement), as.numeric(newt_GA3$measurement), use="complete.obs")

newt_GA3$info[newt_GA3$location %in% list(1, 2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,20, 39, 40, 59, 60, 79, 81,82,83,84,85,86,87,88,89,90,91,92,93,94,95,96,97,98)]   <- "edge"
newt_GA3$info[newt_GA3$location %in% list(0, 19, 80, 99)] <- "corner"
newt_GA3$info[is.na(newt_GA3$info)]<- "middle"

extra_plot <- ggplot(newt_GA3)+
  ggtitle("Num of Ind") +
  xlab("Newt") + ylab("Snake")+
  geom_point(aes(x = as.numeric(measurement), y = as.numeric(snake_value), color=info), size = 1 ) +
  theme_bw()+
   theme(text = element_text(size=10),title =element_text(size=8),
         legend.text = element_text(size = 8),
         legend.position ="bottom",
         legend.key.size = unit(0.1, "lines"),
         legend.title=element_blank())+
  guides(shape = guide_legend(override.aes = list(size = 1)),
         color = guide_legend(override.aes = list(size = 1)))

#try to arrage the grids

main_plot +  inset_element(extra_plot, left = 1., bottom = -0.12, right = 1.35, top = 0.5)

```