### 2022-08-31 graphs_newt_snakes/corcompare.Rmd

---
title: "Cor Compare"
output: html_document
date: '2022-08-26'
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE,echo = FALSE)

library(ggplot2)
library(viridis)
library(colorspace)
library(gridExtra)
library(tidyverse)
require(reshape2)
library(plyr)
library(gridGraphics)
library(grid)
library(knitr)
library("reshape")

```

## Corilation Compare!

I have run four separate experiments and now I will compare their different correlations to see if I can tell the difference between each of the four experiments.

Experiments:

- tall map with 5 sections
- tall map with 7 sections
- tall map with 5 sections no heritability
- tall map with 5 sections no interaction

I am investigating the correlation calculations. I predicted that though co-evolution (interaction & evolution) species' phenotypes would become spatially correlated. I saw some very high correlations (around 0.5) in some of these simulations. To see if there is a difference between the correlations in my experiment I will create some comparative histograms.

Predictions:

-  5 sections & 7 sections will have *similar* correlation distributions
- no heritability & no interaction correlation distributions will look *very different* from 5 sections & 7 sections

```{r, loading in the data}

GA1_cor_tall_file<- read.csv("~/Desktop/GA1_tall_redo/all_cor.txt" , header = TRUE)
GA1_cor_tall_file$group_rep <- paste0(GA1_cor_tall_file$snake_mu_rate, "_", GA1_cor_tall_file$snake_mu_effect_sd,"_",GA1_cor_tall_file$newt_mu_rate, "_", GA1_cor_tall_file$newt_mu_effect_sd, "_", GA1_cor_tall_file$rep)
GA1_cor_tall_file$group <- paste0(GA1_cor_tall_file$snake_mu_rate, "_", GA1_cor_tall_file$snake_mu_effect_sd,"_",GA1_cor_tall_file$newt_mu_rate, "_", GA1_cor_tall_file$newt_mu_effect_sd)
GA1_cor_tall_file$sim=LETTERS[factor(GA1_cor_tall_file$group)]
GA1_cor_tall_file$experiment = "A"

GA1_cor_tall_7_file<- read.csv("~/Desktop/GA1_tall_redo/all_cor_7.txt" , header = TRUE)
GA1_cor_tall_7_file$group_rep <- paste0(GA1_cor_tall_7_file$snake_mu_rate, "_", GA1_cor_tall_7_file$snake_mu_effect_sd,"_",GA1_cor_tall_7_file$newt_mu_rate, "_", GA1_cor_tall_7_file$newt_mu_effect_sd, "_", GA1_cor_tall_7_file$rep)
GA1_cor_tall_7_file$group <- paste0(GA1_cor_tall_7_file$snake_mu_rate, "_", GA1_cor_tall_7_file$snake_mu_effect_sd,"_",GA1_cor_tall_7_file$newt_mu_rate, "_", GA1_cor_tall_7_file$newt_mu_effect_sd)
GA1_cor_tall_7_file$sim=LETTERS[factor(GA1_cor_tall_7_file$group)]
GA1_cor_tall_7_file$experiment = "B"

her_cor_tall_file<- read.csv(paste0(filelocation,"/all_cor.txt") , header = TRUE)
her_cor_tall_file$group_rep <- paste0(her_cor_tall_file$snake_mu_rate, "_", her_cor_tall_file$snake_mu_effect_sd,"_",her_cor_tall_file$newt_mu_rate, "_", her_cor_tall_file$newt_mu_effect_sd, "_", her_cor_tall_file$rep)
her_cor_tall_file$group <- paste0(her_cor_tall_file$snake_mu_rate, "_", her_cor_tall_file$snake_mu_effect_sd,"_",her_cor_tall_file$newt_mu_rate, "_", her_cor_tall_file$newt_mu_effect_sd)
her_cor_tall_file$sim=LETTERS[factor(her_cor_tall_file$group)]
her_cor_tall_file$experiment = "C"

inter_cor_tall_file<- read.csv(paste0(filelocation,"/all_cor.txt") , header = TRUE)
inter_cor_tall_file$group_rep <- paste0(inter_cor_tall_file$snake_mu_rate, "_", inter_cor_tall_file$snake_mu_effect_sd,"_",inter_cor_tall_file$newt_mu_rate, "_", inter_cor_tall_file$newt_mu_effect_sd, "_", inter_cor_tall_file$rep)
inter_cor_tall_file$group <- paste0(inter_cor_tall_file$snake_mu_rate, "_", inter_cor_tall_file$snake_mu_effect_sd,"_",inter_cor_tall_file$newt_mu_rate, "_", inter_cor_tall_file$newt_mu_effect_sd)
inter_cor_tall_file$sim=LETTERS[factor(inter_cor_tall_file$group)]
inter_cor_tall_file$experiment = "D"

all_cor_exp = rbind(GA1_cor_tall_file, GA1_cor_tall_7_file, her_cor_tall_file, inter_cor_tall_file)

```

```{r, making the fig}

hist_cors_by_gen_col_all_exp_rep <- function(file, title, subt, x_label, y_label, min_gen, max_gen){
 snewt$mean_res = snewt$mean_toxin+snewt$dif
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
   file$snake_newt <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  #unique(file$sim)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=as.numeric(mean_newt_pheno_By_mean_snake_pheno),fill=factor(experiment)), alpha=0.4, position="identity")+
  geom_vline(xintercept=real_snewt_cor, linetype="dashed")+
  geom_vline(xintercept=0, linetype="solid")+
  #geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  #geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  #geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  theme(plot.title = element_text(size=5)) +
  xlim(-1.1, 1.1)+
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8),
                    strip.background = element_blank(), strip.text.x = element_blank(),
                   title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                    axis.title.y = element_blank())

  return(plot)
}

dens_cors_by_gen_col_all_exp_rep <- function(file, title, subt, x_label, y_label, min_gen, max_gen){
 snewt$mean_res = snewt$mean_toxin+snewt$dif
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
   file$snake_newt <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  #unique(file$sim)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_density(aes(x=as.numeric(mean_newt_pheno_By_mean_snake_pheno),fill=factor(experiment)), alpha=0.4, position="identity")+
  geom_vline(xintercept=real_snewt_cor, linetype="dashed")+
  geom_vline(xintercept=0, linetype="solid")+
  #geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  #geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  #geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  theme(plot.title = element_text(size=5)) +
  xlim(-1.1, 1.1)+
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8),
                    strip.background = element_blank(), strip.text.x = element_blank(),
                   title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                    axis.title.y = element_blank())

  return(plot)
}
```

```{r, cor com figs}

ginmin = 0
ginmax = 5000
suppressMessages(print(hist_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

ginmin = 5000
ginmax = 10000
suppressMessages(print(hist_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

ginmin = 10000
ginmax = 15000
suppressMessages(print(hist_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

ginmin = 15000
ginmax = 20000
suppressMessages(print(hist_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

ginmin = 20000
ginmax = 25000
suppressMessages(print(hist_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

ginmin = 25000
ginmax = 30000
suppressMessages(print(hist_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

ginmin = 30000
ginmax = 35000
suppressMessages(print(hist_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

ginmin = 35000
ginmax = 40000
suppressMessages(print(hist_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

ginmin = 40000
ginmax = 45000
suppressMessages(print(hist_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

ginmin = 45000
ginmax = 50000
suppressMessages(print(hist_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

```

```{r, density com figs}

ginmin = 0
ginmax = 5000
suppressMessages(print(dens_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

ginmin = 5000
ginmax = 10000
suppressMessages(print(dens_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

ginmin = 10000
ginmax = 15000
suppressMessages(print(dens_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

ginmin = 15000
ginmax = 20000
suppressMessages(print(dens_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

ginmin = 20000
ginmax = 25000
suppressMessages(print(dens_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

ginmin = 25000
ginmax = 30000
suppressMessages(print(dens_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

ginmin = 30000
ginmax = 35000
suppressMessages(print(dens_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

ginmin = 35000
ginmax = 40000
suppressMessages(print(dens_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

ginmin = 40000
ginmax = 45000
suppressMessages(print(dens_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

ginmin = 45000
ginmax = 50000
suppressMessages(print(dens_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

```

```{r, one big one}

ginmin = 0
ginmax = 50000
suppressMessages(print(hist_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

suppressMessages(print(dens_cors_by_gen_col_all_exp_rep(all_cor_exp, "Simulation Correlation Comparied to Real Correlation", "inter", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))

```

# next step is to count the high and the lows 
like make a lower limit like must be above 0.3