### 2022-10-20 graphs_newt_snakes/interaction_rate_inv.Rmd
Looked at more data and added maps to my slim script! Need to fix corcompare to look nicer and spend a lot of time writing about all of these new results.

---
title: "Different Interaction Rates"
output: html_document
date: '2022-10-17'
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE,echo = FALSE)

library(ggplot2)
library(viridis)
library(colorspace)
library(gridExtra)
library(tidyverse)
require(reshape2)
library(plyr)
library(gridGraphics)
library(grid)
library(knitr)
library(kableExtra)
library("reshape")
library(ggplot2)
library(ggpubr)
theme_set(theme_pubr())
library(stringr)
library(patchwork)
```

## Goal:

Think, explore, & write about what the co-evolutionary interaction between newts & snakes with different genetic architectures (GAs, combination of mutation rate & mutation effect size) can lead to. This markdown is investigating

### Questions:

## Background

## Experiment

I created a simulation study to observe the co-evolutionary outcome of the newt-snake interaction with different genetic architectures (GAs) in a spatial setting. I hypothesized that we would see an interaction (co-evolutionary arms race) between newt and snake phenotype under some GA combinations when newts and snakes were evolving over geographical space. Each GA is paired with another GA creating 16 combinations.

GA1 experiment values:

- 1e-8*(0.005)^2 = 2.5e-13
- 1e-9*(0.05)^2 = 2.5e-12
- 1e-10*(0.5)^2 = 2.5e-11
- 1e-11*(5.0)^2 = 2.5e-10

Landscape: A tall map!: 35*4 H, 35 W

I tested different interaction rates:

  - 0.01
  - 0.025
  - 0.05
  - 0.075
  - 0.10
  - 0.125
  - 0.15
  - 0.20
  - 0.25
  - 0.30
  - 0.35

Each GA combination, trial, and interaction rate has its own msprime simulation.

Sometimes a higher interaction rate will lead to a species extinction, so ending values are approximate

### The data

## Preping the data by making singular files

```{bash, data file gathering}
#!/usr/bin/env bash

# Run these commands while in The the folder that contains the file you would like to condence
# Goal is to condense all the text files into one file 
# Requirements: need to have the same name and need to add additional columns
thepath=/Users/victoria/Desktop/
foldername=interaction_rate
thefolder=${thepath}${foldername}/*

add_in_column=",rep,snake_mu_rate,snake_mu_effect_sd,newt_mu_rate,newt_mu_effect_sd,interrate"

lit_head_file=$(find ${thepath}${foldername} -name "*_lit_*" | head -n 1)
lit_head=$(head -n 1 "${lit_head_file}")

cor_head_file=$(find ${thepath}${foldername} -name "*_cor_*" | head -n 1)
cor_head=$(head -n 1 "${cor_head_file}")

grid_head_file=$(find ${thepath}${foldername} -name "*_grid_*" | head -n 1)
grid_head=$(head -n 1 "${grid_head_file}")

if [[ -f ${thepath}${foldername}/all_lit.txt ]] && [[ -f ${thepath}${foldername}/all_grid.txt ]] && [[ -f ${thepath}${foldername}/all_cor.txt ]]
then
  echo "All cor, lit, and grid files exist!"
  echo "This program will now end!"
  exit 0
fi

#creating the files
echo ${lit_head}${add_in_column} > ${thepath}${foldername}/all_lit.txt
echo ${interrate_head}${add_in_column} > ${thepath}${foldername}/all_grid.txt
echo ${cor_head}${add_in_column} > ${thepath}${foldername}/all_cor.txt

for filename in $thefolder*.text;
  do 
#getting the variables

my_type=$(echo ${filename} |sed 's/.*GA205_//; s/_.*//')
snake_mu_rate=$(echo ${filename} |sed 's/.*snake_mu_rate_//; s/_.*//')
snake_mu_effect_sd=$(echo ${filename} |sed 's/.*snake_mu_effect_sd_//; s/_.*//')
newt_mu_rate=$(echo ${filename} |sed 's/.*newt_mu_rate_//; s/_.*//')
newt_mu_effect_sd=$(echo ${filename} |sed 's/.*newt_mu_effect_sd_//; s/_.*//')
interrate=$(echo ${filename} |sed 's/.*_interaction_rate_//; s/_.*//')
rep=$(echo ${filename} | sed "s/.*rep_//; s/_.*//")
#echo ${newt_mu_rate}	
# echo $my_type
# echo $my_other_type
# these new columns adding them to the data
sed -i '' -e 's/$/,'${rep}'/' $filename
sed -i '' -e 's/$/,'${snake_mu_rate}'/' $filename
sed -i '' -e 's/$/,'${snake_mu_effect_sd}'/' $filename
sed -i '' -e 's/$/,'${newt_mu_rate}'/' $filename
sed -i '' -e 's/$/,'${newt_mu_effect_sd}'/' $filename
sed -i '' -e 's/$/,'${interrate}'/' $filename
#infile >outfile
if [[ $my_type == "lit" ]]
then
    tail -n +2 -q $filename >> ${thepath}${foldername}/all_lit.txt
fi

if [[ $my_type == "grid" ]]
then
    tail -n +2 -q $filename >> ${thepath}${foldername}/all_grid.txt
fi

if [[ $my_type == "cor" ]]
then
    tail -n +2 -q $filename >> ${thepath}${foldername}/all_cor.txt
fi

done

```

```{r, read in the files}
snewt <- read.csv("~/Desktop/snake_newt_real_data.csv", header = TRUE)

GA1_lit_interrate_file<- read.csv("~/Desktop/interaction_rate/all_lit.txt" , header = TRUE)
GA1_lit_interrate_file$group_rep <- paste0(GA1_lit_interrate_file$snake_mu_rate, "_", GA1_lit_interrate_file$snake_mu_effect_sd,"_",GA1_lit_interrate_file$newt_mu_rate, "_", GA1_lit_interrate_file$newt_mu_effect_sd, "_", GA1_lit_interrate_file$rep)
GA1_lit_interrate_file$group <- paste0(GA1_lit_interrate_file$snake_mu_rate, "_", GA1_lit_interrate_file$snake_mu_effect_sd,"_",GA1_lit_interrate_file$newt_mu_rate, "_", GA1_lit_interrate_file$newt_mu_effect_sd)
GA1_lit_interrate_file$sim=LETTERS[factor(GA1_lit_interrate_file$group)]
GA1_lit_interrate_file$dif=as.numeric(GA1_lit_interrate_file$Snake_mean_Pheno)-as.numeric(GA1_lit_interrate_file$Newt_mean_Pheno)

GA1_lit_interrate_file$Snake_mean_Pheno<-as.numeric(GA1_lit_interrate_file$Snake_mean_Pheno)
GA1_lit_interrate_file$generation<-as.numeric(GA1_lit_interrate_file$generation)
GA1_lit_interrate_file$Newt_mean_Pheno<-as.numeric(GA1_lit_interrate_file$Newt_mean_Pheno)

GA1_cor_interrate_file<- read.csv("~/Desktop/interaction_rate/all_cor.txt" , header = TRUE)
GA1_cor_interrate_file$group_rep <- paste0(GA1_cor_interrate_file$snake_mu_rate, "_", GA1_cor_interrate_file$snake_mu_effect_sd,"_",GA1_cor_interrate_file$newt_mu_rate, "_", GA1_cor_interrate_file$newt_mu_effect_sd, "_", GA1_cor_interrate_file$rep)
GA1_cor_interrate_file$group <- paste0(GA1_cor_interrate_file$snake_mu_rate, "_", GA1_cor_interrate_file$snake_mu_effect_sd,"_",GA1_cor_interrate_file$newt_mu_rate, "_", GA1_cor_interrate_file$newt_mu_effect_sd)
GA1_cor_interrate_file$sim=LETTERS[factor(GA1_cor_interrate_file$group)]

unique(GA1_lit_interrate_file$rep)

GA1_lit_interrate_file <- GA1_lit_interrate_file[which(GA1_lit_interrate_file$rep==0),]
GA1_cor_interrate_file <- GA1_cor_interrate_file[which(GA1_cor_interrate_file$rep==0),]

```

```{r, cor box set up}

boxplot_cors_by_gen <- function(file, title, subt, x_label, y_label, min_gen, max_gen){
 snewt$mean_res = snewt$mean_toxin+snewt$dif
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
  file$snake_newt <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  #unique(file$sim)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ylim(-1.1, 1.1)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor(interrate), y=as.numeric(mean_newt_pheno_By_mean_snake_pheno), color = factor(interrate)))+
  geom_hline(yintercept=real_snewt_cor, linetype="dashed")+
    geom_hline(yintercept=0, linetype="solid")+
  #geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  #geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  #geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  theme(plot.title = element_text(size=5)) +
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

  return(plot)
}

boxplot_cor_grid_by_gen <- function(file, title, subt, x_label, y_label, min_gen, max_gen){
 snewt$mean_res = snewt$mean_toxin+snewt$dif
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
  file$snake_newt <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  unique(file$sim)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ylim(-1.1, 1.1)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor(grid_num), y=as.numeric(mean_newt_pheno_By_mean_snake_pheno), color = factor(grid_num)))+
  geom_hline(yintercept=real_snewt_cor, linetype="dashed")+
    geom_hline(yintercept=0, linetype="solid")+
  #geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  #geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  #geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  theme(plot.title = element_text(size=5)) +
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

  return(plot)
}

hist_cors_by_gen_col_rep <- function(file, title, subt, x_label, y_label, min_gen, max_gen){
 snewt$mean_res = snewt$mean_toxin+snewt$dif
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
   file$snake_newt <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  #unique(file$sim)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=as.numeric(mean_newt_pheno_By_mean_snake_pheno),fill=factor(rep)))+
  geom_vline(xintercept=real_snewt_cor, linetype="dashed")+
  geom_vline(xintercept=0, linetype="solid")+
  #geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  #geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  #geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  theme(plot.title = element_text(size=5)) +
  xlim(-1.1, 1.1)+
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

  return(plot)
}

hist_cors_by_gen_col_rep_nofill <- function(file, title, subt, x_label, y_label, min_gen, max_gen){
 snewt$mean_res = snewt$mean_toxin+snewt$dif
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
  file$snake_newt <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  unique(file$sim)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=as.numeric(mean_newt_pheno_By_mean_snake_pheno)))+
  geom_vline(xintercept=real_snewt_cor, linetype="dashed")+
  geom_vline(xintercept=0, linetype="solid")+
  #geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  #geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  #geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  theme(plot.title = element_text(size=5)) +
  #stat_bin(bins=30)+
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

suppressMessages(print(plot))

  return(plot)
}

var_cor <- function(file, cfile, title, subt, x_label, y_label, min_gen, max_gen){
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
  real_snewt_var = var(snewt$dif)
  var_cor_table=data.frame()

  for(i in 1:length(unique(file$snake_newt_rep))){
    pattern<-unique(file$snake_newt_rep)[i]
    temp_file<-file[which(file$gen>=min_gen & file$gen<=max_gen & file$snake_newt_rep==pattern),]
    temp_cor_file<-cfile[which(cfile$gen>=min_gen & cfile$gen<=max_gen & cfile$snake_newt_rep==pattern),]

    var_cor_table_temp<- data.frame("pattern"=pattern, "var"= var(temp_file$dif),"cor" = cor(temp_file$newt_toxin, temp_file$snake_res), "mean_corr" = mean(temp_cor_file$mean_newt_pheno_By_mean_snake_pheno))
    var_cor_table <- rbind(var_cor_table, var_cor_table_temp)
  }

  var_cor_table_temp<- data.frame("pattern"="real", "var"= real_snewt_var, "cor" = real_snewt_cor, "mean_corr" = real_snewt_cor)
  var_cor_table <- rbind(var_cor_table, var_cor_table_temp)

  plot <- ggplot(var_cor_table) +
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=as.numeric(var), y=as.numeric(cor), color = pattern), shape=17, size=3)+
  geom_point(aes(x=as.numeric(var), y=as.numeric(mean_corr), color = pattern), size=3)+
  theme(plot.title = element_text(size=5)) +
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))

  return(plot)
}

cor_trends_by_gen <- function(file, cfile, title, subt, x_label, x_label_cor, y_label, min_gen, max_gen, pattern){
#file=GA1_lit_interrate_file
#cfile=GA1_cor_interrate_file 
 # pattern="1e-11_5_1e-11_5_0"
 # pattern="1e-11_5_1e-11_5_0"  
 file$snake_newt_rep <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd,"_", file$rep)
  cfile$snake_newt_rep <- paste0(cfile$snake_mu_rate, "_", cfile$snake_mu_effect_sd, "_", cfile$newt_mu_rate, "_", cfile$newt_mu_effect_sd, "_", cfile$rep)

  plot_1 <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen & file$snake_newt_rep==pattern),]) +
    #geom_line(aes(generation, dif, group=rep, col="dif"))+
    ggtitle(title, subtitle = paste0(subt," pattern ", pattern, " From ", min_gen, " gen to ", max_gen, " gen"))+
    geom_line(aes(generation, Snake_mean_Pheno, group=interrate, color="Snake_mean_Pheno"),)+
    geom_line(aes(generation, Newt_mean_Pheno, group=interrate, color="Newt_mean_Pheno"))+
    #ylim(0,10)+
    #xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake_mean_Pheno" = "steelblue", "Newt_mean_Pheno" = "darkred"),
labels = c("Snake", "Newt"))+
    theme_bw() +theme(legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank(), axis.ticks.x = element_blank(), axis.text.x = element_blank(),
                      plot.margin=unit(c(1,1,+0.3,1), "cm"))
cor_plot <- ggplot(cfile[which(cfile$gen>=min_gen & cfile$gen<=max_gen & cfile$snake_newt_rep==pattern),]) +
    geom_line(aes(generation, mean_newt_pheno_By_mean_snake_pheno, col=factor(interrate)))+
    #ggtitle(title, subtitle = subt) +
    ylim(-1.1, 1.1)+
    #xlab(x_label_cor) + ylab(y_label)+
    theme_bw() +theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank(),
                      plot.margin=unit(c(-0.5,1,1,1), "cm"))

  figure <- ggarrange(plot_1, cor_plot + font("x.text", size = 10),heights = c(2, 1.5),ncol = 1, nrow = 2, align = "v")

  return(figure)
}

pop_size_trends_by_gen <- function(file, cfile, title, subt, x_label, x_label_cor, y_label, min_gen, max_gen, pattern){
#file=GA1_lit_interrate_file
#cfile=GA1_cor_interrate_file 
 # pattern="1e-11_5_1e-11_5_0"
 # pattern="1e-11_5_1e-11_5_0"  
 file$snake_newt_rep <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd,"_", file$rep)
  cfile$snake_newt_rep <- paste0(cfile$snake_mu_rate, "_", cfile$snake_mu_effect_sd, "_", cfile$newt_mu_rate, "_", cfile$newt_mu_effect_sd, "_", cfile$rep)

  plot_1 <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen & file$snake_newt_rep==pattern),]) +
    #geom_line(aes(generation, dif, group=rep, col="dif"))+
    ggtitle(title, subtitle = paste0(subt," pattern ", pattern, " From ", min_gen, " gen to ", max_gen, " gen"))+
    geom_line(aes(generation, Snake_pop_size, group=interrate, color="Snake"),)+
    geom_line(aes(generation, Newt_pop_size, group=interrate, color="Newt"))+
    #ylim(0,10)+
    #xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake" = "steelblue", "Newt" = "darkred"),
labels = c("Snake", "Newt"))+
    theme_bw() +theme(legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank(), axis.ticks.x = element_blank(), axis.text.x = element_blank(),
                      plot.margin=unit(c(1,1,+0.3,1), "cm"))
cor_plot <- ggplot(cfile[which(cfile$gen>=min_gen & cfile$gen<=max_gen & cfile$snake_newt_rep==pattern),]) +
    geom_line(aes(generation, num_newts_By_num_snakes, col=factor(interrate)))+
    #ggtitle(title, subtitle = subt) +
    ylim(-1.1, 1.1)+
    #xlab(x_label_cor) + ylab(y_label)+
    theme_bw() +theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank(),
                      plot.margin=unit(c(-0.5,1,1,1), "cm"))

  figure <- ggarrange(plot_1, cor_plot + font("x.text", size = 10),heights = c(2, 1.5),ncol = 1, nrow = 2, align = "v")

  return(figure)
}

```

```{r, general file and graph making}

#the functions to make the params file and the summary and heatmap figures

create_parm_file <- function(file, min_gen, max_gen, interrate){
params <- data_frame()

for(i in 1:length(unique(file$snake_newt_rep))){

  pattern=unique(file$snake_newt_rep)[i]
  sumitup <- file[which(file$gen>=min_gen & file$gen<=max_gen & file$snake_newt_rep ==pattern & file$interrate==interrate),]

  params_temp<- data_frame(Newt_mean_Pheno=mean(sumitup$Newt_mean_Pheno),
                           Snake_mean_Pheno=mean(sumitup$Snake_mean_Pheno), 
                           Newt_mean_Pheno_sd=sd(sumitup$Newt_mean_Pheno),
                           Snake_mean_Pheno_sd=sd(sumitup$Snake_mean_Pheno),
                           Newt_pop_size=mean(sumitup$Newt_pop_size),
                           Snake_pop_size=mean(sumitup$Snake_pop_size),
                           dif=mean(sumitup$dif),
                           snake_mu_rate=sumitup$snake_mu_rate[1],
                           newt_mu_rate=sumitup$newt_mu_rate[1],
                           snake_mu_effect_sd=sumitup$snake_mu_effect_sd[1],
                           newt_mu_effect_sd=sumitup$newt_mu_effect_sd[1],
                           rep=sumitup$rep[1])

  params <- rbind(params, params_temp)
}

params$snake <- paste0(params$snake_mu_rate, "_",params$snake_mu_effect_sd)
params$newt <- paste0(params$newt_mu_rate, "_",params$newt_mu_effect_sd)
params$snake_newt <- paste0(params$snake, "_", params$newt)
return(params)
}

 create_parm_file_nc <- function(file, min_gen, max_gen){

   file$test_factor <- paste0(file$interrate, "_", file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$rep)

params <- data_frame()

for(i in 1:length(unique(file$test_factor))){

  pattern=unique(file$test_factor)[i]

  sumitup <- file[which(file$gen>=min_gen & file$gen<=max_gen & file$test_factor ==pattern),]

  params_temp<- data_frame(Newt_mean_Pheno=mean(sumitup$Newt_mean_Pheno),
                           Snake_mean_Pheno=mean(sumitup$Snake_mean_Pheno), 
                           Newt_mean_Pheno_sd=sd(sumitup$Newt_mean_Pheno),
                           Snake_mean_Pheno_sd=sd(sumitup$Snake_mean_Pheno),
                           Newt_pop_size=mean(sumitup$Newt_pop_size),
                           Snake_pop_size=mean(sumitup$Snake_pop_size),
                           dif=mean(sumitup$dif),
                           snake_mu_rate=unlist(strsplit(pattern, "_"))[2],
                           newt_mu_rate=unlist(strsplit(pattern, "_"))[4],
                           snake_mu_effect_sd=unlist(strsplit(pattern, "_"))[3],
                           newt_mu_effect_sd=unlist(strsplit(pattern, "_"))[5],
                           rep=unlist(strsplit(pattern, "_"))[6],
                           interrate=unlist(strsplit(pattern, "_"))[1],
                           test_factor=pattern)

  params <- rbind(params, params_temp)
}

params$snake <- paste0(params$snake_mu_rate, "_",params$snake_mu_effect_sd)
params$newt <- paste0(params$newt_mu_rate, "_",params$newt_mu_effect_sd)
params$snake_newt <- paste0(params$snake, "_", params$newt)
return(params)
}

GA_summary <- function(file, min_gen, max_gen, x_line, y_line, title, subt, y_label, x_label, interrate){

file$snake_newt_rep <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$rep)

params<-create_parm_file(file, min_gen, max_gen, interrate)

my_plot <- ggplot(params) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=as.numeric(get(x_line)), y=as.numeric(get(y_line)), color = factor(snake), shape=factor(newt)),size=3)+
#  geom_point(aes(x=as.numeric(Newt_pop_size_20), y=as.numeric(Snake_pop_size_20), color = factor(snake), shape=factor(newt)))+
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  #facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text( hjust=1, size = 8))

return(my_plot)
}

build_heatmap <- function(file, min_gen, max_gen, fill_line, fill_line_text, title, subt, y_label, x_label, red_blue=F){

  params<-create_parm_file_nc(file, min_gen, max_gen)

 my_plot <- ggplot(params, aes(newt,  paste0(interrate, "_", snake, "_", rep))) +    # Create heatmap with ggplot2
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_tile(aes(fill = get(fill_line)))+theme(axis.text.x = element_text(angle = 45, hjust=1, size = 10), axis.text.y = element_text(size = 10))

 if(red_blue==FALSE){
  my_plot<- my_plot + scale_fill_viridis_c(name = fill_line_text) + theme(legend.text=element_text(angle = 45, hjust=1, size=10))
  }
 if(red_blue==TRUE){
  my_plot<- my_plot + scale_fill_gradient2(low = "red",mid = "beige", high = "blue", name = fill_line_text) + theme(legend.text=element_text(size=5))
 }

  return(my_plot)
}

trends_by_gen <- function(file, title, subt, x_label, y_label, min_gen, max_gen, rep){
 file$generation<-as.numeric(file$generation)
 file$Snake_mean_Pheno<-as.numeric(file$Snake_mean_Pheno)
 file$Newt_mean_Pheno<-as.numeric(file$Newt_mean_Pheno)
  plot <- ggplot(file[ which(file$generation>=min_gen & file$generation<=max_gen),]) +
    facet_wrap(~sim)+
    geom_line(aes(generation, as.numeric(dif), group=interrate, col="dif"))+
    #ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
    geom_line(aes(generation, Snake_mean_Pheno, group=interrate, color="Snake_mean_Pheno"),)+
    geom_line(aes(generation, Newt_mean_Pheno, group=interrate, color="Newt_mean_Pheno"))+
    geom_hline(yintercept=0, linetype="dashed", color="gray54")+
    #ylim(-1.,2.5)+
    xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake_mean_Pheno" = "steelblue", "Newt_mean_Pheno" = "darkred", "dif" = "black"),
labels = c("Snake", "Newt", "Dif"))+
    theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())
  return(plot)
}

whale_gen <- function(file, title, subt, x_label, y_label,min_gen, max_gen, rep){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)
dif <- ggplot(file[ which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep),]) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen, rep ",rep)) +
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=as.numeric(Newt_pop_size), y=as.numeric(Snake_pop_size), color = dif))+
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      # strip.background = element_blank(), 
                      # strip.text.x = element_blank(), 
                      # panel.grid.major = element_blank(), 
                      # panel.grid.minor = element_blank(),
                      # panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

whale_gen_by_interrate <- function(file, title, subt, x_label, y_label,min_gen, max_gen, rep, interrate){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)
dif <- ggplot(file[ which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep & file$interrate==interrate),]) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen, rep ",rep, " interrate ", interrate)) +
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=as.numeric(Newt_pop_size), y=as.numeric(Snake_pop_size), color = dif))+
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      # strip.background = element_blank(), 
                      # strip.text.x = element_blank(), 
                      # panel.grid.major = element_blank(), 
                      # panel.grid.minor = element_blank(),
                      # panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

pop_size_hist <- function(file, title, subt, x_label, y_label,min_gen, max_gen, rep){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  file<-file[ which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep),]
  themin <- min(file$Snake_pop_size-file$Newt_pop_size) 
  themax <- max(file$Snake_pop_size-file$Newt_pop_size)
  true_max <- max(abs(themin), abs(themax))
dif <- ggplot(file) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(Snake_pop_size-Newt_pop_size), fill="darkolivegreen4")+
  xlim((-1*themax)-5, themax+5)+
  theme(plot.title = element_text(size=5)) +
  facet_wrap(~sim)+
  geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      # strip.background = element_blank(), 
                      # strip.text.x = element_blank(), 
                      # panel.grid.major = element_blank(), 
                      # panel.grid.minor = element_blank(),
                      # panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)
}

pheno_dif_hist <- function(file, title, subt, x_label, y_label,min_gen, max_gen, rep){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  file<-file[ which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep),]

  themin <- min(file$dif) 
  themax <- max(file$dif)
  true_max <- max(abs(themin), abs(themax))
dif <- ggplot(file) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen, rep ",rep)) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(dif), fill="aquamarine4")+
  xlim((-1*true_max)-1, true_max+1)+

  # y=as.numeric(Snake_pop_size-Newt_pop_size), color = snake_GA, shape = factor(newt_GA)
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      # strip.background = element_blank(), 
                      # strip.text.x = element_blank(), 
                      # panel.grid.major = element_blank(), 
                      # panel.grid.minor = element_blank(),
                      # panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

two_hist <- function(file, title, subt, x_label, y_label,min_gen, max_gen, rep){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)

  file<-file[ which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep),]

  thepopmin <- min(file$Snake_pop_size-file$Newt_pop_size)/100 
  thepopmax <- max(file$Snake_pop_size-file$Newt_pop_size)/100
  thedifmin <- min(file$dif) 
  thedifmax <- max(file$dif)
  true_max <- max(abs(thepopmin), abs(thepopmax),abs(thedifmin), abs(thedifmax))

  data_points <- nrow(file[which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep &  file$snake_GA==file$snake_GA[1] &  file$newt_GA==file$newt_GA[1]),])
dif <- ggplot(file) +

  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen, rep ",rep)) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=dif, y=..count../data_points), fill="#8B008B",  alpha = 0.65)+
  geom_histogram(aes(x=(Snake_pop_size-Newt_pop_size)/100, y=..count../data_points), fill="#008B00",  alpha = 0.65)+
  scale_y_continuous(labels = scales::percent)+
  xlim((-1*true_max)-1, true_max+1)+
  # y=as.numeric(Snake_pop_size-Newt_pop_size), color = snake_GA, shape = factor(newt_GA)
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      # strip.background = element_blank(), 
                      # strip.text.x = element_blank(), 
                      # panel.grid.major = element_blank(), 
                      # panel.grid.minor = element_blank(),
                      # panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

two_hist_all_rep <- function(file, title, subt, x_label, y_label,min_gen, max_gen){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)

  file<-file[ which(file$generation>=min_gen & file$generation<=max_gen),]

  thepopmin <- min(file$Snake_pop_size-file$Newt_pop_size)/100 
  thepopmax <- max(file$Snake_pop_size-file$Newt_pop_size)/100
  thedifmin <- min(file$dif) 
  thedifmax <- max(file$dif)
  true_max <- max(abs(thepopmin), abs(thepopmax),abs(thedifmin), abs(thedifmax))

  data_points <- nrow(file[which(file$generation>=min_gen & file$generation<=max_gen  &  file$snake_GA==file$snake_GA[1] &  file$newt_GA==file$newt_GA[1]),])

dif <- ggplot(file) +

  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=dif, y=..count../data_points), fill="#8B008B",  alpha = 0.65, bins = 20)+ #purple
  geom_histogram(aes(x=(Snake_pop_size-Newt_pop_size)/100, y=..count../data_points), fill="#008B00",  alpha = 0.65, bins = 20)+ #green
  scale_y_continuous(labels = scales::percent)+
  xlim((-1*true_max)-1, true_max+1)+
  # y=as.numeric(Snake_pop_size-Newt_pop_size), color = snake_GA, shape = factor(newt_GA)
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      #strip.background = element_blank(), 
                      strip.text.x = element_blank(), 
                      #panel.grid.major = element_blank(), 
                      #panel.grid.minor = element_blank(),
                      #panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

two_hist_death_all_rep <- function(file, title, subt, x_label, y_label,min_gen, max_gen){
  #file=GA1_lit_interrate_file
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)

  file<-file[ which(file$generation>=min_gen & file$generation<=max_gen),]

  thepopmin <- min(file$newt_deaths)
  thepopmax <- max(file$newt_deaths)
  thedifmin <- min(file$snake_deaths) 
  thedifmax <- max(file$snake_deaths)
  true_max <- max(abs(thepopmin), abs(thepopmax),abs(thedifmin), abs(thedifmax))

  data_points <- nrow(file[which(file$generation>=min_gen & file$generation<=max_gen  &  file$snake_GA==file$snake_GA[1] &  file$newt_GA==file$newt_GA[1]),])

dif <- ggplot(file) +

  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=newt_deaths), fill="red",  alpha = 0.65, bins = 20)+
  geom_histogram(aes(x=snake_deaths), fill="blue",  alpha = 0.65, bins = 20)+
  #scale_y_continuous(labels = scales::percent)+
  #xlim((-1*true_max)-1, true_max+1)+
  # y=as.numeric(Snake_pop_size-Newt_pop_size), color = snake_GA, shape = factor(newt_GA)
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  #geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      #strip.background = element_blank(), 
                      strip.text.x = element_blank(), 
                      #panel.grid.major = element_blank(), 
                      #panel.grid.minor = element_blank(),
                      #panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

grid_plot<- function(file_name, spegen, file, type, row_num, col_num, interrate){

file$snake_newt_rep <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd,"_",file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$rep)

GA1<-file[which(file$snake_newt_rep==file_name & file$generation==spegen & file$interrate==interrate),]
GA1_type = paste0("generation|", type)

GA1_type = GA1[,grepl(GA1_type, names(GA1)) ]
GA1_type <- GA1_type[-1]

data_long_type_comp <- gather(GA1_type, condition, measurement, factor_key=TRUE)
data_long_type_comp[,c("species","type", "type2","location")]<- str_split_fixed(data_long_type_comp$condition, "_", 4)

data_long_type_comp$y_loc<- rep(sort(c(rep.int(row_num:1, 2)), decreasing = TRUE),col_num)

data_long_type_comp$x_loc<- sort(rep.int(1:col_num, row_num*2))

main_plot<-ggplot(data_long_type_comp)+

  geom_point(data = data_long_type_comp[which(data_long_type_comp$species=="snake"),], aes(x = x_loc, y = y_loc, fill = as.numeric(measurement)), size = 6, shape = 22, ) +
  geom_point(data = data_long_type_comp[which(data_long_type_comp$species=="newt"),], aes(x = x_loc, y = y_loc, fill = as.numeric(measurement)), size = 4, shape = 21, ) +
  scale_fill_viridis_c()+
  #ylim(0, row_num)+
  #xlim(0, col_num)+
  theme_bw()+
  theme(legend.justification = "top")

if(type != "num_ind"){
main_plot<- main_plot + labs(title = paste0(str_to_title(type)," Phenotype of Locations"),
              subtitle = paste0(file_name, ", Generation: ", spegen),
              fill = "Toxicity/Resistance")+
 scale_fill_viridis_c()
}
if(type == "num_ind"){
main_plot<- main_plot + labs(title = paste0(str_to_title(type)," Phenotype of Locations"),
              subtitle = paste0(file_name, ", Generation: ", spegen),
              fill = "Newt/Snake Phenotype")+
scale_fill_viridis_c(option = "plasma")
}

snake<-data_long_type_comp[which(data_long_type_comp$species=="snake"),]
newt<-data_long_type_comp[which(data_long_type_comp$species=="newt"),]
newt$snake_value <- snake$measurement
print(cor(as.numeric(snake$measurement), as.numeric(newt$measurement), use="complete.obs"))

newt$info <- "middle"
newt$info[newt$location %in% c(0, col_num-1, col_num*row_num-1-col_num, col_num*row_num-1)] <- "corner"

newt$info[newt$location %in% c(c(1:(col_num-2)), seq(col_num, (col_num*row_num-col_num-2), col_num), seq((col_num-1+col_num), (col_num*row_num-1-col_num-col_num), col_num),  c((col_num*row_num-col_num):(col_num*row_num-2)))]   <- "edge"

extra_plot <- ggplot(newt)+
  ggtitle("Mean Phenotype") +
  xlab("Newt") + ylab("Snake")+
  geom_point(aes(x = as.numeric(measurement), y = as.numeric(snake_value), color=info), size = 1 ) +
  theme_bw()+
   theme(text = element_text(size=10),title =element_text(size=8),
         legend.text = element_text(size = 8),
         legend.position ="bottom",
         legend.key.size = unit(0.1, "lines"),
         legend.title=element_blank())+
  guides(shape = guide_legend(override.aes = list(size = 1)),
         color = guide_legend(override.aes = list(size = 1)))

final_plot <- main_plot +  inset_element(extra_plot, left = 1., bottom = -0.10, right = 1.35, top = 0.5)
return(final_plot)}

```

generations

```{r, graphing by gen}

trends_by_gen(GA1_lit_interrate_file, title="Space Tall: Newt VS Snake Population Size GA1", subt="Gen 16sims", y_label="Mean Phenotype", x_label="Generations", 100, 50000, rep=0)

```

```{r, whale}
min_gen=5000
max_gen=10000
rep=0

whale_gen(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Phenotype", min_gen=min_gen, max_gen=max_gen, rep=rep )
```

```{r, combo myplots}

ginmin = 0
ginmax = 50000
#thecolors=c("#F8766D", "#B79F00", "#00BA38", "#00BFC4", "#619CFF", "#F564E3")

boxplot_cors_by_gen(GA1_cor_interrate_file, "interrate: Simulation Correlation By Grid Number Comparied to Real Correlation Trials", "GA1", "newt/snake GAs&rep", "Cor", ginmin, ginmax)

```

## Correlation Histograms  {.tabset}

In order to understand how spatial correlations where changing with time I took 5,000 generation time slices to look at all four trials correlation values. Each color is a different trial per GA combination. The histogram values are stacked. This section only looks at the 5 sections.

### Plot 1
```{r, GA1 cor by dif windows1}
ginmin = 0
ginmax = 5000
boxplot_cors_by_gen(GA1_cor_interrate_file, "interrate: Simulation Correlation By Grid Number Comparied to Real Correlation Trials", "GA1", "newt/snake GAs&rep", "Cor", ginmin, ginmax)
```

### Plot 2

```{r, GA1 cor by dif windows 2}
ginmin = 5000
ginmax = 10000
boxplot_cors_by_gen(GA1_cor_interrate_file, "interrate: Simulation Correlation By Grid Number Comparied to Real Correlation Trials", "GA1", "newt/snake GAs&rep", "Cor", ginmin, ginmax)
```

### Plot 3
```{r, GA1 cor by dif windows 3}
ginmin = 10000
ginmax = 15000
boxplot_cors_by_gen(GA1_cor_interrate_file, "interrate: Simulation Correlation By Grid Number Comparied to Real Correlation Trials", "GA1", "newt/snake GAs&rep", "Cor", ginmin, ginmax)
```

### Plot 4
```{r, GA1 cor by dif windows 4}
ginmin = 15000
ginmax = 20000
boxplot_cors_by_gen(GA1_cor_interrate_file, "interrate: Simulation Correlation By Grid Number Comparied to Real Correlation Trials", "GA1", "newt/snake GAs&rep", "Cor", ginmin, ginmax)
```

### Plot 5
```{r, GA1 cor by dif windows 5}
ginmin = 20000
ginmax = 25000
boxplot_cors_by_gen(GA1_cor_interrate_file, "interrate: Simulation Correlation By Grid Number Comparied to Real Correlation Trials", "GA1", "newt/snake GAs&rep", "Cor", ginmin, ginmax)
```

### Plot 6
```{r, GA1 cor by dif windows 6}
ginmin = 25000
ginmax = 30000
boxplot_cors_by_gen(GA1_cor_interrate_file, "interrate: Simulation Correlation By Grid Number Comparied to Real Correlation Trials", "GA1", "newt/snake GAs&rep", "Cor", ginmin, ginmax)
```

### Plot 7
```{r, GA1 cor by dif windows 7}
ginmin = 30000
ginmax = 35000
boxplot_cors_by_gen(GA1_cor_interrate_file, "interrate: Simulation Correlation By Grid Number Comparied to Real Correlation Trials", "GA1", "newt/snake GAs&rep", "Cor", ginmin, ginmax)
```

### Plot 8
```{r, GA1 cor by dif windows 8}
ginmin = 35000
ginmax = 40000
boxplot_cors_by_gen(GA1_cor_interrate_file, "interrate: Simulation Correlation By Grid Number Comparied to Real Correlation Trials", "GA1", "newt/snake GAs&rep", "Cor", ginmin, ginmax)
```

### Plot 9
```{r, GA1 cor by dif windows 9}
ginmin = 40000
ginmax = 45000
boxplot_cors_by_gen(GA1_cor_interrate_file, "interrate: Simulation Correlation By Grid Number Comparied to Real Correlation Trials", "GA1", "newt/snake GAs&rep", "Cor", ginmin, ginmax)
```

### Plot 10
```{r, GA1 cor by dif windows 10}
ginmin = 45000
ginmax = 50000
boxplot_cors_by_gen(GA1_cor_interrate_file, "interrate: Simulation Correlation By Grid Number Comparied to Real Correlation Trials", "GA1", "newt/snake GAs&rep", "Cor", ginmin, ginmax)
```

## {-}

## Phenotype Correlation across time {.tabset}

Next, we will examine three randomly chosen plots from both the 5 section and 7 section experiment. Time (in generations) in on the x-axis and both mean phenotype and phenotype spatial correlation in on the y-axis. Newt whole population mean phenotype is red, while snake mean phenotype is blue. The pink line is the phenotype spatial correlation.

### Random 1 (5 sec)
```{r, GA1 cor and pheno}
GA1_pattern <- unique(paste0(GA1_lit_interrate_file$snake_mu_rate, "_", GA1_lit_interrate_file$snake_mu_effect_sd,"_",GA1_lit_interrate_file$newt_mu_rate, "_", GA1_lit_interrate_file$newt_mu_effect_sd, "_", GA1_lit_interrate_file$rep))

myrandomlist=sample.int(length(GA1_pattern), 3)

cor_trends_by_gen(GA1_lit_interrate_file, GA1_cor_interrate_file, "test", "GA1", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[myrandomlist[1]])

```

### Random 2 (5 sec)
```{r, random cor-pheno plot 2}

cor_trends_by_gen(GA1_lit_interrate_file, GA1_cor_interrate_file, "test", "GA1", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[myrandomlist[2]])
```

### Random 3 (5 sec)
```{r, random cor-pheno plot 3}

cor_trends_by_gen(GA1_lit_interrate_file, GA1_cor_interrate_file, "test", "GA1", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[myrandomlist[3]])
```

## {-}

## Population Size Correlation across time {.tabset}

### Random 1 
```{r, GA1 pop size 1}
pop_size_trends_by_gen(GA1_lit_interrate_file, GA1_cor_interrate_file, "test", "GA1", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[myrandomlist[1]])

```

### Random 2 
```{r, random pop size plot 2}

pop_size_trends_by_gen(GA1_lit_interrate_file, GA1_cor_interrate_file, "test", "GA1", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[myrandomlist[2]])
```

### Random 3 
```{r, random pop size plot 3}

pop_size_trends_by_gen(GA1_lit_interrate_file, GA1_cor_interrate_file, "test", "GA1", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[myrandomlist[3]])
```

## {-}

## What happens over time (looking at the beginning and late part of my simulations) {.tabset}

### Pheno Beginning 0.01

```{r, early 0.01 plots}

min_gen=100
max_gen=5000
interrate=0.01

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno Beginning 0.025

```{r, early 0.025 plots}

min_gen=100
max_gen=5000
interrate=0.025

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno Beginning 0.05

```{r, early 0.05 plots}

min_gen=100
max_gen=5000
interrate=0.05

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno Beginning 0.075

```{r, early 0.075 plots}

min_gen=100
max_gen=5000
interrate=0.075

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno Beginning 0.10

```{r, early 0.10 plots}

min_gen=100
max_gen=5000
interrate=0.10

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno Beginning 0.125

```{r, early 0.125 plots}

min_gen=100
max_gen=5000
interrate=0.125

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno Beginning 0.15

```{r, early 0.15 plots}

min_gen=100
max_gen=5000
interrate=0.15

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno Beginning 0.20

```{r, early 0.20 plots}

min_gen=100
max_gen=5000
interrate=0.20

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno Beginning 0.25

```{r, early 0.25 plots}

min_gen=100
max_gen=5000
interrate=0.25

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno Beginning 0.30

```{r, early 0.30 plots}

min_gen=100
max_gen=5000
interrate=0.30

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno Beginning 0.35

```{r, early 0.35 plots}

min_gen=100
max_gen=5000
interrate=0.35

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno End 0.01

```{r, late 0.01 plots}

min_gen=45000
max_gen=50000
interrate=0.01

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno End 0.025

```{r, late 0.025 plots}

min_gen=45000
max_gen=50000
interrate=0.025

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno End 0.05

```{r, late 0.05 plots}

min_gen=45000
max_gen=50000
interrate=0.05

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno End 0.075

```{r, late 0.075 plots}

min_gen=45000
max_gen=50000
interrate=0.075

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno End 0.10

```{r, late 0.10 plots}

min_gen=45000
max_gen=50000
interrate=0.10

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno End 0.125

```{r, late 0.125 plots}

min_gen=45000
max_gen=50000
interrate=0.125

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno End 0.15

```{r, late 0.15 plots}

min_gen=45000
max_gen=50000
interrate=0.15

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno End 0.20

```{r, late 0.20 plots}

min_gen=45000
max_gen=50000
interrate=0.20

#whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno End 0.25

```{r, late 0.25 plots}

min_gen=45000
max_gen=50000
interrate=0.25

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno End 0.30

```{r, late 0.30 plots}

min_gen=45000
max_gen=50000
interrate=0.30

#whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

### Pheno End 0.35

```{r, late 0.35 plots}

min_gen=45000
max_gen=50000
interrate=0.35

whale_gen_by_interrate(GA1_lit_interrate_file, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(GA1_lit_interrate_file$rep)[1], interrate)
```

## {-}

## Heatmap {.tabset}

### interrate: Population Size (Early)  
```{r, interrate earlyplots Heatmap1, fig.height=4}
min_gen=100
max_gen=5000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

```

### interrate: Population Size (EM1)  
```{r, interrate earlyplots Heatmap2, fig.height=4}
min_gen=5000
max_gen=10000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

```

### interrate: Population Size (EM2)  
```{r, interrate earlyplots Heatmap3, fig.height=4}
min_gen=10000
max_gen=15000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

```

### interrate: Population Size (EM3)  
```{r, interrate earlyplots Heatmap4, fig.height=4}
min_gen=15000
max_gen=20000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

```

### interrate: Population Size (EM4)  
```{r, interrate earlyplots Heatmap5, fig.height=4}
min_gen=20000
max_gen=25000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

```

### interrate: Population Size (Mid)  
```{r, interrate midplots Heatmap6, fig.height=4}
min_gen=25000
max_gen=30000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

```

### interrate: Population Size (ML1)  
```{r, interrate midplots Heatmap7, fig.height=4}
min_gen=30000
max_gen=35000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

```

### interrate: Population Size (ML2)  
```{r, interrate midplots Heatmap8, fig.height=4}
min_gen=35000
max_gen=40000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

```

### interrate: Population Size (ML3)  
```{r, interrate midplots Heatmap9, fig.height=4}
min_gen=40000
max_gen=45000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

```

### interrate: Population Size (Late)  
```{r, interrate plots Heatmap10, fig.height=4}
min_gen=45000
max_gen=50000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

```

### interrate: Phenotype (Early) 
```{r, interrate earlyplots Heatmap1 pheno, fig.height=4}

min_gen=100
max_gen=5000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### interrate: Phenotype (EM1) 
```{r, interrate earlyplots Heatmap2 pheno, fig.height=4}

min_gen=5000
max_gen=10000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### interrate: Phenotype (EM2) 
```{r, interrate earlyplots Heatmap3 pheno, fig.height=4}

min_gen=10000
max_gen=15000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### interrate: Phenotype (EM3) 
```{r, interrate earlyplots Heatmap4 pheno, fig.height=4}

min_gen=15000
max_gen=20000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### interrate: Phenotype (EM4) 
```{r, interrate earlyplots Heatmap5 pheno, fig.height=4}

min_gen=20000
max_gen=25000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### interrate: Phenotype (Mid) 
```{r, interrate earlyplots Heatmap6 pheno, fig.height=4}

min_gen=25000
max_gen=30000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### interrate: Phenotype (ML1) 
```{r, interrate earlyplots Heatmap7 pheno, fig.height=4}

min_gen=30000
max_gen=35000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### interrate: Phenotype (ML2) 
```{r, interrate earlyplots Heatmap8 pheno, fig.height=4}

min_gen=35000
max_gen=40000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### interrate: Phenotype (ML3) 
```{r, interrate earlyplots Heatmap9 pheno, fig.height=4}

min_gen=40000
max_gen=45000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### interrate: Phenotype (Late) 
```{r, interrate plots Heatmap10 pheno, fig.height=4}

min_gen=45000
max_gen=50000

build_heatmap(GA1_lit_interrate_file, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

## {-}

##Grid {.tabset}

```{r, red in grid files}
require(readr)
GA1_grid_interrate_file<- read.csv("~/Desktop/interaction_rate/all_grid.txt" , header = TRUE) %>%  filter(between(generation, 10000, 15000))
GA1_grid_interrate_file$group_rep <- paste0(GA1_grid_interrate_file$snake_mu_rate, "_", GA1_grid_interrate_file$snake_mu_effect_sd,"_",GA1_grid_interrate_file$newt_mu_rate, "_", GA1_grid_interrate_file$newt_mu_effect_sd, "_", GA1_grid_interrate_file$rep)
GA1_grid_interrate_file$group <- paste0(GA1_grid_interrate_file$snake_mu_rate, "_", GA1_grid_interrate_file$snake_mu_effect_sd,"_",GA1_grid_interrate_file$newt_mu_rate, "_", GA1_grid_interrate_file$newt_mu_effect_sd)
GA1_grid_interrate_file$sim=LETTERS[factor(GA1_grid_interrate_file$group)]
GA1_grid_interrate_file$grid_num = 5

GA1_grid_interrate_file$snake_newt_rep <- paste0(GA1_grid_interrate_file$snake_mu_rate, "_", GA1_grid_interrate_file$snake_mu_effect_sd,"_",GA1_grid_interrate_file$newt_mu_rate, "_", GA1_grid_interrate_file$newt_mu_effect_sd, "_", GA1_grid_interrate_file$rep)

```

## Grid Figs {.tabset}

### 0.01
```{r, cor 0.01 high cor mean}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[1]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.01

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),])
if(value!=0){
grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) 
}
```

### 0.025
```{r, cor 0.025 high cor mean}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[1]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.025

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),])
if(value!=0){
 grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) 
}
```

### 0.05
```{r, cor 0.05 high cor mean}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[1]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.05

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.075
```{r, cor 0.075 high cor mean}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[1]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.075

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.10
```{r, cor 0.10 high cor mean1}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[1]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.10

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),])
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.125
```{r, cor 0.125 high cor mean 1}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[1]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.125

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ 
  grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) 
  }
```

### 0.15
```{r, cor 0.15 high cor mean 1}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[1]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.15

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.20
```{r, cor 0.20 high cor mean 1}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[1]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.20

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.25
```{r, cor 0.25 high cor mean 1}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[1]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.25

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.30
```{r, cor 0.30 high cor mean 1}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[1]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.30

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.35
```{r, cor 0.35 high cor mean 1}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[1]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.35

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

## {-}

## Grid Figs 2 {.tabset}

### 0.01
```{r, cor 0.01 high cor mean2}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[2]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.01

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.025
```{r, cor 0.025 high cor mean2}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[2]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.025

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.05
```{r, cor 0.05 high cor mea2}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[2]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.05

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.075
```{r, cor 0.075 high cor mean2}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[2]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.075

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.10
```{r, cor 0.10 high cor mean2}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[2]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.10

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.125
```{r, cor 0.125 high cor mean2}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[2]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.125

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.15
```{r, cor 0.15 high cor mean2}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[2]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.15

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.20
```{r, cor 0.20 high cor mean2}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[2]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.20

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),])

if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.25
```{r, cor 0.25 high cor mean2}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[2]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.25

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.30
```{r, cor 0.30 high cor mean2}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[2]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.30

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.35
```{r, cor 0.35 high cor mean2}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[2]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.35

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

## {-}

## Grid Figs {.tabset}

### 0.01
```{r, cor 0.01 high cor mean3}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[3]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.01

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.025
```{r, cor 0.025 high cor mean3}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[3]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.025

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.05
```{r, cor 0.05 high cor mean3}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[3]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.05

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.075
```{r, cor 0.075 high cor mean3}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[3]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.075

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.10
```{r, cor 0.10 high cor mean3}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[3]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.10

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.125
```{r, cor 0.125 high cor mean3}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[3]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.125

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.15
```{r, cor 0.15 high cor mean3}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[3]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.15

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.20
```{r, cor 0.20 high cor mean3}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[3]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.20

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.25
```{r, cor 0.25 high cor mean3}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[3]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.25

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.30
```{r, cor 0.30 high cor mean3}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[3]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.30

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),])
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

### 0.35
```{r, cor 0.35 high cor mean3}

file=GA1_grid_interrate_file
file_name=GA1_pattern[myrandomlist[3]]
spegen=14981
type="mean"
col_num=5
row_num=col_num*4
interrate=0.35

value<-nrow(GA1_grid_interrate_file[which(GA1_grid_interrate_file$snake_newt_rep==file_name & GA1_grid_interrate_file$generation==spegen & GA1_grid_interrate_file$interrate==interrate),]) 
if(value!=0){ grid_plot(file_name, spegen, file, type, row_num, col_num, interrate) }
```

## {-}