### 2022-12-06 graphs_newt_snakes/points.Rmd

---
title: "byindvidual"
output: html_document
date: '2022-08-03'
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE,echo = FALSE)

library(ggplot2)
library(viridis)
library(colorspace)
library(gridExtra)
library(tidyverse)
require(reshape2)
library(plyr)
library(gridGraphics)
library(grid)
library(knitr)
library(kableExtra)
library("reshape")
library(ggplot2)
library(ggpubr)
theme_set(theme_pubr())
library(stringr)
library(patchwork)
```

```{r, data gathering}

info <- read.table("~/Desktop/test.txt", header=TRUE)

f <- function(x, y, ph, xbreaks, ybreaks){
  xb <- cut(x, xbreaks)
  yb <- cut(y, ybreaks)
  phb <- tapply(ph, list(xb, yb), mean)
  return(phb)
}

```

```{r, data plotting tall}

width=35
height=35*4
n=5
m=20
xbreaks <- seq(0, width, length.out=n)
ybreaks <- seq(0, height, length.out=m)

s <- with(info[info$species=="S",], f(loc_x, loc_y, phenotype, xbreaks, ybreaks))
n <- with(info[info$species=="N",], f(loc_x, loc_y, phenotype, xbreaks, ybreaks))

stopifnot(all(colnames(s)==colnames(n)) && all(rownames(s)==rownames(n)))

cor(as.vector(s),as.vector(n), use="pairwise")

n=7
m=28
xbreaks <- seq(0, width, length.out=n)
ybreaks <- seq(0, height, length.out=m)

s <- with(info[info$species=="S",], f(loc_x, loc_y, phenotype, xbreaks, ybreaks))
n <- with(info[info$species=="N",], f(loc_x, loc_y, phenotype, xbreaks, ybreaks))

stopifnot(all(colnames(s)==colnames(n)) && all(rownames(s)==rownames(n)))

cor(as.vector(s),as.vector(n), use="pairwise")

image(ybreaks,xbreaks, t(n))
with(info[info$species=="N",], points(loc_y, loc_x, pch=20,cex=0.5, col=rev(heat.colors(24))[as.numeric(cut(phenotype,24))]))

image(ybreaks,xbreaks, t(s))
with(info[info$species=="S",], points(loc_y, loc_x, pch=20,cex=0.5, col=rev(heat.colors(24))[as.numeric(cut(phenotype,24))]))

```

```{r, plotting long}

width=35
height=35*4
n=5
m=20
xbreaks <- seq(0, width, length.out=n)
ybreaks <- seq(0, height, length.out=m)

s <- with(info[info$species=="S",], f(loc_x, loc_y, phenotype, xbreaks, ybreaks))
n <- with(info[info$species=="N",], f(loc_x, loc_y, phenotype, xbreaks, ybreaks))

stopifnot(all(colnames(s)==colnames(n)) && all(rownames(s)==rownames(n)))

cor(as.vector(s),as.vector(n), use="pairwise")

image(ybreaks,xbreaks, t(n))
with(info[info$species=="N",], points(loc_y, loc_x, pch=20,cex=0.5, col=rev(heat.colors(24))[as.numeric(cut(phenotype,24))]))

image(ybreaks,xbreaks, t(s))
with(info[info$species=="S",], points(loc_y, loc_x, pch=20,cex=0.5, col=rev(heat.colors(24))[as.numeric(cut(phenotype,24))]))

```

```{r, find the nearest neghbor}

```