### 2023-01-18 graphs_newt_snakes/high_to_low_v.Rmd

---
title: "high to low v"
output: html_document
date: '2023-01-13'
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE,echo = FALSE)

library(ggplot2)
library(viridis)
library(colorspace)
library(gridExtra)
library(tidyverse)
require(reshape2)
library(plyr)
library(gridGraphics)
library(grid)
library(knitr)
library(kableExtra)
library("reshape")
library(ggplot2)
library(ggpubr)
theme_set(theme_pubr())
library(stringr)
library(patchwork)
```

## Goal:

### Questions:

What does it mean to be correlated?

## Background

## Experiment

I created a simulation study to observe the co-evolutionary outcome of the newt-snake interaction with different genetic architectures (GAs) in a spatial setting. I hypothesized that we would see an interaction (co-evolutionary arms race) between newt and snake phenotype under some GA combinations when newts and snakes were evolving over geographical space. Each GA is paired with another GA creating 16 combinations.

Experiment values:

### The data

many of the experiments did not run to completion (i will run this part 3 times)

```{bash, data file gathering 1}
#!/usr/bin/env bash

# Run these comands while in The the folder that contains the file you would like to condence
# Goal is to condence all the text files into one file 
# Requirements: need to have the same name and need to add additional columns
thepath=/Users/victoria/Desktop/
foldername=low_to_high_v/mu_a
thefolder=${thepath}${foldername}/*

add_in_column=",rep,snake_mu_rate,snake_mu_effect_sd,newt_mu_rate,newt_mu_effect_sd"

cor_head_file=$(find ${thepath}${foldername} -name "*_GA205_cor_*" | head -n 1)
cor_head=$(head -n 1 "${cor_head_file}")

grid_head_file=$(find ${thepath}${foldername} -name "*_GA205_grid_*" | head -n 1)
grid_head=$(head -n 1 "${grid_head_file}")

lit_head_file=$(find ${thepath}${foldername} -name "*_GA205_lit_*" | head -n 1)
lit_head=$(head -n 1 "${lit_head_file}")

#test -f /Users/victoria/Desktop/bash_test/all_lit.txt && echo "all lit file exists."
#test -f /Users/victoria/Desktop/bash_test/all_grid.txt && echo "all grid file exists."
#test -f /Users/victoria/Desktop/bash_test/all_cor.txt && echo "all cor file exists."

if [[ -f ${thepath}${foldername}/all_lit.txt ]] && [[ -f ${thepath}${foldername}/all_grid.txt ]] && [[ -f ${thepath}${foldername}/all_cor.txt ]]
then
  echo "All cor, lit, and grid files exist!"
  echo "This program will now end!"
  exit 0
fi

#creating the files
echo ${lit_head}${add_in_column} > ${thepath}${foldername}/all_lit.txt
echo ${grid_head}${add_in_column} > ${thepath}${foldername}/all_grid.txt
echo ${cor_head}${add_in_column} > ${thepath}${foldername}/all_cor.txt

for filename in $thefolder*.text;
  do 
#getting the variables 
my_type=$(echo ${filename} |sed 's/.*GA205_//; s/_.*//')
snake_mu_rate=$(echo ${filename} |sed 's/.*snake_mu_rate_//; s/_.*//')
snake_mu_effect_sd=$(echo ${filename} |sed 's/.*snake_mu_effect_sd_//; s/_.*//')
newt_mu_rate=$(echo ${filename} |sed 's/.*newt_mu_rate_//; s/_.*//')
newt_mu_effect_sd=$(echo ${filename} |sed 's/.*newt_mu_effect_sd_//; s/_.*//')
rep=$(echo ${filename} | sed "s/.*rep_//; s/_.*//")
#echo ${newt_mu_rate}	
# echo $my_type
# echo $my_other_type
# these new columns adding them to the data
sed -i '' -e 's/$/,'${rep}'/' $filename
sed -i '' -e 's/$/,'${snake_mu_rate}'/' $filename
sed -i '' -e 's/$/,'${snake_mu_effect_sd}'/' $filename
sed -i '' -e 's/$/,'${newt_mu_rate}'/' $filename
sed -i '' -e 's/$/,'${newt_mu_effect_sd}'/' $filename
#infile >outfile
if [[ $my_type == "lit" ]]
then
    tail -n +2 -q $filename >> ${thepath}${foldername}/all_lit.txt
fi

if [[ $my_type == "grid" ]]
then
    tail -n +2 -q $filename >> ${thepath}${foldername}/all_grid.txt
fi

if [[ $my_type == "cor" ]]
then
    tail -n +2 -q $filename >> ${thepath}${foldername}/all_cor.txt
fi

done

```

```{bash, data file gathering 2}
#!/usr/bin/env bash

# Run these comands while in The the folder that contains the file you would like to condence
# Goal is to condence all the text files into one file 
# Requirements: need to have the same name and need to add additional columns
thepath=/Users/victoria/Desktop/
foldername=low_to_high_v/mu_b
thefolder=${thepath}${foldername}/*

add_in_column=",rep,snake_mu_rate,snake_mu_effect_sd,newt_mu_rate,newt_mu_effect_sd"

cor_head_file=$(find ${thepath}${foldername} -name "*_GA205_cor_*" | head -n 1)
cor_head=$(head -n 1 "${cor_head_file}")

grid_head_file=$(find ${thepath}${foldername} -name "*_GA205_grid_*" | head -n 1)
grid_head=$(head -n 1 "${grid_head_file}")

lit_head_file=$(find ${thepath}${foldername} -name "*_GA205_lit_*" | head -n 1)
lit_head=$(head -n 1 "${lit_head_file}")

#test -f /Users/victoria/Desktop/bash_test/all_lit.txt && echo "all lit file exists."
#test -f /Users/victoria/Desktop/bash_test/all_grid.txt && echo "all grid file exists."
#test -f /Users/victoria/Desktop/bash_test/all_cor.txt && echo "all cor file exists."

if [[ -f ${thepath}${foldername}/all_lit.txt ]] && [[ -f ${thepath}${foldername}/all_grid.txt ]] && [[ -f ${thepath}${foldername}/all_cor.txt ]]
then
  echo "All cor, lit, and grid files exist!"
  echo "This program will now end!"
  exit 0
fi

#creating the files
echo ${lit_head}${add_in_column} > ${thepath}${foldername}/all_lit.txt
echo ${grid_head}${add_in_column} > ${thepath}${foldername}/all_grid.txt
echo ${cor_head}${add_in_column} > ${thepath}${foldername}/all_cor.txt

for filename in $thefolder*.text;
  do 
#getting the variables 
my_type=$(echo ${filename} |sed 's/.*GA205_//; s/_.*//')
snake_mu_rate=$(echo ${filename} |sed 's/.*snake_mu_rate_//; s/_.*//')
snake_mu_effect_sd=$(echo ${filename} |sed 's/.*snake_mu_effect_sd_//; s/_.*//')
newt_mu_rate=$(echo ${filename} |sed 's/.*newt_mu_rate_//; s/_.*//')
newt_mu_effect_sd=$(echo ${filename} |sed 's/.*newt_mu_effect_sd_//; s/_.*//')
rep=$(echo ${filename} | sed "s/.*rep_//; s/_.*//")
#echo ${newt_mu_rate}	
# echo $my_type
# echo $my_other_type
# these new columns adding them to the data
sed -i '' -e 's/$/,'${rep}'/' $filename
sed -i '' -e 's/$/,'${snake_mu_rate}'/' $filename
sed -i '' -e 's/$/,'${snake_mu_effect_sd}'/' $filename
sed -i '' -e 's/$/,'${newt_mu_rate}'/' $filename
sed -i '' -e 's/$/,'${newt_mu_effect_sd}'/' $filename
#infile >outfile
if [[ $my_type == "lit" ]]
then
    tail -n +2 -q $filename >> ${thepath}${foldername}/all_lit.txt
fi

if [[ $my_type == "grid" ]]
then
    tail -n +2 -q $filename >> ${thepath}${foldername}/all_grid.txt
fi

if [[ $my_type == "cor" ]]
then
    tail -n +2 -q $filename >> ${thepath}${foldername}/all_cor.txt
fi

done

```

```{bash, data file gathering 3}
#!/usr/bin/env bash

# Run these comands while in The the folder that contains the file you would like to condence
# Goal is to condence all the text files into one file 
# Requirements: need to have the same name and need to add additional columns
thepath=/Users/victoria/Desktop/
foldername=low_to_high_v/mu_c
thefolder=${thepath}${foldername}/*

add_in_column=",rep,snake_mu_rate,snake_mu_effect_sd,newt_mu_rate,newt_mu_effect_sd"

cor_head_file=$(find ${thepath}${foldername} -name "*_GA205_cor_*" | head -n 1)
cor_head=$(head -n 1 "${cor_head_file}")

grid_head_file=$(find ${thepath}${foldername} -name "*_GA205_grid_*" | head -n 1)
grid_head=$(head -n 1 "${grid_head_file}")

lit_head_file=$(find ${thepath}${foldername} -name "*_GA205_lit_*" | head -n 1)
lit_head=$(head -n 1 "${lit_head_file}")

#test -f /Users/victoria/Desktop/bash_test/all_lit.txt && echo "all lit file exists."
#test -f /Users/victoria/Desktop/bash_test/all_grid.txt && echo "all grid file exists."
#test -f /Users/victoria/Desktop/bash_test/all_cor.txt && echo "all cor file exists."

if [[ -f ${thepath}${foldername}/all_lit.txt ]] && [[ -f ${thepath}${foldername}/all_grid.txt ]] && [[ -f ${thepath}${foldername}/all_cor.txt ]]
then
  echo "All cor, lit, and grid files exist!"
  echo "This program will now end!"
  exit 0
fi

#creating the files
echo ${lit_head}${add_in_column} > ${thepath}${foldername}/all_lit.txt
echo ${grid_head}${add_in_column} > ${thepath}${foldername}/all_grid.txt
echo ${cor_head}${add_in_column} > ${thepath}${foldername}/all_cor.txt

for filename in $thefolder*.text;
  do 
#getting the variables 
my_type=$(echo ${filename} |sed 's/.*GA205_//; s/_.*//')
snake_mu_rate=$(echo ${filename} |sed 's/.*snake_mu_rate_//; s/_.*//')
snake_mu_effect_sd=$(echo ${filename} |sed 's/.*snake_mu_effect_sd_//; s/_.*//')
newt_mu_rate=$(echo ${filename} |sed 's/.*newt_mu_rate_//; s/_.*//')
newt_mu_effect_sd=$(echo ${filename} |sed 's/.*newt_mu_effect_sd_//; s/_.*//')
rep=$(echo ${filename} | sed "s/.*rep_//; s/_.*//")
#echo ${newt_mu_rate}	
# echo $my_type
# echo $my_other_type
# these new columns adding them to the data
sed -i '' -e 's/$/,'${rep}'/' $filename
sed -i '' -e 's/$/,'${snake_mu_rate}'/' $filename
sed -i '' -e 's/$/,'${snake_mu_effect_sd}'/' $filename
sed -i '' -e 's/$/,'${newt_mu_rate}'/' $filename
sed -i '' -e 's/$/,'${newt_mu_effect_sd}'/' $filename
#infile >outfile
if [[ $my_type == "lit" ]]
then
    tail -n +2 -q $filename >> ${thepath}${foldername}/all_lit.txt
fi

if [[ $my_type == "grid" ]]
then
    tail -n +2 -q $filename >> ${thepath}${foldername}/all_grid.txt
fi

if [[ $my_type == "cor" ]]
then
    tail -n +2 -q $filename >> ${thepath}${foldername}/all_cor.txt
fi

done

```

```{bash, data file gathering 4}
#!/usr/bin/env bash

# Run these comands while in The the folder that contains the file you would like to condence
# Goal is to condence all the text files into one file 
# Requirements: need to have the same name and need to add additional columns
thepath=/Users/victoria/Desktop/
foldername=low_to_high_v/mu_d
thefolder=${thepath}${foldername}/*

add_in_column=",rep,snake_mu_rate,snake_mu_effect_sd,newt_mu_rate,newt_mu_effect_sd"

cor_head_file=$(find ${thepath}${foldername} -name "*_GA205_cor_*" | head -n 1)
cor_head=$(head -n 1 "${cor_head_file}")

grid_head_file=$(find ${thepath}${foldername} -name "*_GA205_grid_*" | head -n 1)
grid_head=$(head -n 1 "${grid_head_file}")

lit_head_file=$(find ${thepath}${foldername} -name "*_GA205_lit_*" | head -n 1)
lit_head=$(head -n 1 "${lit_head_file}")

#test -f /Users/victoria/Desktop/bash_test/all_lit.txt && echo "all lit file exists."
#test -f /Users/victoria/Desktop/bash_test/all_grid.txt && echo "all grid file exists."
#test -f /Users/victoria/Desktop/bash_test/all_cor.txt && echo "all cor file exists."

if [[ -f ${thepath}${foldername}/all_lit.txt ]] && [[ -f ${thepath}${foldername}/all_grid.txt ]] && [[ -f ${thepath}${foldername}/all_cor.txt ]]
then
  echo "All cor, lit, and grid files exist!"
  echo "This program will now end!"
  exit 0
fi

#creating the files
echo ${lit_head}${add_in_column} > ${thepath}${foldername}/all_lit.txt
echo ${grid_head}${add_in_column} > ${thepath}${foldername}/all_grid.txt
echo ${cor_head}${add_in_column} > ${thepath}${foldername}/all_cor.txt

for filename in $thefolder*.text;
  do 
#getting the variables 
my_type=$(echo ${filename} |sed 's/.*GA205_//; s/_.*//')
snake_mu_rate=$(echo ${filename} |sed 's/.*snake_mu_rate_//; s/_.*//')
snake_mu_effect_sd=$(echo ${filename} |sed 's/.*snake_mu_effect_sd_//; s/_.*//')
newt_mu_rate=$(echo ${filename} |sed 's/.*newt_mu_rate_//; s/_.*//')
newt_mu_effect_sd=$(echo ${filename} |sed 's/.*newt_mu_effect_sd_//; s/_.*//')
rep=$(echo ${filename} | sed "s/.*rep_//; s/_.*//")
#echo ${newt_mu_rate}	
# echo $my_type
# echo $my_other_type
# these new columns adding them to the data
sed -i '' -e 's/$/,'${rep}'/' $filename
sed -i '' -e 's/$/,'${snake_mu_rate}'/' $filename
sed -i '' -e 's/$/,'${snake_mu_effect_sd}'/' $filename
sed -i '' -e 's/$/,'${newt_mu_rate}'/' $filename
sed -i '' -e 's/$/,'${newt_mu_effect_sd}'/' $filename
#infile >outfile
if [[ $my_type == "lit" ]]
then
    tail -n +2 -q $filename >> ${thepath}${foldername}/all_lit.txt
fi

if [[ $my_type == "grid" ]]
then
    tail -n +2 -q $filename >> ${thepath}${foldername}/all_grid.txt
fi

if [[ $my_type == "cor" ]]
then
    tail -n +2 -q $filename >> ${thepath}${foldername}/all_cor.txt
fi

done

```

```{r, data gathering}

snewt <- read.csv("~/Desktop/snake_newt_real_data.csv", header = TRUE)

mu_a_lit<- read.csv("~/Desktop/low_to_high_v/mu_a/all_lit.txt" , header = TRUE)

mu_a_lit$group_rep <- paste0(mu_a_lit$snake_mu_rate, "_", mu_a_lit$snake_mu_effect_sd,"_",mu_a_lit$newt_mu_rate, "_", mu_a_lit$newt_mu_effect_sd, "_", mu_a_lit$rep)

mu_a_lit$group <- paste0(mu_a_lit$snake_mu_rate, "_", mu_a_lit$snake_mu_effect_sd,"_",mu_a_lit$newt_mu_rate, "_", mu_a_lit$newt_mu_effect_sd)

mu_a_lit$sim=LETTERS[factor(mu_a_lit$group)]
mu_a_lit$dif <- mu_a_lit$Snake_mean_Pheno-mu_a_lit$Newt_mean_Pheno

mu_a_cor<- read.csv("~/Desktop/low_to_high_v/mu_a/all_cor.txt" , header = TRUE)
mu_a_cor$group_rep <- paste0(mu_a_cor$snake_mu_rate, "_", mu_a_cor$snake_mu_effect_sd,"_",mu_a_cor$newt_mu_rate, "_", mu_a_cor$newt_mu_effect_sd, "_", mu_a_cor$rep)
mu_a_cor$group <- paste0(mu_a_cor$snake_mu_rate, "_", mu_a_cor$snake_mu_effect_sd,"_",mu_a_cor$newt_mu_rate, "_", mu_a_cor$newt_mu_effect_sd)
mu_a_cor$sim=LETTERS[factor(mu_a_cor$group)]

# Mu B
mu_b_lit<- read.csv("~/Desktop/low_to_high_v/mu_b/all_lit.txt" , header = TRUE)

mu_b_lit$group_rep <- paste0(mu_b_lit$snake_mu_rate, "_", mu_b_lit$snake_mu_effect_sd,"_",mu_b_lit$newt_mu_rate, "_", mu_b_lit$newt_mu_effect_sd, "_", mu_b_lit$rep)

mu_b_lit$group <- paste0(mu_b_lit$snake_mu_rate, "_", mu_b_lit$snake_mu_effect_sd,"_",mu_b_lit$newt_mu_rate, "_", mu_b_lit$newt_mu_effect_sd)

mu_b_lit$sim=LETTERS[factor(mu_b_lit$group)]
mu_b_lit$dif <- mu_b_lit$Snake_mean_Pheno-mu_b_lit$Newt_mean_Pheno

mu_b_cor<- read.csv("~/Desktop/low_to_high_v/mu_b/all_cor.txt" , header = TRUE)
mu_b_cor$group_rep <- paste0(mu_b_cor$snake_mu_rate, "_", mu_b_cor$snake_mu_effect_sd,"_",mu_b_cor$newt_mu_rate, "_", mu_b_cor$newt_mu_effect_sd, "_", mu_b_cor$rep)
mu_b_cor$group <- paste0(mu_b_cor$snake_mu_rate, "_", mu_b_cor$snake_mu_effect_sd,"_",mu_b_cor$newt_mu_rate, "_", mu_b_cor$newt_mu_effect_sd)
mu_b_cor$sim=LETTERS[factor(mu_b_cor$group)]

# no exp

mu_c_lit<- read.csv("~/Desktop/low_to_high_v/mu_c/all_lit.txt" , header = TRUE)

mu_c_lit$group_rep <- paste0(mu_c_lit$snake_mu_rate, "_", mu_c_lit$snake_mu_effect_sd,"_",mu_c_lit$newt_mu_rate, "_", mu_c_lit$newt_mu_effect_sd, "_", mu_c_lit$rep)

mu_c_lit$group <- paste0(mu_c_lit$snake_mu_rate, "_", mu_c_lit$snake_mu_effect_sd,"_",mu_c_lit$newt_mu_rate, "_", mu_c_lit$newt_mu_effect_sd)

mu_c_lit$sim=LETTERS[factor(mu_c_lit$group)]
mu_c_lit$dif <- mu_c_lit$Snake_mean_Pheno-mu_c_lit$Newt_mean_Pheno

mu_c_cor<- read.csv("~/Desktop/low_to_high_v/mu_c/all_cor.txt" , header = TRUE)
mu_c_cor$group_rep <- paste0(mu_c_cor$snake_mu_rate, "_", mu_c_cor$snake_mu_effect_sd,"_",mu_c_cor$newt_mu_rate, "_", mu_c_cor$newt_mu_effect_sd, "_", mu_c_cor$rep)
mu_c_cor$group <- paste0(mu_c_cor$snake_mu_rate, "_", mu_c_cor$snake_mu_effect_sd,"_",mu_c_cor$newt_mu_rate, "_", mu_c_cor$newt_mu_effect_sd)
mu_c_cor$sim=LETTERS[factor(mu_c_cor$group)]

# mu d

mu_d_lit<- read.csv("~/Desktop/low_to_high_v/mu_d/all_lit.txt" , header = TRUE)

mu_d_lit$group_rep <- paste0(mu_d_lit$snake_mu_rate, "_", mu_d_lit$snake_mu_effect_sd,"_",mu_d_lit$newt_mu_rate, "_", mu_d_lit$newt_mu_effect_sd, "_", mu_d_lit$rep)

mu_d_lit$group <- paste0(mu_d_lit$snake_mu_rate, "_", mu_d_lit$snake_mu_effect_sd,"_",mu_d_lit$newt_mu_rate, "_", mu_d_lit$newt_mu_effect_sd)

mu_d_lit$sim=LETTERS[factor(mu_d_lit$group)]
mu_d_lit$dif <- mu_d_lit$Snake_mean_Pheno-mu_d_lit$Newt_mean_Pheno

mu_d_cor<- read.csv("~/Desktop/low_to_high_v/mu_d/all_cor.txt" , header = TRUE)
mu_d_cor$group_rep <- paste0(mu_d_cor$snake_mu_rate, "_", mu_d_cor$snake_mu_effect_sd,"_",mu_d_cor$newt_mu_rate, "_", mu_d_cor$newt_mu_effect_sd, "_", mu_d_cor$rep)
mu_d_cor$group <- paste0(mu_d_cor$snake_mu_rate, "_", mu_d_cor$snake_mu_effect_sd,"_",mu_d_cor$newt_mu_rate, "_", mu_d_cor$newt_mu_effect_sd)
mu_d_cor$sim=LETTERS[factor(mu_d_cor$group)]

```

```{r, cor box set up}

boxplot_cors_by_gen <- function(file, title, subt, x_label, y_label, min_gen, max_gen){
 snewt$mean_res = snewt$mean_toxin+snewt$dif
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
  file$snake_newt <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  unique(file$sim)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ylim(-1.1, 1.1)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor(rep), y=as.numeric(mean_newt_pheno_By_mean_snake_pheno), color = "CA"))+
  geom_hline(yintercept=real_snewt_cor, linetype="dashed")+
    geom_hline(yintercept=0, linetype="solid")+
  #geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  #geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  #geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  theme(plot.title = element_text(size=5)) +
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

  return(plot)
}

hist_cors_by_gen_col_rep <- function(file, title, subt, x_label, y_label, min_gen, max_gen){
 snewt$mean_res = snewt$mean_toxin+snewt$dif
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
   file$snake_newt <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  #unique(file$sim)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=as.numeric(mean_newt_pheno_By_mean_snake_pheno),fill=factor(rep)))+
  geom_vline(xintercept=real_snewt_cor, linetype="dashed")+
  geom_vline(xintercept=0, linetype="solid")+
  #geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  #geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  #geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  theme(plot.title = element_text(size=5)) +
  xlim(-1.1, 1.1)+
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

  return(plot)
}

hist_cors_by_gen_col_rep_nofill <- function(file, title, subt, x_label, y_label, min_gen, max_gen){
 snewt$mean_res = snewt$mean_toxin+snewt$dif
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
  file$snake_newt <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  unique(file$sim)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=as.numeric(mean_newt_pheno_By_mean_snake_pheno)))+
  geom_vline(xintercept=real_snewt_cor, linetype="dashed")+
  geom_vline(xintercept=0, linetype="solid")+
  #geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  #geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  #geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  theme(plot.title = element_text(size=5)) +
  #stat_bin(bins=30)+
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

suppressMessages(print(plot))

  return(plot)
}

var_cor <- function(file, cfile, title, subt, x_label, y_label, min_gen, max_gen){
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
  real_snewt_var = var(snewt$dif)
  var_cor_table=data.frame()

  for(i in 1:length(unique(file$snake_newt_rep))){
    pattern<-unique(file$snake_newt_rep)[i]
    temp_file<-file[which(file$gen>=min_gen & file$gen<=max_gen & file$snake_newt_rep==pattern),]
    temp_cor_file<-cfile[which(cfile$gen>=min_gen & cfile$gen<=max_gen & cfile$snake_newt_rep==pattern),]

    var_cor_table_temp<- data.frame("pattern"=pattern, "var"= var(temp_file$dif),"cor" = cor(temp_file$newt_toxin, temp_file$snake_res), "mean_corr" = mean(temp_cor_file$mean_newt_pheno_By_mean_snake_pheno))
    var_cor_table <- rbind(var_cor_table, var_cor_table_temp)
  }

  var_cor_table_temp<- data.frame("pattern"="real", "var"= real_snewt_var, "cor" = real_snewt_cor, "mean_corr" = real_snewt_cor)
  var_cor_table <- rbind(var_cor_table, var_cor_table_temp)

  plot <- ggplot(var_cor_table) +
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=as.numeric(var), y=as.numeric(cor), color = pattern), shape=17, size=3)+
  geom_point(aes(x=as.numeric(var), y=as.numeric(mean_corr), color = pattern), size=3)+
  theme(plot.title = element_text(size=5)) +
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8))

  return(plot)
}

cor_trends_by_gen <- function(file, cfile, title, subt, x_label, x_label_cor, y_label, min_gen, max_gen, pattern){
#file=mu_a_lit
#cfile=mu_a_cor 
 # pattern="1.0e-10_0.05_1.0e-08_0.005_3"
 file$snake_newt_rep <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd,"_", file$rep)
  cfile$snake_newt_rep <- paste0(cfile$snake_mu_rate, "_", cfile$snake_mu_effect_sd, "_", cfile$newt_mu_rate, "_", cfile$newt_mu_effect_sd, "_", cfile$rep)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen & file$snake_newt_rep==pattern),]) +
    #geom_line(aes(generation, dif, group=rep, col="dif"))+
      ggtitle(title, subtitle = paste0(subt," pattern ", pattern, " From ", min_gen, " gen to ", max_gen, " gen"))+
    geom_line(aes(generation, Snake_mean_Pheno, group=rep, color="Snake_mean_Pheno"),)+
    geom_line(aes(generation, Newt_mean_Pheno, group=rep, color="Newt_mean_Pheno"))+
    #ylim(0,10)+
    #xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake_mean_Pheno" = "steelblue", "Newt_mean_Pheno" = "darkred"),
labels = c("Snake", "Newt"))+
    theme_bw() +theme(legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank(), axis.ticks.x = element_blank(), axis.text.x = element_blank(),
                      plot.margin=unit(c(1,1,+0.3,1), "cm"))

  cor_plot <- ggplot(cfile[which(cfile$gen>=min_gen & cfile$gen<=max_gen & cfile$snake_newt_rep==pattern),]) +
    geom_line(aes(generation, mean_newt_pheno_By_mean_snake_pheno, col="CA"))+
    #ggtitle(title, subtitle = subt) +
    ylim(-1.1, 1.1)+
    #xlab(x_label_cor) + ylab(y_label)+
    theme_bw() +theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank(),
                      plot.margin=unit(c(-0.5,1,1,1), "cm"))

  file<-file[which(file$gen>=min_gen & file$gen<=max_gen & file$snake_newt_rep==pattern),]
  cfile<-cfile[which(cfile$gen>=min_gen & cfile$gen<=max_gen & cfile$snake_newt_rep==pattern),]

  if(nrow(cfile)>nrow(file)){
    toDelete <- seq(0, nrow(cfile), 2)
    cfile_short <-  cfile[-toDelete, ]
  }else{
    cfile_short <-  cfile
  }

  if(nrow(cfile_short)>nrow(file)){
      toDelete <- seq(0, nrow(cfile_short), 2)
      cfile_short <-  cfile_short[-toDelete, ]
    }
  figure <- ggarrange(plot, cor_plot + font("x.text", size = 10),heights = c(2, 1.5),ncol = 1, nrow = 2, align = "v")
    print(paste0("pattern ", pattern))
  print(paste0("Cor between average snake pheno and local cor ", cor(file$Snake_mean_Pheno, cfile_short$mean_newt_pheno_By_mean_snake_pheno)))
  print(paste0("Cor between average newt pheno and local cor ", cor(file$Newt_mean_Pheno, cfile_short$mean_newt_pheno_By_mean_snake_pheno)))
  print(paste0("Cor between average dif pheno and local cor ", cor(file$dif, cfile_short$mean_newt_pheno_By_mean_snake_pheno)))
  print(paste0("Cor between newt pheno and snake ", cor(file$Newt_mean_Pheno, file$Snake_mean_Pheno)))
  return(figure)
}

four_types_boxplot_cors_by_gen <- function(file,title, subt, x_label, y_label, min_gen, max_gen){
  file$snake_newt <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  plot <- ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ylim(-1.1, 1.1)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor("CA"), y=as.numeric(mean_newt_pheno_By_mean_snake_pheno), color = "CA"))+
  geom_boxplot(aes(x=factor("CB"), y=as.numeric(num_newts_By_num_snakes), color = "CB"))+
  geom_boxplot(aes(x=factor("CC"), y=as.numeric(sum_newt_pheno_By_num_snake), color = "CC"))+
  geom_boxplot(aes(x=factor("CD"), y=as.numeric(sum_snake_pheno_By_num_newt), color = "CD"))+
  geom_hline(yintercept=0, linetype="solid")+
  theme(plot.title = element_text(size=5)) +
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

  return(plot)
}

```

```{r, general file and graph making}

#the functions to make the params file and the summary and heatmap figures

create_parm_file <- function(file, min_gen, max_gen){
params <- data_frame()

for(i in 1:length(unique(file$snake_newt_rep))){

  pattern=unique(file$snake_newt_rep)[i]
  sumitup <- file[which(file$gen>=min_gen & file$gen<=max_gen & file$snake_newt_rep ==pattern),]

  params_temp<- data_frame(Newt_mean_Pheno=mean(sumitup$Newt_mean_Pheno),
                           Snake_mean_Pheno=mean(sumitup$Snake_mean_Pheno), 
                           Newt_mean_Pheno_sd=sd(sumitup$Newt_mean_Pheno),
                           Snake_mean_Pheno_sd=sd(sumitup$Snake_mean_Pheno),
                           Newt_pop_size=mean(sumitup$Newt_pop_size),
                           Snake_pop_size=mean(sumitup$Snake_pop_size),
                           dif=mean(sumitup$dif),
                           snake_mu_rate=sumitup$snake_mu_rate[1],
                           newt_mu_rate=sumitup$newt_mu_rate[1],
                           snake_mu_effect_sd=sumitup$snake_mu_effect_sd[1],
                           newt_mu_effect_sd=sumitup$newt_mu_effect_sd[1],
                           rep=sumitup$rep[1])

  params <- rbind(params, params_temp)
}

params$snake <- paste0(params$snake_mu_rate, "_",params$snake_mu_effect_sd)
params$newt <- paste0(params$newt_mu_rate, "_",params$newt_mu_effect_sd)
params$snake_newt <- paste0(params$snake, "_", params$newt)
return(params)
}

 create_parm_file_nc <- function(file, min_gen, max_gen){

   file$test_factor <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$rep)

params <- data_frame()

for(i in 1:length(unique(file$test_factor))){

  pattern=unique(file$test_factor)[i]

  sumitup <- file[which(file$gen>=min_gen & file$gen<=max_gen & file$test_factor ==pattern),]

  params_temp<- data_frame(Newt_mean_Pheno=mean(sumitup$Newt_mean_Pheno),
                           Snake_mean_Pheno=mean(sumitup$Snake_mean_Pheno), 
                           Newt_mean_Pheno_sd=sd(sumitup$Newt_mean_Pheno),
                           Snake_mean_Pheno_sd=sd(sumitup$Snake_mean_Pheno),
                           Newt_pop_size=mean(sumitup$Newt_pop_size),
                           Snake_pop_size=mean(sumitup$Snake_pop_size),
                           dif=mean(sumitup$dif),
                           snake_mu_rate=unlist(strsplit(pattern, "_"))[2],
                           newt_mu_rate=unlist(strsplit(pattern, "_"))[4],
                           snake_mu_effect_sd=unlist(strsplit(pattern, "_"))[3],
                           newt_mu_effect_sd=unlist(strsplit(pattern, "_"))[5],
                           rep=unlist(strsplit(pattern, "_"))[6],
                           test_factor=pattern)

  params <- rbind(params, params_temp)
}

params$snake <- paste0(params$snake_mu_rate, "_",params$snake_mu_effect_sd)
params$newt <- paste0(params$newt_mu_rate, "_",params$newt_mu_effect_sd)
params$snake_newt <- paste0(params$snake, "_", params$newt)
return(params)
}

GA_summary <- function(file, min_gen, max_gen, x_line, y_line, title, subt, y_label, x_label ){

file$snake_newt_rep <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$rep)

params<-create_parm_file(file, min_gen, max_gen)
params = na.omit(params)
my_plot <- ggplot(params) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=as.numeric(get(x_line)), y=as.numeric(get(y_line)), color = factor(snake), shape=factor(newt)),size=3)+
#  geom_point(aes(x=as.numeric(Newt_pop_size_20), y=as.numeric(Snake_pop_size_20), color = factor(snake), shape=factor(newt)))+
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  #facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text( hjust=1, size = 8))

return(my_plot)
}

GA_summary_by_species <- function(file, min_gen, max_gen, x_line, y_line, title, subt, y_label, x_label){

file$snake_newt_rep <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$rep)

params<-create_parm_file(file, min_gen, max_gen)
params = na.omit(params)
my_plot <- ggplot(params) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=(get(x_line)), y=(get(y_line)), color = factor(snake), shape=factor(newt)),size=3)+
#  geom_point(aes(x=as.numeric(Newt_pop_size_20), y=as.numeric(Snake_pop_size_20), color = factor(snake), shape=factor(newt)))+
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  #facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text( hjust=1, size = 8))

return(my_plot)
}

build_heatmap <- function(file, min_gen, max_gen, fill_line, fill_line_text, title, subt, y_label, x_label, red_blue=F){
  file$snake_newt_rep <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$rep)
  params<-create_parm_file(file, min_gen, max_gen)
  params = na.omit(params)

 my_plot <- ggplot(params, aes(newt, paste0(snake, "_", rep))) +    # Create heatmap with ggplot2
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_tile(aes(fill = get(fill_line)))+theme(axis.text.x = element_text(angle = 45, hjust=1, size = 10), axis.text.y = element_text(size = 10))

 if(red_blue==FALSE){
  my_plot<- my_plot + scale_fill_viridis_c(name = fill_line_text) + theme(legend.text=element_text(angle = 45, hjust=1, size=10))
  }
 if(red_blue==TRUE){
  my_plot<- my_plot + scale_fill_gradient2(low = "red",mid = "beige", high = "blue", name = fill_line_text) + theme(legend.text=element_text(size=5))
 }

  return(my_plot)
}

trends_by_gen <- function(file, title, subt, x_label, y_label, min_gen, max_gen, rep){

  plot <- ggplot(file[ which(file$generation>=min_gen & file$generation<=max_gen),]) +
    facet_wrap(~sim)+
    geom_line(aes(generation, dif, group=rep, col="dif"))+
    ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
    geom_line(aes(generation, Snake_mean_Pheno, group=rep, color="Snake_mean_Pheno"),)+
    geom_line(aes(generation, Newt_mean_Pheno, group=rep, color="Newt_mean_Pheno"))+
    geom_hline(yintercept=0, linetype="dashed", color="gray54")+
    #ylim(-1.,2.5)+
    xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake_mean_Pheno" = "steelblue", "Newt_mean_Pheno" = "darkred", "dif" = "black"),
labels = c("Snake", "Newt", "Dif"))+
    theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())
  return(plot)
}

whale_gen <- function(file, title, subt, x_label, y_label,min_gen, max_gen, rep){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)
dif <- ggplot(file[ which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep),]) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen, rep ",rep)) +
  xlab(x_label) + ylab(y_label)+
  geom_point(aes(x=as.numeric(Newt_pop_size), y=as.numeric(Snake_pop_size), color = dif))+
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      # strip.background = element_blank(), 
                      # strip.text.x = element_blank(), 
                      # panel.grid.major = element_blank(), 
                      # panel.grid.minor = element_blank(),
                      # panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

pop_size_hist <- function(file, title, subt, x_label, y_label,min_gen, max_gen, rep){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  file<-file[ which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep),]
  themin <- min(file$Snake_pop_size-file$Newt_pop_size) 
  themax <- max(file$Snake_pop_size-file$Newt_pop_size)
  true_max <- max(abs(themin), abs(themax))
dif <- ggplot(file) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(Snake_pop_size-Newt_pop_size), fill="darkolivegreen4")+
  xlim((-1*themax)-5, themax+5)+
  theme(plot.title = element_text(size=5)) +
  facet_wrap(~sim)+
  geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      # strip.background = element_blank(), 
                      # strip.text.x = element_blank(), 
                      # panel.grid.major = element_blank(), 
                      # panel.grid.minor = element_blank(),
                      # panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)
}

pheno_dif_hist <- function(file, title, subt, x_label, y_label,min_gen, max_gen, rep){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)
  file<-file[ which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep),]

  themin <- min(file$dif) 
  themax <- max(file$dif)
  true_max <- max(abs(themin), abs(themax))
dif <- ggplot(file) +
  #xlim(0, 1000)+
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen, rep ",rep)) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(dif), fill="aquamarine4")+
  xlim((-1*true_max)-1, true_max+1)+

  # y=as.numeric(Snake_pop_size-Newt_pop_size), color = snake_GA, shape = factor(newt_GA)
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      # strip.background = element_blank(), 
                      # strip.text.x = element_blank(), 
                      # panel.grid.major = element_blank(), 
                      # panel.grid.minor = element_blank(),
                      # panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

two_hist <- function(file, title, subt, x_label, y_label,min_gen, max_gen, rep){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)

  file<-file[ which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep),]

  thepopmin <- min(file$Snake_pop_size-file$Newt_pop_size)/100 
  thepopmax <- max(file$Snake_pop_size-file$Newt_pop_size)/100
  thedifmin <- min(file$dif) 
  thedifmax <- max(file$dif)
  true_max <- max(abs(thepopmin), abs(thepopmax),abs(thedifmin), abs(thedifmax))

  data_points <- nrow(file[which(file$generation>=min_gen & file$generation<=max_gen & file$rep==rep &  file$snake_GA==file$snake_GA[1] &  file$newt_GA==file$newt_GA[1]),])
dif <- ggplot(file) +

  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen, rep ",rep)) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=dif, y=..count../data_points), fill="#8B008B",  alpha = 0.65)+
  geom_histogram(aes(x=(Snake_pop_size-Newt_pop_size)/100, y=..count../data_points), fill="#008B00",  alpha = 0.65)+
  scale_y_continuous(labels = scales::percent)+
  xlim((-1*true_max)-1, true_max+1)+
  # y=as.numeric(Snake_pop_size-Newt_pop_size), color = snake_GA, shape = factor(newt_GA)
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      # strip.background = element_blank(), 
                      # strip.text.x = element_blank(), 
                      # panel.grid.major = element_blank(), 
                      # panel.grid.minor = element_blank(),
                      # panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

two_hist_all_rep <- function(file, title, subt, x_label, y_label,min_gen, max_gen){
  #file$rep==rep
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)

  file<-file[ which(file$generation>=min_gen & file$generation<=max_gen),]

  thepopmin <- min(file$Snake_pop_size-file$Newt_pop_size)/100 
  thepopmax <- max(file$Snake_pop_size-file$Newt_pop_size)/100
  thedifmin <- min(file$dif) 
  thedifmax <- max(file$dif)
  true_max <- max(abs(thepopmin), abs(thepopmax),abs(thedifmin), abs(thedifmax))

  data_points <- nrow(file[which(file$generation>=min_gen & file$generation<=max_gen  &  file$snake_GA==file$snake_GA[1] &  file$newt_GA==file$newt_GA[1]),])

dif <- ggplot(file) +

  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=dif, y=..count../data_points), fill="#8B008B",  alpha = 0.65, bins = 20)+ #purple
  geom_histogram(aes(x=(Snake_pop_size-Newt_pop_size)/100, y=..count../data_points), fill="#008B00",  alpha = 0.65, bins = 20)+ #green
  scale_y_continuous(labels = scales::percent)+
  xlim((-1*true_max)-1, true_max+1)+
  # y=as.numeric(Snake_pop_size-Newt_pop_size), color = snake_GA, shape = factor(newt_GA)
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      #strip.background = element_blank(), 
                      strip.text.x = element_blank(), 
                      #panel.grid.major = element_blank(), 
                      #panel.grid.minor = element_blank(),
                      #panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

two_hist_death_all_rep <- function(file, title, subt, x_label, y_label,min_gen, max_gen){
  #file=mu_a_lit
  file$snake_GA<-paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd)
  file$newt_GA<-paste0(file$newt_mu_rate, "_", file$newt_mu_effect_sd)

  file<-file[ which(file$generation>=min_gen & file$generation<=max_gen),]

  thepopmin <- min(file$newt_deaths)
  thepopmax <- max(file$newt_deaths)
  thedifmin <- min(file$snake_deaths) 
  thedifmax <- max(file$snake_deaths)
  true_max <- max(abs(thepopmin), abs(thepopmax),abs(thedifmin), abs(thedifmax))

  data_points <- nrow(file[which(file$generation>=min_gen & file$generation<=max_gen  &  file$snake_GA==file$snake_GA[1] &  file$newt_GA==file$newt_GA[1]),])

dif <- ggplot(file) +

  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab(x_label) + ylab(y_label)+
  geom_histogram(aes(x=newt_deaths), fill="red",  alpha = 0.65, bins = 20)+
  geom_histogram(aes(x=snake_deaths), fill="blue",  alpha = 0.65, bins = 20)+
  #scale_y_continuous(labels = scales::percent)+
  #xlim((-1*true_max)-1, true_max+1)+
  # y=as.numeric(Snake_pop_size-Newt_pop_size), color = snake_GA, shape = factor(newt_GA)
  #paste0(snake_mu_rate, "_", snake_mu_effect_sd), shape = paste0(newt_mu_rate, "_", newt_mu_effect_sd), group=T)
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  facet_wrap(~sim)+
  #geom_vline(xintercept = 0)+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      #strip.background = element_blank(), 
                      strip.text.x = element_blank(), 
                      #panel.grid.major = element_blank(), 
                      #panel.grid.minor = element_blank(),
                      #panel.background = element_blank(), 
                      axis.line = element_line(colour = "black"))

return(dif)

}

#two_hist_death_all_rep(mu_a_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=20000, max_gen=25000)

grid_plot<- function(file_name, spegen, file, type, row_num, col_num){

file$snake_newt_rep <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd,"_",file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$rep)

GA1<-file[which(file$snake_newt_rep==file_name & file$generation==spegen),]
GA1_type = paste0("generation|", type)

GA1_type = GA1[,grepl(GA1_type, names(GA1)) ]
GA1_type <- GA1_type[-1]

data_long_type_comp <- gather(GA1_type, condition, measurement, factor_key=TRUE)
data_long_type_comp[,c("species","type", "type2","location")]<- str_split_fixed(data_long_type_comp$condition, "_", 4)

data_long_type_comp$y_loc<- rep(sort(c(rep.int(row_num:1, 2)), decreasing = TRUE),col_num)

data_long_type_comp$x_loc<- sort(rep.int(1:col_num, row_num*2))

main_plot<-ggplot(data_long_type_comp)+

  geom_point(data = data_long_type_comp[which(data_long_type_comp$species=="snake"),], aes(x = x_loc, y = y_loc, fill = as.numeric(measurement)), size = 6, shape = 22, ) +
  geom_point(data = data_long_type_comp[which(data_long_type_comp$species=="newt"),], aes(x = x_loc, y = y_loc, fill = as.numeric(measurement)), size = 4, shape = 21, ) +
  scale_fill_viridis_c()+
  #ylim(0, row_num)+
  #xlim(0, col_num)+
  theme_bw()+
  theme(legend.justification = "top")

if(type != "num_ind"){
main_plot<- main_plot + labs(title = paste0(str_to_title(type)," Phenotype of Locations"),
              subtitle = paste0(file_name, ", Generation: ", spegen),
              fill = "Toxicity/Resistance")+
 scale_fill_viridis_c()
}
if(type == "num_ind"){
main_plot<- main_plot + labs(title = paste0("Number of Individuals at Locations ", str_to_title(type)),
              subtitle = paste0(file_name, ", Generation: ", spegen),
              fill = "Newt/Snake PopSize")+
scale_fill_viridis_c(option = "plasma")
}

snake<-data_long_type_comp[which(data_long_type_comp$species=="snake"),]
newt<-data_long_type_comp[which(data_long_type_comp$species=="newt"),]
newt$snake_value <- snake$measurement
print(cor(as.numeric(snake$measurement), as.numeric(newt$measurement), use="complete.obs"))

newt$info <- "middle"

#need to look into, coloring problem
newt$info[newt$location %in% c(0, row_num-1, col_num*row_num-row_num, col_num*row_num-1)] <- "corner"

newt$info[newt$location %in% c(c(1:(row_num-2)), seq(row_num, (col_num*row_num-row_num-2), row_num), seq((row_num-1+row_num), (col_num*row_num-1-row_num), row_num),  c((col_num*row_num-row_num+1):(col_num*row_num-2)))]   <- "edge"

extra_plot <- ggplot(newt)+
  ggtitle("Mean Phenotype") +
  xlab("Newt") + ylab("Snake")+
  geom_point(aes(x = as.numeric(measurement), y = as.numeric(snake_value), color=info), size = 1 ) +
  theme_bw()+
   theme(text = element_text(size=10),title =element_text(size=8),
         legend.text = element_text(size = 8),
         legend.position ="bottom",
         legend.key.size = unit(0.1, "lines"),
         legend.title=element_blank())+
  guides(shape = guide_legend(override.aes = list(size = 1)),
         color = guide_legend(override.aes = list(size = 1)))

final_plot <- main_plot +  inset_element(extra_plot, left = 1., bottom = -0.10, right = 1.35, top = 0.5)
return(final_plot)}

```

## Mean Phenotype Whole Simulation {.tabset}

### Phenotype differences (mu_a)
```{r, graphing by gen}
trends_by_gen(mu_a_lit, title="Space Tall: Newt VS Snake Population Size GA1", subt="Gen 16sims", y_label="Mean Phenotype", x_label="Generations", 100, 50000, rep=1)

#aggregate(mu_a_lit[which(mu_a_lit$generation>=100 & mu_a_lit$generation<=50000),]$dif, list(mu_a_lit[which(mu_a_lit$generation>=100 & mu_a_lit$generation<=50000),]$group_rep), FUN=mean)

```

### Table of average Differences (mu_a)

```{r, average difference}
aggregate(mu_a_lit[which(mu_a_lit$generation>=100 & mu_a_lit$generation<=50000),]$dif, list(mu_a_lit[which(mu_a_lit$generation>=100 & mu_a_lit$generation<=50000),]$group), FUN=mean) 
```

### Phenotype differences (mu_b)
```{r, graphing by gen lw}
trends_by_gen(mu_b_lit, title="Space Tall: Newt VS Snake Population Size GA1", subt="Gen 16sims", y_label="Mean Phenotype", x_label="Generations", 100, 50000, rep=1)

```

### Table of average Differences (mu_b)

```{r, average difference lw}
aggregate(mu_b_lit[which(mu_b_lit$generation>=100 & mu_b_lit$generation<=50000),]$dif, list(mu_b_lit[which(mu_b_lit$generation>=100 & mu_b_lit$generation<=50000),]$group), FUN=mean) 
```

### Phenotype differences (mu_c)
```{r, graphing by gen ne}
trends_by_gen(mu_c_lit, title="Space Tall: Newt VS Snake Population Size GA1", subt="Gen 16sims", y_label="Mean Phenotype", x_label="Generations", 100, 50000, rep=1)

```

### Table of average Differences (mu_c)

```{r, average difference ne}
aggregate(mu_c_lit[which(mu_c_lit$generation>=100 & mu_c_lit$generation<=50000),]$dif, list(mu_c_lit[which(mu_c_lit$generation>=100 & mu_c_lit$generation<=50000),]$group), FUN=mean) 
```

### Phenotype differences (mu_d)
```{r, graphing by gen mud}
trends_by_gen(mu_d_lit, title="Space Tall: Newt VS Snake Population Size GA1", subt="Gen 16sims", y_label="Mean Phenotype", x_label="Generations", 100, 50000, rep=1)

```

### Table of average Differences (mu_d)

```{r, average difference mud}
aggregate(mu_d_lit[which(mu_d_lit$generation>=100 & mu_d_lit$generation<=50000),]$dif, list(mu_d_lit[which(mu_d_lit$generation>=100 & mu_d_lit$generation<=50000),]$group), FUN=mean) 
```

## {-}

## Connection between higher phenotype and population{.tabset}

### Phenotype differences (mu_a)
```{r, whale}
min_gen=5000
max_gen=10000
rep=1

whale_gen(mu_a_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Phenotype", min_gen=min_gen, max_gen=max_gen, rep=rep )
```

### Phenotype & Populationsize differences (mu_a)
```{r, popsize and pheno}

suppressMessages(print(two_hist(mu_a_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Popsize/1000 (green) & Snake-Newt Phenotype (purple)",  min_gen=min_gen, max_gen=max_gen, rep=rep )))

```

### Phenotype differences (mu_b)
```{r, whale lw}
min_gen=5000
max_gen=10000
rep=1

whale_gen(mu_b_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Phenotype", min_gen=min_gen, max_gen=max_gen, rep=rep )
```

### Phenotype & Populationsize differences (mu_b)
```{r, popsize and pheno lw}

suppressMessages(print(two_hist(mu_b_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Popsize/1000 (green) & Snake-Newt Phenotype (purple)",  min_gen=min_gen, max_gen=max_gen, rep=rep )))

```

### Phenotype differences (mu_c)
```{r, whale ne}
min_gen=5000
max_gen=10000
rep=1

whale_gen(mu_c_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Phenotype", min_gen=min_gen, max_gen=max_gen, rep=rep )
```

### Phenotype & Populationsize differences (mu_c)
```{r, popsize and pheno ne}

suppressMessages(print(two_hist(mu_c_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Popsize/1000 (green) & Snake-Newt Phenotype (purple)",  min_gen=min_gen, max_gen=max_gen, rep=rep )))

```

### Phenotype differences (mu_d)
```{r, whale mud}
min_gen=5000
max_gen=10000
rep=1

whale_gen(mu_d_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Phenotype", min_gen=min_gen, max_gen=max_gen, rep=rep )
```

### Phenotype & Populationsize differences (mu_d)
```{r, popsize and pheno mud}

suppressMessages(print(two_hist(mu_d_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Popsize/1000 (green) & Snake-Newt Phenotype (purple)",  min_gen=min_gen, max_gen=max_gen, rep=rep )))

```

## {-}

## Correlation{.tabset}

### cor (mu_a)
```{r, combo myplots}

ginmin = 10000
ginmax = 15000

boxplot_cors_by_gen(mu_a_cor, "Simulation Correlation Comparied to Real Correlation Trials", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)

```

### cor (mu_b)
```{r, combo myplots lw}

ginmin = 10000
ginmax = 15000

boxplot_cors_by_gen(mu_b_cor, "Simulation Correlation Comparied to Real Correlation Trials", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)

```

### cor (mu_c)
```{r, combo myplots ne}

ginmin = 10000
ginmax = 15000

boxplot_cors_by_gen(mu_c_cor, "Simulation Correlation Comparied to Real Correlation Trials", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)

```

### cor (mu_d)
```{r, combo myplots mud}

ginmin = 10000
ginmax = 15000

boxplot_cors_by_gen(mu_d_cor, "Simulation Correlation Comparied to Real Correlation Trials", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)

```

## {-}

## Correlation Histograms Mu A{.tabset}

### Plot 1
```{r, GA1 cor by dif windows}
ginmin = 0
ginmax = 5000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_a_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 2

```{r, GA1 cor by dif windows 2}
ginmin = 5000
ginmax = 10000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_a_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 3
```{r, GA1 cor by dif windows 3}
ginmin = 10000
ginmax = 15000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_a_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 4
```{r, GA1 cor by dif windows 4}
ginmin = 15000
ginmax = 20000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_a_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 5
```{r, GA1 cor by dif windows 5}
ginmin = 20000
ginmax = 25000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_a_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 6
```{r, GA1 cor by dif windows 6}
ginmin = 25000
ginmax = 30000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_a_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 7
```{r, GA1 cor by dif windows 7}
ginmin = 30000
ginmax = 35000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_a_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 8
```{r, GA1 cor by dif windows 8}
ginmin = 35000
ginmax = 40000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_a_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 9
```{r, GA1 cor by dif windows 9}
ginmin = 40000
ginmax = 45000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_a_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 10
```{r, GA1 cor by dif windows 10}
ginmin = 45000
ginmax = 50000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_a_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

## {-}

## Correlation Histograms Mu B {.tabset}

### Plot 1
```{r, GA1 cor by dif windows lw}
ginmin = 0
ginmax = 5000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_b_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 2

```{r, GA1 cor by dif windows 2 lw}
ginmin = 5000
ginmax = 10000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_b_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 3
```{r, GA1 cor by dif windows 3 lw}
ginmin = 10000
ginmax = 15000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_b_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 4
```{r, GA1 cor by dif windows 4 lw}
ginmin = 15000
ginmax = 20000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_b_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 5
```{r, GA1 cor by dif windows 5 lw}
ginmin = 20000
ginmax = 25000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_b_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 6
```{r, GA1 cor by dif windows 6 lw}
ginmin = 25000
ginmax = 30000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_b_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 7
```{r, GA1 cor by dif windows 7 lw}
ginmin = 30000
ginmax = 35000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_b_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 8
```{r, GA1 cor by dif windows 8 lw}
ginmin = 35000
ginmax = 40000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_b_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 9
```{r, GA1 cor by dif windows 9 lw}
ginmin = 40000
ginmax = 45000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_b_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 10
```{r, GA1 cor by dif windows 10 lw}
ginmin = 45000
ginmax = 50000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_b_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

## {-}

## Correlation Histograms Mu C {.tabset}

### Plot 1
```{r, GA1 cor by dif windows ne}
ginmin = 0
ginmax = 5000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_c_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 2

```{r, GA1 cor by dif windows 2 ne}
ginmin = 5000
ginmax = 10000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_c_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 3
```{r, GA1 cor by dif windows 3 ne}
ginmin = 10000
ginmax = 15000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_c_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 4
```{r, GA1 cor by dif windows 4 ne}
ginmin = 15000
ginmax = 20000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_c_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 5
```{r, GA1 cor by dif windows 5 ne}
ginmin = 20000
ginmax = 25000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_c_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 6
```{r, GA1 cor by dif windows 6 ne}
ginmin = 25000
ginmax = 30000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_c_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 7
```{r, GA1 cor by dif windows 7 ne}
ginmin = 30000
ginmax = 35000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_c_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 8
```{r, GA1 cor by dif windows 8 ne}
ginmin = 35000
ginmax = 40000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_c_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 9
```{r, GA1 cor by dif windows 9 ne}
ginmin = 40000
ginmax = 45000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_c_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 10
```{r, GA1 cor by dif windows 10 ne}
ginmin = 45000
ginmax = 50000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_c_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

## {-}

## Correlation Histograms Mu D {.tabset}

### Plot 1
```{r, GA1 cor by dif windows mud}
ginmin = 0
ginmax = 5000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_d_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 2

```{r, GA1 cor by dif windows 2 mud}
ginmin = 5000
ginmax = 10000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_d_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 3
```{r, GA1 cor by dif windows 3 mud}
ginmin = 10000
ginmax = 15000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_d_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 4
```{r, GA1 cor by dif windows 4 mud}
ginmin = 15000
ginmax = 20000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_d_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 5
```{r, GA1 cor by dif windows 5 mud}
ginmin = 20000
ginmax = 25000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_d_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 6
```{r, GA1 cor by dif windows 6 mud}
ginmin = 25000
ginmax = 30000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_d_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 7
```{r, GA1 cor by dif windows 7 mud}
ginmin = 30000
ginmax = 35000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_d_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 8
```{r, GA1 cor by dif windows 8 mud}
ginmin = 35000
ginmax = 40000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_d_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 9
```{r, GA1 cor by dif windows 9 mud}
ginmin = 40000
ginmax = 45000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_d_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

### Plot 10
```{r, GA1 cor by dif windows 10 mud}
ginmin = 45000
ginmax = 50000
suppressMessages(print(hist_cors_by_gen_col_rep(mu_d_cor, "Simulation Correlation Comparied to Real Correlation", "mu = NUM", "newt/snake GAs&rep", "Cor", ginmin, ginmax)))
```

## {-}

## Correlation across time (mu_a)  {.tabset}

### Random 1 (mu_a)  
```{r, GA1 cor and pheno}
GA1_pattern <- unique(paste0(mu_a_lit$snake_mu_rate, "_", mu_a_lit$snake_mu_effect_sd,"_",mu_a_lit$newt_mu_rate, "_", mu_a_lit$newt_mu_effect_sd, "_", mu_a_lit$rep))

myrandomlist=sample.int(length(GA1_pattern), 3)

# for(i in myrandomlist){
#   myine<-cor_trends_by_gen(mu_a_lit, mu_a_cor, "test", "mu = NUM", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[i])
#   plot(myine)
# }

cor_trends_by_gen(mu_a_lit, mu_a_cor, "test", "mu = NUM", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[myrandomlist[1]])

```

### Random 2  (mu_a)
```{r, random cor-pheno plot 2}

cor_trends_by_gen(mu_a_lit, mu_a_cor, "test", "mu = NUM", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[myrandomlist[2]])
```

### Random 3  (mu_a)
```{r, random cor-pheno plot 3}

cor_trends_by_gen(mu_a_lit, mu_a_cor, "test", "mu = NUM", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[myrandomlist[3]])
```

## {-}

## Correlation across time (mu_b)  {.tabset}

### Random 1 (mu_b)  
```{r, GA1 cor and pheno  lw}
GA1_pattern <- unique(paste0(mu_b_lit$snake_mu_rate, "_", mu_b_lit$snake_mu_effect_sd,"_",mu_b_lit$newt_mu_rate, "_", mu_b_lit$newt_mu_effect_sd, "_", mu_b_lit$rep))

myrandomlist=sample.int(length(GA1_pattern), 3)

# for(i in myrandomlist){
#   myine<-cor_trends_by_gen(mu_b_lit, mu_b_cor, "test", "mu = NUM", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[i])
#   plot(myine)
# }

cor_trends_by_gen(mu_b_lit, mu_b_cor, "test", "mu = NUM", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[myrandomlist[1]])

```

### Random 2  (mu_b)
```{r, random cor-pheno plot 2  lw}

cor_trends_by_gen(mu_b_lit, mu_b_cor, "test", "mu = NUM", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[myrandomlist[2]])
```

### Random 3  (mu_b)
```{r, random cor-pheno plot 3  lw}

cor_trends_by_gen(mu_b_lit, mu_b_cor, "test", "mu = NUM", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[myrandomlist[3]])
```

## {-}

## Correlation across time (mu_c)  {.tabset}

### Random 1 (mu_c)  
```{r, GA1 cor and pheno ne}
GA1_pattern <- unique(paste0(mu_c_lit$snake_mu_rate, "_", mu_c_lit$snake_mu_effect_sd,"_",mu_c_lit$newt_mu_rate, "_", mu_c_lit$newt_mu_effect_sd, "_", mu_c_lit$rep))

myrandomlist=sample.int(length(GA1_pattern), 3)

# for(i in myrandomlist){
#   myine<-cor_trends_by_gen(mu_c_lit, mu_c_cor, "test", "mu = NUM", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[i])
#   plot(myine)
# }

cor_trends_by_gen(mu_c_lit, mu_c_cor, "test", "mu = NUM", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[myrandomlist[1]])

```

### Random 2  (mu_c)
```{r, random cor-pheno plot 2 ne}

cor_trends_by_gen(mu_c_lit, mu_c_cor, "test", "mu = NUM", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[myrandomlist[2]])
```

### Random 3  (mu_c)
```{r, random cor-pheno plot 3 ne}

cor_trends_by_gen(mu_c_lit, mu_c_cor, "test", "mu = NUM", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[myrandomlist[3]])
```

## {-}

## Correlation across time (mu_d)  {.tabset}

### Random 1 (mu_d)  
```{r, GA1 cor and pheno  mud}
GA1_pattern <- unique(paste0(mu_d_lit$snake_mu_rate, "_", mu_d_lit$snake_mu_effect_sd,"_",mu_d_lit$newt_mu_rate, "_", mu_d_lit$newt_mu_effect_sd, "_", mu_d_lit$rep))

myrandomlist=sample.int(length(GA1_pattern), 3)

# for(i in myrandomlist){
#   myine<-cor_trends_by_gen(mu_d_lit, mu_d_cor, "test", "mu = NUM", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[i])
#   plot(myine)
# }

cor_trends_by_gen(mu_d_lit, mu_d_cor, "test", "mu = NUM", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[myrandomlist[1]])

```

### Random 2  (mu_d)
```{r, random cor-pheno plot 2  mud}

cor_trends_by_gen(mu_d_lit, mu_d_cor, "test", "mu = NUM", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[myrandomlist[2]])
```

### Random 3  (mu_d)
```{r, random cor-pheno plot 3  mud}

cor_trends_by_gen(mu_d_lit, mu_d_cor, "test", "mu = NUM", "newt/snake GAs&rep","other", "Cor", min_gen=100, max_gen=50000, pattern=GA1_pattern[myrandomlist[3]])
```

## {-}

## What happens over time (looking at the beginning, middle, and late part of my simulations) (mu_a) {.tabset}

### Pheno Beginning (mu_a)

```{r, early plots}

min_gen=100
max_gen=5000

# for(i in 1:length(unique(mu_a_lit$rep))){
#   myine<-whale_gen(mu_a_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_a_lit$rep)[i])
#   plot(myine)
# 
# }
whale_gen(mu_a_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_a_lit$rep)[1])
```

### Pheno Middle (mu_a)

```{r, middle plots}

min_gen=25000
max_gen=30000

# for(i in 1:length(unique(mu_a_lit$rep))){
#   myine<-whale_gen(mu_a_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_a_lit$rep)[i])
#   plot(myine)
# 
# }
whale_gen(mu_a_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_a_lit$rep)[1])
```

### Pheno End (mu_a)

```{r, end plots}

min_gen=45000
max_gen=50000

# for(i in 1:length(unique(mu_a_lit$rep))){
#   myine<-whale_gen(mu_a_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_a_lit$rep)[i])
#   plot(myine)
# 
# }
whale_gen(mu_a_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_a_lit$rep)[1])
```

### Dif Beginning (mu_a)
```{r, early plots 1a}

min_gen=100
max_gen=5000

# for(i in 1:length(unique(mu_a_lit$rep))){
#   myine<-two_hist(mu_a_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_a_lit$rep)[i] )
#   suppressMessages(plot(myine))
# }

two_hist(mu_a_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_a_lit$rep)[1] )

```

### Dif Middle (mu_a)
```{r, early plots 2a}

min_gen=25000
max_gen=30000

# for(i in 1:length(unique(mu_a_lit$rep))){
#   myine<-two_hist(mu_a_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_a_lit$rep)[i] )
#   suppressMessages(plot(myine))
# }

two_hist(mu_a_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_a_lit$rep)[1] )

```

### Dif End (mu_a)
```{r, early plots 3a}

min_gen=45000
max_gen=50000

# for(i in 1:length(unique(mu_a_lit$rep))){
#   myine<-two_hist(mu_a_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_a_lit$rep)[i] )
#   suppressMessages(plot(myine))
# }

two_hist(mu_a_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_a_lit$rep)[1] )

```

## {-}

## What happens over time (looking at the beginning, middle, and late part of my simulations) (mu_b) {.tabset}

### Pheno Beginning (mu_b)

```{r, early plots lw}

min_gen=100
max_gen=5000

# for(i in 1:length(unique(mu_b_lit$rep))){
#   myine<-whale_gen(mu_b_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_b_lit$rep)[i])
#   plot(myine)
# 
# }
whale_gen(mu_b_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_b_lit$rep)[1])
```

### Pheno Middle (mu_b)

```{r, middle plots lw}

min_gen=25000
max_gen=30000

# for(i in 1:length(unique(mu_b_lit$rep))){
#   myine<-whale_gen(mu_b_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_b_lit$rep)[i])
#   plot(myine)
# 
# }
whale_gen(mu_b_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_b_lit$rep)[1])
```

### Pheno End (mu_b)

```{r, end plots lw}

min_gen=45000
max_gen=50000

# for(i in 1:length(unique(mu_b_lit$rep))){
#   myine<-whale_gen(mu_b_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_b_lit$rep)[i])
#   plot(myine)
# 
# }
whale_gen(mu_b_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_b_lit$rep)[1])
```

### Dif Beginning (mu_b)
```{r, early plots 1a lw}

min_gen=100
max_gen=5000

# for(i in 1:length(unique(mu_b_lit$rep))){
#   myine<-two_hist(mu_b_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_b_lit$rep)[i] )
#   suppressMessages(plot(myine))
# }

two_hist(mu_b_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_b_lit$rep)[1] )

```

### Dif Middle (mu_b)
```{r, early plots 2a lw}

min_gen=25000
max_gen=30000

# for(i in 1:length(unique(mu_b_lit$rep))){
#   myine<-two_hist(mu_b_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_b_lit$rep)[i] )
#   suppressMessages(plot(myine))
# }

two_hist(mu_b_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_b_lit$rep)[1] )

```

### Dif End (mu_b)
```{r, early plots 3a lw}

min_gen=45000
max_gen=50000

# for(i in 1:length(unique(mu_b_lit$rep))){
#   myine<-two_hist(mu_b_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_b_lit$rep)[i] )
#   suppressMessages(plot(myine))
# }

two_hist(mu_b_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_b_lit$rep)[1] )

```

## {-}

## What happens over time (looking at the beginning, middle, and late part of my simulations) (mu_c) {.tabset}

### Pheno Beginning (mu_c)

```{r, early plots ne}

min_gen=100
max_gen=5000

# for(i in 1:length(unique(mu_c_lit$rep))){
#   myine<-whale_gen(mu_c_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_c_lit$rep)[i])
#   plot(myine)
# 
# }
whale_gen(mu_c_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_c_lit$rep)[1])
```

### Pheno Middle (mu_c)

```{r, middle plots ne}

min_gen=25000
max_gen=30000

# for(i in 1:length(unique(mu_c_lit$rep))){
#   myine<-whale_gen(mu_c_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_c_lit$rep)[i])
#   plot(myine)
# 
# }
whale_gen(mu_c_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_c_lit$rep)[1])
```

### Pheno End (mu_c)

```{r, end plots ne}

min_gen=45000
max_gen=50000

# for(i in 1:length(unique(mu_c_lit$rep))){
#   myine<-whale_gen(mu_c_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_c_lit$rep)[i])
#   plot(myine)
# 
# }
whale_gen(mu_c_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_c_lit$rep)[1])
```

### Dif Beginning (mu_c)
```{r, early plots 1a ne}

min_gen=100
max_gen=5000

# for(i in 1:length(unique(mu_c_lit$rep))){
#   myine<-two_hist(mu_c_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_c_lit$rep)[i] )
#   suppressMessages(plot(myine))
# }

two_hist(mu_c_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_c_lit$rep)[1] )

```

### Dif Middle (mu_c)
```{r, early plots 2a ne}

min_gen=25000
max_gen=30000

# for(i in 1:length(unique(mu_c_lit$rep))){
#   myine<-two_hist(mu_c_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_c_lit$rep)[i] )
#   suppressMessages(plot(myine))
# }

two_hist(mu_c_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_c_lit$rep)[1] )

```

### Dif End (mu_c)
```{r, early plots 3a ne}

min_gen=45000
max_gen=50000

# for(i in 1:length(unique(mu_c_lit$rep))){
#   myine<-two_hist(mu_c_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_c_lit$rep)[i] )
#   suppressMessages(plot(myine))
# }

two_hist(mu_c_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_c_lit$rep)[1] )

```

## {-}

## What happens over time (looking at the beginning, middle, and late part of my simulations) (mu_d) {.tabset}

### Pheno Beginning (mu_d)

```{r, early plots mud}

min_gen=100
max_gen=5000

# for(i in 1:length(unique(mu_d_lit$rep))){
#   myine<-whale_gen(mu_d_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_d_lit$rep)[i])
#   plot(myine)
# 
# }
whale_gen(mu_d_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_d_lit$rep)[1])
```

### Pheno Middle (mu_d)

```{r, middle plots mud}

min_gen=25000
max_gen=30000

# for(i in 1:length(unique(mu_d_lit$rep))){
#   myine<-whale_gen(mu_d_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_d_lit$rep)[i])
#   plot(myine)
# 
# }
whale_gen(mu_d_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_d_lit$rep)[1])
```

### Pheno End (mu_d)

```{r, end plots mud}

min_gen=45000
max_gen=50000

# for(i in 1:length(unique(mu_d_lit$rep))){
#   myine<-whale_gen(mu_d_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_d_lit$rep)[i])
#   plot(myine)
# 
# }
whale_gen(mu_d_lit, title="Space: Newt VS Snake Population Size Color=S-N Phenotype", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize", min_gen=min_gen, max_gen=max_gen, rep=unique(mu_d_lit$rep)[1])
```

### Dif Beginning (mu_d)
```{r, early plots 1a mud}

min_gen=100
max_gen=5000

# for(i in 1:length(unique(mu_d_lit$rep))){
#   myine<-two_hist(mu_d_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_d_lit$rep)[i] )
#   suppressMessages(plot(myine))
# }

two_hist(mu_d_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_d_lit$rep)[1] )

```

### Dif Middle (mu_d)
```{r, early plots 2a mud}

min_gen=25000
max_gen=30000

# for(i in 1:length(unique(mu_d_lit$rep))){
#   myine<-two_hist(mu_d_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_d_lit$rep)[i] )
#   suppressMessages(plot(myine))
# }

two_hist(mu_d_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_d_lit$rep)[1] )

```

### Dif End (mu_d)
```{r, early plots 3a mud}

min_gen=45000
max_gen=50000

# for(i in 1:length(unique(mu_d_lit$rep))){
#   myine<-two_hist(mu_d_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_d_lit$rep)[i] )
#   suppressMessages(plot(myine))
# }

two_hist(mu_d_lit, title="Snake-Newt Winning Hist", subt="Gen 16sims", y_label="Count", x_label="Snake-Newt Phenotype (purple) & Snake-Newt Popsize/100 (green)",  min_gen=min_gen, max_gen=max_gen, rep=unique(mu_d_lit$rep)[1] )

```

## {-}

## Summary (mu_a) {.tabset}

### Early-Sim Population Size Summary

```{r, early plots summary}

min_gen=100
max_gen=5000
GA_summary(mu_a_lit, min_gen=min_gen, max_gen=max_gen, x_line="Newt_pop_size", y_line="Snake_pop_size", title="Space: Newt VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize")
```

### Mid-Sim Population Size Summary
```{r, plots summary 2}
# summary

min_gen=25000
max_gen=30000

# summary
GA_summary(mu_a_lit, min_gen=min_gen, max_gen=max_gen, x_line="Newt_pop_size", y_line="Snake_pop_size", title="Space: Newt VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize")
```

### Late-Sim Population Size Summary
```{r, plots summary 3}
min_gen=45000
max_gen=50000

# summary
GA_summary(mu_a_lit, min_gen=min_gen, max_gen=max_gen, x_line="Newt_pop_size", y_line="Snake_pop_size", title="Space: Newt VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize")

```

### Early Difference Summary
```{r, dif plots summary 1}

min_gen=100
max_gen=5000

GA_summary(mu_a_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="Snake_pop_size", title="Space: Dif VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Dif")
```

### Mid Difference Summary
```{r, dif plots summary 2}
min_gen=25000
max_gen=30000

GA_summary(mu_a_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="Snake_pop_size", title="Space: Dif VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Dif")

```

### Late Difference Summary
```{r, dif plots summary 3}
min_gen=45000
max_gen=50000
GA_summary(mu_a_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="Snake_pop_size", title="Space: Dif VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Dif")

```

### By Snake GA (Early)
```{r, snakega plots summary 1}
min_gen=100
max_gen=5000
#

GA_summary_by_species(mu_a_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="snake", title="Space: Dif VS Snake GA", subt="Gen 16sims", y_label="Snake GA", x_label="Dif")

```

### By Snake GA (Mid)
```{r, snakega plots summary 2}
min_gen=25000
max_gen=30000
#

GA_summary_by_species(mu_a_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="snake", title="Space: Dif VS Snake GA", subt="Gen 16sims", y_label="Snake GA", x_label="Dif")

```

### By Snake GA (Late)
```{r, snakega plots summary 3}
min_gen=45000
max_gen=50000
#

GA_summary_by_species(mu_a_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="snake", title="Space: Dif VS Snake GA", subt="Gen 16sims", y_label="Snake GA", x_label="Dif")

```

### By Newt GA (Early)
```{r, newtga plots summary 1}
min_gen=100
max_gen=5000
#

GA_summary_by_species(mu_a_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="newt", title="Space: Dif VS Newt GA", subt="Gen 16sims", y_label="Newt GA", x_label="Dif")

```

### By Newt GA (Mid)
```{r, newtga plots summary 2}
min_gen=25000
max_gen=30000
#

GA_summary_by_species(mu_a_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="newt", title="Space: Dif VS Newt GA", subt="Gen 16sims", y_label="Newt GA", x_label="Dif")

```

### By Newt GA (Late)
```{r, newtga plots summary 3}
min_gen=45000
max_gen=50000
#

GA_summary_by_species(mu_a_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="newt", title="Space: Dif VS Newt GA", subt="Gen 16sims", y_label="Newt GA", x_label="Dif")

```

## {-}

## Summary (mu_b) {.tabset}

### Early-Sim Population Size Summary

```{r, early plots summary lw}

min_gen=100
max_gen=5000
GA_summary(mu_b_lit, min_gen=min_gen, max_gen=max_gen, x_line="Newt_pop_size", y_line="Snake_pop_size", title="Space: Newt VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize")
```

### Mid-Sim Population Size Summary
```{r, plots summary 2 lw}
# summary

min_gen=25000
max_gen=30000

# summary
GA_summary(mu_b_lit, min_gen=min_gen, max_gen=max_gen, x_line="Newt_pop_size", y_line="Snake_pop_size", title="Space: Newt VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize")
```

### Late-Sim Population Size Summary
```{r, plots summary 3 lw}
min_gen=45000
max_gen=50000

# summary
GA_summary(mu_b_lit, min_gen=min_gen, max_gen=max_gen, x_line="Newt_pop_size", y_line="Snake_pop_size", title="Space: Newt VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize")

```

### Early Difference Summary
```{r, dif plots summary 1 lw}

min_gen=100
max_gen=5000

GA_summary(mu_b_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="Snake_pop_size", title="Space: Dif VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Dif")
```

### Mid Difference Summary
```{r, dif plots summary 2 lw}
min_gen=25000
max_gen=30000

GA_summary(mu_b_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="Snake_pop_size", title="Space: Dif VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Dif")

```

### Late Difference Summary
```{r, dif plots summary 3 lw}
min_gen=45000
max_gen=50000
GA_summary(mu_b_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="Snake_pop_size", title="Space: Dif VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Dif")

```

### By Snake GA (Early)
```{r, snakega plots summary 1 lw}
min_gen=100
max_gen=5000
#

GA_summary_by_species(mu_b_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="snake", title="Space: Dif VS Snake GA", subt="Gen 16sims", y_label="Snake GA", x_label="Dif")

```

### By Snake GA (Mid)
```{r, snakega plots summary 2 lw}
min_gen=25000
max_gen=30000
#

GA_summary_by_species(mu_b_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="snake", title="Space: Dif VS Snake GA", subt="Gen 16sims", y_label="Snake GA", x_label="Dif")

```

### By Snake GA (Late)
```{r, snakega plots summary 3 lw}
min_gen=45000
max_gen=50000
#

GA_summary_by_species(mu_b_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="snake", title="Space: Dif VS Snake GA", subt="Gen 16sims", y_label="Snake GA", x_label="Dif")

```

### By Newt GA (Early)
```{r, newtga plots summary 1 lw}
min_gen=100
max_gen=5000
#

GA_summary_by_species(mu_b_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="newt", title="Space: Dif VS Newt GA", subt="Gen 16sims", y_label="Newt GA", x_label="Dif")

```

### By Newt GA (Mid)
```{r, newtga plots summary 2 lw}
min_gen=25000
max_gen=30000
#

GA_summary_by_species(mu_b_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="newt", title="Space: Dif VS Newt GA", subt="Gen 16sims", y_label="Newt GA", x_label="Dif")

```

### By Newt GA (Late)
```{r, newtga plots summary 3 lw}
min_gen=45000
max_gen=50000
#

GA_summary_by_species(mu_b_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="newt", title="Space: Dif VS Newt GA", subt="Gen 16sims", y_label="Newt GA", x_label="Dif")

```

## {-}

## Summary (mu_c) {.tabset}

### Early-Sim Population Size Summary

```{r, early plots summary ne}

min_gen=100
max_gen=5000
GA_summary(mu_c_lit, min_gen=min_gen, max_gen=max_gen, x_line="Newt_pop_size", y_line="Snake_pop_size", title="Space: Newt VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize")
```

### Mid-Sim Population Size Summary
```{r, plots summary 2 ne}
# summary

min_gen=25000
max_gen=30000

# summary
GA_summary(mu_c_lit, min_gen=min_gen, max_gen=max_gen, x_line="Newt_pop_size", y_line="Snake_pop_size", title="Space: Newt VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize")
```

### Late-Sim Population Size Summary
```{r, plots summary 3 ne}
min_gen=45000
max_gen=50000

# summary
GA_summary(mu_c_lit, min_gen=min_gen, max_gen=max_gen, x_line="Newt_pop_size", y_line="Snake_pop_size", title="Space: Newt VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize")

```

### Early Difference Summary
```{r, dif plots summary 1 ne}

min_gen=100
max_gen=5000

GA_summary(mu_c_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="Snake_pop_size", title="Space: Dif VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Dif")
```

### Mid Difference Summary
```{r, dif plots summary 2 ne}
min_gen=25000
max_gen=30000

GA_summary(mu_c_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="Snake_pop_size", title="Space: Dif VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Dif")

```

### Late Difference Summary
```{r, dif plots summary 3 ne}
min_gen=45000
max_gen=50000
GA_summary(mu_c_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="Snake_pop_size", title="Space: Dif VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Dif")

```

### By Snake GA (Early)
```{r, snakega plots summary 1 ne}
min_gen=100
max_gen=5000
#

GA_summary_by_species(mu_c_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="snake", title="Space: Dif VS Snake GA", subt="Gen 16sims", y_label="Snake GA", x_label="Dif")

```

### By Snake GA (Mid)
```{r, snakega plots summary 2 ne}
min_gen=25000
max_gen=30000
#

GA_summary_by_species(mu_c_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="snake", title="Space: Dif VS Snake GA", subt="Gen 16sims", y_label="Snake GA", x_label="Dif")

```

### By Snake GA (Late)
```{r, snakega plots summary 3 ne}
min_gen=45000
max_gen=50000
#

GA_summary_by_species(mu_c_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="snake", title="Space: Dif VS Snake GA", subt="Gen 16sims", y_label="Snake GA", x_label="Dif")

```

### By Newt GA (Early)
```{r, newtga plots summary 1 ne}
min_gen=100
max_gen=5000
#

GA_summary_by_species(mu_c_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="newt", title="Space: Dif VS Newt GA", subt="Gen 16sims", y_label="Newt GA", x_label="Dif")

```

### By Newt GA (Mid)
```{r, newtga plots summary 2 ne}
min_gen=25000
max_gen=30000
#

GA_summary_by_species(mu_c_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="newt", title="Space: Dif VS Newt GA", subt="Gen 16sims", y_label="Newt GA", x_label="Dif")

```

### By Newt GA (Late)
```{r, newtga plots summary 3 ne}
min_gen=45000
max_gen=50000
#

GA_summary_by_species(mu_c_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="newt", title="Space: Dif VS Newt GA", subt="Gen 16sims", y_label="Newt GA", x_label="Dif")

```

## {-}

## Summary (mu_d) {.tabset}

### Early-Sim Population Size Summary

```{r, early plots summary mud}

min_gen=100
max_gen=5000
GA_summary(mu_d_lit, min_gen=min_gen, max_gen=max_gen, x_line="Newt_pop_size", y_line="Snake_pop_size", title="Space: Newt VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize")
```

### Mid-Sim Population Size Summary
```{r, plots summary 2 mud}
# summary

min_gen=25000
max_gen=30000

# summary
GA_summary(mu_d_lit, min_gen=min_gen, max_gen=max_gen, x_line="Newt_pop_size", y_line="Snake_pop_size", title="Space: Newt VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize")
```

### Late-Sim Population Size Summary
```{r, plots summary 3 mud}
min_gen=45000
max_gen=50000

# summary
GA_summary(mu_d_lit, min_gen=min_gen, max_gen=max_gen, x_line="Newt_pop_size", y_line="Snake_pop_size", title="Space: Newt VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Newt Popsize")

```

### Early Difference Summary
```{r, dif plots summary 1 mud}

min_gen=100
max_gen=5000

GA_summary(mu_d_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="Snake_pop_size", title="Space: Dif VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Dif")
```

### Mid Difference Summary
```{r, dif plots summary 2 mud}
min_gen=25000
max_gen=30000

GA_summary(mu_d_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="Snake_pop_size", title="Space: Dif VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Dif")

```

### Late Difference Summary
```{r, dif plots summary 3 mud}
min_gen=45000
max_gen=50000
GA_summary(mu_d_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="Snake_pop_size", title="Space: Dif VS Snake Population Size", subt="Gen 16sims", y_label="Snake Popsize", x_label="Dif")

```

### By Snake GA (Early)
```{r, snakega plots summary 1 mud}
min_gen=100
max_gen=5000
#

GA_summary_by_species(mu_d_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="snake", title="Space: Dif VS Snake GA", subt="Gen 16sims", y_label="Snake GA", x_label="Dif")

```

### By Snake GA (Mid)
```{r, snakega plots summary 2 mud}
min_gen=25000
max_gen=30000
#

GA_summary_by_species(mu_d_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="snake", title="Space: Dif VS Snake GA", subt="Gen 16sims", y_label="Snake GA", x_label="Dif")

```

### By Snake GA (Late)
```{r, snakega plots summary 3 mud}
min_gen=45000
max_gen=50000
#

GA_summary_by_species(mu_d_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="snake", title="Space: Dif VS Snake GA", subt="Gen 16sims", y_label="Snake GA", x_label="Dif")

```

### By Newt GA (Early)
```{r, newtga plots summary 1 mud}
min_gen=100
max_gen=5000
#

GA_summary_by_species(mu_d_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="newt", title="Space: Dif VS Newt GA", subt="Gen 16sims", y_label="Newt GA", x_label="Dif")

```

### By Newt GA (Mid)
```{r, newtga plots summary 2 mud}
min_gen=25000
max_gen=30000
#

GA_summary_by_species(mu_d_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="newt", title="Space: Dif VS Newt GA", subt="Gen 16sims", y_label="Newt GA", x_label="Dif")

```

### By Newt GA (Late)
```{r, newtga plots summary 3 mud}
min_gen=45000
max_gen=50000
#

GA_summary_by_species(mu_d_lit, min_gen=min_gen, max_gen=max_gen, x_line="dif", y_line="newt", title="Space: Dif VS Newt GA", subt="Gen 16sims", y_label="Newt GA", x_label="Dif")

```

## {-}

## Heatmap (mu_a) {.tabset}

### Population Size (Early)

```{r, early plots Heatmap}

min_gen=100
max_gen=5000
#heatmap
build_heatmap(mu_a_lit, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)
```

### Population Size (Mid)
```{r, early plots Heatmap 2}
min_gen=25000
max_gen=30000
#heatmap
build_heatmap(mu_a_lit, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

```

### Population Size (Late)
```{r, early plots Heatmap 3}
min_gen=45000
max_gen=50000
#heatmap
build_heatmap(mu_a_lit, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)
```

### Phenotype (Early)
```{r, early plots Heatmap pheno 1}
min_gen=100
max_gen=5000
build_heatmap(mu_a_lit, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### Phenotype (Mid)
```{r, early plots Heatmap pheno 2}

min_gen=25000
max_gen=30000
#heatmap

build_heatmap(mu_a_lit, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### Phenotype (Late)
```{r, early plots Heatmap pheno 3}
min_gen=45000
max_gen=50000
#heatmap

build_heatmap(mu_a_lit, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)

```

## {-}

## Heatmap (mu_b) {.tabset}

### Population Size (Early)

```{r, early plots Heatmap lw}

min_gen=100
max_gen=5000
#heatmap
build_heatmap(mu_b_lit, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)
```

### Population Size (Mid)
```{r, early plots Heatmap 2 lw}
min_gen=25000
max_gen=30000
#heatmap
build_heatmap(mu_b_lit, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

```

### Population Size (Late)
```{r, early plots Heatmap 3 lw}
min_gen=45000
max_gen=50000
#heatmap
build_heatmap(mu_b_lit, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)
```

### Phenotype (Early)
```{r, early plots Heatmap pheno 1 lw}
min_gen=100
max_gen=5000
build_heatmap(mu_b_lit, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### Phenotype (Mid)
```{r, early plots Heatmap pheno 2 lw}

min_gen=25000
max_gen=30000
#heatmap

build_heatmap(mu_b_lit, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### Phenotype (Late)
```{r, early plots Heatmap pheno 3 lw}
min_gen=45000
max_gen=50000
#heatmap

build_heatmap(mu_b_lit, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)

```

## {-}

## Heatmap (mu_c) {.tabset}

### Population Size (Early)

```{r, early plots Heatmap ne}

min_gen=100
max_gen=5000
#heatmap
build_heatmap(mu_c_lit, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)
```

### Population Size (Mid)
```{r, early plots Heatmap 2 ne}
min_gen=25000
max_gen=30000
#heatmap
build_heatmap(mu_c_lit, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

```

### Population Size (Late)
```{r, early plots Heatmap 3 ne}
min_gen=45000
max_gen=50000
#heatmap
build_heatmap(mu_c_lit, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)
```

### Phenotype (Early)
```{r, early plots Heatmap pheno 1 ne}
min_gen=100
max_gen=5000
build_heatmap(mu_c_lit, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### Phenotype (Mid)
```{r, early plots Heatmap pheno 2 ne}

min_gen=25000
max_gen=30000
#heatmap

build_heatmap(mu_c_lit, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### Phenotype (Late)
```{r, early plots Heatmap pheno 3 ne}
min_gen=45000
max_gen=50000
#heatmap

build_heatmap(mu_c_lit, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)

```

## {-}

## Heatmap (mu_d) {.tabset}

### Population Size (Early)

```{r, early plots Heatmap mud}

min_gen=100
max_gen=5000
#heatmap
build_heatmap(mu_d_lit, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)
```

### Population Size (Mid)
```{r, early plots Heatmap 2 mud}
min_gen=25000
max_gen=30000
#heatmap
build_heatmap(mu_d_lit, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)

```

### Population Size (Late)
```{r, early plots Heatmap 3 mud}
min_gen=45000
max_gen=50000
#heatmap
build_heatmap(mu_d_lit, min_gen=min_gen, max_gen=max_gen, fill_line="Snake_pop_size", fill_line_text="Snake Population Size", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA", red_blue=F)
```

### Phenotype (Early)
```{r, early plots Heatmap pheno 1 mud}
min_gen=100
max_gen=5000
build_heatmap(mu_d_lit, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### Phenotype (Mid)
```{r, early plots Heatmap pheno 2 mud}

min_gen=25000
max_gen=30000
#heatmap

build_heatmap(mu_d_lit, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)
```

### Phenotype (Late)
```{r, early plots Heatmap pheno 3 mud}
min_gen=45000
max_gen=50000
#heatmap

build_heatmap(mu_d_lit, min_gen=min_gen, max_gen=max_gen, fill_line="dif", fill_line_text="Snake-Newt", title="Newt GA by Snake GA", subt="16 Sim Heatmap", y_label="Snake GA", x_label="Newt GA",red_blue=T)

```

## {-}

```{r, red in grid files}
require(readr)
mu_a_grid<- read.csv("~/Desktop/low_to_high_v/mu_a/all_grid.txt" , header = TRUE) %>%  filter(between(generation, 9000, 10000))
mu_a_grid$group_rep <- paste0(mu_a_grid$snake_mu_rate, "_", mu_a_grid$snake_mu_effect_sd,"_",mu_a_grid$newt_mu_rate, "_", mu_a_grid$newt_mu_effect_sd, "_", mu_a_grid$rep)
mu_a_grid$group <- paste0(mu_a_grid$snake_mu_rate, "_", mu_a_grid$snake_mu_effect_sd,"_",mu_a_grid$newt_mu_rate, "_", mu_a_grid$newt_mu_effect_sd)
mu_a_grid$sim=LETTERS[factor(mu_a_grid$group)]
mu_a_grid$grid_num = 5

```

```{r, red in grid files lw}
require(readr)
mu_b_grid<- read.csv("~/Desktop/low_to_high_v/mu_b/all_grid.txt" , header = TRUE) %>%  filter(between(generation, 9000, 10000))
mu_b_grid$group_rep <- paste0(mu_b_grid$snake_mu_rate, "_", mu_b_grid$snake_mu_effect_sd,"_",mu_b_grid$newt_mu_rate, "_", mu_b_grid$newt_mu_effect_sd, "_", mu_b_grid$rep)
mu_b_grid$group <- paste0(mu_b_grid$snake_mu_rate, "_", mu_b_grid$snake_mu_effect_sd,"_",mu_b_grid$newt_mu_rate, "_", mu_b_grid$newt_mu_effect_sd)
mu_b_grid$sim=LETTERS[factor(mu_b_grid$group)]
mu_b_grid$grid_num = 5

```

```{r, red in grid files ne}
require(readr)
mu_c_grid<- read.csv("~/Desktop/low_to_high_v/mu_c/all_grid.txt" , header = TRUE) %>%  filter(between(generation, 9000, 10000))
mu_c_grid$group_rep <- paste0(mu_c_grid$snake_mu_rate, "_", mu_c_grid$snake_mu_effect_sd,"_",mu_c_grid$newt_mu_rate, "_", mu_c_grid$newt_mu_effect_sd, "_", mu_c_grid$rep)
mu_c_grid$group <- paste0(mu_c_grid$snake_mu_rate, "_", mu_c_grid$snake_mu_effect_sd,"_",mu_c_grid$newt_mu_rate, "_", mu_c_grid$newt_mu_effect_sd)
mu_c_grid$sim=LETTERS[factor(mu_c_grid$group)]
mu_c_grid$grid_num = 5

```

```{r, red in grid files mud}
require(readr)
mu_d_grid<- read.csv("~/Desktop/low_to_high_v/mu_d/all_grid.txt" , header = TRUE) %>%  filter(between(generation, 9000, 10000))
mu_d_grid$group_rep <- paste0(mu_d_grid$snake_mu_rate, "_", mu_d_grid$snake_mu_effect_sd,"_",mu_d_grid$newt_mu_rate, "_", mu_d_grid$newt_mu_effect_sd, "_", mu_d_grid$rep)
mu_d_grid$group <- paste0(mu_d_grid$snake_mu_rate, "_", mu_d_grid$snake_mu_effect_sd,"_",mu_d_grid$newt_mu_rate, "_", mu_d_grid$newt_mu_effect_sd)
mu_d_grid$sim=LETTERS[factor(mu_d_grid$group)]
mu_d_grid$grid_num = 5

```

## What is up with the correlations

## Early Simulation Correlation (mu_a)  {.tabset}

### Mean
```{r, cor grid high cor mean}

file=mu_a_grid
file_name_a="1e-08_0.05_1e-08_0.0158_1"
spegen=9501
type="mean"
row_num=20
col_num=5

grid_plot(file_name_a, spegen, file, type, row_num, col_num)
```

### Max
```{r, cor grid high cor max}

file=mu_a_grid
spegen=9501
type="max"
row_num=20
col_num=5

grid_plot(file_name_a, spegen, file, type, row_num, col_num)
```

### Min
```{r, cor grid high cor min}

file=mu_a_grid
spegen=9501
type="min"
row_num=20
col_num=5

grid_plot(file_name_a, spegen, file, type, row_num, col_num)
```

### popsize
```{r, cor grid high cor popsize}

file=mu_a_grid
spegen=9501
type="num_ind"
row_num=20
col_num=5
grid_plot(file_name_a, spegen, file, type, row_num, col_num)

```

## {-}

## Early Simulation Correlation (mu_b)  {.tabset}

### Mean
```{r, cor grid high cor mean lw}

file=mu_b_grid
file_name_b="1e-09_0.158_1e-09_0.5_1"
spegen=9501
type="mean"
row_num=20
col_num=5

grid_plot(file_name_b, spegen, file, type, row_num, col_num)
```

### Max
```{r, cor grid high cor max lw}

file=mu_b_grid
spegen=9501
type="max"
row_num=20
col_num=5

grid_plot(file_name_b, spegen, file, type, row_num, col_num)
```

### Min
```{r, cor grid high cor min lw}

file=mu_b_grid
spegen=9501
type="min"
row_num=20
col_num=5

grid_plot(file_name_b, spegen, file, type, row_num, col_num)
```

### popsize
```{r, cor grid high cor popsize lw}

file=mu_b_grid
spegen=9501
type="num_ind"
row_num=20
col_num=5
grid_plot(file_name_b, spegen, file, type, row_num, col_num)

```

## {-}

## Early Simulation Correlation (mu_c)  {.tabset}

### Mean
```{r, cor grid high cor mean ne}

file=mu_c_grid
file_name_c="1e-10_0.5_1e-10_1.58_1"
spegen=9501
type="mean"
row_num=20
col_num=5

grid_plot(file_name_c, spegen, file, type, row_num, col_num)
```

### Max
```{r, cor grid high cor max  ne}

file=mu_c_grid
spegen=9501
type="max"
row_num=20
col_num=5

grid_plot(file_name_c, spegen, file, type, row_num, col_num)
```

### Min
```{r, cor grid high cor min ne}

file=mu_c_grid
spegen=9501
type="min"
row_num=20
col_num=5

grid_plot(file_name_c, spegen, file, type, row_num, col_num)
```

### popsize
```{r, cor grid high cor popsize ne}

file=mu_c_grid
spegen=9501
type="num_ind"
row_num=20
col_num=5
grid_plot(file_name_c, spegen, file, type, row_num, col_num)

```

## {-}

## Early Simulation Correlation (mu_d)  {.tabset}

### Mean
```{r, cor grid high cor mean vd}

file=mu_d_grid
file_name_d="1e-11_1.58_1e-11_5_1"
spegen=9501
type="mean"
row_num=20
col_num=5

grid_plot(file_name_d, spegen, file, type, row_num, col_num)
```

### Max
```{r, cor grid high cor max vd}

file=mu_d_grid
spegen=9501
type="max"
row_num=20
col_num=5

grid_plot(file_name_d, spegen, file, type, row_num, col_num)
```

### Min
```{r, cor grid high cor min vd}

file=mu_d_grid
spegen=9501
type="min"
row_num=20
col_num=5

grid_plot(file_name_d, spegen, file, type, row_num, col_num)
```

### popsize
```{r, cor grid high cor popsize vd}

file=mu_d_grid
spegen=9501
type="num_ind"
row_num=20
col_num=5
grid_plot(file_name_d, spegen, file, type, row_num, col_num)

```

## {-}

```{r, red in grid files 2}
require(readr)
mu_a_grid<- read.csv("~/Desktop/low_to_high_v/mu_a/all_grid.txt" , header = TRUE) %>%  filter(between(generation, 40000, 50000))
mu_a_grid$group_rep <- paste0(mu_a_grid$snake_mu_rate, "_", mu_a_grid$snake_mu_effect_sd,"_",mu_a_grid$newt_mu_rate, "_", mu_a_grid$newt_mu_effect_sd, "_", mu_a_grid$rep)
mu_a_grid$group <- paste0(mu_a_grid$snake_mu_rate, "_", mu_a_grid$snake_mu_effect_sd,"_",mu_a_grid$newt_mu_rate, "_", mu_a_grid$newt_mu_effect_sd)
mu_a_grid$sim=LETTERS[factor(mu_a_grid$group)]
mu_a_grid$grid_num = 5

```

```{r, red in grid files 2 lw}
require(readr)
mu_b_grid<- read.csv("~/Desktop/low_to_high_v/mu_b/all_grid.txt" , header = TRUE) %>%  filter(between(generation, 40000, 50000))
mu_b_grid$group_rep <- paste0(mu_b_grid$snake_mu_rate, "_", mu_b_grid$snake_mu_effect_sd,"_",mu_b_grid$newt_mu_rate, "_", mu_b_grid$newt_mu_effect_sd, "_", mu_b_grid$rep)
mu_b_grid$group <- paste0(mu_b_grid$snake_mu_rate, "_", mu_b_grid$snake_mu_effect_sd,"_",mu_b_grid$newt_mu_rate, "_", mu_b_grid$newt_mu_effect_sd)
mu_b_grid$sim=LETTERS[factor(mu_b_grid$group)]
mu_b_grid$grid_num = 5

```

```{r, red in grid files 2 ne}
require(readr)
mu_c_grid<- read.csv("~/Desktop/low_to_high_v/mu_c/all_grid.txt" , header = TRUE) %>%  filter(between(generation, 40000, 50000))
mu_c_grid$group_rep <- paste0(mu_c_grid$snake_mu_rate, "_", mu_c_grid$snake_mu_effect_sd,"_",mu_c_grid$newt_mu_rate, "_", mu_c_grid$newt_mu_effect_sd, "_", mu_c_grid$rep)
mu_c_grid$group <- paste0(mu_c_grid$snake_mu_rate, "_", mu_c_grid$snake_mu_effect_sd,"_",mu_c_grid$newt_mu_rate, "_", mu_c_grid$newt_mu_effect_sd)
mu_c_grid$sim=LETTERS[factor(mu_c_grid$group)]
mu_c_grid$grid_num = 5

```

```{r, red in grid files 2 mud}
require(readr)
mu_d_grid<- read.csv("~/Desktop/low_to_high_v/mu_d/all_grid.txt" , header = TRUE) %>%  filter(between(generation, 40000, 50000))
mu_d_grid$group_rep <- paste0(mu_d_grid$snake_mu_rate, "_", mu_d_grid$snake_mu_effect_sd,"_",mu_d_grid$newt_mu_rate, "_", mu_d_grid$newt_mu_effect_sd, "_", mu_d_grid$rep)
mu_d_grid$group <- paste0(mu_d_grid$snake_mu_rate, "_", mu_d_grid$snake_mu_effect_sd,"_",mu_d_grid$newt_mu_rate, "_", mu_d_grid$newt_mu_effect_sd)
mu_d_grid$sim=LETTERS[factor(mu_d_grid$group)]
mu_d_grid$grid_num = 5

```

## Late Simulation Correlation (mu_a) {.tabset}

### Mean
```{r, cor grid high cor mean late}
#1e-11_5_1e-10_0.5_0
file=mu_a_grid
spegen=44501
type="mean"
row_num=20
col_num=5

grid_plot(file_name_a, spegen, file, type, row_num, col_num)
```

### Max
```{r, cor grid high cor max late}

file=mu_a_grid
spegen=44501
type="max"
row_num=20
col_num=5

grid_plot(file_name_a, spegen, file, type, row_num, col_num)
```

### Min
```{r, cor grid high cor min late}

file=mu_a_grid
spegen=44501
type="min"
row_num=20
col_num=5

grid_plot(file_name_a, spegen, file, type, row_num, col_num)
```

### Pop size
```{r, cor grid high cor popsize late}

file=mu_a_grid
spegen=44501
type="num_ind"
row_num=20
col_num=5
grid_plot(file_name_a, spegen, file, type, row_num, col_num)

```

## {-}

## Late Simulation Correlation (mu_b) {.tabset}

### Mean
```{r, cor grid high cor mean late lw}

file=mu_b_grid
spegen=44501
type="mean"
row_num=20
col_num=5

grid_plot(file_name_b, spegen, file, type, row_num, col_num)
```

### Max
```{r, cor grid high cor max late lw}

file=mu_b_grid
spegen=44501
type="max"
row_num=20
col_num=5

grid_plot(file_name_b, spegen, file, type, row_num, col_num)
```

### Min
```{r, cor grid high cor min late lw}

file=mu_b_grid
spegen=44501
type="min"
row_num=20
col_num=5

grid_plot(file_name_b, spegen, file, type, row_num, col_num)
```

### Pop size
```{r, cor grid high cor popsize late lw}

file=mu_b_grid
spegen=44501
type="num_ind"
row_num=20
col_num=5
grid_plot(file_name_b, spegen, file, type, row_num, col_num)

```

## {-}

## Late Simulation Correlation (mu_c) {.tabset}

### Mean
```{r, cor grid high cor mean late ne}

file=mu_c_grid
spegen=44501
type="mean"
row_num=20
col_num=5

grid_plot(file_name_c, spegen, file, type, row_num, col_num)
```

### Max
```{r, cor grid high cor max late ne}

file=mu_c_grid
spegen=44501
type="max"
row_num=20
col_num=5

grid_plot(file_name_c, spegen, file, type, row_num, col_num)
```

### Min
```{r, cor grid high cor min late ne}

file=mu_c_grid
spegen=44501
type="min"
row_num=20
col_num=5

grid_plot(file_name_c, spegen, file, type, row_num, col_num)
```

### Pop size
```{r, cor grid high cor popsize late ne}

file=mu_c_grid
spegen=44501
type="num_ind"
row_num=20
col_num=5
grid_plot(file_name_c, spegen, file, type, row_num, col_num)

```

## {-}

## Late Simulation Correlation (mu_d) {.tabset}

### Mean
```{r, cor grid high cor mean late mud}

file=mu_d_grid
spegen=44501
type="mean"
row_num=20
col_num=5

grid_plot(file_name_d, spegen, file, type, row_num, col_num)
```

### Max
```{r, cor grid high cor max late mud}

file=mu_d_grid
spegen=44501
type="max"
row_num=20
col_num=5

grid_plot(file_name_d, spegen, file, type, row_num, col_num)
```

### Min
```{r, cor grid high cor min late mud}

file=mu_d_grid
spegen=44501
type="min"
row_num=20
col_num=5

grid_plot(file_name_d, spegen, file, type, row_num, col_num)
```

### Pop size
```{r, cor grid high cor popsize late mud}

file=mu_d_grid
spegen=44501
type="num_ind"
row_num=20
col_num=5
grid_plot(file_name_d, spegen, file, type, row_num, col_num)

```

## {-}