### 2023-02-22 graphs_newt_snakes/Paper_figs.Rmd
making fig improvements

---
title: "Paper Figs"
output: html_document
date: '2023-01-30'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE,echo = FALSE)

library(ggplot2)
library(viridis)
library(colorspace)
library(gridExtra)
library(gridtext)
library(tidyverse)
require(reshape2)
library(plyr)
library(gridGraphics)
library(grid)
library(knitr)
library(kableExtra)
library(reshape2)
library(ggplot2)
library(ggpubr)
library(stringr)
library(patchwork)
```

```{r, data gathering}
move_together_lit<- read.csv("~/Desktop/move_together/all_lit.txt" , header = TRUE)

move_together_lit$group_rep <- paste0(move_together_lit$snake_mu_rate, "_", move_together_lit$snake_mu_effect_sd,"_",move_together_lit$newt_mu_rate, "_", move_together_lit$newt_mu_effect_sd, "_", move_together_lit$rep)

move_together_lit$group <- paste0(move_together_lit$snake_mu_rate, "_", move_together_lit$snake_mu_effect_sd,"_",move_together_lit$newt_mu_rate, "_", move_together_lit$newt_mu_effect_sd)

move_together_lit$sim=LETTERS[factor(move_together_lit$group)]
move_together_lit$dif <- move_together_lit$Snake_mean_Pheno-move_together_lit$Newt_mean_Pheno

```

```{r, red in grid files 2}
require(readr)
move_together_grid<- read.csv("~/Desktop/move_together/all_grid.txt" , header = TRUE) %>%  filter(between(generation, 40000, 50000))
move_together_grid$group_rep <- paste0(move_together_grid$snake_mu_rate, "_", move_together_grid$snake_mu_effect_sd,"_",move_together_grid$newt_mu_rate, "_", move_together_grid$newt_mu_effect_sd, "_", move_together_grid$rep)
move_together_grid$group <- paste0(move_together_grid$snake_mu_rate, "_", move_together_grid$snake_mu_effect_sd,"_",move_together_grid$newt_mu_rate, "_", move_together_grid$newt_mu_effect_sd)
move_together_grid$sim=LETTERS[factor(move_together_grid$group)]
move_together_grid$grid_num = 5

cost_grade_grid<- read.csv("~/Desktop/cost_grade/all_grid.txt" , header = TRUE) %>%  filter(between(generation, 40000, 50000))
cost_grade_grid$group_rep <- paste0(cost_grade_grid$snake_mu_rate, "_", cost_grade_grid$snake_mu_effect_sd,"_",cost_grade_grid$newt_mu_rate, "_", cost_grade_grid$newt_mu_effect_sd, "_", cost_grade_grid$rep)
cost_grade_grid$group <- paste0(cost_grade_grid$snake_mu_rate, "_", cost_grade_grid$snake_mu_effect_sd,"_",cost_grade_grid$newt_mu_rate, "_", cost_grade_grid$newt_mu_effect_sd)
cost_grade_grid$sim=LETTERS[factor(cost_grade_grid$group)]
cost_grade_grid$grid_num = 5

```

```{r, finding the slope of coevolution}

newt_first_mean_pheno = setNames(aggregate(move_together_lit[which(move_together_lit$generation>=100 & move_together_lit$generation<=1000),]$Newt_mean_Pheno, list(move_together_lit[which(move_together_lit$generation>=100 & move_together_lit$generation<=1000),]$group_rep), FUN=mean), c("group_rep", "N_pheno_start" ))

snake_first_mean_pheno = setNames(aggregate(move_together_lit[which(move_together_lit$generation>=100 & move_together_lit$generation<=1000),]$Snake_mean_Pheno, list(move_together_lit[which(move_together_lit$generation>=100 & move_together_lit$generation<=1000),]$group_rep), FUN=mean) , c("group_rep", "S_pheno_start" ))

newt_last_mean_pheno = setNames(aggregate(move_together_lit[which(move_together_lit$generation>=(max(move_together_lit$generation)-1000) & move_together_lit$generation<=max(move_together_lit$generation)),]$Newt_mean_Pheno, list(move_together_lit[which(move_together_lit$generation>=(max(move_together_lit$generation)-1000) & move_together_lit$generation<=max(move_together_lit$generation)),]$group_rep), FUN=mean), c("group_rep", "N_pheno_end" ))

snake_last_mean_pheno = setNames(aggregate(move_together_lit[which(move_together_lit$generation>=(max(move_together_lit$generation)-1000) & move_together_lit$generation<=max(move_together_lit$generation)),]$Snake_mean_Pheno, list(move_together_lit[which(move_together_lit$generation>=(max(move_together_lit$generation)-1000) & move_together_lit$generation<=max(move_together_lit$generation)),]$group_rep), FUN=mean), c("group_rep", "S_pheno_end" ))

mean_phenotype_of_group = cbind(newt_first_mean_pheno, newt_last_mean_pheno[2], snake_first_mean_pheno[2], snake_last_mean_pheno[2])

move_together_lit$N_pheno_start = mean_phenotype_of_group[match(move_together_lit$group_rep, mean_phenotype_of_group$group_rep),"N_pheno_start"]

move_together_lit$S_pheno_start = mean_phenotype_of_group[match(move_together_lit$group_rep, mean_phenotype_of_group$group_rep),"S_pheno_start"]

move_together_lit$N_pheno_end = mean_phenotype_of_group[match(move_together_lit$group_rep, mean_phenotype_of_group$group_rep),"N_pheno_end"]

move_together_lit$S_pheno_end = mean_phenotype_of_group[match(move_together_lit$group_rep, mean_phenotype_of_group$group_rep),"S_pheno_end"]

N_str_gen = setNames(aggregate(move_together_lit[which(move_together_lit$Newt_mean_Pheno<=(move_together_lit$N_pheno_start)) ,]$generation, list(move_together_lit[which(move_together_lit$Newt_mean_Pheno<=(move_together_lit$N_pheno_start)),]$group_rep), FUN=max), c("group_rep", "N_str_gen"))

S_str_gen = setNames(aggregate(move_together_lit[which(move_together_lit$Snake_mean_Pheno<=(move_together_lit$S_pheno_start)) ,]$generation, list(move_together_lit[which(move_together_lit$Snake_mean_Pheno<=(move_together_lit$S_pheno_start)),]$group_rep), FUN=max), c("group_rep", "S_str_gen"))

move_together_lit$N_str_gen <- N_str_gen[match(move_together_lit$group_rep, N_str_gen$group_rep),'N_str_gen']
move_together_lit$S_str_gen <- S_str_gen[match(move_together_lit$group_rep, S_str_gen$group_rep),'S_str_gen']

N_end_gen = aggregate(move_together_lit[which(move_together_lit$generation>=move_together_lit$N_str_gen  & move_together_lit$Newt_mean_Pheno>=(.98*(move_together_lit$N_pheno_end))) ,]$generation, list(move_together_lit[which( move_together_lit$generation>=move_together_lit$N_str_gen  & move_together_lit$Newt_mean_Pheno>=(.98*move_together_lit$N_pheno_end)),]$group_rep), FUN=min)

S_end_gen = aggregate(move_together_lit[which(move_together_lit$generation>=move_together_lit$S_str_gen  & move_together_lit$Snake_mean_Pheno>=(.98*move_together_lit$S_pheno_end)) ,]$generation, list(move_together_lit[which(move_together_lit$generation>=move_together_lit$S_str_gen  & move_together_lit$Snake_mean_Pheno>=(.98*move_together_lit$S_pheno_end)),]$group_rep), FUN=min)

move_together_lit$N_end_gen <- N_end_gen[match(move_together_lit$group_rep, N_end_gen$Group.1),'x']
move_together_lit$S_end_gen <- S_end_gen[match(move_together_lit$group_rep, S_end_gen$Group.1),'x']

# get the difference in generation time and the difference in phenotype

npt<-ggplot(move_together_lit) +
geom_boxplot(aes(sim, N_end_gen), color="darkred")
spt<-ggplot(move_together_lit) +
  geom_boxplot(aes(sim, S_end_gen), color="steelblue")
#  geom_point(aes(group_rep, N_end_gen), color="darkred")+
 # geom_point(aes(group_rep, S_end_gen), color="steelblue")
grid.arrange(npt, spt, ncol=1)

ggplot(move_together_lit) +
  facet_wrap(~sim)+
geom_point(aes(rep, N_end_gen), color="darkred")+
geom_point(aes(rep, S_end_gen), color="steelblue")

npt2<-ggplot(move_together_lit) + # How long did coevolution last?
geom_boxplot(aes(sim, N_end_gen-N_str_gen), color="darkred")
spt2<-ggplot(move_together_lit) +
  geom_boxplot(aes(sim, S_end_gen-S_str_gen), color="steelblue")

grid.arrange(npt2, spt2, ncol=1)

npt3<-ggplot(move_together_lit) + #change in phenotype
geom_boxplot(aes(sim, N_pheno_end-N_pheno_start), color="darkred")
spt3<-ggplot(move_together_lit) +
  geom_boxplot(aes(sim, S_pheno_end-S_pheno_start), color="steelblue")

grid.arrange(npt3, spt3, ncol=1)

ggplot(move_together_lit) +
geom_boxplot(aes(sim, S_end_gen-N_end_gen), color="black")

```

```{r, summary of slopw}

create_parm_file <- function(file, newt_x_line_start, newt_x_line_end, snake_x_line_start, snake_x_line_end, newt_y_line_start, newt_y_line_end, snake_y_line_start, snake_y_line_end){
params <- data_frame()

for(i in 1:length(unique(file$snake_newt_rep))){

  pattern=unique(file$snake_newt_rep)[i]
  sumitup <- file[which(file$snake_newt_rep ==pattern),]

  params_temp<- data_frame(Newt_mean_Pheno=mean(sumitup$Newt_mean_Pheno),
                           Snake_mean_Pheno=mean(sumitup$Snake_mean_Pheno), 
                           Newt_mean_Pheno_sd=sd(sumitup$Newt_mean_Pheno),
                           Snake_mean_Pheno_sd=sd(sumitup$Snake_mean_Pheno),
                           snake_mu_rate=sumitup$snake_mu_rate[1],
                           newt_mu_rate=sumitup$newt_mu_rate[1],
                           snake_mu_effect_sd=sumitup$snake_mu_effect_sd[1],
                           newt_mu_effect_sd=sumitup$newt_mu_effect_sd[1],
                           rep=sumitup$rep[1],
                           newt_x_line_start=sumitup[[newt_x_line_start]][1], 
                           newt_x_line_end=sumitup[[newt_x_line_end]][1], 
                           snake_x_line_start=sumitup[[snake_x_line_start]][1], 
                           snake_x_line_end=sumitup[[snake_x_line_end]][1], 
                           newt_y_line_start=sumitup[[newt_y_line_start]][1], 
                           newt_y_line_end=sumitup[[newt_y_line_end]][1],
                           snake_y_line_start=sumitup[[snake_y_line_start]][1], 
                           snake_y_line_end=sumitup[[snake_y_line_end]][1])

  params <- rbind(params, params_temp)
}

params$snake <- paste0(params$snake_mu_rate, "_",params$snake_mu_effect_sd)
params$newt <- paste0(params$newt_mu_rate, "_",params$newt_mu_effect_sd)
params$snake_newt <- paste0(params$snake, "_", params$newt)
return(params)
}

GA_summary_first_species <- function(file, newt_x_line_start, newt_x_line_end, snake_x_line_start, snake_x_line_end, newt_y_line_start, newt_y_line_end, snake_y_line_start, snake_y_line_end, title, subt, y_label, x_label, time){

file$snake_newt_rep <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$rep)

params<-create_parm_file(file, newt_x_line_start, newt_x_line_end, snake_x_line_start, snake_x_line_end, newt_y_line_start, newt_y_line_end, snake_y_line_start, snake_y_line_end)

my_plot <- 
  ggplot(params) +

  #ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab("Newt") + ylab(" ")+
  geom_point(aes(x=(as.numeric((newt_x_line_end)) - as.numeric((newt_x_line_start))), y=(as.numeric((newt_y_line_end)) - as.numeric((newt_y_line_start))), color = factor(newt), shape=factor(snake)),size=3)+

  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  #facet_wrap(~sim)+
  theme_bw() +  theme(legend.position='none', axis.text.x = element_text( hjust=1, size = 8))

my_plot2 <- 
  ggplot(params) +

  #ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab("Snake") + ylab(" ")+
  #geom_point(aes(x=(as.numeric(newt_x_line_end) - as.numeric((newt_x_line_start))), y=(as.numeric((newt_y_line_end)) - as.numeric((newt_y_line_start))), color = factor(snake), shape=factor(newt)),size=3)+
geom_point(aes(x=(as.numeric((snake_x_line_end)) - as.numeric((snake_x_line_start))), y=(as.numeric((snake_y_line_end)) - as.numeric((snake_y_line_start))), color = factor(snake), shape=factor(newt)),size=3)+
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  #facet_wrap(~sim)+
  labs(color = "GA 1", shape = "GA 2")+
  theme_bw() +  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(), axis.text.x = element_text( hjust=1, size = 8))

if(time=="ony"){
  my_plot <- my_plot+ylim(0, 50000)
  my_plot2 <-my_plot2+ ylim(0, 50000)
}

if(time=="onx"){
  my_plot <- my_plot+xlim(0, 50000)
  my_plot2 <- my_plot2+xlim(0, 50000)
}

final_plot = grid.arrange(my_plot, my_plot2, ncol=2, widths = c(1,1.2),
                           top = title, left = y_label, bottom = x_label)

return(final_plot)
}

GA_summary_second_species <- function(file, newt_x_line_start, newt_x_line_end, snake_x_line_start, snake_x_line_end, newt_y_line_start, newt_y_line_end, snake_y_line_start, snake_y_line_end, title, subt, y_label, x_label, time){

file$snake_newt_rep <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$rep)

params<-create_parm_file(file, newt_x_line_start, newt_x_line_end, snake_x_line_start, snake_x_line_end, newt_y_line_start, newt_y_line_end, snake_y_line_start, snake_y_line_end)

my_plot <- 
  ggplot(params) +

  #ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab("Newt") + ylab(" ")+
  geom_point(aes(x=(as.numeric((newt_x_line_end)) - as.numeric((newt_x_line_start))), y=(as.numeric((newt_y_line_end)) - as.numeric((newt_y_line_start))), color = factor(snake), shape=factor(newt)),size=3)+

  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  #facet_wrap(~sim)+
  theme_bw() +  theme(legend.position='none', axis.text.x = element_text( hjust=1, size = 8))

my_plot2 <- 
  ggplot(params) +

  #ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
  xlab("Snake") + ylab(" ")+
  #geom_point(aes(x=(as.numeric(newt_x_line_end) - as.numeric((newt_x_line_start))), y=(as.numeric((newt_y_line_end)) - as.numeric((newt_y_line_start))), color = factor(snake), shape=factor(newt)),size=3)+
geom_point(aes(x=(as.numeric((snake_x_line_end)) - as.numeric((snake_x_line_start))), y=(as.numeric((snake_y_line_end)) - as.numeric((snake_y_line_start))), color = factor(newt), shape=factor(snake)),size=3)+
  theme(plot.title = element_text(size=5)) +
  #scale_colour_gradient2(high = "steelblue", mid = "beige", low = "darkred")+
  #labs(colour = color_label) +
  #facet_wrap(~sim)+
  labs(color = "GA 2", shape = "GA 1")+
  theme_bw() +  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(), axis.text.x = element_text( hjust=1, size = 8))

if(time=="ony"){
  my_plot <- my_plot+ylim(0, 50000)
  my_plot2 <-my_plot2+ ylim(0, 50000)
}

if(time=="onx"){
  my_plot <- my_plot+xlim(0, 50000)
  my_plot2 <- my_plot2+xlim(0, 50000)
}

final_plot = grid.arrange(my_plot, my_plot2, ncol=2, widths = c(1,1.2),
                           top = title, left = y_label, bottom = x_label)

return(final_plot)
}

file = move_together_lit
newt_x_line_start = "N_pheno_start"
newt_x_line_end = "N_pheno_end"
newt_y_line_start = "N_str_gen"
newt_y_line_end = "N_end_gen"
snake_x_line_start = "S_pheno_start"
snake_x_line_end = "S_pheno_end"
snake_y_line_start = "S_str_gen"
snake_y_line_end = "S_end_gen"
title = "GA slope summary Focus on GA of Center Species"
subt = "test"
y_label = "# Gen Coevolving"
x_label = "Change in phenotype"
time ="ony"

GA_summary_first_species(file, newt_x_line_start, newt_x_line_end, snake_x_line_start, snake_x_line_end, newt_y_line_start, newt_y_line_end, snake_y_line_start, snake_y_line_end, title, subt, y_label, x_label, time)

title = "GA slope summary Focus on GA of Second Species"
GA_summary_second_species(file, newt_x_line_start, newt_x_line_end, snake_x_line_start, snake_x_line_end, newt_y_line_start, newt_y_line_end, snake_y_line_start, snake_y_line_end, title, subt, y_label, x_label, time)

file = move_together_lit
newt_y_line_start = "N_pheno_start"
newt_y_line_end = "N_pheno_end"
newt_x_line_start = "N_str_gen"
newt_x_line_end = "N_end_gen"
snake_y_line_start = "S_pheno_start"
snake_y_line_end = "S_pheno_end"
snake_x_line_start = "S_str_gen"
snake_x_line_end = "S_end_gen"
title = "GA slope summary Focus on GA of Center Species"
subt = "test"
x_label = "# Gen Coevolving"
y_label = "Change in phenotype"
time ="onx"

GA_summary_second_species(file, newt_x_line_start, newt_x_line_end, snake_x_line_start, snake_x_line_end, newt_y_line_start, newt_y_line_end, snake_y_line_start, snake_y_line_end, title, subt, y_label, x_label, time)

title = "GA slope summary Focus on GA of Second Species"
GA_summary_second_species(file, newt_x_line_start, newt_x_line_end, snake_x_line_start, snake_x_line_end, newt_y_line_start, newt_y_line_end, snake_y_line_start, snake_y_line_end, title, subt, y_label, x_label, time)

```

```{r, making newts and snakes in grid form look nice}

duo_grid_plot<- function(file_name, spegen, file, type, row_num, col_num){

file$snake_newt_rep <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd,"_",file$newt_mu_rate, "_", file$newt_mu_effect_sd, "_", file$rep)

GA1<-file[which(file$snake_newt_rep==file_name & file$generation==spegen),]
GA1_type = paste0("generation|", type)

GA1_type = GA1[,grepl(GA1_type, names(GA1)) ]
GA1_type <- GA1_type[-1]

data_long_type_comp <- gather(GA1_type, condition, measurement, factor_key=TRUE)
data_long_type_comp[,c("species","type", "type2","location")]<- str_split_fixed(data_long_type_comp$condition, "_", 4)

data_long_type_comp$y_loc<- rep(sort(c(rep.int(row_num:1, 2)), decreasing = TRUE),col_num)

data_long_type_comp$x_loc<- sort(rep.int(1:col_num, row_num*2))

main_plot<-ggplot(data_long_type_comp, aes(x = x_loc, y = y_loc))+
  geom_tile(aes(fill = as.numeric(measurement)))+
  facet_wrap(~species)+
  coord_fixed() +
  theme_bw()+
  #margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")+
  theme( legend.justification = "top", 
         plot.margin = margin(t = 0.3,  # Top margin
                             r = 0,  # Right margin
                             b = 0.3,  # Bottom margin
                             l = 0,  # Left margin
                             unit = "cm")  ) # legend.position ="bottom"
#plot.margin = unit(c(1, 1, 3, 1), "cm"),
if(type != "num_ind"){
  main_plot<- main_plot + labs(title = paste0(str_to_title(type)," Phenotype of Locations"),
                               subtitle = paste0(file_name, ", Generation: ", spegen),
                               fill = "Toxicity/Resistance")+
    scale_fill_viridis_c()
}
if(type == "num_ind"){
  main_plot<- main_plot + labs(title = paste0("Number of Individuals at Locations ", str_to_title(type)),
                               subtitle = paste0(file_name, ", Generation: ", spegen),
                               fill = "Newt/Snake PopSize")+
    scale_fill_viridis_c(option = "plasma")
}

snake<-data_long_type_comp[which(data_long_type_comp$species=="snake"),]
newt<-data_long_type_comp[which(data_long_type_comp$species=="newt"),]
newt$snake_value <- snake$measurement
print(cor(as.numeric(snake$measurement), as.numeric(newt$measurement), use="complete.obs"))

newt$info <- "middle"

#need to look into, coloring problem
newt$info[newt$location %in% c(0, row_num-1, col_num*row_num-row_num, col_num*row_num-1)] <- "corner"

newt$info[newt$location %in% c(c(1:(row_num-2)), seq(row_num, (col_num*row_num-row_num-2), row_num), seq((row_num-1+row_num), (col_num*row_num-1-row_num), row_num),  c((col_num*row_num-row_num+1):(col_num*row_num-2)))]   <- "edge"

extra_plot <- ggplot(newt)+
  ggtitle("Mean Phenotype") +
  xlab("Newt") + ylab("Snake")+
  geom_point(aes(x = as.numeric(measurement), y = as.numeric(snake_value), color=info), size = 1 ) +
  theme_bw()+
  theme(text = element_text(size=10),title =element_text(size=8),
        legend.text = element_text(size = 8),
        legend.position ="bottom",
        legend.key.size = unit(0.1, "lines"),
        legend.title=element_blank())+
  guides(shape = guide_legend(override.aes = list(size = 1)),
         color = guide_legend(override.aes = list(size = 1)))

bloan <- ggplot() + theme_void()
final_plot <- grid.arrange(main_plot, arrangeGrob(bloan, extra_plot), ncol = 2)

grid.arrange(main_plot, bloan, extra_plot, ncol = 2, 
             layout_matrix=cbind(c(1,1),c(2,3)))

#final_plot <- main_plot +  inset_element(extra_plot, left = 1., bottom = -0.10, right = 1.35, top = 0.5)

return(final_plot)}

file=move_together_grid
file_name="1e-11_5_1e-10_0.5_0"
spegen=44501
type="mean"
row_num=20
col_num=5

duo_grid_plot(file_name, spegen, file, type, row_num, col_num)

file=cost_grade_grid
file_name="1e-11_5_1e-10_0.5_0"
spegen=44501
type="mean"
row_num=20
col_num=5

duo_grid_plot(file_name, spegen, file, type, row_num, col_num)

```

```{r, looking at the slop}

file = move_together_lit
title = "test"
subt = "test"
x_label ="xx"
y_label = "yy"
min_gen = 100
max_gen = 50000

  ggplot(file[ which(file$generation>=min_gen & file$generation<=max_gen),]) +
    facet_wrap(~sim)+
    #geom_line(aes(generation, dif, group=rep, col="dif"))+
    ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen")) +
    geom_line(aes(generation, Snake_mean_Pheno, group=rep, color="Snake_mean_Pheno"),)+
    geom_line(aes(generation, Newt_mean_Pheno, group=rep, color="Newt_mean_Pheno"))+
    #geom_hline(yintercept=0, linetype="dashed", color="gray54")+
    #coord_trans(y="log2")+
    scale_y_continuous(trans = "log2")+
    #ylim(-1.,2.5)+
    xlab(x_label) + ylab(y_label)+
    scale_color_manual(name = "Type",
values = c( "Snake_mean_Pheno" = "steelblue", "Newt_mean_Pheno" = "darkred", "dif" = "black"),
labels = c("Snake", "Newt", "Dif"))+
    theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), legend.position="none",
                      strip.background = element_blank(), strip.text.x = element_blank(),
                      title =element_text(size=8, face='bold'), axis.title.x = element_blank(),
                      axis.title.y = element_blank())

```

```{r, see the correlations}

snewt <- read.csv("~/Desktop/snake_newt_real_data.csv", header = TRUE)

GA1_cor<- read.csv("~/Desktop/GA1_tall_redo/all_cor.txt" , header = TRUE)
GA1_cor$group_rep <- paste0(GA1_cor$snake_mu_rate, "_", GA1_cor$snake_mu_effect_sd,"_",GA1_cor$newt_mu_rate, "_", GA1_cor$newt_mu_effect_sd, "_", GA1_cor$rep)
GA1_cor$group <- paste0(GA1_cor$snake_mu_rate, "_", GA1_cor$snake_mu_effect_sd,"_",GA1_cor$newt_mu_rate, "_", GA1_cor$newt_mu_effect_sd)
GA1_cor$sim=LETTERS[factor(GA1_cor$group)]
GA1_cor$exp="Flat Map"

cost_grade_cor<- read.csv("~/Desktop/cost_grade/all_cor.txt" , header = TRUE)
cost_grade_cor$group_rep <- paste0(cost_grade_cor$snake_mu_rate, "_", cost_grade_cor$snake_mu_effect_sd,"_",cost_grade_cor$newt_mu_rate, "_", cost_grade_cor$newt_mu_effect_sd, "_", cost_grade_cor$rep)
cost_grade_cor$group <- paste0(cost_grade_cor$snake_mu_rate, "_", cost_grade_cor$snake_mu_effect_sd,"_",cost_grade_cor$newt_mu_rate, "_", cost_grade_cor$newt_mu_effect_sd)
cost_grade_cor$sim=LETTERS[factor(cost_grade_cor$group)]
cost_grade_cor$exp="Cost Gradent Map"

interrate_grade<- read.csv("~/Desktop/interrate_grade/all_cor.txt" , header = TRUE)
interrate_grade$group_rep <- paste0(interrate_grade$snake_mu_rate, "_", interrate_grade$snake_mu_effect_sd,"_",interrate_grade$newt_mu_rate, "_", interrate_grade$newt_mu_effect_sd, "_", interrate_grade$rep)
interrate_grade$group <- paste0(interrate_grade$snake_mu_rate, "_", interrate_grade$snake_mu_effect_sd,"_",interrate_grade$newt_mu_rate, "_", interrate_grade$newt_mu_effect_sd)
interrate_grade$sim=LETTERS[factor(interrate_grade$group)]
interrate_grade$exp="Interaction Gradent Map"

boxplot_cors_by_gen_col_rep <- function(file1, file2, file3, title, subt, x_label, y_label, min_gen, max_gen){
 snewt$mean_res = snewt$mean_toxin+snewt$dif
  real_snewt_cor = cor(snewt$mean_toxin, snewt$mean_res)
   file1$snake_newt <- paste0(file1$snake_mu_rate, "_", file1$snake_mu_effect_sd, "_", file1$newt_mu_rate, "_", file1$newt_mu_effect_sd)
   file2$snake_newt <- paste0(file2$snake_mu_rate, "_", file2$snake_mu_effect_sd, "_", file2$newt_mu_rate, "_", file2$newt_mu_effect_sd)
   file3$snake_newt <- paste0(file3$snake_mu_rate, "_", file3$snake_mu_effect_sd, "_", file3$newt_mu_rate, "_", file3$newt_mu_effect_sd)
file <- rbind(file1, file2, file3)

file$sim <- str_replace_all(file$sim, "[ABCDEIM]", 'Sims')
#c("A"='Low Mu Var', "B"='Low Mu Var', "C"='Low Mu Var', "D"='Low Mu Var', "E"='Low Mu Var', "I"='Low Mu Var', "M"='Low Mu Var'))

file$cor_score <- ifelse(file$sim == "Sims", "False","True")

#file$cor_score <- ifelse(file$sim == "A" | file$sim == "B"| file$sim == "C" | file$sim == "D"| file$sim == "E"| file$sim == "I"| file$sim == "M", "False","True")

file$exp = factor(file$exp, levels=c("Flat Map", "Cost Gradent Map", "Interaction Gradent Map"))

  plot <-  ggplot(file[which(file$gen>=min_gen & file$gen<=max_gen),]) +
  ggtitle(title, subtitle = paste0(subt, " From ", min_gen, " gen to ", max_gen, " gen"))+
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor(sim, level=c('Sims', "F","G","H","J","K","L","N","O","P")), y=as.numeric(mean_newt_pheno_By_mean_snake_pheno), fill=factor(cor_score)))+
  theme(plot.title = element_text(size=5)) +
  #xlim(-1.1, 1.1)+
    geom_hline(yintercept=real_snewt_cor, linetype="dashed")+
  facet_wrap(~exp, ncol=1)+
  scale_fill_manual(values =c("darkolivegreen3", "cadetblue2"), labels=c('Low Mutatinal Variance', 'Coevolution'))+
    labs(fill = " ")+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      title =element_text(size=8, face='bold'))

  return(plot)
}

file1 = GA1_cor[which(1:nrow(GA1_cor) %% 2 != 0) , ]
file2 = cost_grade_cor
file3 = interrate_grade
title = "Spatial Phenotype Correlation Across Different Simulation Conditions"
subt = " "
exp_labeles=c("Flat Map", "Cost Gradent Map", "Interaction Gradent Map")
x_label = "Simulations"
y_label = "Spatila Correlation"
min_gen = 40000
max_gen = 50000

boxplot_cors_by_gen_col_rep(file1, file2, file3, title, subt, x_label, y_label, min_gen, max_gen)

```

```{r, slope function}

find_coevo_slope <- function(file, coevo_begin_start_gen, coevo_begin_end_gen, coevo_endlength_gen, close_percent){

  # coevo_begin_start_gen - number of generations to skip as coevo could begin 
newt_first_mean_pheno = setNames(aggregate(file[which(file$generation>=coevo_begin_start_gen & file$generation<=coevo_begin_end_gen),]$Newt_mean_Pheno, list(file[which(file$generation>=coevo_begin_start_gen & file$generation<=coevo_begin_end_gen),]$group_rep), FUN=mean), c("group_rep", "N_pheno_start" ))

snake_first_mean_pheno = setNames(aggregate(file[which(file$generation>=coevo_begin_start_gen & file$generation<=coevo_begin_end_gen),]$Snake_mean_Pheno, list(file[which(file$generation>=coevo_begin_start_gen & file$generation<=coevo_begin_end_gen),]$group_rep), FUN=mean) , c("group_rep", "S_pheno_start" ))

newt_last_mean_pheno = setNames(aggregate(file[which(file$generation>=(max(file$generation)-coevo_endlength_gen) & file$generation<=max(file$generation)),]$Newt_mean_Pheno, list(file[which(file$generation>=(max(file$generation)-coevo_endlength_gen) & file$generation<=max(file$generation)),]$group_rep), FUN=mean), c("group_rep", "N_pheno_end" ))

snake_last_mean_pheno = setNames(aggregate(file[which(file$generation>=(max(file$generation)-coevo_endlength_gen) & file$generation<=max(file$generation)),]$Snake_mean_Pheno, list(file[which(file$generation>=(max(file$generation)-coevo_endlength_gen) & file$generation<=max(file$generation)),]$group_rep), FUN=mean), c("group_rep", "S_pheno_end" ))

mean_phenotype_of_group = cbind(newt_first_mean_pheno, newt_last_mean_pheno[2], snake_first_mean_pheno[2], snake_last_mean_pheno[2])

file$N_pheno_start = mean_phenotype_of_group[match(file$group_rep, mean_phenotype_of_group$group_rep),"N_pheno_start"]

file$S_pheno_start = mean_phenotype_of_group[match(file$group_rep, mean_phenotype_of_group$group_rep),"S_pheno_start"]

file$N_pheno_end = mean_phenotype_of_group[match(file$group_rep, mean_phenotype_of_group$group_rep),"N_pheno_end"]

file$S_pheno_end = mean_phenotype_of_group[match(file$group_rep, mean_phenotype_of_group$group_rep),"S_pheno_end"]

N_str_gen = setNames(aggregate(file[which(file$Newt_mean_Pheno<=(file$N_pheno_start)) ,]$generation, list(file[which(file$Newt_mean_Pheno<=(file$N_pheno_start)),]$group_rep), FUN=max), c("group_rep", "N_str_gen"))

S_str_gen = setNames(aggregate(file[which(file$Snake_mean_Pheno<=(file$S_pheno_start)) ,]$generation, list(file[which(file$Snake_mean_Pheno<=(file$S_pheno_start)),]$group_rep), FUN=max), c("group_rep", "S_str_gen"))

file$N_str_gen <- N_str_gen[match(file$group_rep, N_str_gen$group_rep),'N_str_gen']
file$S_str_gen <- S_str_gen[match(file$group_rep, S_str_gen$group_rep),'S_str_gen']

N_end_gen = aggregate(file[which(file$generation>=file$N_str_gen  & file$Newt_mean_Pheno>=(close_percent*(file$N_pheno_end))) ,]$generation, list(file[which( file$generation>=file$N_str_gen  & file$Newt_mean_Pheno>=(close_percent*file$N_pheno_end)),]$group_rep), FUN=min)

S_end_gen = aggregate(file[which(file$generation>=file$S_str_gen  & file$Snake_mean_Pheno>=(close_percent*file$S_pheno_end)) ,]$generation, list(file[which(file$generation>=file$S_str_gen  & file$Snake_mean_Pheno>=(close_percent*file$S_pheno_end)),]$group_rep), FUN=min)

file$N_end_gen <- N_end_gen[match(file$group_rep, N_end_gen$Group.1),'x']
file$S_end_gen <- S_end_gen[match(file$group_rep, S_end_gen$Group.1),'x']

return(file)

}

boxplot_coevo_slope <- function(file1, file2, file3, title, subt, x_label, y_label){

   file1$snake_newt <- paste0(file1$snake_mu_rate, "_", file1$snake_mu_effect_sd, "_", file1$newt_mu_rate, "_", file1$newt_mu_effect_sd)
   file2$snake_newt <- paste0(file2$snake_mu_rate, "_", file2$snake_mu_effect_sd, "_", file2$newt_mu_rate, "_", file2$newt_mu_effect_sd)
   file3$snake_newt <- paste0(file3$snake_mu_rate, "_", file3$snake_mu_effect_sd, "_", file3$newt_mu_rate, "_", file3$newt_mu_effect_sd)
file <- rbind(file1, file2, file3)
file$newt_pheno_change=file$N_pheno_end - file$N_pheno_start
file$snake_pheno_change=file$S_pheno_end - file$S_pheno_start
#file$newt_coevo_slope = (file$N_pheno_end - file$N_pheno_start)/ (file$N_end_gen - file$N_str_gen)
#file$snake_coevo_slope = (file$S_pheno_end - file$S_pheno_start)/ (file$S_end_gen - file$S_str_gen)

file$coevo_score <- ifelse(file$sim == "A" | file$sim == "B"| file$sim == "C" | file$sim == "D"| file$sim == "E"| file$sim == "I"| file$sim == "M", "False","True")

file$exp = factor(file$exp, levels=c("No Interaction", "No Heritability", "Coevolution Flat Map"))

file_long <- tidyr::pivot_longer(file, c(newt_pheno_change, snake_pheno_change))

  plot <-  ggplot(file_long) +
  ggtitle(title, subtitle = paste0(subt))+
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor(sim, level=c("A","B", "C", "D", "E", "I", "M", "F","G","H","J","K","L","N","O","P")), y=as.numeric(value), fill=factor(name)))+
  theme(plot.title = element_text(size=5)) +
  #xlim(-1.1, 1.1)+
  facet_wrap(~exp, ncol=1)+
  scale_fill_manual(values =c("darkred", "steelblue"), labels=c('Newt', 'Snake'))+
    labs(fill = " ")+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      title =element_text(size=8, face='bold'))

  return(plot)
}

boxplot_coevo_slope_spec_avg <- function(file1, file2, file3, title, subt, x_label, y_label){

   file1$snake_newt <- paste0(file1$snake_mu_rate, "_", file1$snake_mu_effect_sd, "_", file1$newt_mu_rate, "_", file1$newt_mu_effect_sd)
   file2$snake_newt <- paste0(file2$snake_mu_rate, "_", file2$snake_mu_effect_sd, "_", file2$newt_mu_rate, "_", file2$newt_mu_effect_sd)
   file3$snake_newt <- paste0(file3$snake_mu_rate, "_", file3$snake_mu_effect_sd, "_", file3$newt_mu_rate, "_", file3$newt_mu_effect_sd)

file <- rbind(file1, file2, file3)
file$newt_pheno_change=file$N_pheno_end - file$N_pheno_start
file$snake_pheno_change=file$S_pheno_end - file$S_pheno_start
#file$newt_coevo_slope = (file$N_pheno_end - file$N_pheno_start)/ (file$N_end_gen - file$N_str_gen)
#file$snake_coevo_slope = (file$S_pheno_end - file$S_pheno_start)/ (file$S_end_gen - file$S_str_gen)

file$cor_score <- ifelse(file$sim == "A" | file$sim == "B"| file$sim == "C" | file$sim == "D"| file$sim == "E"| file$sim == "I"| file$sim == "M", "False","True")
file$exp = factor(file$exp, levels=c("No Interaction", "No Heritability", "Coevolution Flat Map"))

file$newt_snake_pheno_avg= (file$newt_pheno_change+file$snake_pheno_change)/2

  plot <-  ggplot(file) +
  ggtitle(title, subtitle = subt)+
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor(sim, level=c("A","B", "C", "D", "E", "I", "M", "F","G","H","J","K","L","N","O","P")), y=as.numeric(newt_snake_pheno_avg), fill=factor(cor_score)))+
  theme(plot.title = element_text(size=5)) +
  #xlim(-1.1, 1.1)+
  facet_wrap(~exp, ncol=1)+
  scale_fill_manual(values =c("darkolivegreen3", "cadetblue2"), labels=c('No Coevolution', 'Coevolution'))+
    labs(fill = " ")+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      title =element_text(size=8, face='bold'))

  return(plot)
}

boxplot_coevo_slope_spec_avg_advantage <- function(file1, file2, file3, title, subt, x_label, y_label){

   file1$snake_newt <- paste0(file1$snake_mu_rate, "_", file1$snake_mu_effect_sd, "_", file1$newt_mu_rate, "_", file1$newt_mu_effect_sd)
   file2$snake_newt <- paste0(file2$snake_mu_rate, "_", file2$snake_mu_effect_sd, "_", file2$newt_mu_rate, "_", file2$newt_mu_effect_sd)
   file3$snake_newt <- paste0(file3$snake_mu_rate, "_", file3$snake_mu_effect_sd, "_", file3$newt_mu_rate, "_", file3$newt_mu_effect_sd)

file <- rbind(file1, file2, file3)
file$newt_pheno_change=file$N_pheno_end - file$N_pheno_start
file$snake_pheno_change=file$S_pheno_end - file$S_pheno_start
#file$newt_coevo_slope = (file$N_pheno_end - file$N_pheno_start)/ (file$N_end_gen - file$N_str_gen)
#file$snake_coevo_slope = (file$S_pheno_end - file$S_pheno_start)/ (file$S_end_gen - file$S_str_gen)

file$exp = factor(file$exp, levels=c("No Interaction", "No Heritability", "Coevolution Flat Map"))

newt_adv= c("D", "C", "B", "H", "G", "L")
mid= c("A", "F", "K", "P")
snake_adv = c("E", "I", "J", "M", "N", "O")

file$coevo_score = 0

file$coevo_score[file$sim %in% newt_adv] = "Newt Advantage"
file$coevo_score[file$sim %in% mid] = "No Advantage"
file$coevo_score[file$sim %in% snake_adv] = "Snake Advantage"
file$newt_snake_pheno_avg= (file$newt_pheno_change+file$snake_pheno_change)/2

  plot <-  ggplot(file) +
  ggtitle(title, subtitle = subt)+
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor(sim, level=c("D", "C", "B", "H", "G", "L", "A", "F", "K", "P", "E", "I", "J", "M", "N", "O")), y=as.numeric(newt_snake_pheno_avg), fill=factor(coevo_score)))+

  theme(plot.title = element_text(size=5)) +
  #xlim(-1.1, 1.1)+
  facet_wrap(~exp, ncol=1)+
  scale_fill_manual(values =c("darkred", "purple", "steelblue"), labels=c('Newt Advantage', 'No Advantage', 'Snake Advantage'))+
    labs(fill = " ")+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      title =element_text(size=8, face='bold'))

  return(plot)
}

boxplot_coevo_slope_flat_map <- function(file, title, subt, x_label, y_label){

   file$snake_newt <- paste0(file$snake_mu_rate, "_", file$snake_mu_effect_sd, "_", file$newt_mu_rate, "_", file$newt_mu_effect_sd)

file$newt_pheno_change=file$N_pheno_end - file$N_pheno_start
file$snake_pheno_change=file$S_pheno_end - file$S_pheno_start
#file$newt_coevo_slope = (file$N_pheno_end - file$N_pheno_start)/ (file$N_end_gen - file$N_str_gen)
#file$snake_coevo_slope = (file$S_pheno_end - file$S_pheno_start)/ (file$S_end_gen - file$S_str_gen)

newt_adv= c("D", "C", "B", "H", "G", "L")
mid= c("A", "F", "K", "P")
snake_adv = c("E", "I", "J", "M", "N", "O")

file$coevo_score = 0

file$coevo_score[file$sim %in% newt_adv] = "Newt Advantage"
file$coevo_score[file$sim %in% mid] = "No Advantage"
file$coevo_score[file$sim %in% snake_adv] = "Snake Advantage"
file$newt_snake_pheno_avg= (file$newt_pheno_change+file$snake_pheno_change)/2

file_long <- tidyr::pivot_longer(file, c(newt_pheno_change, snake_pheno_change))

  plot <-  ggplot(file_long) +
  ggtitle(title, subtitle = subt)+
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor(sim, level=c("D", "C", "B", "H", "G", "L", "A", "F", "K", "P", "E", "J", "I", "O", "N", "M")), y=as.numeric(value), fill=factor(name)))+
  theme(plot.title = element_text(size=5)) +
  #xlim(-1.1, 1.1)+
  scale_fill_manual(values =c("darkred", "steelblue"), labels=c('Newt', 'Snake'))+
    labs(fill = " ")+
    #scale_fill_manual(values =c("darkred", "purple", "steelblue"), labels=c('Newt Advantage', 'No Advantage', 'Snake Advantage'))+
    labs(fill = " ")+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      title =element_text(size=8, face='bold'))

  return(plot)
}

boxplot_coevo_slope_small <- function(file1, file2, file3, title, subt, x_label, y_label){

   file1$snake_newt <- paste0(file1$snake_mu_rate, "_", file1$snake_mu_effect_sd, "_", file1$newt_mu_rate, "_", file1$newt_mu_effect_sd)
   file2$snake_newt <- paste0(file2$snake_mu_rate, "_", file2$snake_mu_effect_sd, "_", file2$newt_mu_rate, "_", file2$newt_mu_effect_sd)
   file3$snake_newt <- paste0(file3$snake_mu_rate, "_", file3$snake_mu_effect_sd, "_", file3$newt_mu_rate, "_", file3$newt_mu_effect_sd)

file <- rbind(file1, file2, file3)
file$newt_pheno_change=file$N_pheno_end - file$N_pheno_start
file$snake_pheno_change=file$S_pheno_end - file$S_pheno_start
#file$newt_coevo_slope = (file$N_pheno_end - file$N_pheno_start)/ (file$N_end_gen - file$N_str_gen)
#file$snake_coevo_slope = (file$S_pheno_end - file$S_pheno_start)/ (file$S_end_gen - file$S_str_gen)

file$exp = factor(file$exp, levels=c("No Interaction", "No Heritability", "Coevolution Flat Map"))

newt_adv= c("D", "C", "B", "H", "G", "L")
mid= c("A", "F", "K", "P")
snake_adv = c("E", "I", "J", "M", "N", "O")

file$coevo_score = 0

file$coevo_score[file$sim %in% newt_adv] = "Newt Advantage"
file$coevo_score[file$sim %in% mid] = "No Advantage"
file$coevo_score[file$sim %in% snake_adv] = "Snake Advantage"

file<- subset(file, !sim %in% c("A", "B", "C", "D", "E", "I", "M"))

file$newt_snake_pheno_avg= (file$newt_pheno_change+file$snake_pheno_change)/2

  plot <-  ggplot(file) +
  ggtitle(title, subtitle = subt)+
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor(sim), y=as.numeric(newt_snake_pheno_avg), fill=factor(coevo_score)))+
  theme(plot.title = element_text(size=5)) +
  #xlim(-1.1, 1.1)+
  facet_wrap(~exp, ncol=1)+
  scale_fill_manual(values =c("darkred", "purple", "steelblue"), labels=c('Newt Advantage', 'No Advantage', 'Snake Advantage'))+
    labs(fill = " ")+
  theme_bw() +  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                      title =element_text(size=8, face='bold'))

  return(plot)
}

file=move_together_lit
coevo_begin_start_gen=100
coevo_begin_end_gen=1000
coevo_endlength_gen=1000 # this is to take the average of the last ___ of generations
close_percent = .98

move_together_lit = find_coevo_slope(file, coevo_begin_start_gen, coevo_begin_end_gen, coevo_endlength_gen, close_percent)

```

```{r, look at the change in phenotyp}

no_interaction_lit<- read.csv("~/Desktop/no_interaction/all_lit.txt" , header = TRUE)

no_interaction_lit$group_rep <- paste0(no_interaction_lit$snake_mu_rate, "_", no_interaction_lit$snake_mu_effect_sd,"_",no_interaction_lit$newt_mu_rate, "_", no_interaction_lit$newt_mu_effect_sd, "_", no_interaction_lit$rep)

no_interaction_lit$group <- paste0(no_interaction_lit$snake_mu_rate, "_", no_interaction_lit$snake_mu_effect_sd,"_",no_interaction_lit$newt_mu_rate, "_", no_interaction_lit$newt_mu_effect_sd)

no_interaction_lit$sim=LETTERS[factor(no_interaction_lit$group)]
no_interaction_lit$dif <- no_interaction_lit$Snake_mean_Pheno-no_interaction_lit$Newt_mean_Pheno
no_interaction_lit$exp="No Interaction"

no_hair_lit<- read.csv("~/Desktop/no_hair/all_lit.txt" , header = TRUE)

no_hair_lit$group_rep <- paste0(no_hair_lit$snake_mu_rate, "_", no_hair_lit$snake_mu_effect_sd,"_",no_hair_lit$newt_mu_rate, "_", no_hair_lit$newt_mu_effect_sd, "_", no_hair_lit$rep)

no_hair_lit$group <- paste0(no_hair_lit$snake_mu_rate, "_", no_hair_lit$snake_mu_effect_sd,"_",no_hair_lit$newt_mu_rate, "_", no_hair_lit$newt_mu_effect_sd)

no_hair_lit$sim=LETTERS[factor(no_hair_lit$group)]
no_hair_lit$dif <- no_hair_lit$Snake_mean_Pheno-no_hair_lit$Newt_mean_Pheno
no_hair_lit$exp= "No Heritability"

GA1_lit<- read.csv("~/Desktop/GA1_tall_redo/all_lit.txt" , header = TRUE)
GA1_lit$group_rep <- paste0(GA1_lit$snake_mu_rate, "_", GA1_lit$snake_mu_effect_sd,"_",GA1_lit$newt_mu_rate, "_", GA1_lit$newt_mu_effect_sd, "_", GA1_lit$rep)

GA1_lit$group <- paste0(GA1_lit$snake_mu_rate, "_", GA1_lit$snake_mu_effect_sd,"_",GA1_lit$newt_mu_rate, "_", GA1_lit$newt_mu_effect_sd)

GA1_lit$sim=LETTERS[factor(GA1_lit$group)]
GA1_lit$dif <- GA1_lit$Snake_mean_Pheno-GA1_lit$Newt_mean_Pheno
GA1_lit$exp="Coevolution Flat Map"

file=no_interaction_lit
file1=find_coevo_slope(file, coevo_begin_start_gen, coevo_begin_end_gen, coevo_endlength_gen, close_percent)
file=no_hair_lit
file2=find_coevo_slope(file, coevo_begin_start_gen, coevo_begin_end_gen, coevo_endlength_gen, close_percent)
file=GA1_lit
file3=find_coevo_slope(file, coevo_begin_start_gen, coevo_begin_end_gen, coevo_endlength_gen, close_percent)

title = "Whole Simulation Phenotype Change"
subt = "Three Experiments"
x_label = "SIM"
y_label = "Change in Phenotype"

boxplot_coevo_slope(file1, file2, file3, title, subt, x_label, y_label)

boxplot_coevo_slope_spec_avg(file1, file2, file3, title, subt, x_label, y_label)

boxplot_coevo_slope_spec_avg_advantage(file1, file2, file3, title, subt, x_label, y_label)

boxplot_coevo_slope_small(file1, file2, file3, title, subt, x_label, y_label)

title = "Phenotype Change Coevolution FLat Map"
boxplot_coevo_slope_flat_map(file3, title, subt, x_label, y_label)
```

```{r, middle four}

# look at coevo slopes for F, G, L, K for same V and high to low V

coevo_begin_start_gen=100
coevo_begin_end_gen=1000
coevo_endlength_gen=1000 # this is to take the average of the last ___ of generations
close_percent = .98

mu_a_lit<- read.csv("~/Desktop/low_to_high_v/mu_a/all_lit.txt" , header = TRUE)

mu_a_lit$group_rep <- paste0(mu_a_lit$snake_mu_rate, "_", mu_a_lit$snake_mu_effect_sd,"_",mu_a_lit$newt_mu_rate, "_", mu_a_lit$newt_mu_effect_sd, "_", mu_a_lit$rep)

mu_a_lit$group <- paste0(mu_a_lit$snake_mu_rate, "_", mu_a_lit$snake_mu_effect_sd,"_",mu_a_lit$newt_mu_rate, "_", mu_a_lit$newt_mu_effect_sd)

mu_a_lit$sim=LETTERS[factor(mu_a_lit$group)]
mu_a_lit$dif <- mu_a_lit$Snake_mean_Pheno-mu_a_lit$Newt_mean_Pheno
mu_a_lit$exp = "Low to High MuV (2.5e-12 to 2.5e-11) Mu=1e-8"
mu_a_lit$change = "yes"
file=mu_a_lit
mu_a_lit = find_coevo_slope(file, coevo_begin_start_gen, coevo_begin_end_gen, coevo_endlength_gen, close_percent)
sub_mu_a_lit = subset(mu_a_lit, mu_a_lit$sim == c("F", "G", "J", "K") )

# Mu B
mu_b_lit<- read.csv("~/Desktop/low_to_high_v/mu_b/all_lit.txt" , header = TRUE)

mu_b_lit$group_rep <- paste0(mu_b_lit$snake_mu_rate, "_", mu_b_lit$snake_mu_effect_sd,"_",mu_b_lit$newt_mu_rate, "_", mu_b_lit$newt_mu_effect_sd, "_", mu_b_lit$rep)

mu_b_lit$group <- paste0(mu_b_lit$snake_mu_rate, "_", mu_b_lit$snake_mu_effect_sd,"_",mu_b_lit$newt_mu_rate, "_", mu_b_lit$newt_mu_effect_sd)

mu_b_lit$sim=LETTERS[factor(mu_b_lit$group)]
mu_b_lit$dif <- mu_b_lit$Snake_mean_Pheno-mu_b_lit$Newt_mean_Pheno
mu_b_lit$exp = "Low to High MuV (2.5e-12 to 2.5e-11) Mu=1e-9"
mu_b_lit$change = "yes"
file=mu_b_lit
mu_b_lit = find_coevo_slope(file, coevo_begin_start_gen, coevo_begin_end_gen, coevo_endlength_gen, close_percent)
sub_mu_b_lit = subset(mu_b_lit, mu_b_lit$sim == c("F", "G", "J", "K") )

# mu c

mu_c_lit<- read.csv("~/Desktop/low_to_high_v/mu_c/all_lit.txt" , header = TRUE)

mu_c_lit$group_rep <- paste0(mu_c_lit$snake_mu_rate, "_", mu_c_lit$snake_mu_effect_sd,"_",mu_c_lit$newt_mu_rate, "_", mu_c_lit$newt_mu_effect_sd, "_", mu_c_lit$rep)

mu_c_lit$group <- paste0(mu_c_lit$snake_mu_rate, "_", mu_c_lit$snake_mu_effect_sd,"_",mu_c_lit$newt_mu_rate, "_", mu_c_lit$newt_mu_effect_sd)

mu_c_lit$sim=LETTERS[factor(mu_c_lit$group)]
mu_c_lit$dif <- mu_c_lit$Snake_mean_Pheno-mu_c_lit$Newt_mean_Pheno
mu_c_lit$exp = "Low to High MuV (2.5e-12 to 2.5e-11) Mu=1e-10"
mu_c_lit$change = "yes"
file=mu_c_lit
mu_c_lit = find_coevo_slope(file, coevo_begin_start_gen, coevo_begin_end_gen, coevo_endlength_gen, close_percent)
sub_mu_c_lit = subset(mu_c_lit, mu_c_lit$sim == c("F", "G", "J", "K") )

# mu d

mu_d_lit<- read.csv("~/Desktop/low_to_high_v/mu_d/all_lit.txt" , header = TRUE)

mu_d_lit$group_rep <- paste0(mu_d_lit$snake_mu_rate, "_", mu_d_lit$snake_mu_effect_sd,"_",mu_d_lit$newt_mu_rate, "_", mu_d_lit$newt_mu_effect_sd, "_", mu_d_lit$rep)

mu_d_lit$group <- paste0(mu_d_lit$snake_mu_rate, "_", mu_d_lit$snake_mu_effect_sd,"_",mu_d_lit$newt_mu_rate, "_", mu_d_lit$newt_mu_effect_sd)

mu_d_lit$sim=LETTERS[factor(mu_d_lit$group)]
mu_d_lit$dif <- mu_d_lit$Snake_mean_Pheno-mu_d_lit$Newt_mean_Pheno
mu_d_lit$exp = "Low to High MuV (2.5e-12 to 2.5e-11) Mu=1e-11"
mu_d_lit$change = "yes"
file=mu_d_lit
mu_d_lit = find_coevo_slope(file, coevo_begin_start_gen, coevo_begin_end_gen, coevo_endlength_gen, close_percent)
sub_mu_d_lit = subset(mu_d_lit, mu_d_lit$sim == c("F", "G", "J", "K") )

v_a_lit<- read.csv("~/Desktop/same_v/v_a/all_lit.txt" , header = TRUE)

v_a_lit$group_rep <- paste0(v_a_lit$snake_mu_rate, "_", v_a_lit$snake_mu_effect_sd,"_",v_a_lit$newt_mu_rate, "_", v_a_lit$newt_mu_effect_sd, "_", v_a_lit$rep)

v_a_lit$group <- paste0(v_a_lit$snake_mu_rate, "_", v_a_lit$snake_mu_effect_sd,"_",v_a_lit$newt_mu_rate, "_", v_a_lit$newt_mu_effect_sd)

v_a_lit$sim=LETTERS[factor(v_a_lit$group)]
v_a_lit$dif <- v_a_lit$Snake_mean_Pheno-v_a_lit$Newt_mean_Pheno
v_a_lit$exp = "MuV = 2.5e-13 (low to high sigma: high to low mu)"
v_a_lit$change = "no"
file=v_a_lit
v_a_lit = find_coevo_slope(file, coevo_begin_start_gen, coevo_begin_end_gen, coevo_endlength_gen, close_percent)
sub_v_a_lit = subset(v_a_lit, v_a_lit$sim == c("F", "G", "J", "K") )

# Mu B
v_b_lit<- read.csv("~/Desktop/same_v/v_b/all_lit.txt" , header = TRUE)

v_b_lit$group_rep <- paste0(v_b_lit$snake_mu_rate, "_", v_b_lit$snake_mu_effect_sd,"_",v_b_lit$newt_mu_rate, "_", v_b_lit$newt_mu_effect_sd, "_", v_b_lit$rep)

v_b_lit$group <- paste0(v_b_lit$snake_mu_rate, "_", v_b_lit$snake_mu_effect_sd,"_",v_b_lit$newt_mu_rate, "_", v_b_lit$newt_mu_effect_sd)

v_b_lit$sim=LETTERS[factor(v_b_lit$group)]
v_b_lit$dif <- v_b_lit$Snake_mean_Pheno-v_b_lit$Newt_mean_Pheno
v_b_lit$exp = "MuV = 2.5e-12 (low to high sigma: high to low mu)"
v_b_lit$change = "no"
file=v_b_lit
v_b_lit = find_coevo_slope(file, coevo_begin_start_gen, coevo_begin_end_gen, coevo_endlength_gen, close_percent)
sub_v_b_lit = subset(v_b_lit, v_b_lit$sim == c("F", "G", "J", "K") )

# Mu C

v_c_lit<- read.csv("~/Desktop/same_v/v_c/all_lit.txt" , header = TRUE)

v_c_lit$group_rep <- paste0(v_c_lit$snake_mu_rate, "_", v_c_lit$snake_mu_effect_sd,"_",v_c_lit$newt_mu_rate, "_", v_c_lit$newt_mu_effect_sd, "_", v_c_lit$rep)

v_c_lit$group <- paste0(v_c_lit$snake_mu_rate, "_", v_c_lit$snake_mu_effect_sd,"_",v_c_lit$newt_mu_rate, "_", v_c_lit$newt_mu_effect_sd)

v_c_lit$sim=LETTERS[factor(v_c_lit$group)]
v_c_lit$dif <- v_c_lit$Snake_mean_Pheno-v_c_lit$Newt_mean_Pheno

v_c_lit$exp = "MuV = 2.5e-11 (low to high sigma: high to low mu)"
v_c_lit$change = "no"
file=v_c_lit
v_c_lit = find_coevo_slope(file, coevo_begin_start_gen, coevo_begin_end_gen, coevo_endlength_gen, close_percent)
sub_v_c_lit = subset(v_c_lit, v_c_lit$sim == c("F", "G", "J", "K") )

# mu d

v_d_lit<- read.csv("~/Desktop/same_v/v_d/all_lit.txt" , header = TRUE)

v_d_lit$group_rep <- paste0(v_d_lit$snake_mu_rate, "_", v_d_lit$snake_mu_effect_sd,"_",v_d_lit$newt_mu_rate, "_", v_d_lit$newt_mu_effect_sd, "_", v_d_lit$rep)

v_d_lit$group <- paste0(v_d_lit$snake_mu_rate, "_", v_d_lit$snake_mu_effect_sd,"_",v_d_lit$newt_mu_rate, "_", v_d_lit$newt_mu_effect_sd)

v_d_lit$sim=LETTERS[factor(v_d_lit$group)]
v_d_lit$dif <- v_d_lit$Snake_mean_Pheno-v_d_lit$Newt_mean_Pheno

v_d_lit$exp = "MuV = 2.5e-10 (low to high sigma: high to low mu)"
v_d_lit$change = "no"
file=v_d_lit
v_d_lit = find_coevo_slope(file, coevo_begin_start_gen, coevo_begin_end_gen, coevo_endlength_gen, close_percent)
sub_v_d_lit = subset(v_d_lit, v_d_lit$sim == c("F", "G", "J", "K") )

GA_stage = rbind(sub_mu_b_lit, sub_mu_c_lit, sub_v_b_lit, sub_v_c_lit)

#rbind(sub_mu_a_lit, sub_mu_b_lit, sub_mu_c_lit, sub_mu_d_lit, sub_v_a_lit, sub_v_b_lit, sub_v_c_lit, sub_v_d_lit)

GA_stage$snake_newt <- paste0(GA_stage$snake_mu_rate, "_", GA_stage$snake_mu_effect_sd, "_", GA_stage$newt_mu_rate, "_", GA_stage$newt_mu_effect_sd)

GA_stage$newt_coevo_slope = (GA_stage$N_pheno_end - GA_stage$N_pheno_start)/ (GA_stage$N_end_gen - GA_stage$N_str_gen)
GA_stage$snake_coevo_slope = (GA_stage$S_pheno_end - GA_stage$S_pheno_start)/ (GA_stage$S_end_gen - GA_stage$S_str_gen)

GA_stage$newt_pheno_change = (GA_stage$N_pheno_end - GA_stage$N_pheno_start)
GA_stage$snake_pheno_change = (GA_stage$S_pheno_end - GA_stage$S_pheno_start)

GA_stage$newt_gen_change = (GA_stage$N_end_gen - GA_stage$N_str_gen)
GA_stage$snake_gen_change = (GA_stage$S_end_gen - GA_stage$S_str_gen)
```

```{r, plotting all of the changes}
title = "Coevolution Slope (Phenotype Change/ Time)"
subt = "Simulation F, J, G, K"
GA_stage_long <- tidyr::pivot_longer(GA_stage, c(newt_coevo_slope, snake_coevo_slope))
y_label="Coevolution Speed"
x_label="Genetic Architectures (red=newt, blue=snake)"
ggplot(GA_stage_long) +
  ggtitle(title, subtitle = paste0(subt))+
  coord_trans(y = "log10")+
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor(sim, level=c("F","J","G","K")), y=(as.numeric(value)), fill=factor(name)), show.legend = FALSE)+
  theme(plot.title = element_text(size=5)) +
  #xlim(-1.1, 1.1)+
  facet_wrap(~exp, ncol=2, dir="v")+
  scale_fill_manual(values =c("darkred", "steelblue"), labels=c('Newt', 'Snake'))+ #
  #  labs(fill = " ")
  scale_x_discrete(labels=c('low/low', 'low/high', 'low/high','high/high'))+
  theme_bw() +  #theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                 #     title =element_text(size=8, face='bold'))+
    theme(legend.position = "none")

title = "Coevolution Phenotype Change"
subt = "Simulation F, J, G, K"
GA_stage_long <- tidyr::pivot_longer(GA_stage, c(newt_pheno_change, snake_pheno_change))
y_label="change in phenotype"
x_label="Genetic Architectures (red=newt, blue=snake)"
ggplot(GA_stage_long) +
  ggtitle(title, subtitle = paste0(subt))+
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor(sim, level=c("F","J","G","K")), y=(as.numeric(value)), fill=factor(name)), show.legend = FALSE)+
  theme(plot.title = element_text(size=5)) +
  #xlim(-1.1, 1.1)+
  facet_wrap(~exp, ncol=2, dir="v")+
  scale_fill_manual(values =c("darkred", "steelblue"), labels=c('Newt', 'Snake'))+ #
  #  labs(fill = " ")
  scale_x_discrete(labels=c('low/low', 'low/high', 'low/high','high/high'))+
  theme_bw() +  #theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                 #     title =element_text(size=8, face='bold'))+
    theme(legend.position = "none")

title = "Coevolution Change in Time"
subt = "Simulation F, J, G, K"
GA_stage_long <- tidyr::pivot_longer(GA_stage, c(newt_gen_change, snake_gen_change))
y_label="change in geneerations"
x_label="Genetic Architectures (red=newt, blue=snake)"
ggplot(GA_stage_long) +
  ggtitle(title, subtitle = paste0(subt))+
  xlab(x_label) + ylab(y_label)+
  geom_boxplot(aes(x=factor(sim, level=c("F","J","G","K")), y=(as.numeric(value)), fill=factor(name)), show.legend = FALSE)+ #
  theme(plot.title = element_text(size=5)) +
  #xlim(-1.1, 1.1)+
  facet_wrap(~exp, ncol=2, dir="v")+
  scale_fill_manual(values =c("darkred", "steelblue"), labels=c('Newt', 'Snake'))+ #
  #  labs(fill = " ")
  scale_x_discrete(labels=c('low/low', 'low/high', 'low/high','high/high'))+
  theme_bw() #+  #theme(axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
                 #     title =element_text(size=8, face='bold'))+
    #theme(legend.position = "none")

```